### 2024-03-05 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@lamalab-org}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2024-03-05 src/tex/acronymns.tex
start using showyourwork

\newacronym{llm}{LLM}{large language model}
\newacronym{ml}{ML}{machine learning}
\newacronym{agi}{AGI}{artificial general intelligence}
\newacronym{api}{API}{application programming interface}
\newacronym{mcq}{MCQ}{multiple-choice question}

### 2024-03-06 src/scripts/compute_basic_statistics.py
start using showyourwork

@@ -10,7 +10,7 @@ def load_data():
     # Load the data
     data = f"{chembench_repo}/data"
 
-    df = classify_questions(data)
+    df = classify_questions(data, debug=True)
 
     total_number_of_questions = len(df)
 
@@ -25,6 +25,9 @@ def load_data():
     manually_generated = df[~df["is_semiautomatically_generated"]]
     with open(output / "manually_generated.txt", "w") as f:
         f.write(str(len(manually_generated)))
+
+    df.to_csv(output / "questions.csv")
+    df.to_pickle(output / "questions.pkl")
 
 
 if __name__ == "__main__":

### 2024-03-06 src/scripts/plot_statistics.py

@@ -1,0 +1,62 @@
+from utils import ONE_COL_GOLDEN_RATIO_HEIGHT_INCH, ONE_COL_WIDTH_INCH
+import matplotlib.pyplot as plt
+from paths import output, figures, scripts
+
+plt.style.use(scripts / "lamalab.mplstyle")
+import pandas as pd
+
+
+import numpy as np
+from plotutils import range_frame
+
+
+def question_count_barplot(df):
+    fig, ax = plt.subplots(
+        1,
+        2,
+        figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH),
+        sharex=True,
+    )
+
+    df_mcq = df[df["is_classification"]]
+    df_not_mcq = df[~df["is_classification"]]
+
+    topics_mcq, counts_mcq = (
+        df_mcq["topic"].value_counts().index,
+        df_mcq["topic"].value_counts().values,
+    )
+    topics_general, counts_general = (
+        df_not_mcq["topic"].value_counts().index,
+        df_not_mcq["topic"].value_counts().values,
+    )
+    # ensure general topics are in the same order as mcq topics. Sort keys and values accordingly
+    topics_general, counts_general = zip(
+        *sorted(
+            zip(topics_general, counts_general), key=lambda x: topics_mcq.get_loc(x[0])
+        )
+    )
+
+    all_counts = np.concatenate([counts_mcq, counts_general])
+
+    ax[0].hlines(topics_mcq, xmin=0, xmax=topics_mcq, linewidth=5, alpha=0.2)
+    ax[0].plot(counts_mcq, topics_mcq, "o", markersize=5, alpha=0.6)
+    ax[0].set_xscale("log")
+    range_frame(ax[0], all_counts, np.arange(len(topics_general)))
+
+    ax[1].hlines(topics_general, xmin=0, xmax=topics_general, linewidth=5, alpha=0.2)
+    ax[1].plot(counts_general, topics_general, "o", markersize=5, alpha=0.6)
+    ax[1].set_xscale("log")
+    range_frame(ax[1], all_counts, np.arange(len(topics_general)))
+
+    ax[0].set_xlabel("Number of Questions")
+
+    fig.savefig(figures / "question_count_barplot.pdf", bbox_inches="tight")
+
+
+def plot_question_statistics():
+    df = pd.read_pickle(output / "questions.pkl")
+    question_count_barplot(df)
+
+
+if __name__ == "__main__":
+    plot_question_statistics()

### 2024-03-06 src/scripts/utils.py

@@ -1,11 +1,27 @@
 import os
+from paths import data
+
+
+from scipy.constants import golden
+
+ONE_COL_WIDTH_INCH = 3.25
+TWO_COL_WIDTH_INCH = 7.2
+
+ONE_COL_GOLDEN_RATIO_HEIGHT_INCH = ONE_COL_WIDTH_INCH / golden
+TWO_COL_GOLDEN_RATIO_HEIGHT_INCH = TWO_COL_WIDTH_INCH / golden
 
 
 def obtain_chembench_repo():
     # if in ../data the repository chem-bench is not found, clone it
     # if it is found, pull the latest version
-    if not os.path.exists("../data/chem-bench"):
-        os.system("git clone https://github.com/lamalab-org/chem-bench-paper.git")
+    cwd = os.getcwd()
+    if not os.path.exists(os.path.join(data, "chem-bench")):
+        os.chdir(data)
+        os.system("git clone https://github.com/lamalab-org/chem-bench.git --depth 1")
+        os.system("git checkout -b humanbench")
     else:
-        os.system("cd ../data/chem-bench; git pull")
-    return "../data/chem-bench"
+        os.chdir(os.path.join(data, "chem-bench"))
+        os.system("git checkout -b humanbench")
+        os.system("git pull")
+    os.chdir(cwd)
+    return os.path.join(data, "chem-bench")

### 2024-03-07 src/tex/safety_data.tex
Create safety_data.tex

\subparagraph{GHS classification, hazard statements and daily allowed intakes}

The GHS classification, hazard statements and daily allowed intake (DAI) data has been extracted from PubChem. 
The former two have been mined through python, while the latter has been manually acquired.
The DAIs have been curated to contain only records approved by the World Health Organization (WHO). 
The chemicals in this class of questions belong to one of the three classes: pesticides (e.g. calcium arsenate), insecticides (e.g. cyfluthrin) or herbicides (e.g. 2,4 - D).

### 2024-03-07 src/tex/safety_data.tex
Update safety_data.tex

@@
 
-The GHS classification, hazard statements and daily allowed intake (DAI) data has been extracted from PubChem. 
-The former two have been mined through python, while the latter has been manually acquired.
+The GHS classification, hazard statements and daily allowed intake (DAI) data have been extracted from PubChem. \cite{pubchem}
+The former two have been mined through python, while the latter has been manually acquired. The questions in their respective category are formatted in the same way.
 The DAIs have been curated to contain only records approved by the World Health Organization (WHO). 

### 2024-03-07 src/scripts/compute_basic_statistics.py
update paper

@@ -15,17 +15,17 @@ def load_data():
     total_number_of_questions = len(df)
 
     with open(output / "total_number_of_questions.txt", "w") as f:
-        f.write(str(total_number_of_questions))
+        f.write(str(total_number_of_questions) + "\endinput")
 
     automatically_generated = df[df["is_semiautomatically_generated"]]
 
     with open(output / "automatically_generated.txt", "w") as f:
-        f.write(str(len(automatically_generated)))
+        f.write(str(len(automatically_generated)) + "\endinput")
 
     manually_generated = df[~df["is_semiautomatically_generated"]]
     with open(output / "manually_generated.txt", "w") as f:
-        f.write(str(len(manually_generated)))
- 
+        f.write(str(len(manually_generated)) + "\endinput")
+
     df.to_csv(output / "questions.csv")
     df.to_pickle(output / "questions.pkl")
 

### 2024-03-07 src/tex/safety_data.tex
Update safety_data.tex
Update src/tex/safety_data.tex

@@
 
-The GHS classification, hazard statements and daily allowed intake (DAI) data have been extracted from PubChem. \cite{pubchem}
-The former two have been mined through python, while the latter has been manually acquired. The questions in their respective category are formatted in the same way.
+The GHS classification, hazard statements and  \gls{dai} data have been extracted from PubChem. \cite{pubchem}
+The former two have been mined through python, while the latter has been manually acquired.
 The DAIs have been curated to contain only records approved by the World Health Organization (WHO). 
 The chemicals in this class of questions belong to one of the three classes: pesticides (e.g. calcium arsenate), insecticides (e.g. cyfluthrin) or herbicides (e.g. 2,4 - D). 
+All data was saved in tabular form and then we programatically created a large set of questions (87 hazard statement definitions, 978 chemical-GHS pictogram matching, 100 DAIs).

### 2024-03-07 src/tex/authors.tex

\author[1]{Nawaf~Alampara~\orcidlink{}}

\author[1]{Adrian~Mirza~\orcidlink{}}

\author[1]{Sreekanth~Kunchapu~\orcidlink{}}

\author[]{Aswanth~Krishnan}

\author[1]{Mara~Willhelmi}

\author[1]{Macjonathan~Oreke}

\author[]{Benedict~Emoekabu}

\author[5]{Tanya~Gupta}

\author[5]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}

\author[]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}

\author[1,2, 3]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
\affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
\affil[2]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
\affil[3]{Helmholtz Institute for Polymers in Energy Applications (HIPOLE), Philosophenweg 7a, 07743 Jena, Germany}
\affil[]{\texttt{mail@kjablonka.com}}

### 2024-03-07 src/tex/safety_data.tex

@@
 The chemicals in this class of questions belong to one of the three classes: pesticides (e.g. calcium arsenate), insecticides (e.g. cyfluthrin) or herbicides (e.g. 2,4 - D). 
-All data was saved in tabular form and then we programatically created a large set of questions (87 hazard statement definitions, 978 chemical-GHS pictogram matching, 100 DAIs).
+All data was saved in tabular form and then we programmatically created questions (87 hazard statement definitions, 978 chemical-GHS pictogram matching, 100 DAIs).

### 2024-03-08 src/scripts/plot_statistics.py
update paper

@@ -4,7 +4,7 @@ from paths import output, figures, scripts
 
 plt.style.use(scripts / "lamalab.mplstyle")
 import pandas as pd
-
+from glob import glob
 
 import numpy as np
 from plotutils import range_frame

### 2024-03-08 src/tex/showyourwork-metadata.tex
rebuild

        \OnGithubActionsfalse
        \def\syw@url{https://github.com/lamalab-org/chembench-paper}
    \def\syw@slug{lamalab-org/chembench-paper}
    \def\syw@sha{9e3e1d5a5f3dc81ac1ab792b79e58cce26f6bbda}
    \def\syw@runid{}

        \AddStampfalse
        \newcommand{\syw@stampX}{0.5}
    \newcommand{\syw@stampY}{0.5}
    \newcommand{\syw@stampSize}{0.75}
    \newcommand{\syw@stampAngle}{-20.0}
    \newcommand{\syw@stampText}{{\faGithub}/lamalab-org/chembench-paper}
    \newcommand{\syw@stampVersion}{0.4.2}

        \addvalue{fig:topic_barplot_script}{src/scripts/plot_statistics.py}
        \addvalue{fig:wordcloud_script}{src/scripts/wordcloud.py}
        \addvalue{fig:human_timing_script}{src/scripts/analyze_human_data.py}
        \addvalue{fig:experience_vs_correctness_script}{src/scripts/analyze_human_data.py}

        \addvalue{output/total_number_of_questions.txt_rule}{Snakefile\#L1}
        \addvalue{output/manually_generated.txt_rule}{Snakefile\#L1}
        \addvalue{output/automatically_generated.txt_rule}{Snakefile\#L1}
        \addvalue{output/number_experts.txt_rule}{Snakefile\#L20}
        \addvalue{output/total_hours.txt_rule}{Snakefile\#L20}
        \addvalue{output/flesch_kincaid_reading_ease.txt_rule}{Snakefile\#L30}
        \addvalue{output/num_h_statements.txt_rule}{Snakefile\#L1}
        \addvalue{output/num_pictograms.txt_rule}{Snakefile\#L1}
        \addvalue{output/num_dai.txt_rule}{Snakefile\#L1}
        \addvalue{output/spearman_experience_score.txt_rule}{Snakefile\#L20}
        \addvalue{output/spearman_experience_score_p.txt_rule}{Snakefile\#L20}

### 2024-03-08 src/scripts/analyze_model_reports.py
start note about tiny benchmarks

@@ -1,0 +1,38 @@
+from chembench.analysis import load_all_reports
+from utils import obtain_chembench_repo
+import os
+
+
+def load_reports():
+    claude2 = load_all_reports(
+        "../reports/claude2/reports/7536581a-d92d-43d3-b946-39e4a7213b7f/"
+    )
+    claude2_zero_t = load_all_reports(
+        "../reports/claude2-zero-T/reports/412657cc-7a11-4d73-80e9-03d6f05cd63e"
+    )
+
+    fb_llama_70b = load_all_reports(
+        "../reports/llama-2-70b-chat/reports/3f8361fa-8fbf-4704-bccd-b0489f401d4e"
+    )
+
+    galactica_120b = load_all_reports(
+        "../reports/galatica_120b/reports/82194130-de35-49b8-8375-ac89e06542ba"
+    )
+
+    gemini_pro = load_all_reports(
+        "../reports/gemini-pro/reports/ebde051c-6d66-456a-a207-a1c65eceaf40"
+    )
+    gemini_pro_zero_t = load_all_reports(
+        "../reports/gemin-pro-zero-T/reports/1e5457ad-96b5-4bc8-bd6c-bad3eb6deb7a"
+    )
+
+    gpt35turbo = load_all_reports(
+        "../reports/gpt-3.5-turbo/reports/f76bf17d-3e12-47c5-b879-9cc0c78be989"
+    )
+
+    gpt4 = load_all_reports(
+        "../reports/gpt-4/reports/76c5bdd4-e893-43d4-b37d-2ade66c20308"
+    )
+    gpt4_zero_t = load_all_reports(
+        "../reports/gpt-4-zero-T/reports/8bbc0722-c86f-4388-9d84-eb8a1a75a72f"
+    )

### 2024-03-08 src/tex/authors.tex
start author cleanup

@@
-\author[1]{Nawaf~Alampara~\orcidlink{}}
+%Conceptualization
+%Data curation
+%Formal analysis - Application of statistical, mathematical, computational, or other formal techniques to analyse or synthesize study data.
+%Funding acquisition
+%Investigation - Conducting a research and investigation process, specifically performing the experiments, or data/evidence collection.
+%Methodology
+%Project administration
+%Resources
+%Software
+%Supervision
+%Validation
+%Visualization
+%Writing -- original draft
+%Writing -- review & editing
 
-\author[1]{Adrian~Mirza~\orcidlink{}}
+\credit{Adrian~Mirza}{0,1,1,0,1,0,0,0,1,0,0,0,0,0}
+\credit{Sreekanth~Kunchapu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Jan Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Lea Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+%ORCID ID:  0000-0002-5968-8706 Affiliation:    Institute for Technical Chemistry and Environmental Chemistry (ITUC) and Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany
+\credit{Mara~Wilhemli}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Macjonathan~Oreke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Michael~Pieler}{1,0,0,0,0,0,0,0,0,0,0,0,a0,0}
+\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,0}
+\credit{Philippe~Schwaller}{1,1,0,1,1,0,0,0,0,1,0,0,0,0}
+\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
+\credit{Berend~Smit}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
+\credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}
 
-\author[1]{Sreekanth~Kunchapu~\orcidlink{}}
+\author[1,*]{Nawaf~Alampara~\orcidlink{}}
 
-\author[]{Aswanth~Krishnan}
+\author[1,*]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
+
+\author[1,*]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
+
+\author[1]{Lea~Klepsch~\orcidlink{0000-0003-4978-4670}}
+\author[1]{Jan~Peschel~\orcidlink{0009-0002-4787-2757}}
+\author[1]{Jakob~Meyer}
+\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
+\author[1]{Maximilian~Greiner}
+\author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
+
+\author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[ituc]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[1]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
+\author[dresden]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
+\author[lsmo]{Amir~Elahi}
+\author[lsmo]{Berend~Smit~\orcidlink{0000-0003-4653-8562}}
+
+\author[qpi]{Aswanth~Krishnan}
 
@@
 
+\author[Occlu]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
+
 \author[]{Benedict~Emoekabu}
 
-\author[5]{Tanya~Gupta}
+\author[nccrcat,liac]{Tanya~Gupta}
 
-\author[5]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
+\author[nccrcat,liac]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
 
-\author[]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[open,stability]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
 
 \author[1,2, 3]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+
 \affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
@@
 \affil[3]{Helmholtz Institute for Polymers in Energy Applications (HIPOLE), Philosophenweg 7a, 07743 Jena, Germany}
+
+\affil[liac]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland.}
+\affil[nccrcat]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland.}
+
+\affil[lsmo]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Valais, Switzerland.}
+\affil[dresden]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
+\affil[open]{OpenBioML.org}
+\affil[stability]{Stability.AI}
+\affil[ituc]{Institute for Technical Chemistry and Environmental Chemistry (ITUC) and Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[qpi]{Qpi volta}
+\affil[occlu]{Occlutech GmbH, Jena, Germany}
+
 \affil[]{\texttt{mail@kjablonka.com}}
+\affil[*]{These authors contributed equally.}

### 2024-03-09 src/tex/ms.tex
blacktex

@@
-% Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[11pt, oneside]{article}
+
+\input{preamble.tex}
+\usepackage{credits}
+\usepackage{orcidlink}
 \usepackage{showyourwork}
 
-% Begin!
+\title{\textsf{Are frontier models superhuman chemists?}}
+
+\input{authors.tex}
 \begin{document}
+\maketitle
 
-% Title
-\title{An open source scientific article}
-
-% Author list
-\author{@lamalab-org}
-
-% Abstract with filler text
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
 \end{abstract}
 
-% Main body with filler text
+\clearpage
+
 \section{Introduction}
-\label{sec:intro}
+\Glspl{llm} are frontier \gls{ml} models trained on massive amounts of text to complete sentences.
+While some see in them \enquote{sparks of \gls{agi}},\cite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\cite{bender2021dangers}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, some even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
+This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can physically perform reactions provided a command in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+The rapid increase in capabilities led to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022}
+Moreover, the use of these models is now also widespread among students\cite{Intelligent.com_2023} as well as research groups.
+For some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
+Unfortunately, apart from anecdotal reports there is little evidence on how \glspl{llm} perform compared to experts.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, a comprehensive analysis is needed.
+For the development of \glspl{llm}, such evaluation is currently mostly performed via standardized benchmarks such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
+The former contains, among 204 tasks, only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
+Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
+While those evaluations can measure how well models can make predictions for very specific tasks, they only give a poor measure of how useful those models might be as a chemical assistant.
 
-\bibliography{bib}
+While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, also those do not satisfy the following basic criteria chemistry benchmarks should satisfy:
+\begin{itemize}
+    \item \emph{End-to-end automation}. For model development, the evaluations will need to be run many times (e.g., on regular intervals of a training run).
+    Approaches that rely on humans scoring the answers of a system\cite{Schulze_Balhorn_2024, ai4science2023impact} can thus not be used.
+    \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\cite{northcutt2021pervasive}
+    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\cite{Frye_2023, Awg}
+    \item \emph{Usable with models that support special treatment of molecules}.
+Some models such as Galactica\cite{taylor2022galactica} use special tokenization or encoding procedures for molecules or equations.
+    To support this, the benchmark system must encode the semantic meaning of various parts of the question or answer.
+    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or even just the raw logits.
+    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
+    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
+    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real world chemistry as well as higher-level university education multiple choice question are seldom utilized.
+    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond the answering of \gls{mcq}.
+    \item \emph{Cover a diverse set of topics}.
+Chemistry, as the \enquote{central science}, bridges multiple disciplines.\cite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
+\end{itemize}
+
+In this work, we report a novel benchmarking framework, chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+It covers a large fraction of the topics taught in undergraduate chemistry curricula at various skill levels and can be used with any system that can return text (i.e., also tool-augmented systems).
+
+To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry (for a total of number of more than \variable{output/total_hours.txt}\unskip) on a subset of the benchmark corpus to be able to compare the performance of current frontier models with the one of humans.
+Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, included safety-related ones, might be very misleading.
+
+\section{Results}
+
+\subsection{Benchmark dataset}
+
+To compile our benchmark corpus we utilized a broad list of sources (\Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
+To ensure maximal interoperability, we curated the data in an extended form of the BigBench format.
+This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
+For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
+
+Importantly, our large pool of questions encompassed a wide range of topics.
+This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different categories (see \Cref{sec:meth-topic} for details on how we assigned topics).
+By design, a focus of our corpus is on safety-related aspects with a (currently) limited sampling of questions from physical or theoretical chemistry, which might be extended in future work.
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/question_count_barplot.pdf}
+    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question.
+The figure shows that not all aspects of chemistry are equally represented in our corpus.
+The corpus currently focuses on safety-related aspects.}
+    \label{fig:topic_barplot}
+    \script{plot_statistics.py}
+\end{figure}
+
+Importantly, the corpus samples both \gls{mcq} and open-ended questions in a balanced way (XX \gls{mcq} questions vs. XX open-ended questions).
+As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).~\cite{kincaid1975derivation}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/wordcloud.pdf}
+    \caption{\textbf{Wordcloud.}}
+    \label{fig:wordcloud}
+    \script{wordcloud.py}
+\end{figure}
+
+\subsection{\enquote{Tiny} subset}
+It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
+For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark.
+
+\subsection{Model evaluation}
+
+\begin{figure}
+    \centering
+    \caption{Caption}
+    \label{fig:enter-label}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
+    \caption{Caption}
+    \label{fig:enter-label}
+\end{figure}
+
+\section{Conclusions}
+
+\subsection{Future work}
+- Advanced prompting techniques such as CoT and test time strategies
+
+- More advanced systems with chemnistry specific tools
+
+- More systematic human baseline: Ask people to come in, do not allow to skip questions
+
+\section{Methods}
+
+\subsection{Curation workflow}\label{sec:curation}
+For our dataset we curated questions from existing exams, but also programmatically created new questions.
+Questions were added via Pull Requests on our GitHub repository and only merged into the main collection after passing manual review.
+
+To ensure that the questions do not enter a training dataset we use the canary string of the Big Bench project.
+This requires that \Gls{llm} developers filter their training dataset for this canary string.
+
+\begin{figure}
+    \includegraphics[width = \textwidth]{figures/chem-bench.pdf}
+        \caption{\textbf{Data curation workflow}.}
+\end{figure}
+
+\paragraph{Manually curated questions}
+
+\paragraph{Semi-programatically generated questions}
+
+\subparagraph{Oxidation states}
+
+\subparagraph{Electron counts}
+
+\paragraph{Programmatically generated questions}
+\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
+We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
+
+\subparagraph{GHS, hazard statements and DAI}
+The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.~\cite{pubchem}
+The former two have been mined via the PubChem \gls{api}, while the latter has been manually compiled.
+The \glspl{dai} have been curated to contain only records approved by the \gls{who}.
+The chemicals in this class of questions belong to one of the three classes: pesticides (e.g., calcium arsenate), insecticides (e.g., cyfluthrin) or herbicides (e.g. 2,4-D).
+All data was saved in tabular form, and then we programmatically created questions (\variable{output/num_h_statements.txt} hazard statement definitions, \variable{output/num_pictograms.txt} chemical-\gls{ghs} pictogram matching, \variable{output/num_dai.txt} \glspl{dai}).
+
+\subparagraph{Number of isomers}
+We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled SMILES from the ZINC dataset.
+To obtain the answers for false options we sampled from the set of number of isomers.
+
+\subsection{Model evaluation workflow}
+
+\paragraph{Prompting}
+
+\subparagraph{Completion models}
+\texttt{Galactica-120b}\cite{taylor2022galactica}
+
+\subparagraph{Instruction-tuned models} \texttt{Claude 2} \texttt{Claude3 (Opus)}\cite{anthropicClaudeModelFamily2024}
+\texttt{GPT-4}\cite{openai2024gpt4}
+\texttt{GPT-3.5-turbo}\cite{brown2020language}
+\texttt{Gemini Pro}\cite{gemini}
+\texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
+\texttt{Llama2-70b}\cite{touvron2023llama}
+\texttt{pplx-7b-chat}
+
+\paragraph{Parsing}
+Our parsing workflow is, by default, multistep and primarily based on regular expressions.
+In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
+In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
+In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
+As initial tests indicated that models sometimes return integers in the form of words, e.g. \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion.
+In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
+The frequency of this fallback being triggered was very different for different models (see XX).
+Manual verification (see \Cref{sec:manually-verified-parsing}) indicates that this \gls{llm}-based parsing is not a relevant error source.
+
+\paragraph{Models}
+
+\subparagraph{Tool augmented models}
+In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
+For this we, on the one hand, investigated the \texttt{online} model of Perplexity.AI and, on the other hand, \texttt{gpt-3.5-turbo} as well as \texttt{claude-2}, with ReAct-style tool augmentation.\cite{yao2023react}
+The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
+We implemented the system using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
+
+\subsection{Human baseline}
+
+\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\cite{nextjs} app router framework.
+This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
+We utilize a MySQL\cite{mysql} database and Prisma \gls{orm}\cite{prisma} for efficient database management.
+The web application is styled with Tailwind CSS\cite{tailwindcss} using the shadcn/ui component library and uses NextAuth\cite{nextauth} for easy and secure user authentication and postMark for sending Emails.
+The application is hosted on the Vercel web hosting platform.
+
+\paragraph{Question selection}
+Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
+For selecting the subset, we decided on addressing two questions:
+\begin{itemize}
+    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
+    \item Are there areas in which the performance of humans is very different from the ones of the models?
+\end{itemize}
+To answer the first question we selected X questions that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
+From those we picked X diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART (see below).
+
+\paragraph{Study design}
+For our initial study we wanted to maximize the response rate given our available resources.
+For this reason, we did not opt for a highly controlled study setting.
+That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+Note that users were also allowed to skip questions.
+
+\subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset we systematically recorded keywords and sources.
+To allow for analysis of the model performance as a function of the topic, we leverage this information together with sequence classification models.
+For questions which can easily be assigned to a topic based on the source (e.g., number of NMR peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
+For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\cite{zeroshotsequence} using the BART model\cite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
+
+\section*{Data and code availability}
+
+\section*{Acknowledgements}
+This work was supported by the Carl Zeiss Foundation and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
+We also want to acknowledge access to the HPC cluster of Stability.AI.
+
+\section*{Conflicts of interest}
+K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.ai.
+
+\section*{Author contributions}
+
+\bibliography{references}
+
+\appendix
+
+\subsection{Benchmark corpus}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+    \caption{Example question.}
+    \label{fig:flesch_kincaid_reading_ease}
+\end{figure}
+
+\subsection{Parsing verification}\label{sec:manually-verified-parsing}
+For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
+
+\subsection{Model performance}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/all_questions_models_completely_correct_radar.pdf}
+    \caption{Caption}
+    \label{fig:all_questions_models_completely_correct_radar}
+    \script{analyze_model_reports.py}
+\end{figure}
+
+\subsection{Human baseline}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/human_timing.pdf}
+    \script{analyze_human_data.py}
+    \label{fig:human_timing}
+    \caption{}
+\end{figure}
+
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_\experience_score.txt}\), \(p \approx \variable{output/spearman_\experience_score_p.txt}\)).
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/experience_vs_correctness.pdf}
+    \script{analyze_human_data.py}
+    \label{fig:experience_vs_correctness}
+    \caption{}
+\end{figure}
 

### 2024-03-09 src/scripts/analyze_model_reports.py

@@ -1,4 +1,8 @@
-from chembench.analysis import load_all_reports
+from chembench.analysis import (
+    load_all_reports,
+    get_human_scored_questions_with_at_least_n_scores,
+    all_correct,
+)
 from utils import obtain_chembench_repo
 import os
 from plotutils import radar_factory, model_color_map
@@ -7,6 +11,196 @@ import matplotlib.pyplot as plt
 from paths import figures, output, scripts
 
 plt.style.use(scripts / "lamalab.mplstyle")
+
+
+def combine_scores_for_model(
+    folder, datafolder, human_baseline_folder=None, min_human_responses: int = 100
+):
+
+    df = load_all_reports(folder, datafolder)
+    if human_baseline_folder is not None:
+        relevant_questions = get_human_scored_questions_with_at_least_n_scores(
+            human_baseline_folder, min_human_responses
+        )
+        df = df[df[("name", 0)].isin(relevant_questions)]
+    df["all_correct"] = df.apply(all_correct, axis=1)
+
+    return df
+
+
+def load_human_aligned_reports():
+    chembench = obtain_chembench_repo()
+    human_baseline_folder = os.path.join(chembench, "reports/humans")
+    datafolder = os.path.join(chembench, "data")
+
+    claude2 = combine_scores_for_model(
+        os.path.join(
+            chembench, "reports/claude2/reports/7536581a-d92d-43d3-b946-39e4a7213b7f"
+        ),
+        datafolder,
+        human_baseline_folder,
+    )
+
+    claude2_react = combine_scores_for_model(
+        os.path.join(
+            chembench,
+            "reports/claude2-react/reports/700ed9b4-995c-4cfd-aa63-0a6e84b3a815",
+        ),
+        datafolder,
+        human_baseline_folder,
+    )
+
+    claude2_zero_t = combine_scores_for_model(
+        os.path.join(
+            chembench,
+            "reports/claude2-zero-T/reports/412657cc-7a11-4d73-80e9-03d6f05cd63e",
+        ),
+        datafolder,
+        human_baseline_folder,
+    )
+
+    claude3 = combine_scores_for_model(
+        os.path.join(
+            chembench, "reports/claude3/reports/702e03be-5cd8-4451-b52c-8d7b9b694304"
+        ),
+        datafolder,
+        human_baseline_folder,
+    )
[…]

### 2024-03-11 src/tex/ms.tex

@@
 
+To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
+We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Graf{sec:parsing}.
+Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
+For example, all the SMILES representations are encapsulated within \texttt{[START_SMILES][\text backslash END_SMILES]} in Galactica\cite{taylor2022galactica}.
+Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \latex syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
+
 \subparagraph{Completion models}

### 2024-03-11 src/tex/ms.tex

@@
 
+\subparagraph{Analytical chemistry}
+The questions are sourced from the master student examination paper at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
+
+\subparagraph{Combustion engineering}
+These questions are derived from the master student examination paper at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
+
+\subparagraph{Functional materials and nano-materials}
+The questions are gathered from a seminar conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
+
+\subparagraph{General chemistry}
+The questions originate from examinations for grade X at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+
+\subparagraph{International Chemical Olympiad (\gls{ICHO})}
+The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova, providing challenging problems for competitive chemistry.
+
+\subparagraph{Material synthesis}
+These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
+
+\subparagraph{Organic reactivity}
+The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
+
+\subparagraph{Periodic table}
+The questions are framed based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
+
+\subparagraph{Polymer chemistry}
+Questions are sourced from examinations held at the University of Hamburg, covering topics related to polymer chemistry and its applications.
+
+\subparagraph{Chemical reactivity}
+These questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
+
+\subparagraph{Chemical safety}
+The questions encompass various aspects of safety, including lab safety, toxicology, material safety data sheets, carcinogens, sourced from reputable safety resources.
+
+\subparagraph{Shape of molecules}
+The questions are sourced from examinations held at the Technical University of Munich, Germany, focusing on determining the shape and geometry of molecules.
+
+\subparagraph{\gls{NMR} spectroscopy}
+Questions are sourced from the twitter page called (\url{https://twitter.com/NMRspectroscopy}), focusing on concepts and applications of nuclear magnetic resonance spectroscopy.
+
 \paragraph{Semi-programatically generated questions}
@@
 \subparagraph{Oxidation states}
-
-\subparagraph{Electron counts}
+The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}.
+These exercises focus on determining the number of isomers for specific molecules.
+To perform these calculations, we utilize the RDKit package \cite{bento2020open}.
+\subparagraph{Electron count of atoms}
+Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}.
+These inquiries involve determining the electron count of atoms in various chemical contexts.
+The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
+\subparagraph{Number of isomers}
+The questions concerning the number of isomers are sourced from the website \url{https://www.cheminfo.org/}.
+These exercises aim to determine the number of isomers for given molecules.
+The calculations for these assessments are conducted using the RDKit package \cite{bento2020open}.
+\subparagraph{Point group for molecules}
+Questions involve predicting the point group for molecules.
+The ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption/tree/main}) is utilized for data generation.
+Given a molecule, predict its point group.
+\subparagraph{SMILES-IUPAC}
+Web scrapped SMILES IUPAC pairs from the PubChem database \cite{pubchem} and then framed questions involve converting SMILES notation to IUPAC nomenclature, utilizing information from the PubChem database.
 

### 2024-03-11 src/tex/ms.tex
feat: update intro and analysis

--- 2024-03-09 src/tex/ms.tex
@@
 While some see in them \enquote{sparks of \gls{agi}},\cite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\cite{bender2021dangers}
+In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they can solve a wide variety of tasks they have not been explicitly trained on.\cite{bommasani2021opportunities, anderljung2023frontier}
+In addition, there is tremendous economic interest and investment in those generative models is said to lead to a market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\cite{bloomberg}
 
@@
 This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can physically perform reactions provided a command in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
-The rapid increase in capabilities led to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022}
-Moreover, the use of these models is now also widespread among students\cite{Intelligent.com_2023} as well as research groups.
-For some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
+But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text. 
+Most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
+Most of these insights are in form of text in scientific publications. 
+Thus, operating on such text might be our best way of \enquote{unlocking} and learning from these insights.
+This might ultimately lead to \enquote{copilots} for chemists that can provide answers to questions or even suggest new experiments.
+However, the rapid increase in capabilities of chemical \gls{ml} models, led to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022}
+To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
+
+It is important to realize that the user base of such models is not only limited to experts in chemistry and materials science. 
+For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\cite{Intelligent.com_2023}
+This also implies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
+Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
 Unfortunately, apart from anecdotal reports there is little evidence on how \glspl{llm} perform compared to experts.
 
-Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, a comprehensive analysis is needed.
+Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, evaluation frameworks are needed that allow to systematically measure progress and mitigate potential harms.
 For the development of \glspl{llm}, such evaluation is currently mostly performed via standardized benchmarks such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
@@
 Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
-While those evaluations can measure how well models can make predictions for very specific tasks, they only give a poor measure of how useful those models might be as a chemical assistant.
+While those evaluations can measure how well models can make predictions for very specific tasks (e.g., predicting solubility for a given molecule reported as \gls{smiles}), they only give a poor measure of how useful those models might be as a chemical assistant.
 
+%ToDo: this is too technical and should be moved to the methods
 While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, also those do not satisfy the following basic criteria chemistry benchmarks should satisfy:
@@
 
-To compile our benchmark corpus we utilized a broad list of sources (\Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-To ensure maximal interoperability, we curated the data in an extended form of the BigBench format.
+To compile our benchmark corpus we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
+To ensure maximal interoperability, we curated the data in an extended form of the widely used BigBench format.
 This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
 
@@
 By design, a focus of our corpus is on safety-related aspects with a (currently) limited sampling of questions from physical or theoretical chemistry, which might be extended in future work.
+For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
 
@@
 
-Importantly, the corpus samples both \gls{mcq} and open-ended questions in a balanced way (XX \gls{mcq} questions vs. XX open-ended questions).
+Importantly, the corpus samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
+\variable{output/non_mcq_questions.txt} open-ended questions).
 As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).~\cite{kincaid1975derivation}
@@
 It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
-For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark.
+For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
+To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
+We also used this subset to seed the app we used for the human baseline study.
 
@@
 
+To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
+We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Cref{sec:parsing}.
+Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
+For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\cite{taylor2022galactica}.
+Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
+
 \subparagraph{Completion models}
@@
 
-\paragraph{Question selection}
+\paragraph{Question selection} \label{sec:subset-selection}
 Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
@@
 \section*{Data and code availability}
+The code and data for chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
+The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
+To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{showyourwork!} framework.\cite{Luger2021}
+The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}.
 
@@
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.ai.
+K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.AI.
 
 \section*{Author contributions}
+
+%\insertcredits
 
@@
 
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_\experience_score.txt}\), \(p \approx \variable{output/spearman_\experience_score_p.txt}\)).
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 

### 2024-03-11 src/tex/ms.tex
feat: add acronyms, fix credits bug, update pipeline
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-03-11 src/tex/ms.tex
@@
 The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
-\subparagraph{Number of isomers}
-The questions concerning the number of isomers are sourced from the website \url{https://www.cheminfo.org/}.
-These exercises aim to determine the number of isomers for given molecules.
-The calculations for these assessments are conducted using the RDKit package \cite{bento2020open}.
 \subparagraph{Point group for molecules}

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{Analytical chemistry}
-The questions are sourced from the master student examination paper at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
+The questions are sourced from master student examination papers at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
 
 \subparagraph{Combustion engineering}
-These questions are derived from the master student examination paper at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
+These questions are derived from master student examination papers at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
 

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{Functional materials and nano-materials}
-The questions are gathered from a seminar conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
+The questions are gathered from a seminar course conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
 

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{General chemistry}
-The questions originate from examinations for grade X at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+The questions originate from 10th grade examinations at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
 
 \subparagraph{International Chemical Olympiad (\gls{ICHO})}
-The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova, providing challenging problems for competitive chemistry.
+The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
 

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{Periodic table}
-The questions are framed based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
+The questions are manually created based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
 

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{Material synthesis}
-These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
+These questions are derived from seminars on material synthesis at the University of Jena (Germany), focusing on the methods and techniques employed in the synthesis of various materials.
 

### 2024-03-11 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 \subparagraph{Organic reactivity}
-The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
+The questions are primarily sourced from tutorial papers from the Technical University of Munich (Germany) exploring reactions and mechanisms in organic chemistry.
 

### 2024-03-12 src/tex/appendix.tex
feat: update manuscript, fix analysis bugs

--- 2024-03-11 src/tex/ms.tex
@@
-\documentclass[11pt, oneside]{article}
+\subsection{Desired properties of a chemistry benchmark} \label{sec:desired-properties}
 
-\input{preamble.tex}
-\usepackage{credits}
-\usepackage{orcidlink}
-\usepackage{showyourwork}
-
-\title{\textsf{Are frontier models superhuman chemists?}}
-
-\input{authors.tex}
-\begin{document}
-\maketitle
-
-\begin{abstract}
-\end{abstract}
-
-\clearpage
-
-\section{Introduction}
-\Glspl{llm} are frontier \gls{ml} models trained on massive amounts of text to complete sentences.
-While some see in them \enquote{sparks of \gls{agi}},\cite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\cite{bender2021dangers}
-
-Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, some even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can physically perform reactions provided a command in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
-The rapid increase in capabilities led to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022}
-Moreover, the use of these models is now also widespread among students\cite{Intelligent.com_2023} as well as research groups.
-For some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
-Unfortunately, apart from anecdotal reports there is little evidence on how \glspl{llm} perform compared to experts.
-
-Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, a comprehensive analysis is needed.
-For the development of \glspl{llm}, such evaluation is currently mostly performed via standardized benchmarks such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
-The former contains, among 204 tasks, only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
-Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
-While those evaluations can measure how well models can make predictions for very specific tasks, they only give a poor measure of how useful those models might be as a chemical assistant.
-
-While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, also those do not satisfy the following basic criteria chemistry benchmarks should satisfy:
 \begin{itemize}
@@
 
-In this work, we report a novel benchmarking framework, chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
-It covers a large fraction of the topics taught in undergraduate chemistry curricula at various skill levels and can be used with any system that can return text (i.e., also tool-augmented systems).
-
-To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry (for a total of number of more than \variable{output/total_hours.txt}\unskip) on a subset of the benchmark corpus to be able to compare the performance of current frontier models with the one of humans.
-Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, included safety-related ones, might be very misleading.
-
-\section{Results}
-
-\subsection{Benchmark dataset}
-
-To compile our benchmark corpus we utilized a broad list of sources (\Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-To ensure maximal interoperability, we curated the data in an extended form of the BigBench format.
+\subsection{Benchmark corpus}
+To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\cite{srivastava2022beyond}
 This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
-
-Importantly, our large pool of questions encompassed a wide range of topics.
-This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different categories (see \Cref{sec:meth-topic} for details on how we assigned topics).
-By design, a focus of our corpus is on safety-related aspects with a (currently) limited sampling of questions from physical or theoretical chemistry, which might be extended in future work.
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/question_count_barplot.pdf}
-    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question.
-The figure shows that not all aspects of chemistry are equally represented in our corpus.
-The corpus currently focuses on safety-related aspects.}
-    \label{fig:topic_barplot}
-    \script{plot_statistics.py}
-\end{figure}
-
-Importantly, the corpus samples both \gls{mcq} and open-ended questions in a balanced way (XX \gls{mcq} questions vs. XX open-ended questions).
-As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).~\cite{kincaid1975derivation}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/wordcloud.pdf}
-    \caption{\textbf{Wordcloud.}}
-    \label{fig:wordcloud}
-    \script{wordcloud.py}
-\end{figure}
-
-\subsection{\enquote{Tiny} subset}
-It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
-For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark.
-
-\subsection{Model evaluation}
-
-\begin{figure}
-    \centering
-    \caption{Caption}
-    \label{fig:enter-label}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
-    \caption{Caption}
-    \label{fig:enter-label}
-\end{figure}
-
-\section{Conclusions}
-
-\subsection{Future work}
-- Advanced prompting techniques such as CoT and test time strategies
-
-- More advanced systems with chemnistry specific tools
-
-- More systematic human baseline: Ask people to come in, do not allow to skip questions
-
-\section{Methods}
-
-\subsection{Curation workflow}\label{sec:curation}
-For our dataset we curated questions from existing exams, but also programmatically created new questions.
-Questions were added via Pull Requests on our GitHub repository and only merged into the main collection after passing manual review.
-
-To ensure that the questions do not enter a training dataset we use the canary string of the Big Bench project.
-This requires that \Gls{llm} developers filter their training dataset for this canary string.
-
-\begin{figure}
-    \includegraphics[width = \textwidth]{figures/chem-bench.pdf}
-        \caption{\textbf{Data curation workflow}.}
-\end{figure}
-
-\paragraph{Manually curated questions}
-
-\subparagraph{Analytical chemistry}
-The questions are sourced from master student examination papers at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
-
-\subparagraph{Combustion engineering}
-These questions are derived from master student examination papers at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
-
-\subparagraph{Functional materials and nano-materials}
-The questions are gathered from a seminar course conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
-
-\subparagraph{General chemistry}
-The questions originate from 10th grade examinations at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
-
-\subparagraph{International Chemical Olympiad (\gls{ICHO})}
-The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
-
-\subparagraph{Material synthesis}
-These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
-
-\subparagraph{Organic reactivity}
-The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
-
-\subparagraph{Periodic table}
-The questions are manually created based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
-
-\subparagraph{Polymer chemistry}
-Questions are sourced from examinations held at the University of Hamburg, covering topics related to polymer chemistry and its applications.
-
-\subparagraph{Chemical reactivity}
-These questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
-
-\subparagraph{Chemical safety}
-The questions encompass various aspects of safety, including lab safety, toxicology, material safety data sheets, carcinogens, sourced from reputable safety resources.
-
-\subparagraph{Shape of molecules}
-The questions are sourced from examinations held at the Technical University of Munich, Germany, focusing on determining the shape and geometry of molecules.
-
-\subparagraph{\gls{NMR} spectroscopy}
-Questions are sourced from the twitter page called (\url{https://twitter.com/NMRspectroscopy}), focusing on concepts and applications of nuclear magnetic resonance spectroscopy.
-
-\paragraph{Semi-programatically generated questions}
-
-\subparagraph{Oxidation states}
-The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}.
-These exercises focus on determining the number of isomers for specific molecules.
-To perform these calculations, we utilize the RDKit package \cite{bento2020open}.
-\subparagraph{Electron count of atoms}
-Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}.
-These inquiries involve determining the electron count of atoms in various chemical contexts.
-The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
-\subparagraph{Point group for molecules}
-Questions involve predicting the point group for molecules.
-The ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption/tree/main}) is utilized for data generation.
-Given a molecule, predict its point group.
-\subparagraph{SMILES-IUPAC}
-Web scrapped SMILES IUPAC pairs from the PubChem database \cite{pubchem} and then framed questions involve converting SMILES notation to IUPAC nomenclature, utilizing information from the PubChem database.
-
-\paragraph{Programmatically generated questions}
-\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
-We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
-
-\subparagraph{GHS, hazard statements and DAI}
-The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.~\cite{pubchem}
-The former two have been mined via the PubChem \gls{api}, while the latter has been manually compiled.
-The \glspl{dai} have been curated to contain only records approved by the \gls{who}.
-The chemicals in this class of questions belong to one of the three classes: pesticides (e.g., calcium arsenate), insecticides (e.g., cyfluthrin) or herbicides (e.g. 2,4-D).
-All data was saved in tabular form, and then we programmatically created questions (\variable{output/num_h_statements.txt} hazard statement definitions, \variable{output/num_pictograms.txt} chemical-\gls{ghs} pictogram matching, \variable{output/num_dai.txt} \glspl{dai}).
-
-\subparagraph{Number of isomers}
-We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled SMILES from the ZINC dataset.
-To obtain the answers for false options we sampled from the set of number of isomers.
-
-\subsection{Model evaluation workflow}
-
-\paragraph{Prompting}
-
-To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
-We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Graf{sec:parsing}.
-Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
-For example, all the SMILES representations are encapsulated within \texttt{[START_SMILES][\text backslash END_SMILES]} in Galactica\cite{taylor2022galactica}.
-Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \latex syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
-
-\subparagraph{Completion models}
-\texttt{Galactica-120b}\cite{taylor2022galactica}
-
-\subparagraph{Instruction-tuned models} \texttt{Claude 2} \texttt{Claude3 (Opus)}\cite{anthropicClaudeModelFamily2024}
-\texttt{GPT-4}\cite{openai2024gpt4}
-\texttt{GPT-3.5-turbo}\cite{brown2020language}
-\texttt{Gemini Pro}\cite{gemini}
-\texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
-\texttt{Llama2-70b}\cite{touvron2023llama}
-\texttt{pplx-7b-chat}
-
-\paragraph{Parsing}
-Our parsing workflow is, by default, multistep and primarily based on regular expressions.
-In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
-In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
-In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
-As initial tests indicated that models sometimes return integers in the form of words, e.g. \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion.
-In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
-The frequency of this fallback being triggered was very different for different models (see XX).
-Manual verification (see \Cref{sec:manually-verified-parsing}) indicates that this \gls{llm}-based parsing is not a relevant error source.
-
-\paragraph{Models}
-
-\subparagraph{Tool augmented models}
-In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this we, on the one hand, investigated the \texttt{online} model of Perplexity.AI and, on the other hand, \texttt{gpt-3.5-turbo} as well as \texttt{claude-2}, with ReAct-style tool augmentation.\cite{yao2023react}
-The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
-We implemented the system using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
-
-\subsection{Human baseline}
-
-\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\cite{nextjs} app router framework.
-This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a MySQL\cite{mysql} database and Prisma \gls{orm}\cite{prisma} for efficient database management.
-The web application is styled with Tailwind CSS\cite{tailwindcss} using the shadcn/ui component library and uses NextAuth\cite{nextauth} for easy and secure user authentication and postMark for sending Emails.
-The application is hosted on the Vercel web hosting platform.
-
-\paragraph{Question selection}
-Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
-For selecting the subset, we decided on addressing two questions:
-\begin{itemize}
-    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
-    \item Are there areas in which the performance of humans is very different from the ones of the models?
-\end{itemize}
-To answer the first question we selected X questions that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-From those we picked X diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART (see below).
-
-\paragraph{Study design}
-For our initial study we wanted to maximize the response rate given our available resources.
-For this reason, we did not opt for a highly controlled study setting.
-That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
-Note that users were also allowed to skip questions.
-
-\subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset we systematically recorded keywords and sources.
-To allow for analysis of the model performance as a function of the topic, we leverage this information together with sequence classification models.
-For questions which can easily be assigned to a topic based on the source (e.g., number of NMR peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
-For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\cite{zeroshotsequence} using the BART model\cite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
-
-\section*{Data and code availability}
-
-\section*{Acknowledgements}
-This work was supported by the Carl Zeiss Foundation and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We also want to acknowledge access to the HPC cluster of Stability.AI.
-
-\section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.ai.
-
-\section*{Author contributions}
-
-\bibliography{references}
-
-\appendix
-
-\subsection{Benchmark corpus}
 
@@
 \end{figure}
-
-\end{document}

### 2024-03-12 src/tex/ms.tex
feat: update manuscript, fix analysis bugs

--- 2024-03-11 src/tex/ms.tex
@@
 \section{Introduction}
-\Glspl{llm} are frontier \gls{ml} models trained on massive amounts of text to complete sentences.
+\Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
+Aggressive scaling of the models has led to a rapid increase in their capabilities,\cite{brown2020language} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\cite{openai2024gpt4}
 While some see in them \enquote{sparks of \gls{agi}},\cite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\cite{bender2021dangers}
+In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability solve a wide variety of tasks they have not been explicitly trained on.\cite{bommasani2021opportunities, anderljung2023frontier}
+This led to tremendous economic interest and investment in those generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\cite{bloomberg}
 
-Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, some even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can physically perform reactions provided a command in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
-The rapid increase in capabilities led to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022}
-Moreover, the use of these models is now also widespread among students\cite{Intelligent.com_2023} as well as research groups.
-For some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
+Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
+This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
+Most of these insights are in form of text in scientific publications. 
+Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
+This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
+Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020}
+However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards}
+To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
+
+It is important to realize that the user base of such models is not only limited to experts in chemistry and materials science that reflect every output such models produce. 
+For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\cite{Intelligent.com_2023}
+This also implies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
+Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
+However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
 Unfortunately, apart from anecdotal reports there is little evidence on how \glspl{llm} perform compared to experts.
 
-Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, a comprehensive analysis is needed.
-For the development of \glspl{llm}, such evaluation is currently mostly performed via standardized benchmarks such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
-The former contains, among 204 tasks, only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
-Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
-While those evaluations can measure how well models can make predictions for very specific tasks, they only give a poor measure of how useful those models might be as a chemical assistant.
+Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, evaluation frameworks are needed to allow us to systematically measure progress and mitigate potential harms.
+For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmarks suites such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
+The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
+Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
+In these cases, the models are  evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, or reactivity) or on predicting the outcome of specific chemical reactions.
+This, however, only gives a very limited view of the general chemical capabilities of the models.
 
-While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, also those do not satisfy the following basic criteria chemistry benchmarks should satisfy:
-\begin{itemize}
-    \item \emph{End-to-end automation}. For model development, the evaluations will need to be run many times (e.g., on regular intervals of a training run).
-    Approaches that rely on humans scoring the answers of a system\cite{Schulze_Balhorn_2024, ai4science2023impact} can thus not be used.
-    \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\cite{northcutt2021pervasive}
-    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\cite{Frye_2023, Awg}
-    \item \emph{Usable with models that support special treatment of molecules}.
-Some models such as Galactica\cite{taylor2022galactica} use special tokenization or encoding procedures for molecules or equations.
-    To support this, the benchmark system must encode the semantic meaning of various parts of the question or answer.
-    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or even just the raw logits.
-    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
-    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
-    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real world chemistry as well as higher-level university education multiple choice question are seldom utilized.
-    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond the answering of \gls{mcq}.
-    \item \emph{Cover a diverse set of topics}.
-Chemistry, as the \enquote{central science}, bridges multiple disciplines.\cite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-\end{itemize}
+While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
+This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
+On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights in how the models compare relative to experts.
 
-In this work, we report a novel benchmarking framework, chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
+In this work, we report a novel benchmarking framework, which we call chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
-It covers a large fraction of the topics taught in undergraduate chemistry curricula at various skill levels and can be used with any system that can return text (i.e., also tool-augmented systems).
+Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also an tool-augmented systems).
 
@@
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, included safety-related ones, might be very misleading.
+The results indicate the limitations that current models still need to overcome to find use in autonomous systems for chemists and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.
 
@@
 
-\subsection{Benchmark dataset}
+\subsection{Benchmark corpus}
 
-To compile our benchmark corpus we utilized a broad list of sources (\Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-To ensure maximal interoperability, we curated the data in an extended form of the BigBench format.
-This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
+To compile our benchmark corpus we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
 For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
 
-Importantly, our large pool of questions encompassed a wide range of topics.
-This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different categories (see \Cref{sec:meth-topic} for details on how we assigned topics).
-By design, a focus of our corpus is on safety-related aspects with a (currently) limited sampling of questions from physical or theoretical chemistry, which might be extended in future work.
+Importantly, our large pool of questions encompasses a wide range of topics.
+This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
+The distribution of topics is also evident from \Cref{fig:pca} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
+In this representation, semantically similar questions are close to each other, and we color the points based on a classification into XX topics. 
+It is clear that a focus of chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.
 
@@
     \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question.
-The figure shows that not all aspects of chemistry are equally represented in our corpus.
-The corpus currently focuses on safety-related aspects.}
+    The figure shows that not all aspects of chemistry are equally represented in our corpus. The corpus currently focuses on safety-related aspects.}
     \label{fig:topic_barplot}
@@
 
-Importantly, the corpus samples both \gls{mcq} and open-ended questions in a balanced way (XX \gls{mcq} questions vs. XX open-ended questions).
-As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).~\cite{kincaid1975derivation}
+While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
+For this reason, chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
+\variable{output/non_mcq_questions.txt} open-ended questions).
+As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
 
-\begin{figure}
-    \centering
-    \includegraphics{figures/wordcloud.pdf}
-    \caption{\textbf{Wordcloud.}}
-    \label{fig:wordcloud}
-    \script{wordcloud.py}
-\end{figure}
-
-\subsection{\enquote{Tiny} subset}
+\paragraph{\enquote{Tiny} subset}
 It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
-For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark.
+For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
+To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
+We also used this subset to seed the app we used for the human baseline study.
 
@@
 
-\subsection{Future work}
-- Advanced prompting techniques such as CoT and test time strategies
+% \subsection{Future work}
+% - Advanced prompting techniques such as CoT and test time strategies
 
-- More advanced systems with chemnistry specific tools
+% - More advanced systems with chemnistry specific tools
 
-- More systematic human baseline: Ask people to come in, do not allow to skip questions
+% - More systematic human baseline: Ask people to come in, do not allow to skip questions
 
@@
 
-\subparagraph{Analytical chemistry}
-The questions are sourced from master student examination papers at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
-
-\subparagraph{Combustion engineering}
-These questions are derived from master student examination papers at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
-
-\subparagraph{Functional materials and nano-materials}
-The questions are gathered from a seminar course conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
-
-\subparagraph{General chemistry}
-The questions originate from 10th grade examinations at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
-
-\subparagraph{International Chemical Olympiad (\gls{ICHO})}
-The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
-
-\subparagraph{Material synthesis}
-These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
-
-\subparagraph{Organic reactivity}
-The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
-
-\subparagraph{Periodic table}
-The questions are manually created based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
-
-\subparagraph{Polymer chemistry}
-Questions are sourced from examinations held at the University of Hamburg, covering topics related to polymer chemistry and its applications.
-
-\subparagraph{Chemical reactivity}
-These questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
-
-\subparagraph{Chemical safety}
-The questions encompass various aspects of safety, including lab safety, toxicology, material safety data sheets, carcinogens, sourced from reputable safety resources.
-
-\subparagraph{Shape of molecules}
-The questions are sourced from examinations held at the Technical University of Munich, Germany, focusing on determining the shape and geometry of molecules.
-
-\subparagraph{\gls{NMR} spectroscopy}
-Questions are sourced from the twitter page called (\url{https://twitter.com/NMRspectroscopy}), focusing on concepts and applications of nuclear magnetic resonance spectroscopy.
-
 \paragraph{Semi-programatically generated questions}
@@
 \subparagraph{Oxidation states}
-The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}.
-These exercises focus on determining the number of isomers for specific molecules.
-To perform these calculations, we utilize the RDKit package \cite{bento2020open}.
-\subparagraph{Electron count of atoms}
-Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}.
-These inquiries involve determining the electron count of atoms in various chemical contexts.
-The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
-\subparagraph{Point group for molecules}
-Questions involve predicting the point group for molecules.
-The ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption/tree/main}) is utilized for data generation.
-Given a molecule, predict its point group.
-\subparagraph{SMILES-IUPAC}
-Web scrapped SMILES IUPAC pairs from the PubChem database \cite{pubchem} and then framed questions involve converting SMILES notation to IUPAC nomenclature, utilizing information from the PubChem database.
+
+\subparagraph{Electron counts}
 
@@
 To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
-We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Graf{sec:parsing}.
+We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Cref{sec:parsing}.
 Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
-For example, all the SMILES representations are encapsulated within \texttt{[START_SMILES][\text backslash END_SMILES]} in Galactica\cite{taylor2022galactica}.
-Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \latex syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
-
-\subparagraph{Completion models}
-\texttt{Galactica-120b}\cite{taylor2022galactica}
-
-\subparagraph{Instruction-tuned models} \texttt{Claude 2} \texttt{Claude3 (Opus)}\cite{anthropicClaudeModelFamily2024}
-\texttt{GPT-4}\cite{openai2024gpt4}
-\texttt{GPT-3.5-turbo}\cite{brown2020language}
-\texttt{Gemini Pro}\cite{gemini}
-\texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
-\texttt{Llama2-70b}\cite{touvron2023llama}
-\texttt{pplx-7b-chat}
+For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\cite{taylor2022galactica}.
+Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
 
@@
 \paragraph{Models}
+\subparagraph{Completion models}
+\texttt{Galactica-120b}\cite{taylor2022galactica}
+
+\subparagraph{Instruction-tuned models} \texttt{Claude 2} \texttt{Claude3 (Opus)}\cite{anthropicClaudeModelFamily2024}
+\texttt{GPT-4}\cite{openai2024gpt4}
+\texttt{GPT-3.5-turbo}\cite{brown2020language}
+\texttt{Gemini Pro}\cite{gemini}
+\texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
+\texttt{Llama2-70b}\cite{touvron2023llama}
+\texttt{pplx-7b-chat}
 
@@
 For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+hat is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
 Note that users were also allowed to skip questions.
@@
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.ai.
+K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.AI.
 
 \section*{Author contributions}
+
+\footnotesize
+\insertcredits
+\normalsize
 
@@
 \appendix
-
-\subsection{Benchmark corpus}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-    \caption{Example question.}
-    \label{fig:flesch_kincaid_reading_ease}
-\end{figure}
-
-\subsection{Parsing verification}\label{sec:manually-verified-parsing}
-For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
-
-\subsection{Model performance}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar.pdf}
-    \caption{Caption}
-    \label{fig:all_questions_models_completely_correct_radar}
-    \script{analyze_model_reports.py}
-\end{figure}
-
-\subsection{Human baseline}
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/human_timing.pdf}
-    \script{analyze_human_data.py}
-    \label{fig:human_timing}
-    \caption{}
-\end{figure}
-
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_\experience_score.txt}\), \(p \approx \variable{output/spearman_\experience_score_p.txt}\)).
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/experience_vs_correctness.pdf}
-    \script{analyze_human_data.py}
-    \label{fig:experience_vs_correctness}
-    \caption{}
-\end{figure}
+\input{appendix.tex}
 

### 2024-03-12 src/tex/ms.tex
work on main text

@@
 
+Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
+For instance, Galactica\cite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
+To support this, chembench is designed to encode the semantic meaning of various parts of the question or answer.  
+For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.
+
+Because many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
+This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
+In those cases, the \gls{llm} that returns the probabilities for various tokens (that are conventionally used for model evaluations) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
+The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).
+
+To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of models on the chembench corpus, including systems augmented with external tools.
+An overview of the results of this evaluation is shown in \Cref{fig:all_questions_models_completely_correct_radar}.
+
+% Bar chart
+
 \begin{figure}
@@
 \end{figure}
+
+% Radar chart
+To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models on different topics.
 

### 2024-03-12 src/tex/ms.tex
include statistics about human baseline

@@
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the app we used for the human baseline study.
+We also used this subset to seed the app we used for the human baseline study. The results shown in the main text are based on this subset. 
+Results for the full corpus can be found in the appendix.
 
@@
 
-To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of models on the chembench corpus, including systems augmented with external tools.
-An overview of the results of this evaluation is shown in \Cref{fig:all_questions_models_completely_correct_radar}.
-
+To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\cite{Huggingface} on the chembench corpus, including systems augmented with external tools.
+An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. In this figure, we show the percentage of questions that the models answered completely correctly.
+We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance with vertical lines.
+The figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric.
+Many other models also outperform the average human performance.
 % Bar chart
@@
 For this reason, we did not opt for a highly controlled study setting. 
-hat is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
 Note that users were also allowed to skip questions.
+
+\paragraph{Participants}
+Users were open to report about their experience in chemistry. 
+\variable{output/num_users_with_education_info.txt} did so. 
+Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree and for \variable{output/num_human_highschool.txt} the highest awarded degree is from high school.
 

### 2024-03-12 src/scripts/analyze_human_data.py

@@ -118,6 +118,37 @@ def make_human_time_score_plot(long_df):
     with open(output / "spearman_experience_score_p.txt", "w") as f:
         f.write(str(np.round(spearman.pvalue, 2)) + "\endinput")
 
+    # find educational background for each user and write this to files
+    edu_dict = {}
+    for user in grouped_by_user.index:
+        edu_dict[user] = long_df[long_df["userid"] == user]["highest_education"].values[
+            0
+        ]
+
+    with open(output / "num_human_phd.txt", "w") as f:
+        f.write(str(list(edu_dict.values()).count("doctorate")) + "\endinput")
+
+    with open(output / "num_human_master.txt", "w") as f:
+        f.write(str(list(edu_dict.values()).count("MSc")) + "\endinput")
+
+    with open(output / "num_human_bachelor.txt", "w") as f:
+        f.write(str(list(edu_dict.values()).count("BSc")) + "\endinput")
+
+    with open(output / "num_human_highschool.txt", "w") as f:
+        f.write(str(list(edu_dict.values()).count("high-school")) + "\endinput")
+
+    with open(output / "num_human_postdoc.txt", "w") as f:
+        f.write(str(list(edu_dict.values()).count("post-doctorate")) + "\endinput")
+
+    with open(output / "num_users_with_education_info.txt", "w") as f:
+        # take those user for which the education info is one of the above categories
+        users_with_education_info = [
+            k
+            for k, v in edu_dict.items()
+            if v in ["doctorate", "MSc", "BSc", "high-school", "post-doctorate"]
+        ]
+        f.write(str(len(users_with_education_info)) + "\endinput")
+
 
 if __name__ == "__main__":
     make_human_performance_plots()

### 2024-03-12 src/scripts/analyze_model_reports.py

@@ -225,8 +225,7 @@ def load_reports():
         os.path.join(chembench, "data"),
     )
 
-    galactica_120b = load_all_reports
-    (
+    galactica_120b = load_all_reports(
         os.path.join(
             chembench,
             "reports/galactica-120b/reports/d7ce25da-bbce-4f06-8a5b-43e6cfb01c30",

### 2024-03-12 src/tex/ms.tex
fix bugs, update text

@@
     \centering
-    \caption{Caption}
-    \label{fig:enter-label}
+    \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly.
+We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
+    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
+    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.}
+    \label{fig:human_vs_models_bar}
 \end{figure}
 
+Given the large interest in tool-augmented systems, the bad performance of these systems in our benchmark is striking. 
+Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
+With the default ReAct setup, however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
+This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+
 % Radar chart
-To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models on different topics.
+To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics.
+We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
 
@@
 
-\section{Conclusions}
+% Calculation vs. no calculation
+Given that some of the questions in chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
+For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.
+
+% Estimate of difficulty
+One might wonder whether the models are capable of estimating if they will be able to answer a question correctly.
+To investigate this, we prompted some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
+
+\begin{figure}
+    \centering
+    %\includegraphics{figures/estimated_difficulty.pdf}
+    \caption{Caption}
+    \label{fig:estimated_difficulty}
+\end{figure}
+
+In \Cref{fig:estimated_difficulty} we show that there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\cite{Li_2023}
+
+\section{Discussion and Conclusions}
+Our findings also lay the groundwork for discussions for the chemical education in the future. 
+Given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
+Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
 
@@
 
-\paragraph{Programmatically generated questions}
 \subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.

### 2024-03-12 src/tex/ms.tex
update manuscript

@@
 Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020}
-However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards}
+However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
 To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).

### 2024-03-12 src/tex/appendix.tex
fix bugs, update text
update text

@@
     \item \emph{End-to-end automation}. For model development, the evaluations will need to be run many times (e.g., on regular intervals of a training run).
-    Approaches that rely on humans scoring the answers of a system\cite{Schulze_Balhorn_2024, ai4science2023impact} can thus not be used.
+    Approaches that rely on humans scoring the answers of a system\cite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
     \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\cite{northcutt2021pervasive}
@@
     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-    \caption{Example question.}
+    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\cite{flesch1948new} is a measure of how easy a text is to read.
+It is calculated based on the average number of syllables per word and the average number of words per sentence.
+The higher the score, the easier the text is to read.
+The distribution of the scores of the questions is shown in the histogram.
+The mean score is \variable{output/flesch_kincaid_mean.txt} and the median score is \variable{output/flesch_kincaid_median.txt}.}
     \label{fig:flesch_kincaid_reading_ease}
@@
 
-\subsection{Parsing verification}\label{sec:manually-verified-parsing}
-For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
+% \subsection{Parsing verification}\label{sec:manually-verified-parsing}
+% For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
 
@@
     \label{fig:human_timing}
-    \caption{}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
 \end{figure}
@@
     \label{fig:experience_vs_correctness}
-    \caption{}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).}
 \end{figure}

### 2024-03-12 src/tex/appendix.tex
add some more analysis leads

@@
 The distribution of the scores of the questions is shown in the histogram.
-The mean score is \variable{output/flesch_kincaid_mean.txt} and the median score is \variable{output/flesch_kincaid_median.txt}.}
+}
     \label{fig:flesch_kincaid_reading_ease}
@@
 \subsection{Model performance}
+We also evaluated the model performance on the entire chembench corpus. 
+\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.
+
+\begin{figure}
+    \label{fig:barplot_all_correct_all_questions}
+\end{figure}
+
+\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the chembench corpus.
 
@@
     \label{fig:experience_vs_correctness}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
 \end{figure}

### 2024-03-12 src/tex/ms.tex
update text
add some more analysis leads

--- 2024-03-12 src/tex/ms.tex
@@
 This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020}
-However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards}
+Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
 To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
@@
 
-Because many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
+Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
 This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
-In those cases, the \gls{llm} that returns the probabilities for various tokens (that are conventionally used for model evaluations) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
+In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\cite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
 The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).
@@
 To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\cite{Huggingface} on the chembench corpus, including systems augmented with external tools.
-An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. In this figure, we show the percentage of questions that the models answered completely correctly.
-We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance with vertical lines.
-The figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric.
+An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
+In this figure, we show the percentage of questions that the models answered completely correctly.
+We use horizontal bars to indicate the performance of various models and highlight statistics (e.g., the average) of the human performance with vertical lines.
+Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly exceeds the average performance of our human scores.
 Many other models also outperform the average human performance.
+It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
 % Bar chart
@@
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
-With the default ReAct setup, however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
+With the default tool augmented setup (using the so-called ReAct method), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
 This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
@@
 To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
-For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics based and hand-crafted rules as well as classifier models.
 We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
@@
     \caption{Caption}
-    \label{fig:enter-label}
+    \label{fig:performance_by_topic}
 \end{figure}
+
+In this spider chart, the worst score for every dimension is 0 (no question answered completely correctly) and the best score is 1 (all questions answered correctly). 
+Thus, a larger colored area indicates a better performance. 
+One can observe that this performance varies widely across models and topics. 
+While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models. 
+While this question is relatively easy () for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+This is mirrored in the questions about point groups of compounds ...
+
+For all topics in chembench, proprietary models outperform their open counterparts.
+
+% Can we say something how this links to changes we have to make to the training data
 
@@
 
+\begin{figure}
+    \centering
+    %\includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
+    \caption{Caption}
+    \label{fig:calculation_performance}
+\end{figure}
+
 % Estimate of difficulty
 One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
+If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
 To investigate this, we prompted some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
@@
 
+\subsection{Confidence estimate}
+To estimate the models' confidence we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the right answer on a scale from 1 to 5.
+
 \subsection{Human baseline}

### 2024-03-12 src/tex/ms.tex
continue analysis leads

@@
 For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\cite{Li_2023}
+For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of XX (on a scale of 1--5) for the XX (out of XX) questions it answered in correctly (XX for the XX questions it answered correctly).
 

### 2024-03-12 src/tex/ms.tex
add caching

@@
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the app we used for the human baseline study. The results shown in the main text are based on this subset. 
-Results for the full corpus can be found in the appendix.
+We also used this subset to seed the app we used for the human baseline study.
 
@@
 This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+
+%% Note also while the performance of the systems is better than what humans do, it is still far from perfect
 
@@
 \subparagraph{Completion models}
-\texttt{Galactica-120b}\cite{taylor2022galactica}
+We used \texttt{Galactica-120b}\cite{taylor2022galactica} with the default settings
 
@@
 Users were open to report about their experience in chemistry. 
-\variable{output/num_users_with_education_info.txt} did so. 
+Overall, \variable{output/num_users_with_education_info.txt} did so. 
 Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
 \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree and for \variable{output/num_human_highschool.txt} the highest awarded degree is from high school.
+
+\paragraph{Comparison with models}
+To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions which at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
+The latter threshold was chosen to limit it to humans who made serious attempts at systematically answering a part of the questions. 
+This analysis might lead to potential biases, most likely in favor of humans as they were allowed to skip questions.
+\variable{output/num_humans_with_204_scores.txt} humans answered more than 200 questions.
 

### 2024-03-12 src/scripts/collect_human_scores.py

@@ -1,0 +1,45 @@
+from glob import glob
+import os
+from pathlib import Path
+import subprocess
+from utils import obtain_chembench_repo
+from paths import output
+
+if __name__ == "__main__":
+    chembench_repo = obtain_chembench_repo()
+    human_files = os.listdir(os.path.join(chembench_repo, "reports", "humans"))
+    human_files = [
+        os.path.join(chembench_repo, "reports", "humans", p) for p in human_files
+    ]
+    # we only consider humans with at least 100 questions
+    human_files = [p for p in human_files if len(glob(os.path.join(p, "*.json"))) > 100]
+
+    # count how often we have 204 (all of the "tiny" benchmark)
+    has_204_scores = [
+        p for p in human_files if len(glob(os.path.join(p, "*.json"))) == 204
+    ]
+
+    with open(output / "num_humans_with_more_than_100_scores.txt", "w") as handle:
+        handle.write(f"{len(human_files)}" + "\endinput")
+
+    with open(output / "num_humans_with_204_scores.txt", "w") as handle:
+        handle.write(f"{len(has_204_scores)}" + "\endinput")
+
+    outpath = os.path.join(output, "human_scores")
+    if not os.path.exists(outpath):
+        os.mkdir(outpath)
+
+    datafolder = os.path.join(chembench_repo, "data")
+    human_baseline_folder = os.path.join(chembench_repo, "reports", "humans")
+    scriptpath = os.path.join(chembench_repo, "scripts", "collect_scores.py")
+
+    for file in human_files:
+        try:
+            p = Path(file).parts[-1]
+            outfile = os.path.join(outpath, p + ".json")
+            subprocess.run(
+                f"python {scriptpath} {file} {outfile} --datafolder={datafolder} --human_baseline_folder={human_baseline_folder}",
+                shell=True,
+            )
+        except Exception:
+            pass

### 2024-03-12 src/scripts/collect_model_scores_human_subset.py

@@ -1,0 +1,29 @@
+from glob import glob
+import os
+from pathlib import Path
+import subprocess
+from utils import obtain_chembench_repo
+from paths import output
+
+if __name__ == "__main__":
+    chembench_repo = obtain_chembench_repo()
+    reports = glob(os.path.join(chembench_repo, "reports", "*.json"))
+    models = [Path(p).parent for p in reports]
+    outpath = os.path.join(output, "human_subset_model_scores")
+    if not os.path.exists(outpath):
+        os.mkdir(outpath)
+
+    datafolder = os.path.join(chembench_repo, "data")
+    human_baseline_folder = os.path.join(chembench_repo, "reports", "humans")
+    scriptpath = os.path.join(chembench_repo, "scripts", "collect_scores.py")
+
+    for file in models:
+        try:
+            p = Path(file).parts[-1]
+            outfile = os.path.join(outpath, p + ".json")
+            subprocess.run(
+                f"python {scriptpath} {file} {outfile} --datafolder={datafolder} --human_baseline_folder={human_baseline_folder}",
+                shell=True,
+            )
+        except Exception:
+            pass

### 2024-03-13 src/scripts/get_model_performance_dicts.py
clean up workflow

@@ -379,9 +379,6 @@ def load_reports():
         & random_baseline_run_names
     )
 
-    with open(output / "intersection.txt", "w") as f:
-        f.write(f"{len(intersection)}" + "\endinput")
-
     # now we filter the dataframes to only contain the runs
     # that are in the intersection
     claude2 = claude2[claude2[("name", 0)].isin(intersection)]

### 2024-03-14 src/tex/ms.tex
clean up workflow
update analysis

@@
 While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models. 
-While this question is relatively easy () for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
-This is mirrored in the questions about point groups of compounds ...
-
-For all topics in chembench, proprietary models outperform their open counterparts.
-
-% Can we say something how this links to changes we have to make to the training data
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \SI{\variable{outputs/subset_scores/is_number_nmr_peaks_gpt-4.txt}}{\percent}).
+While this question is relatively easy (\SI{\variable{outputs/human_subset_scores/is_number_nmr_peaks.txt}}{\percent}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+This is mirrored in the questions about point groups of compounds, which is models struggle with more than humans ().
 
@@
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}. 
+To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
 

### 2024-03-14 src/scripts/analyze_model_reports.py

@@ -30,8 +30,7 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
         df_dict["gpt35turbo_zero_t"].groupby("topic")["all_correct_"].mean()
     )
     gpt35turbo_react_mean = (
-        
-        
+        df_dict["gpt35turbo_react"].groupby("topic")["all_correct_"].mean()
     )
     gpt4_mean = df_dict["gpt4"].groupby("topic")["all_correct_"].mean()
     # gpt4zero_t_mean = df_dict["gpt4zero_t"].groupby("topic")["all_correct_"].mean()
@@ -44,112 +43,117 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
     random_baseline_mean = (
         df_dict["random_baseline"].groupby("topic")["all_correct_"].mean()
     )
+
     if human_dicts is None:
         # Sort the data based on the maximum value in each group
         sorted_data = sorted(
-            zip(
-                [
-                    gpt4_mean,
-                    claude2_mean,
-                    claude2_react_mean,
-                    claude3_mean,
-                    gemini_pro_mean,
-                    gpt35turbo_mean,
-                    gpt35turbo_react_mean,
-                    llama70b_mean,
-                    galactica_120b_mean,
-                    mixtral_mean,
-                    pplx7b_chat_mean,
-                    pplx7b_online_mean,
-                    random_baseline_mean,
-                ],
-                [
-                    "GPT-4",
-                    "Claude2",
-                    "Claude2-ReAct",
-                    "Claude3",
-                    "Gemini-Pro",
-                    "GPT-3.5-Turbo",
-                    "GPT-3.5-Turbo-ReAct",
-                    "LLAMA-2-70B",
-                    "Galactica-120B",
-                    "MixTRAL-8x7B",
-                    "PPLX-7B-Chat",
-                    "PPLX-7B-Online",
-                    "Random Baseline",
-                ],
-                [
-                    model_color_map["gpt4"],
-                    model_color_map["claude2"],
-                    model_color_map["claude2_react"],
-                    model_color_map["claude3"],
-                    model_color_map["gemini_pro"],
-                    model_color_map["gpt35turbo"],
-                    model_color_map["gpt35turbo_react"],
-                    model_color_map["llama70b"],
-                    model_color_map["galactica_120b"],
-                    model_color_map["mixtral"],
-                    model_color_map["pplx7b_chat"],
-                    model_color_map["pplx7b_online"],
[…]

### 2024-03-14 src/scripts/get_model_performance_dicts.py

@@ -357,50 +357,69 @@ def load_reports(topic_frame):
     pplx7b_online_run_names = set(pplx7b_online[("name", 0)])
     random_baseline_run_names = set(random_baseline[("name", 0)])
 
-    # now we take the intersection of these sets
-    intersection = (
-        claude2_run_names
-        & claude2_react_run_names
-        & claude2_zero_t_run_names
-        & claude3_run_names
-        & galactica_120b_run_names
-        & gemini_pro_zero_t_run_names
-        & gemini_pro_run_names
-        & gpt35turbo_run_names
-        & gpt35turbo_zero_t_run_names
-        & gpt35turbo_react_run_names
-        & gpt4_run_names
-        & llama70b_run_names
-        & mixtral_run_names
-        & pplx7b_chat_run_names
-        & pplx7b_online_run_names
-        & random_baseline_run_names
+    print(
+        "RUN COUNTS",
+        len(claude2_run_names),
+        len(claude2_react_run_names),
+        len(claude2_zero_t_run_names),
+        len(claude3_run_names),
+        len(galactica_120b_run_names),
+        len(gemini_pro_zero_t_run_names),
+        len(gemini_pro_run_names),
+        len(gpt35turbo_run_names),
+        len(gpt35turbo_zero_t_run_names),
+        len(gpt35turbo_react_run_names),
+        len(gpt4_run_names),
+        len(llama70b_run_names),
+        len(mixtral_run_names),
+        len(pplx7b_chat_run_names),
+        len(pplx7b_online_run_names),
+        len(random_baseline_run_names),
     )
+    # now we take the intersection of these sets
+    # intersection = (
+    #     claude2_run_names
+    #     & claude2_react_run_names
+    #     & claude2_zero_t_run_names
+    #     & claude3_run_names
+    #     & galactica_120b_run_names
+    #     & gemini_pro_zero_t_run_names
+    #     & gemini_pro_run_names
+    #     & gpt35turbo_run_names
+    #     & gpt35turbo_zero_t_run_names
+    #     & gpt35turbo_react_run_names
+    #     & gpt4_run_names
+    #     & llama70b_run_names
+    #     & mixtral_run_names
+    #     & pplx7b_chat_run_names
+    #     & pplx7b_online_run_names
[…]

### 2024-03-14 src/tex/appendix.tex
update authors and prepare confidence score distribution

@@
 \end{figure}
+
+\subsection{Confidence estimates}
+
+Since it is important to understand if models can provide an indication of whether their answer might likely be incorrect, we prompted some of our top performing \glspl{llm} to return the confidence in providing a correct answer on an ordinal scale.
+
+\Cref{XX} plots the distribution of those scores.
+
+\begin{figure} 
+    \centering
+    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.}} 
+    \label{fig:confidence_score_distributions}
+\end{figure}

### 2024-03-14 src/tex/ms.tex
update authors and prepare confidence score distribution

@@
 We also want to acknowledge access to the HPC cluster of Stability.AI.
+M.A.
+expresses gratitude to the European Research Council (ERC) for evaluating the project with the reference number 101106377 titled ‘‘CLARIFIER’’ and accepting it for funding under the HORIZON TMA MSCA Postdoctoral Fellowships - European Fellowships.
+Furthermore, M.A.
+acknowledges the funding provided by UK Research and Innovation (UKRI) under the UK government’s Horizon Europe funding guarantee (Grant Reference: EP/Y023447/1; Organization Reference: 101106377).
 

### 2024-03-14 src/tex/ms.tex
work on abstract

@@
 \begin{abstract}
+    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly been trained on. 
+    This is relevant for the chemical sciences, which face the problem of learning from many small and diverse datasets that are frequently in the form of text.
+    Thus, chemists have started to adapt LLMs to make predictions, optimize reactions, or even autonomously design and perform experiments.
+
+    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    In this work, we report an automatic framework, chembench, that we used to probe the chemical capabilities of LLMs compared to human chemists.
+
+    % Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge
+    We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4, and found that the best models, on average, outperform the best human chemists in our study.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g. about the safety profile of chemicals.
+
+    % One or two sentences to put the results into a more general context.
+
 \end{abstract}

### 2024-03-14 src/tex/ms.tex
update abstract

@@
 
-    % Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge
     We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4, and found that the best models, on average, outperform the best human chemists in our study.
@@
 
-    % One or two sentences to put the results into a more general context.
-
+    Our results demonstrate that while frontier models show outstanding performance on chemical tasks, additional research is required to ensure useful and safe chemical language models. 
+    The findings also suggest that chemistry curricula might need to adapt and that continued development of evaluation frameworks is critical for systematic improvement of LLMs.
 \end{abstract}

### 2024-03-14 src/tex/ms.tex
update abstract

--- 2024-03-14 src/tex/ms.tex
@@
     This is relevant for the chemical sciences, which face the problem of learning from many small and diverse datasets that are frequently in the form of text.
-    Thus, chemists have started to adapt LLMs to make predictions, optimize reactions, or even autonomously design and perform experiments.
+    LLMs are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
 
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    In this work, we report an automatic framework, chembench, that we used to probe the chemical capabilities of LLMs compared to human chemists.
+    Here, we introduce \enquote{chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
 
-    % Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge
     We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4, and found that the best models, on average, outperform the best human chemists in our study.
@@
 
-    % One or two sentences to put the results into a more general context.
-
+    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, there is a critical need for further research to enhance their safety and utility in chemical sciences.
+    Our findings also indicate a need for adaptation to chemistry curricula and highlight the importance of continued development of evaluation frameworks for systematic improvement of safe and useful LLMs.
 \end{abstract}

### 2024-03-14 src/tex/appendix.tex
update confidence score estimates
update plotting workflow

@@
     \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar.pdf}
+    \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
     \caption{Caption}
-    \label{fig:all_questions_models_completely_correct_radar}
+    \label{fig:all_questions_models_completely_correct_radar_overall}
     \script{analyze_model_reports.py}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/performance_per_topic.pdf}
+    \caption{Caption}
+    \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
 \end{figure}
@@
 
-\subsection{Confidence estimates}
+\subsection{Confidence estimates} \label{sec:confidence_estimates}
 
@@
 
-\Cref{XX} plots the distribution of those scores.
+\Cref{fig:confidence_score_distributions} plots the distribution of those scores.
 
@@
     \centering
+    \includegraphics{figures/confidence_score_distributions.pdf}
     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.}} 
     \label{fig:confidence_score_distributions}
+    \script{plot_confidence_score_distributions.py}
 \end{figure}

### 2024-03-14 src/tex/ms.tex
update plotting workflow

@@
 While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \SI{\variable{outputs/subset_scores/is_number_nmr_peaks_gpt-4.txt}}{\percent}).
-While this question is relatively easy (\SI{\variable{outputs/human_subset_scores/is_number_nmr_peaks.txt}}{\percent}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
-This is mirrored in the questions about point groups of compounds, which is models struggle with more than humans ().
+% In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{outputs/subset_scores/is_number_nmr_peaks_gpt-4.txt}.
+% While this question is relatively easy (\variable{outputs/human_subset_scores/is_number_nmr_peaks.txt}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+% This is mirrored in the questions about point groups of compounds, which is models struggle with more than humans ().
+
+% GFK performance -> would the models obtain a chemical license
 

### 2024-03-14 src/scripts/analyze_model_reports.py

@@ -52,14 +52,14 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                     [
                         gpt4_mean,
                         claude2_mean,
-                        claude2_react_mean,
+                        # claude2_react_mean,
                         claude3_mean,
                         gemini_pro_mean,
                         gpt35turbo_mean,
-                        gpt35turbo_react_mean,
-                        llama70b_mean,
-                        galactica_120b_mean,
-                        mixtral_mean,
+                        # gpt35turbo_react_mean,
+                        # llama70b_mean,
+                        # galactica_120b_mean,
+                        # mixtral_mean,
                         pplx7b_chat_mean,
                         pplx7b_online_mean,
                         random_baseline_mean,
@@ -67,14 +67,14 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                     [
                         "GPT-4",
                         "Claude2",
-                        "Claude2-ReAct",
+                        # "Claude2-ReAct",
                         "Claude3",
                         "Gemini-Pro",
                         "GPT-3.5-Turbo",
-                        "GPT-3.5-Turbo-ReAct",
-                        "LLAMA-2-70B",
-                        "Galactica-120B",
-                        "MixTRAL-8x7B",
+                        # "GPT-3.5-Turbo-ReAct",
+                        # "LLAMA-2-70B",
+                        # "Galactica-120B",
+                        # "MixTRAL-8x7B",
                         "PPLX-7B-Chat",
                         "PPLX-7B-Online",
                         "Random Baseline",
@@ -82,14 +82,14 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                     [
                         model_color_map["gpt4"],
                         model_color_map["claude2"],
-                        model_color_map["claude2_react"],
+                        # model_color_map["claude2_react"],
                         model_color_map["claude3"],
                         model_color_map["gemini_pro"],
                         model_color_map["gpt35turbo"],
-                        model_color_map["gpt35turbo_react"],
-                        model_color_map["llama70b"],
-                        model_color_map["galactica_120b"],
-                        model_color_map["mixtral"],
+                        # model_color_map["gpt35turbo_react"],
+                        # model_color_map["llama70b"],
+                        # model_color_map["galactica_120b"],
[…]

### 2024-03-14 src/scripts/plot_confidence_score_distributions.py
update analysis code

@@ -11,7 +11,7 @@ from plotutils import range_frame, model_color_map
 import os
 import numpy as np
 
-plt.style.use(scripts / "lamalab.org")
+plt.style.use(scripts / "lamalab.mplstyle")
 
 
 def plot():

### 2024-03-14 src/tex/appendix.tex
feat: add initial discussion about related work
co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>

@@
 \end{itemize}
+
+\subsection{Related work}
+Existing benchmarks such as those from \citet{guo2023large}, \cite{sun2023scieval}, \citet{Schulze_Balhorn_2024}, \citet{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}).
+While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
+ChemBench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols.
+Moreover, we believe that our robust framework will engage the community in open-source contributions.
 

### 2024-03-15 src/tex/ms.tex
update paper
update analysis code
added safety description

@@
 
+% overall performance profile of various model is comparable
 \begin{figure}
@@
 % GFK performance -> would the models obtain a chemical license 
+% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
 
@@
 
+\subparagraph{Analytical chemistry}
+The questions are sourced from the master student examination paper at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
+
+\subparagraph{Combustion engineering}
+These questions are derived from the master student examination paper at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
+
+\subparagraph{Functional materials and nano-materials}
+The questions are gathered from a seminar conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
+
+\subparagraph{General chemistry}
+The questions originate from examinations for grade X at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+
+\subparagraph{International Chemical Olympiad (\gls{ICHO})}
+The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova, providing challenging problems for competitive chemistry.
+
+\subparagraph{Material synthesis}
+These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
+
+\subparagraph{Organic reactivity}
+The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
+
+\subparagraph{Periodic table}
+The questions are framed based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
+
+\subparagraph{Polymer chemistry}
+Questions are sourced from examinations held at the University of Hamburg, covering topics related to polymer chemistry and its applications.
+
+\subparagraph{Chemical reactivity}
+These questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
+
+\subparagraph{Chemical safety}
+This category comprises a total of 4792 questions, covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \cite{pubchem}, contributing 976 questions.
+Additionally, it explores crucial topics like materials compatibility, with 3200 questions, essential for understanding the hazardous impact of mixing certain materials and chemicals. 
+Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of 90 questions, focusing on enhancing federal oversight of chemical safety and security by promoting coordination between federal agencies like OSHA, EPA, and DHS, as well as state governments to ensure compliance with safety and security requirements for chemical facilities.
+Moreover, this category addresses questions related to chemical storage and compatibility, totaling 297 questions, and covers topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets.
+The chemical safety category stands out as one of the primary focus areas within the chembench dataset.
+
+\subparagraph{Shape of molecules}
+The questions are sourced from examinations held at the Technical University of Munich, Germany, focusing on determining the shape and geometry of molecules.
+
+\subparagraph{\gls{NMR} spectroscopy}
+Questions are sourced from the twitter page called (\url{https://twitter.com/NMRspectroscopy}), focusing on concepts and applications of nuclear magnetic resonance spectroscopy.
+
 \paragraph{Semi-programatically generated questions}

### 2024-03-15 src/scripts/analyze_performance_per_source.py

@@ -1,0 +1,108 @@
+from paths import output
+import pickle
+import os
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+from paths import scripts, figures, data
+from loguru import logger
+
+plt.style.use(scripts / "lamalab.mplstyle")
+
+
+def obtain_score_for_subset(df, subset):
+    print(df[df[subset]]["all_correct_"])
+    return df[df[subset]]["all_correct_"].astype(int).mean()
+
+
+subsets = [
+    "is_point_group",
+    "is_number_of_isomers",
+    "is_number_nmr_peaks",
+    "is_gfk",
+    "is_dai",
+    "is_pictograms",
+    "is_name",
+    "is_smiles_name",
+    "is_organic_reactivity",
+    "is_electron_counts",
+    "is_chemical_compatibility",
+    "is_materials_compatibility",
+    "is_oup",
+    "is_olympiad",
+    "is_toxicology",
+    "is_polymer_chemistry",
+]
+
+
+def obtain_subset_scores(data_dict, outdir):
+    all_scores = []
+    for subset in subsets:
+        for model, scores in data_dict.items():
+            try:
+                score = obtain_score_for_subset(data_dict[model], subset)
+
+                with open(os.path.join(outdir, f"{subset}_{model}.txt"), "w") as handle:
+                    handle.write(str(int(np.round(score * 100, 0))) + "\endinput")
+
+                all_scores.append({"model": model, "score": score, "subset": subset})
+            except Exception as e:
+                logger.warning(f"Failed for {model} {subset} due to {e}")
+                pass
+
+    return all_scores
+
+
+def obtain_subset_scores_humans(data_dict, outdir):
+    all_scores = []
+    for subset in subsets:
+        subset_scores = []
+        try:
+            for human, scores in data_dict.items():
+                score = obtain_score_for_subset(data_dict[human], subset)
+                subset_scores.append(score)
+            mean_subset_score = np.mean(subset_scores)
+            with open(os.path.join(outdir, f"{subset}.txt"), "w") as handle:
+                handle.write(
+                    str(int(np.round(mean_subset_score * 100, 0))) + "\endinput"
+                )
+        except Exception as e:
+            logger.warning(f"Failed for {subset} due to {e}")
+    all_scores.append({"model": "human", "score": score, "subset": subset})
[…]

### 2024-03-15 src/scripts/embed_questions.py

@@ -1,0 +1,43 @@
+from chembench.classification import build_question_df, QClassifier
+from glob import glob
+from langchain.globals import set_llm_cache
+from dotenv import load_dotenv
+from langchain.cache import SQLiteCache
+from utils import obtain_chembench_repo
+import os
+import pickle
+from paths import data
+
+load_dotenv("../.env")
+set_llm_cache(SQLiteCache(database_path=".langchain.db"))
+
+if __name__ == "__main__":
+
+    repo = obtain_chembench_repo
+    JSON_PATHS = glob(os.path.join(repo, "data/**/*.json"), recursive=True)
+
+    df = build_question_df(JSON_PATHS)
+
+    TOPICS = [
+        "general chemistry (e.g., oxidation states, stoichiometry, point groups, electron counts, balancing reactions, atomic structure, electronic configuration, periodic table, periodic trends, etc.)",
+        "organic chemistry (e.g., isomers of organic compounds, photochemistry, organic reactions, organic reactivity, asymmetric centers, etc.)",
+        "inorganic chemistry (e.g., coordination compounds, inorganic compounds, energetics, etc.)",
+        "physical chemistry (e.g. thermodynamics, kinetics, etc.)",
+        "computational chemistry (e.g., molecular modeling, molecular dynamics, etc.)",
+        "analytical chemistry (e.g., chromatography, spectroscopy, etc.)",
+        "macromolecular chemistry (e.g., polymers, etc.)",
+        "materials science (e.g., nanomaterials, metallurgy, formulations etc.)",
+        "toxicology (e.g., toxicological effects, poisons, etc.)",
+        "chemical safety (e.g., compatibility of chemicals and materials, GHS, irritation, safety data sheets, chemical hazards, handling of compounds, pyrophoric materials, etc.)",
+        "biochemistry (e.g., metabolism, enzymes, organisms, DNA, amino acids, etc.)",
+    ]
+
+    # BART example
+    q_classifier_ = QClassifier(embedding_model="bart", topics=TOPICS)
+
+    embeddings = q_classifier_.get_embedding(df["formatted"].to_list())
+
+    names = df["name"].to_list()
+
+    with open(data / "embeddings.pkl") as handle:
+        pickle.dumps({"names": names, "embeddings": embeddings}, handle)

### 2024-03-15 src/scripts/plot_confidence_score_distributions.py

@@ -1,0 +1,71 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+from paths import scripts, figures
+import seaborn as sns
+from utils import (
+    obtain_chembench_repo,
+    ONE_COL_WIDTH_INCH,
+    ONE_COL_GOLDEN_RATIO_HEIGHT_INCH,
+)
+from plotutils import range_frame, model_color_map
+import os
+import numpy as np
+
+plt.style.use(scripts / "lamalab.mplstyle")
+
+
+def plot():
+    chemnbench = obtain_chembench_repo()
+    gpt = pd.read_csv(
+        os.path.join(
+            chemnbench,
+            "reports",
+            "confidence_estimates",
+            "confidence_estimates",
+            "results_gpt-4.csv",
+        )
+    )
+    gpt["model"] = "GPT-4"
+
+    claude_2 = pd.read_csv(
+        os.path.join(
+            chemnbench,
+            "reports",
+            "confidence_estimates",
+            "confidence_estimates",
+            "results_claude2.csv",
+        )
+    )
+
+    claude_2["model"] = "Claude 2"
+
+    claude_3 = pd.read_csv(
+        os.path.join(
+            chemnbench,
+            "reports",
+            "confidence_estimates",
+            "confidence_estimates",
+            "results_claude3.csv",
+        )
+    )
+
+    claude_3["model"] = "Claude 3"
+
+    all_results = pd.concat([gpt, claude_2, claude_3])
+
+    fig, ax = plt.subplots(
+        figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
+    )
+    sns.violinplot(data=all_results, x="estimate", y="model", ax=ax)
+    sns.swarmplot(data=all_results, x="estimate", y="model", ax=ax)
+
+    ax.set_xlabel("confidence estimate return by the model")
+    ax.set_ylabel("model")
+
+    range_frame(ax, np.array([1, 5], np.array([0, 2])))
+
+    fig.savefig(figures / "confidence_score_distributions.pdf", bbox_inches="tight")
+
+
+if __name__ == "__main__":
+    plot()

### 2024-03-15 src/tex/appendix.tex
add command for chembench

@@
 While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
-ChemBench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols.
-Moreover, we believe that our robust framework will engage the community in open-source contributions.
+\chembench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols.
+Moreover, we believe that our robust framework, including the platform \url{chembench.org}, will engage the community in open-source contributions.
 
@@
 \subsection{Model performance}
-We also evaluated the model performance on the entire chembench corpus. 
+We also evaluated the model performance on the entire \chembench corpus. 
 \Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.
@@
 
-\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the chembench corpus.
+\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the \chembench corpus.
 

### 2024-03-15 src/tex/ms.tex
add command for chembench

--- 2024-03-14 src/tex/ms.tex
@@
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
 
@@
 
-In this work, we report a novel benchmarking framework, which we call chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
+In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
@@
 In this representation, semantically similar questions are close to each other, and we color the points based on a classification into XX topics. 
-It is clear that a focus of chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.
+It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.
 
@@
 While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-For this reason, chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
+For this reason, \chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
 \variable{output/non_mcq_questions.txt} open-ended questions).
@@
 For instance, Galactica\cite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
-To support this, chembench is designed to encode the semantic meaning of various parts of the question or answer.  
+To support this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
 For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.
 
-Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
+Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
 This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
@@
 % GFK performance -> would the models obtain a chemical license 
+% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
 
 % Calculation vs. no calculation
-Given that some of the questions in chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
+Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
 For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.
@@
 Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
+
+% given that LLMs have parsed almost the entire internet, it is perhaps not surprising to seem them perform well on many of the task which often require 
+% specific knowledge humans might not have at hand without consulting an additional knowledge bank 
+% when it comes to reasoning tasks (NMR, point groups) the humans tend to beat LLMs
 
@@
 
+Another aspect of requiring unsupervised question answering is that in real life humans have tools and are able to use them for answering any question of interest. 
+In our current study, we prompted users to not use those tools.
+
 \paragraph{Participants}
@@
 \section*{Data and code availability}
-The code and data for chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
+The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 

### 2024-03-15 src/scripts/plot_confidence_score_distributions.py
manuscript compiles

@@ -39,30 +39,31 @@ def plot():
 
     claude_2["model"] = "Claude 2"
 
-    claude_3 = pd.read_csv(
-        os.path.join(
-            chemnbench,
-            "reports",
-            "confidence_estimates",
-            "confidence_estimates",
-            "results_claude3.csv",
-        )
-    )
+    # claude_3 = pd.read_csv(
+    #     os.path.join(
+    #         chemnbench,
+    #         "reports",
+    #         "confidence_estimates",
+    #         "confidence_estimates",
+    #         "results_claude3.csv",
+    #     )
+    # )
 
-    claude_3["model"] = "Claude 3"
+    # claude_3["model"] = "Claude 3"
 
-    all_results = pd.concat([gpt, claude_2, claude_3])
+    all_results = pd.concat([gpt, claude_2]).reset_index(drop=True)
 
     fig, ax = plt.subplots(
         figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
     )
+    print(all_results)
     sns.violinplot(data=all_results, x="estimate", y="model", ax=ax)
-    sns.swarmplot(data=all_results, x="estimate", y="model", ax=ax)
+    # sns.swarmplot(data=all_results, x="estimate", y="model", ax=ax)
 
-    ax.set_xlabel("confidence estimate return by the model")
+    ax.set_xlabel("confidence estimate returned by the model")
     ax.set_ylabel("model")
 
-    range_frame(ax, np.array([1, 5], np.array([0, 2])))
+    range_frame(ax, np.array([1, 5]), np.array([0, 2]))
 
     fig.savefig(figures / "confidence_score_distributions.pdf", bbox_inches="tight")
 

### 2024-03-16 src/tex/ms.tex
manuscript compiles
pin env to make embeddings work

@@
 
+\begin{figure}
+    \centering
+    \includegraphics{figures/question_diversity.pdf}
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.}}
+    \label{fig:question_diversity}
+    \script{plot_question_diversity.py}
+\end{figure}
+
 While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-For this reason, \chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
-\variable{output/non_mcq_questions.txt} open-ended questions).
+For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
 As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
@@
 
+\appendix
+\section{Appendix}
+\input{appendix.tex}
+
 \bibliography{references}
 
-\appendix
-\input{appendix.tex}
-
 \end{document}

### 2024-03-16 src/scripts/embed_questions.py

@@ -7,13 +7,14 @@ from utils import obtain_chembench_repo
 import os
 import pickle
 from paths import data
+import numpy as np
 
 load_dotenv("../.env")
 set_llm_cache(SQLiteCache(database_path=".langchain.db"))
 
 if __name__ == "__main__":
 
-    repo = obtain_chembench_repo
+    repo = obtain_chembench_repo()
     JSON_PATHS = glob(os.path.join(repo, "data/**/*.json"), recursive=True)
 
     df = build_question_df(JSON_PATHS)
@@ -34,10 +35,9 @@ if __name__ == "__main__":
 
     # BART example
     q_classifier_ = QClassifier(embedding_model="bart", topics=TOPICS)
-
+    q_classifier_.batch_size = 32
     embeddings = q_classifier_.get_embedding(df["formatted"].to_list())
 
     names = df["name"].to_list()
 
-    with open(data / "embeddings.pkl") as handle:
-        pickle.dumps({"names": names, "embeddings": embeddings}, handle)
+    np.save(data / "embeddings.npy", embeddings)

### 2024-03-17 src/scripts/count_json_files.py
update the counting scripts

@@ -1,8 +1,7 @@
 import os
-from glob import glob
 from utils import obtain_chembench_repo
 from paths import output as output_path
-
+from loguru import logger 
 
 def count_json_files_in_directory(directory):
     json_files = []
@@ -23,17 +22,22 @@ def process_directory(directory):
             elif sub_dir == 'icho':
                 num_json_files = count_json_files_in_directory(sub_dir_path)
                 total_json_files += num_json_files
-                with open(output_path / "json_file_counts.txt", 'a') as f:
-                    f.write(f'{sub_dir} has {num_json_files} .json files\n')
+                with open(output_path / f"json_file_counts_{sub_dir}.txt", 'a') as f:
+                    f.write(f'{sub_dir}' + '\endinput')
+
+                logger.info(f"Number of json files in {sub_dir}: {num_json_files}")
             else:
                 num_json_files = count_json_files_in_directory(sub_dir_path)
                 total_json_files += num_json_files
-                with open(output_path / "json_file_counts.txt", 'a') as f:
-                    f.write(f'{sub_dir} has {num_json_files} .json files\n')
+                with open(output_path / f"json_file_counts_{sub_dir}.txt", 'a') as f:
+                    f.write(f'{sub_dir}' + '\endinput')
 
+                logger.info(f"Number of json files in {sub_dir}: {num_json_files}")
+    
     with open(output_path / "json_file_counts.txt", 'a') as f:
-        f.write(f'Total number of JSON files in {directory}: {total_json_files}\n')
+        f.write(f'{total_json_files}' + '\endinput')
 
+    logger.info(f"Total number of json files: {total_json_files}")
 
 def process_safety_directory(safety_dir):
     total_safety_json_files = 0
@@ -42,8 +46,10 @@ def process_safety_directory(safety_dir):
         if os.path.isdir(safety_subdir_path):
             num_json_files = count_json_files_in_directory(safety_subdir_path)
             total_safety_json_files += num_json_files
-            with open(output_path / "json_file_counts.txt", 'a') as f:
-                f.write(f'{safety_subdir} of safety has {num_json_files} .json files\n')
+            with open(output_path / f"json_file_counts_{safety_subdir}.txt", 'a') as f:
+                f.write(f'{safety_subdir}' + '\endinput')
+
+            logger.info(f"Number of json files in {safety_subdir}: {num_json_files}")
 
     return total_safety_json_files
 

### 2024-03-17 src/tex/ms.tex

--- 2024-03-15 src/tex/ms.tex
@@
 
+\begin{figure}
+    \centering
+    \includegraphics{figures/question_diversity.pdf}
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.}}
+    \label{fig:question_diversity}
+    \script{plot_question_diversity.py}
+\end{figure}
+
 While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-For this reason, chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs.
-\variable{output/non_mcq_questions.txt} open-ended questions).
+For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
 As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
@@
 % Calculation vs. no calculation
-Given that some of the questions in chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
+Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
 For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.
@@
 These exercises focus on determining the number of isomers for specific molecules.
-To perform these calculations, we utilize the RDKit package \cite{bento2020open}.
+To perform these calculations, we utilize the RDKit package \cite{rdkit}.
 \subparagraph{Electron count of atoms}
@@
 These inquiries involve determining the electron count of atoms in various chemical contexts.
-The calculations for these tasks are facilitated using the RDKit package \cite{bento2020open}.
+The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
 \subparagraph{Number of isomers}
@@
 These exercises aim to determine the number of isomers for given molecules.
-The calculations for these assessments are conducted using the RDKit package \cite{bento2020open}.
+The calculations for these assessments are conducted using the RDKit package \cite{rdkit}.
 \subparagraph{Point group for molecules}
@@
 \subparagraph{SMILES-IUPAC}
-Web scrapped SMILES IUPAC pairs from the PubChem database \cite{pubchem} and then framed questions involve converting SMILES notation to IUPAC nomenclature, utilizing information from the PubChem database.
+Web scrapped \gls{smiles} \gls{iupac} pairs from the PubChem database \cite{pubchem} and then framed questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature, utilizing information from the PubChem database.
 
-\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
+\paragraph{Programmatically generated questions} 
+\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\cite{Irwin_2012} and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
 We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.

### 2024-03-17 src/tex/ms.tex

@@
 \subparagraph{Analytical chemistry}
-The questions are sourced from the master student examination paper at the University of Jena, Germany. These questions focus on the principles and techniques employed in analytical chemistry.
+The questions are based on master student examination papers at the University of Jena (Germany). 
+These questions focus on the principles and techniques employed in analytical chemistry.
 
 \subparagraph{Combustion engineering}
-These questions are derived from the master student examination paper at the University of Magdeburg, Germany. They explore topics related to combustion processes and engineering principles.
+These questions are based on master student examination papers at the University of Magdeburg (Germany). 
+They explore topics related to combustion processes and engineering principles.
 
-\subparagraph{Functional materials and nano-materials}
-The questions are gathered from a seminar conducted at the University of Jena, delving into topics concerning functional materials and nano-materials, exploring their properties and applications.
+\subparagraph{Functional materials and nanomaterials}
+The questions based on exercises in a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
 
 \subparagraph{General chemistry}
-The questions originate from examinations for grade X at Marist Comprehensive Academy Uturu, Nigeria. They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+The questions originate from examinations for grade 10 at Marist Comprehensive Academy Uturu (Nigeria). 
+They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+In addition, we included questions based on the general chemistry courses of the Bachelor of Science in Chemistry at the Technical University of Munich (Germany).
 
 \subparagraph{International Chemical Olympiad (\gls{ICHO})}
-The questions are sourced from various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova, providing challenging problems for competitive chemistry.
+The questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
 
 \subparagraph{Material synthesis}
-These questions are derived from seminars on material synthesis at the University of Jena, focusing on the methods and techniques employed in the synthesis of various materials.
+These questions are based on seminars on material synthesis at the University of Jena (Germany), focusing on the methods and techniques employed in the synthesis of various materials.
 
 \subparagraph{Organic reactivity}
-The questions are primarily sourced from tutorial papers from the Technical University of Munich, Germany, exploring reactions and mechanisms in organic chemistry.
+The questions are primarily sourced from tutorials and exam papers at for the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), exploring reactions and mechanisms in organic chemistry.
 
 \subparagraph{Periodic table}
-The questions are framed based on the reference periodic table from the Royal Society of Chemistry (\url{https://www.rsc.org/periodic-table/}), covering various aspects and trends of the periodic table.
+The questions are manually created to cover various aspects and trends of the periodic table.
 
 \subparagraph{Polymer chemistry}
-Questions are sourced from examinations held at the University of Hamburg, covering topics related to polymer chemistry and its applications.
+Questions are based on examinations held at the University of Hamburg (Germany), covering topics related to polymer chemistry and its applications.
 
@@
 This category comprises a total of 4792 questions, covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \cite{pubchem}, contributing 976 questions.
-Additionally, it explores crucial topics like materials compatibility, with 3200 questions, essential for understanding the hazardous impact of mixing certain materials and chemicals. 
-Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of 90 questions, focusing on enhancing federal oversight of chemical safety and security by promoting coordination between federal agencies like OSHA, EPA, and DHS, as well as state governments to ensure compliance with safety and security requirements for chemical facilities.
+Additionally, it explores crucial topics like materials' compatibility, with 3200 questions, essential for understanding the hazardous impact of mixing certain materials and chemicals. 
+Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of 90 questions.
+These questions are based on the question bank used for the \enquote{Sachkundeprüfung} (expertise examination) according to \§ 11 of the \enquote{Chemikalien-Verbotsverordnung} (Chemical Prohibition Ordinance).
 Moreover, this category addresses questions related to chemical storage and compatibility, totaling 297 questions, and covers topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets.
-The chemical safety category stands out as one of the primary focus areas within the chembench dataset.
-
-\subparagraph{Shape of molecules}
-The questions are sourced from examinations held at the Technical University of Munich, Germany, focusing on determining the shape and geometry of molecules.
+The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
 
@@
 The questions regarding oxidation states are extracted from the website \url{https://www.cheminfo.org/}. 
-These exercises focus on determining the number of isomers for specific molecules.
-To perform these calculations, we utilize the RDKit package \cite{rdkit}.
+These exercises focus on determining the number of isomers for specific molecules. To perform these calculations, we utilize the RDKit package \cite{rdkit}.
+
 \subparagraph{Electron count of atoms}
 Questions pertaining to the electron count of atoms are gathered from the website \url{https://www.cheminfo.org/}. 
-These inquiries involve determining the electron count of atoms in various chemical contexts.
-The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
+These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
+
 \subparagraph{Number of isomers}
 The questions concerning the number of isomers are sourced from the website \url{https://www.cheminfo.org/}. 
-These exercises aim to determine the number of isomers for given molecules.
-The calculations for these assessments are conducted using the RDKit package \cite{rdkit}.
+These exercises aim to determine the number of isomers for given molecules. The calculations for these assessments are conducted using the RDKit package \cite{rdkit}.
+
 \subparagraph{Point group for molecules}
@@
 \subparagraph{Number of isomers}
-We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled SMILES from the ZINC dataset.
+We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled \gls{smiles} from the ZINC dataset.\cite{Irwin_2012}
 To obtain the answers for false options we sampled from the set of number of isomers.

### 2024-03-17 src/tex/ms.tex
update question description

--- 2024-03-16 src/tex/ms.tex
@@
 
+\subparagraph{Analytical chemistry}
+The \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the University of Jena (Germany). 
+These questions focus on the principles and techniques employed in analytical chemistry.
+
+\subparagraph{Combustion engineering}
+These \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} questions are based on master student examination papers at the University of Magdeburg (Germany). 
+They explore topics related to combustion processes and engineering principles.
+
+\subparagraph{Functional materials and nanomaterials}
+The \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
+
+\subparagraph{General chemistry}
+\variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} questions originate from examinations for grade 10 at Marist Comprehensive Academy Uturu (Nigeria). 
+They cover fundamental concepts in chemistry suitable for students at the secondary school level.
+In addition, we included questions based on the general chemistry courses of the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), focusing on inorganic chemistry (\variable{output/question_count_per_dir/json_file_counts_ac_faessler_tum.txt}) and principles of chemistry (\variable{output/question_count_per_dir/json_file_counts_pum_tum.txt}).
+
+\subparagraph{International Chemical Olympiad (\gls{ICHO})}
+\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
+
+\subparagraph{Material synthesis}
+\variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} questions are based on seminars on material synthesis at the University of Jena (Germany), focusing on the methods and techniques employed in the synthesis of various materials.
+
+\subparagraph{Organic reactivity}
+\variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt}  questions are primarily sourced from tutorials and exam papers at for the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), exploring reactions and mechanisms in organic chemistry.
+
+\subparagraph{Periodic table}
+\variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} questions are manually created to cover various aspects and trends of the periodic table.
+
+\subparagraph{Polymer chemistry}
+\variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} questions are based on examinations held at the University of Hamburg (Germany), covering topics related to polymer chemistry and its applications.
+
+\subparagraph{Chemical reactivity}
+\variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
+
+\subparagraph{Chemical safety} 
+This category comprises a total of \variable{output/question_count_per_dir/json_file_counts_safety.txt} questions, including semi-programmatically created questions covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \cite{pubchem} (see below for details).
+Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals.
+Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions. 
+These questions are based on the question bank used for the \enquote{Sachkundeprüfung} (expertise examination) according to \§ 11 of the \enquote{Chemikalien-Verbotsverordnung} (Chemical Prohibition Ordinance).
+Moreover, this category addresses questions lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
+The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
+
+\subparagraph{\gls{NMR} spectroscopy}
+Questions are based on images shared via the Twitter account \url{https://twitter.com/NMRspectroscopy}, focusing choosing appropriate \gls{nmr} experiments.
+
 \paragraph{Semi-programatically generated questions}
@@
 \subparagraph{Oxidation states}
+\variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} questions regarding oxidation states are based on data from the website \url{https://www.cheminfo.org/}.
 
-\subparagraph{Electron counts}
+\subparagraph{Total electron count of molecules}
+\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} questions pertaining to the total electron count of molecules are based on data from the website \url{https://www.cheminfo.org/}.
+These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
 
-\subparagraph{Number of NMR peaks} To generate tasks about the number of \gls{nmr} peaks, we randomly sampled SMILES from the ZINC dataset and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
+\subparagraph{Number of isomers}
+We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled \gls{smiles} from the ZINC dataset.\cite{Irwin_2012}
+To obtain the answers for false options we sampled from the set of number of isomers. In total, we generated \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} questions.
+
+\subparagraph{Point group for molecules}
+\variable{output/question_count_per_dir/json_file_counts_point_group.txt} questions involve predicting the point group for molecules.
+Our ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption}) is utilized for data generation.
+Given a molecule, it uses spglib\cite{spglib} to identify the point group. 
+Assigned point groups were manually verified to only include well-defined cases.
+
+\subparagraph{SMILES-IUPAC}
+We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \cite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature.
+\variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt} questions involve the reverse task, i.e., converting \gls{iupac} nomenclature to \gls{smiles} notation.
+
+\subparagraph{Number of NMR peaks} 
+To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\cite{Irwin_2012} and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
 We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
@@
 \subparagraph{GHS, hazard statements and DAI}
-The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.~\cite{pubchem}
+The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.\cite{pubchem}
 The former two have been mined via the PubChem \gls{api}, while the latter has been manually compiled. 
+In total, we generated \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} questions about the \gls{ghs} classification of chemicals, \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} questions about definition of hazard statements, and \variable{output/question_count_per_dir/json_file_counts_dai.txt} questions about \glspl{dai}.
 The \glspl{dai} have been curated to contain only records approved by the \gls{who}.
 The chemicals in this class of questions belong to one of the three classes: pesticides (e.g., calcium arsenate), insecticides (e.g., cyfluthrin) or herbicides (e.g. 2,4-D).
-All data was saved in tabular form, and then we programmatically created questions (\variable{output/num_h_statements.txt} hazard statement definitions, \variable{output/num_pictograms.txt} chemical-\gls{ghs} pictogram matching, \variable{output/num_dai.txt} \glspl{dai}).
-
-\subparagraph{Number of isomers}
-We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled SMILES from the ZINC dataset.
-To obtain the answers for false options we sampled from the set of number of isomers.
 

### 2024-03-17 src/tex/acronymns.tex
start using showyourwork
update paper
rebuild
start note about tiny benchmarks
feat: add acronyms, fix credits bug, update pipeline
added safety description
fix issues with acronyms

@@
 \newacronym{mcq}{MCQ}{multiple-choice question}
+\newacronym{rest}{REST}{Representational State Transfer}
+\newacronym{orm}{ORM}{object relational mapping}
+\newacronym{dai}{DAI}{daily allowed intake}
+\newacronym{ghs}{GHS}{Globally Harmonized System of Classification and Labelling of Chemicals}
+\newacronym{who}{WHO}{World Health Organization}
+\newacronym{nmr}{NMR}{nuclear magnetic resonance}
+\newacronym{helm}{HELM}{Holistic Evaluation of Language Models}
+\newacronym{smiles}{SMILES}{Simplified Molecular Input Line-Entry System}
+\newacronym{pca}{PCA}{principal component analysis}
+\newacronym{iupac}{IUPAC}{International Union of Pure and Applied Chemistry}

### 2024-03-17 src/tex/ms.tex
fix issues with acronyms

@@
 
-\subparagraph{International Chemical Olympiad (\gls{ICHO})}
+\subparagraph{International Chemical Olympiad}
 \variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
@@
 Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions. 
-These questions are based on the question bank used for the \enquote{Sachkundeprüfung} (expertise examination) according to \§ 11 of the \enquote{Chemikalien-Verbotsverordnung} (Chemical Prohibition Ordinance).
+These questions are based on the question bank used for the \enquote{Sachkundeprüfung} (expertise examination) according to §11 of the \enquote{Chemikalien-Verbotsverordnung} (Chemical Prohibition Ordinance).
 Moreover, this category addresses questions lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
@@
 
-\subparagraph{\gls{NMR} spectroscopy}
+\subparagraph{NMR spectroscopy}
 Questions are based on images shared via the Twitter account \url{https://twitter.com/NMRspectroscopy}, focusing choosing appropriate \gls{nmr} experiments.

### 2024-03-17 src/tex/ms.tex
update contributions

@@
 Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions. 
-These questions are based on the question bank used for the \enquote{Sachkundeprüfung} (expertise examination) according to §11 of the \enquote{Chemikalien-Verbotsverordnung} (Chemical Prohibition Ordinance).
+These questions are based on the question bank used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
 Moreover, this category addresses questions lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 

### 2024-03-17 src/tex/ms.tex
make compile and embedding script

@@
     We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4, and found that the best models, on average, outperform the best human chemists in our study.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g. about the safety profile of chemicals.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g., about the safety profile of chemicals.
 
@@
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
-Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also an tool-augmented systems).
+Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
 
@@
 The distribution of topics is also evident from \Cref{fig:pca} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
-In this representation, semantically similar questions are close to each other, and we color the points based on a classification into XX topics. 
+In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
 It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.
@@
     \includegraphics{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.}}
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+We then project the embeddings into a two-dimensional space using \gls{pca}.
+We color the points based on a classification into topics.}
     \label{fig:question_diversity}
@@
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
+With the default tool augmented setup (using the so-called ReAct method\cite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
 This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
@@
 
-In this spider chart, the worst score for every dimension is 0 (no question answered completely correctly) and the best score is 1 (all questions answered correctly). 
+In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
 Thus, a larger colored area indicates a better performance. 

### 2024-03-17 src/scripts/plot_question_diversity.py

@@ -1,8 +1,9 @@
 import numpy as np
 import matplotlib.pyplot as plt
-
+import pandas as pd
 
 from sklearn.decomposition import PCA
+from sklearn.preprocessing import StandardScaler
 import seaborn as sns
 from paths import data, scripts, figures
 from utils import ONE_COL_GOLDEN_RATIO_HEIGHT_INCH, ONE_COL_WIDTH_INCH
@@ -12,8 +13,11 @@ plt.style.use(scripts / "lamalab.mplstyle")
 
 def load_embeddings_and_labels():
     # read npy file
-    embeddings = np.load(data / "gpt_embeddings/embeddings.npy")
-    labels = np.load(data / "gpt_embeddings/labels.npy")
+    embeddings = np.load(data / "embeddings.npy")
+    scaler = StandardScaler()
+    embeddings = scaler.fit_transform(embeddings)
+    df = pd.read_pickle(data / "questions.pkl")
+    labels = df["topic"].values
     return embeddings, labels
 
 

### 2024-03-18 src/tex/ms.tex
cleaned up references

@@
 In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
-The frequency of this fallback being triggered was very different for different models (see XX).
-Manual verification (see \Cref{sec:manually-verified-parsing}) indicates that this \gls{llm}-based parsing is not a relevant error source.
+Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
 

### 2024-03-18 src/scripts/wordcloud.py

@@ -1,7 +1,7 @@
 from wordcloud import WordCloud
 import matplotlib.pyplot as plt
 from utils import ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH
-from paths import figures, output
+from paths import figures, output, data
 import textacy
 from textacy import text_stats as ts
 from chembench.utils import (
@@ -79,6 +79,6 @@ def plot_reading_ease(df):
 if __name__ == "__main__":
     import pandas as pd
 
-    df = pd.read_pickle(output / "questions.pkl")
+    df = pd.read_pickle(data / "questions.pkl")
     create_wordcloud(df)
     plot_reading_ease(df)

### 2024-03-18 src/scripts/plot_question_diversity.py
update performance plot code

@@ -38,7 +38,7 @@ def plot_pca_map():
     # place legend outside of the plot
 
     plt.axis("off")
-    sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=100, alpha=0.7)
+    sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=10, alpha=0.7)
     f.savefig(figures / "question_diversity.pdf", bbox_inches="tight")
 
 

### 2024-03-18 src/tex/ms.tex
update performance plot code
add references

@@
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
@@
     \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question.
-    The figure shows that not all aspects of chemistry are equally represented in our corpus. The corpus currently focuses on safety-related aspects.}
+    The figure shows that not all aspects of chemistry are equally represented in our corpus.
+The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}.
+This figure represents the combined count of \glstext{mcq} and open-ended questions.}
     \label{fig:topic_barplot}
@@
     \centering
+    %\includegraphics{figures/human_vs_models_bar.pdf}
     \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly.
@@
     \label{fig:human_vs_models_bar}
+    \script{plot_overview_performance_plot.py}
 \end{figure}

### 2024-03-18 src/tex/ms.tex
implemented subset analysis

@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
-As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
+%As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
 
@@
 \subparagraph{Completion models}
-We used \texttt{Galactica-120b}\cite{taylor2022galactica} with the default settings
+We used \texttt{Galactica-120b}\cite{taylor2022galactica} with the default settings.
 
-\subparagraph{Instruction-tuned models} \texttt{Claude 2} \texttt{Claude3 (Opus)}\cite{anthropicClaudeModelFamily2024}
-\texttt{GPT-4}\cite{openai2024gpt4}
-\texttt{GPT-3.5-turbo}\cite{brown2020language}
-\texttt{Gemini Pro}\cite{gemini}
+\subparagraph{Instruction-tuned models} In addition, we used \texttt{Claude 2}, \texttt{Claude3 (Opus)},\cite{anthropicClaudeModelFamily2024} \texttt{GPT-4},\cite{openai2024gpt4} \texttt{GPT-3.5-turbo},\cite{brown2020language}
+\texttt{Gemini Pro},\cite{gemini}
 \texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
-\texttt{Llama2-70b}\cite{touvron2023llama}
+\texttt{Llama2-70b},\cite{touvron2023llama}
 \texttt{pplx-7b-chat}
@@
 
+The authors thank Julian Kimmig for feedback on the web app.
+
 \section*{Conflicts of interest}

### 2024-03-18 src/tex/ms.tex
update data description

@@
 \subparagraph{Analytical chemistry}
-The \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the University of Jena (Germany). 
+\variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the University of Jena (Germany). 
 These questions focus on the principles and techniques employed in analytical chemistry.
@@
 \subparagraph{Combustion engineering}
-These \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} questions are based on master student examination papers at the University of Magdeburg (Germany). 
+\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} questions are based on master student examination papers at the University of Magdeburg (Germany). 
 They explore topics related to combustion processes and engineering principles.
@@
 \subparagraph{Functional materials and nanomaterials}
-The \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
+\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
 
@@
 \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
+
+\subparagraph{Textbook questions}
+\variable{output/question_count_per_dir/json_file_counts_oup.txt} questions are based on questions from various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography.
 

### 2024-03-18 src/tex/ms.tex
update analysis

@@
     \centering
-    %\includegraphics{figures/human_vs_models_bar.pdf}
+    \includegraphics{figures/human_subset_performance.pdf}
     \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly.
@@
 While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
-% In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{outputs/subset_scores/is_number_nmr_peaks_gpt-4.txt}.
-% While this question is relatively easy (\variable{outputs/human_subset_scores/is_number_nmr_peaks.txt}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
-% This is mirrored in the questions about point groups of compounds, which is models struggle with more than humans ().
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4)
+While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs.
+\variable{output/human_subset_scores/is_number_of_isomers.txt}).
+These findings also shin an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+On those questions the models tend to perform better than on some our semi-automatically constructed tasks.
 

### 2024-03-18 src/scripts/analyze_performance_per_source.py

@@ -115,7 +115,7 @@ if __name__ == "__main__":
     if not os.path.exists(outdir_humans):
         os.makedirs(outdir, exist_ok=True)
 
-    human_scores = obtain_subset_scores_humans(human_scores["raw_scores"], outdir)
+    human_scores = obtain_subset_scores_humans(human_scores["raw_scores"], outdir_humans)
 
     all_scores = pd.DataFrame(model_scores + human_scores)
 

### 2024-03-18 src/scripts/plot_confidence_score_distributions.py

@@ -39,18 +39,20 @@ def plot():
 
     claude_2["model"] = "Claude 2"
 
-    claude_3 = pd.read_csv(
-        os.path.join(
-            chembench,
-            "reports",
-            "confidence_estimates",
-            "confidence_estimates",
-            "results_claude3.csv",
-        )
-    )
+    # claude_3 = pd.read_csv(
+    #     os.path.join(
+    #         chembench,
+    #         "reports",
+    #         "confidence_estimates",
+    #         "confidence_estimates",
+    #         "results_claude3.csv",
+    #     )
+    # )
 
-    claude_3["model"] = "Claude 3"
-    all_results = pd.concat([gpt, claude_2, claude_3]).reset_index(drop=True)
+    # claude_3["model"] = "Claude 3"
+    # all_results = pd.concat([gpt, claude_2, claude_3]).reset_index(drop=True)
+
+    all_results = pd.concat([gpt, claude_2]).reset_index(drop=True)
 
     fig, ax = plt.subplots(
         figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
@@ -73,7 +75,7 @@ def plot():
             xmax=counts,
             linewidth=5,
             alpha=0.5,
-            color=model_color_map[model],
+            # color=model_color_map[model],
         )
 
         ax.plot(
@@ -81,9 +83,11 @@ def plot():
             counts.index,
             "o",
             markersize=5,
-            alpha=0.5,
-            color=model_color_map[model],
+            alpha=1.0,
+            # color=model_color_map[model],
         )
+
+    range_frame(ax, np.array(counts), np.array([0.5, 5.5]))
 
     ax.set_xlabel("Number of Questions")
     ax.set_ylabel("confidence estimate")
@@ -140,7 +144,7 @@ def plot_violin():
     ax.set_xlabel("confidence estimate returned by the model")
     ax.set_ylabel("model")
 
-    range_frame(ax, np.array([1, 5]), np.array([0, 2]))
+    range_frame(ax, np.array([0.5, 5.5]), np.array([0, 2]))
 
     fig.savefig(
         figures / "confidence_score_distributions_violins.pdf", bbox_inches="tight"

### 2024-03-18 src/scripts/plot_question_diversity.py

@@ -6,7 +6,11 @@ from sklearn.decomposition import PCA
 from sklearn.preprocessing import StandardScaler
 import seaborn as sns
 from paths import data, scripts, figures
-from utils import ONE_COL_GOLDEN_RATIO_HEIGHT_INCH, ONE_COL_WIDTH_INCH
+from utils import (
+    ONE_COL_GOLDEN_RATIO_HEIGHT_INCH,
+    ONE_COL_WIDTH_INCH,
+    TWO_COL_WIDTH_INCH,
+)
 
 plt.style.use(scripts / "lamalab.mplstyle")
 
@@ -34,11 +38,14 @@ def plot_pca_map():
     pca_1 = pca_1[shuffled_indices]
     pca_2 = pca_2[shuffled_indices]
     labels = np.array(labels)[shuffled_indices]
-    f = plt.figure(figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH))
+    f, ax = plt.subplots(figsize=(TWO_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH))
     # place legend outside of the plot
+    # make legend in two columns
 
     plt.axis("off")
-    sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=10, alpha=0.7)
+    sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=10, alpha=0.7, ax=ax)
+    ax.legend(loc="center left", ncol=3, bbox_to_anchor=(0, 1.2))
+    f.tight_layout()
     f.savefig(figures / "question_diversity.pdf", bbox_inches="tight")
 
 

### 2024-03-18 src/tex/ms.tex
update plots

@@
 In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+Our benchmark consists of \variable{output/total_number_of_questions.txt} question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
 Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
@@
     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.}
+    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
+    \Cref{fig:barplot_all_correct_all_questions} gives an overview of the performance of various models on the entire corpus.
+    }
     \label{fig:human_vs_models_bar}
@@
 We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
-
-% overall performance profile of various model is comparable
-\begin{figure}
-    \centering
-    \includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
-    \caption{Caption}
-    \label{fig:performance_by_topic}
-\end{figure}
 
@@
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
-To investigate this, we prompted some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
+To investigate this, we prompted\cite{xiong2023llms} some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
 
@@
 % - More systematic human baseline: Ask people to come in, do not allow to skip questions
+
+% clearly defined metrics important for deep learning https://towardsdatascience.com/the-road-to-biology-2-0-will-pass-through-black-box-data-bbd00fabf959
 

### 2024-03-18 src/scripts/analyze_performance_per_source.py

@@ -85,10 +85,14 @@ def obtain_subset_scores_humans(data_dict, outdir):
                 subset_scores.append(score)
         except Exception as e:
             logger.warning(f"Failed for {subset} due to {e}")
-        mean_subset_score = np.nanmean(subset_scores)
-        with open(os.path.join(outdir, f"{subset}.txt"), "w") as handle:
-            handle.write(str(int(np.round(mean_subset_score * 100, 0))) + "\endinput")
-
+        try:
+            mean_subset_score = np.nanmean(subset_scores)
+            with open(os.path.join(outdir, f"{subset}.txt"), "w") as handle:
+                handle.write(
+                    str(int(np.round(mean_subset_score * 100, 0))) + "\endinput"
+                )
+        except Exception as e:
+            logger.warning(f"Failed for {subset} due to {e}")
     all_scores.append({"model": "human", "score": score, "subset": subset})
 
     return all_scores

### 2024-03-18 src/scripts/collect_human_scores.py

@@ -9,8 +9,11 @@ if __name__ == "__main__":
     chembench_repo = obtain_chembench_repo()
     human_files = os.listdir(os.path.join(chembench_repo, "reports", "humans"))
     human_files = [
-        os.path.join(chembench_repo, "reports", "humans", p) for p in human_files
+        os.path.join(chembench_repo, "reports", "humans", p)
+        for p in human_files
+        if not "clr6ugeta0000i708dr5c308o" in p
     ]
+
     # we only consider humans with at least 100 questions
     human_files = [p for p in human_files if len(glob(os.path.join(p, "*.json"))) > 100]
 

### 2024-03-18 src/tex/ms.tex
update plots

@@
 In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt} question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
 Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
@@
 This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
-The distribution of topics is also evident from \Cref{fig:pca} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
+The distribution of topics is also evident from \Cref{fig:question_diversity} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
 In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.
+It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as large distinct clusters.
 
@@
 In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4)
-While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt}) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
 This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs.
-\variable{output/human_subset_scores/is_number_of_isomers.txt}).
+\variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
 These findings also shin an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 

### 2024-03-19 src/tex/ms.tex
plot confidence vs. performance in main text
ensure links for all variables

@@
     \centering
-    %\includegraphics{figures/estimated_difficulty.pdf}
-    \caption{Caption}
-    \label{fig:estimated_difficulty}
+    \includegraphics{figures/confidence_vs_performance_overall.pdf}
+    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.}}
+    \label{fig:confidence_vs_performance}
+    \script{joint_analysis_confidence_performance.py}
 \end{figure}
@@
 To ensure that the questions do not enter a training dataset we use the canary string of the Big Bench project.
-This requires that \Gls{llm} developers filter their training dataset for this canary string.
+This requires that \Gls{llm} developers filter their training dataset for this canary string.\cite{openai2024gpt4, srivastava2022beyond}
 

### 2024-03-19 src/tex/authors.tex
feat: add acronyms, fix credits bug, update pipeline
update authors
update authors and prepare confidence score distribution
added safety description
pin env to make embeddings work
set author ordering
clean up credit order
update contributions
make compile and embedding script
update affiliations

@@
 
-\credit{Adrian~Mirza}{0,1,1,0,1,0,0,0,1,0,0,0,0,0}
-\credit{Sreekanth~Kunchapu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+% Main contributors
+\credit{Adrian~Mirza}{1,1,1,0,1,0,0,0,1,0,1,1,0,1}
+\credit{Nawaf~Alampara}{1,1,1,0,1,0,0,0,1,0,1,0,0,1}
+\credit{Sreekanth~Kunchapu}{1,1,0,0,1,0,0,0,0,0,1,0,0,1}
+
+% Question contributors
+\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,0}
 \credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Jan Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Lea Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-%ORCID ID:  0000-0002-5968-8706 Affiliation:    Institute for Technical Chemistry and Environmental Chemistry (ITUC) and Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany
 \credit{Mara~Wilhemli}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
 \credit{Macjonathan~Oreke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Michael~Pieler}{1,0,0,0,0,0,0,0,0,0,0,0,a0,0}
-\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,0}
+
+% Question answerers
+\credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Lea Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Jan Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
+\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+
+% senior authors 
+\credit{Michael~Pieler}{1,0,0,0,0,0,0,0,0,0,0,0,0,0}
 \credit{Philippe~Schwaller}{1,1,0,1,1,0,0,0,0,1,0,0,0,0}
-\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
-\credit{Berend~Smit}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
 \credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}
 
-\author[1,*]{Nawaf~Alampara~\orcidlink{}}
-
-\author[1,*]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
-
-\author[1,*]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
-
-\author[1]{Lea~Klepsch~\orcidlink{0000-0003-4978-4670}}
-\author[1]{Jan~Peschel~\orcidlink{0009-0002-4787-2757}}
-\author[1]{Jakob~Meyer}
-\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
-\author[1]{Maximilian~Greiner}
-\author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
-
-\author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
-\author[ituc]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
-\author[1]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
-\author[dresden]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
-\author[lsmo]{Amir~Elahi}
-\author[lsmo]{Berend~Smit~\orcidlink{0000-0003-4653-8562}}
-
-\author[qpi]{Aswanth~Krishnan}
-
+\author[1,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
+\author[1,$\star$]{Nawaf~Alampara~\orcidlink{}}
+\author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
+\author[]{Benedict~Emoekabu}
+\author[qpi]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
+\author[nccrcat,liac]{Tanya~Gupta}
 \author[1]{Mara~Willhelmi}
-
 \author[1]{Macjonathan~Oreke}
 
-\author[Occlu]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
+\author[cambridge]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
+\author[1]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
+\author[lsmo]{Amir~Elahi}
+\author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
+\author[1]{Maximilian~Greiner}
+\author[1]{Lea~Klepsch~\orcidlink{0000-0003-4978-4670}}
+\author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
+\author[ituc, 2]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[1]{Jakob~Meyer}
+\author[1]{Jan~Peschel~\orcidlink{0009-0002-4787-2757}}
+\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
 
-\author[]{Benedict~Emoekabu}
+\author[1]{Johanna Schreiber}
+\author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[dresden]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
 
-\author[nccrcat,liac]{Tanya~Gupta}
-
+% Senior authors
+\author[open,stability]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
 \author[nccrcat,liac]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-
-\author[open,stability]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
-
 \author[1,2, 3]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
@@
 \affil[stability]{Stability.AI}
-\affil[ituc]{Institute for Technical Chemistry and Environmental Chemistry (ITUC) and Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[qpi]{Qpi volta}
-\affil[occlu]{Occlutech GmbH, Jena, Germany}
+\affil[ituc]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[qpi]{QpiVolta Technologies Pvt Ltd}
+\affil[cambridge]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, UK}
 
 \affil[]{\texttt{mail@kjablonka.com}}
-\affil[*]{These authors contributed equally.}
+\affil[$\star$]{These authors contributed equally.}

### 2024-03-19 src/tex/ms.tex
update affiliations
cleanup

@@
     \centering
-    \includegraphics{figures/question_diversity.pdf}
+    \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
     \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
@@
 \variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
-These findings also shin an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+These findings also shine an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
 On those questions the models tend to perform better than on some our semi-automatically constructed tasks.
@@
     \includegraphics{figures/confidence_vs_performance_overall.pdf}
-    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.}}
+    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model.
+To obtain those estimates, we prompted the models to return a confidence score on an ordinal scale.
+We then plot the correctness of the answers (which can be either completely correct or completely incorrect) against the confidence score.
+        The stripplots show the individual data points, the lines show the mean.}
     \label{fig:confidence_vs_performance}
@@
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
-To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{showyourwork!} framework.\cite{Luger2021}
-The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}. 
+To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\\showyourwork} framework.\cite{Luger2021}
+The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
 To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

### 2024-03-19 src/tex/appendix.tex
update plots
cleanup
nature bibstyle

@@
     \label{fig:flesch_kincaid_reading_ease}
+    \script{wordcloud.py}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of multiple choice questions and general questions per topic.}
+    \script{plot_statistics.py}
 \end{figure}
@@
 \begin{figure}
+    \centering
+    \includegraphics{figures/overall_performance.pdf}
+    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
+Scores computed on the entire \chembench corpus.}
     \label{fig:barplot_all_correct_all_questions}
+    \script{plot_overview_performance_plot.py}
 \end{figure}
@@
 \end{figure}
+
+\printnoidxglossary[type=\acronymtype]  % https://github.com/tectonic-typesetting/tectonic/issues/704

### 2024-03-19 src/tex/ms.tex
nature bibstyle

@@
 To obtain those estimates, we prompted the models to return a confidence score on an ordinal scale.
-We then plot the correctness of the answers (which can be either completely correct or completely incorrect) against the confidence score.
-        The stripplots show the individual data points, the lines show the mean.}
+We then plot the correctness of the answers [which can be either completely correct (1) or completely incorrect (0)] against the confidence score.
+        The stripplots show the individual data points, the lines show the mean and the errorbars show the standard error of the mean.
+The figure shows that the reliability of the confidence estimates varies widely across models.
+While the ones for GPT-4 and Claude-2 are not particularly reliable, the ones for Claude-3 follow the expected trend.}
     \label{fig:confidence_vs_performance}

### 2024-03-19 src/scripts/plot_confidence_score_distributions.py

@@ -69,14 +69,14 @@ def plot():
         # ensure that the counts are sorted by the index
         counts = counts.sort_index()
 
-        ax.hlines(
-            counts.index,
-            xmin=0,
-            xmax=counts,
-            linewidth=5,
-            alpha=0.5,
-            # color=model_color_map[model],
-        )
+        # ax.hlines(
+        #     counts.index,
+        #     xmin=0,
+        #     xmax=counts,
+        #     linewidth=5,
+        #     alpha=0.5,
+        #     # color=model_color_map[model],
+        # )
 
         ax.plot(
             counts,

### 2024-03-19 src/tex/appendix.tex
modernize latex

@@
 \subsection{Related work}
-Existing benchmarks such as those from \citet{guo2023large}, \cite{sun2023scieval}, \citet{Schulze_Balhorn_2024}, \citet{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}).
+Existing benchmarks such as those from \textcite{guo2023large}, \cite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}).
 While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
@@
     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of multiple choice questions and general questions per topic.}
+    \label{fig:question_count_barplot_mcq_vs_general}
     \script{plot_statistics.py}

### 2024-03-19 src/tex/ms.tex
modernize latex

@@
 It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
-For instance,~\citet{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
+For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
@@
 
+\printbibliography
+
+\clearpage
 \appendix
@@
 
-\bibliography{references}
-
+\printbibliography
 \end{document}

### 2024-03-19 src/tex/ms.tex
refsection for appendix

@@
 \clearpage
-
+\begin{refsection}
 \section{Introduction}
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of the models has led to a rapid increase in their capabilities,\cite{brown2020language} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\cite{openai2024gpt4}
-While some see in them \enquote{sparks of \gls{agi}},\cite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\cite{bender2021dangers}
-In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability solve a wide variety of tasks they have not been explicitly trained on.\cite{bommasani2021opportunities, anderljung2023frontier}
-This led to tremendous economic interest and investment in those generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\cite{bloomberg}
+Aggressive scaling of the models has led to a rapid increase in their capabilities,\autocite{brown2020language} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+While some see in them \enquote{sparks of \gls{agi}},\autocite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
+In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+This led to tremendous economic interest and investment in those generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
 
-Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
+This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober} or generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
@@
 This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
-However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
+Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
 To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
@@
 It is important to realize that the user base of such models is not only limited to experts in chemistry and materials science that reflect every output such models produce. 
-For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\cite{Intelligent.com_2023}
+For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
 This also implies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
@@
 Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, evaluation frameworks are needed to allow us to systematically measure progress and mitigate potential harms.
-For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmarks suites such as BigBench\cite{srivastava2022beyond} or the LM Eval Harness.\cite{eval-harness}
+For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmarks suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
 The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
-Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\cite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\cite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\cite{wu2018moleculenet} Therapeutic Data Commons\cite{huang2021therapeutics} or MatBench.\cite{dunn2020benchmarking}
+Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
 In these cases, the models are  evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, or reactivity) or on predicting the outcome of specific chemical reactions.
@@
 
-While some benchmark based on university entrance exams\cite{Zaki_2024, arora2023llms} or automatic text mining\cite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
+While some benchmark based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
 This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
-%As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}
+%As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\autocite{kincaid1975derivation}
 
 \paragraph{\enquote{Tiny} subset}
-It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\cite{polo2024tinybenchmarks}
+It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
@@
 Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
-For instance, Galactica\cite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
+For instance, Galactica\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
 To support this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
@@
 Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
-This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
-In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\cite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
+This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
+In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
 The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).
 
-To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\cite{Huggingface} on the chembench corpus, including systems augmented with external tools.
+To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the chembench corpus, including systems augmented with external tools.
 An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
@@
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\cite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
+With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
 This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
@@
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
-To investigate this, we prompted\cite{xiong2023llms} some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
+To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
 
@@
 In \Cref{fig:estimated_difficulty} we show that there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\cite{Li_2023}
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of XX (on a scale of 1--5) for the XX (out of XX) questions it answered in correctly (XX for the XX questions it answered correctly).
@@
 To ensure that the questions do not enter a training dataset we use the canary string of the Big Bench project.
-This requires that \Gls{llm} developers filter their training dataset for this canary string.\cite{openai2024gpt4, srivastava2022beyond}
+This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
 
@@
 \subparagraph{Chemical safety} 
-This category comprises a total of \variable{output/question_count_per_dir/json_file_counts_safety.txt} questions, including semi-programmatically created questions covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \cite{pubchem} (see below for details).
+This category comprises a total of \variable{output/question_count_per_dir/json_file_counts_safety.txt} questions, including semi-programmatically created questions covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \autocite{pubchem} (see below for details).
 Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals.
@@
 \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} questions pertaining to the total electron count of molecules are based on data from the website \url{https://www.cheminfo.org/}.
-These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \cite{rdkit}.
+These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \autocite{rdkit}.
 
 \subparagraph{Number of isomers}
-We used MAYGEN\cite{Yirik_2021} to compute the number of isomers for a set of randomly sampled \gls{smiles} from the ZINC dataset.\cite{Irwin_2012}
+We used MAYGEN\autocite{Yirik_2021} to compute the number of isomers for a set of randomly sampled \gls{smiles} from the ZINC dataset.\autocite{Irwin_2012}
 To obtain the answers for false options we sampled from the set of number of isomers. In total, we generated \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} questions.
@@
 Our ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption}) is utilized for data generation.
-Given a molecule, it uses spglib\cite{spglib} to identify the point group. 
+Given a molecule, it uses spglib\autocite{spglib} to identify the point group. 
 Assigned point groups were manually verified to only include well-defined cases.
@@
 \subparagraph{SMILES-IUPAC}
-We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \cite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature.
+We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \autocite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature.
 \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt} questions involve the reverse task, i.e., converting \gls{iupac} nomenclature to \gls{smiles} notation.
@@
 \subparagraph{Number of NMR peaks} 
-To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\cite{Irwin_2012} and then used OpenChemLib\cite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
+To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
 We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
@@
 \subparagraph{GHS, hazard statements and DAI}
-The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.\cite{pubchem}
+The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.\autocite{pubchem}
 The former two have been mined via the PubChem \gls{api}, while the latter has been manually compiled. 
@@
 Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
-For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\cite{taylor2022galactica}.
+For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
 Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
@@
 \subparagraph{Completion models}
-We used \texttt{Galactica-120b}\cite{taylor2022galactica} with the default settings.
+We used \texttt{Galactica-120b}\autocite{taylor2022galactica} with the default settings.
 
-\subparagraph{Instruction-tuned models} In addition, we used \texttt{Claude 2}, \texttt{Claude3 (Opus)},\cite{anthropicClaudeModelFamily2024} \texttt{GPT-4},\cite{openai2024gpt4} \texttt{GPT-3.5-turbo},\cite{brown2020language}
-\texttt{Gemini Pro},\cite{gemini}
-\texttt{Mixtral-8x7b}\cite{jiang2024mixtral}
-\texttt{Llama2-70b},\cite{touvron2023llama}
+\subparagraph{Instruction-tuned models} In addition, we used \texttt{Claude 2}, \texttt{Claude3 (Opus)},\autocite{anthropicClaudeModelFamily2024} \texttt{GPT-4},\autocite{openai2024gpt4} \texttt{GPT-3.5-turbo},\autocite{brown2020language}
+\texttt{Gemini Pro},\autocite{gemini}
+\texttt{Mixtral-8x7b}\autocite{jiang2024mixtral}
+\texttt{Llama2-70b},\autocite{touvron2023llama}
 \texttt{pplx-7b-chat}
@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this we, on the one hand, investigated the \texttt{online} model of Perplexity.AI and, on the other hand, \texttt{gpt-3.5-turbo} as well as \texttt{claude-2}, with ReAct-style tool augmentation.\cite{yao2023react}
+For this we, on the one hand, investigated the \texttt{online} model of Perplexity.AI and, on the other hand, \texttt{gpt-3.5-turbo} as well as \texttt{claude-2}, with ReAct-style tool augmentation.\autocite{yao2023react}
 The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
@@
 
-\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\cite{nextjs} app router framework.
+\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
 This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a MySQL\cite{mysql} database and Prisma \gls{orm}\cite{prisma} for efficient database management.
-The web application is styled with Tailwind CSS\cite{tailwindcss} using the shadcn/ui component library and uses NextAuth\cite{nextauth} for easy and secure user authentication and postMark for sending Emails.
+We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
 The application is hosted on the Vercel web hosting platform.
@@
 For questions which can easily be assigned to a topic based on the source (e.g., number of NMR peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
-For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\cite{zeroshotsequence} using the BART model\cite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
+For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
 
@@
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
-To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\\showyourwork} framework.\cite{Luger2021}
+To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\\showyourwork} framework.\autocite{Luger2021}
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
@@
 \normalsize
-
 \printbibliography
+\end{refsection}
 
 \clearpage
+\begin{refsection}
 \appendix
@@
 
-\printbibliography
+\printbibliography[heading=subbibintoc]
+\end{refsection}
 \end{document}

### 2024-03-19 src/tex/appendix.tex
updates

@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
     \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
-    \caption{Caption}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none of the questions were answered completely correctly.
+    }
     \label{fig:all_questions_models_completely_correct_radar_overall}

### 2024-03-19 src/tex/ms.tex
updates

@@
 
-\begin{figure}
+\begin{figure}[htb]
     \includegraphics[width = \textwidth]{figures/chem-bench.pdf}
-    \caption{\textbf{Data curation workflow}.}
+    \caption{\textbf{Overview of the workflow for the assembly of the \chembench corpus}.
+To assemble the \chembench corpus, we first collected questions from various sources.
+Some tasks were manually curated, others semi-programmatically.
+For all questions we added semantic annotations to make them compatible with systems that use special processing for modalities that are not conventional natural text.
+Before adding the questions to the corpus, we reviewed with manual as well as automatic methods.}
+    \label{fig:curation_workflow}
 \end{figure}

### 2024-03-19 src/scripts/analyze_human_data.py

@@ -93,6 +93,9 @@ def make_human_time_score_plot(long_df):
         .mean()
     )
     grouped_by_user.dropna(inplace=True)
+    grouped_by_user = grouped_by_user[
+        grouped_by_user["all_correct"] != 1
+    ]  # exclude cheater
     fig, ax = plt.subplots(
         1, 1, figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
     )

### 2024-03-19 src/tex/ms.tex
try to fix the requires-calculation subset

@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of the models has led to a rapid increase in their capabilities,\autocite{brown2020language} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+Aggressive scaling of the models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
 While some see in them \enquote{sparks of \gls{agi}},\autocite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}

### 2024-03-19 src/tex/appendix.tex
one more author

@@
 \subsection{Related work}
-Existing benchmarks such as those from \textcite{guo2023large}, \cite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}).
+Existing benchmarks such as those from \textcite{guo2023large}, \cite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
 While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
@@
 
-\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the \chembench corpus.
+\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
 
@@
     \includegraphics{figures/performance_per_topic.pdf}
-    \caption{Caption}
+    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by the models for each data source.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none of the questions were answered completely correctly.}
     \label{fig:performance_per_topic}
@@
     \includegraphics{figures/confidence_score_distributions.pdf}
-    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.}} 
+    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
+The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
+The barplots show how many questions were answered with each confidence score.}
     \label{fig:confidence_score_distributions}

### 2024-03-19 src/tex/ms.tex
one more author

@@
 
-\begin{figure}
-    \centering
-    %\includegraphics{figures/chem-bench-completley-correct-calculation-vs-no-calculation.pdf}
-    \caption{Caption}
-    \label{fig:calculation_performance}
-\end{figure}
-
 % Estimate of difficulty
@@
 
-In \Cref{fig:estimated_difficulty} we show that there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
+In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}

### 2024-03-19 src/scripts/analyze_model_reports.py

@@ -110,14 +110,14 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                         claude3_mean,
                         gemini_pro_mean,
                         gpt35turbo_mean,
-                        gpt35turbo_react_mean,
+                        # gpt35turbo_react_mean,
                         # llama70b_mean,
                         # galactica_120b_mean,
                         # mixtral_mean,
                         pplx7b_chat_mean,
                         pplx7b_online_mean,
                         random_baseline_mean,
-                        human_dicts,
+                        # human_dicts,
                     ],
                     [
                         "GPT-4",
@@ -133,7 +133,7 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                         "PPLX-7B-Chat",
                         "PPLX-7B-Online",
                         "Random Baseline",
-                        "Average Human",
+                        #  "Average Human",
                     ],
                     [
                         model_color_map["gpt4"],
@@ -158,7 +158,6 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
         )
     theta = radar_factory(len(claude2_mean), frame="polygon")
 
-    print(sorted_data)
     # Adjust the layout to leave space for labels
     fig, ax = plt.subplots(1, 1, figsize=(6, 6), subplot_kw=dict(projection="radar"))
     for data_, label, color in sorted_data:
@@ -490,9 +489,9 @@ if __name__ == "__main__":
     make_overall_performance_radar_plot(df_dicts["overall"], "overall")
     radarplot_requires_calculation(df_dicts["overall"], "overall")
 
-    # make_overall_performance_radar_plot(
-    #     df_dicts["human_aligned"], "human", human_dicts["topic_mean"]
-    # )
+    make_overall_performance_radar_plot(
+        df_dicts["human_aligned"], "human", human_dicts["topic_mean"]
+    )
     # radarplot_requires_calculation(
     #     df_dicts["human_aligned"], "human", human_dicts["topic_mean"]
     # )

### 2024-03-19 src/tex/ms.tex
update analysis

@@
 For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
-For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of XX (on a scale of 1--5) for the XX (out of XX) questions it answered in correctly (XX for the XX questions it answered correctly).
+For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/gpt4_is_pictograms_num_correct_overall.txt} questions it answered in correctly and \variable{output/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
-We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Cref{sec:parsing}.
+We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
 Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 

### 2024-03-19 src/tex/ms.tex
write reading ease to file

--- 2024-03-19 src/tex/ms.tex
@@
 For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
-For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of XX (on a scale of 1--5) for the XX (out of XX) questions it answered in correctly (XX for the XX questions it answered correctly).
+For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered in correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
-As initial tests indicated that models sometimes return integers in the form of words, e.g. \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion.
+As initial tests indicated that models sometimes return integers in the form of words, e.g.
+\enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
 In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.

### 2024-03-19 src/scripts/wordcloud.py

@@ -77,6 +77,12 @@ def plot_reading_ease(df):
             + "\endinput"
         )
 
+    reading_ease_frame = pd.DataFrame(
+        {"flesch_kincaid_reading_ease": flesch_kincaid_reading_ease, "name": df["name"]}
+    )
+
+    reading_ease_frame.to_pickle(output / "reading_ease.pkl")
+
 
 if __name__ == "__main__":
     import pandas as pd

### 2024-03-19 src/scripts/reading_ease_vs_model_performance.py
correlate reading ease with performance

@@ -1,0 +1,41 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from paths import scripts, output, figures, data
+import pickle
+
+from utils import ONE_COL_GOLDEN_RATIO_HEIGHT_INCH, ONE_COL_WIDTH_INCH
+import seaborn as sns
+
+from plotutils import range_frame, model_color_map
+
+plt.style.use(scripts / "lamalab.mplstyle")
+
+
+def align_model_scores_reading_ease(model_scores, reading_ease_frame):
+    out_frame = pd.merge(
+        model_scores, reading_ease_frame, left_on="name", right_on="name"
+    )
+    return out_frame
+
+
+if __name__ == "__main__":
+    with open(data / "model_score_dicts.pkl", "rb") as handle:
+        model_scores = pickle.load(handle)["overall"]
+
+    reading_ease_frame = pd.read_pickle(output / "reading_ease.pkl")
+
+    all_aligned = []
+    for model, frame in model_scores.items():
+        aligned = align_model_scores_reading_ease(frame, reading_ease_frame)
+        aligned["model"] = model
+        all_aligned.append(aligned)
+
+    all_aligned = pd.concat(all_aligned)
+
+    fig, ax = plt.subplots(1, 1, figsize=(5, 5))
+    sns.scatterplot(
+        data=all_aligned, x="reading_ease", y="all_correct", hue="model", ax=ax
+    )
+
+    fig.savefig(figures / "reading_ease_vs_model_performance.pdf", bbox_inches="tight")

### 2024-03-19 src/tex/appendix.tex
correlate reading ease with performance
update reading ease figure

@@
 
+\begin{figure}
+    \centering
+    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
+    \caption{\textbf{Model performance as a function of reading ease.} The violinplots show the distribution of reading ease scores for questions that were answered completely correctly and those that were not.
+We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
+}
+    \script{reading_ease_vs_model_performance.py}
+\end{figure}
+
 \subsection{Human baseline}

### 2024-03-19 src/tex/ms.tex
add main text radar

@@
 
+\begin{figure}
+    \centering
+    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
+    \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.}
+    \label{fig:all_questions_models_completely_correct_radar_human}
+    \script{analyze_model_reports.py}
+\end{figure}
+
 In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
@@
 % Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
+% in contrast to LLM the human performance is also less spiky
 

### 2024-03-19 src/tex/ms.tex
try heatmap

@@
     \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.}
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    This figure shows the performance on the subset of questions that was answered by humans, too. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    }
     \label{fig:all_questions_models_completely_correct_radar_human}

### 2024-03-19 src/scripts/analyze_performance_per_source.py

@@ -54,6 +54,23 @@ subset_clean_names = [
     "polymer chemistry",
 ]
 
+relevant_models = [
+    "claude2",
+    "claude2_react",
+    "claude2_zero_t" "claude3",
+    "galactica_120b"
+    "gemini_pro"
+    "gemini_pro_zero_t"
+    "gpt35turbo"
+    "gpt35turbo_react"
+    "gpt4"
+    "lama70b"
+    "mixtral"
+    "pplx7b_chat"
+    "pplx7b_online"
+    "random_baseline",
+]
+
 rename_dict = dict(zip(subsets, subset_clean_names))
 
 
@@ -122,6 +139,7 @@ if __name__ == "__main__":
         human_scores["raw_scores"], outdir_humans
     )
 
+    model_scores = model_scores[model_scores["model"].isin(relevant_models)]
     all_scores = pd.DataFrame(model_scores + human_scores)
 
     all_scores["subset"] = all_scores["subset"].map(rename_dict)

### 2024-03-20 src/tex/appendix.tex
add main text radar
we need to create the heatmap on subsets

@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
     \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by the models for each data source.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
-A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none of the questions were answered completely correctly.}
+A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
+    }
     \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic_tiny.pdf}
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic_tiny}
     \script{analyze_performance_per_source.py}
@@
     \centering
+    \hspace*{-1cm}
     \includegraphics{figures/reading_ease_vs_model_performance.pdf}
@@
     \script{reading_ease_vs_model_performance.py}
+    \label{fig:reading_ease_vs_model_performance}
 \end{figure}

### 2024-03-20 src/scripts/reading_ease_vs_model_performance.py

@@ -60,6 +60,7 @@ if __name__ == "__main__":
         hue="model",
         ax=ax,
         palette=model_color_map,
+        cut=0,
     )
 
     # change labels for the colors
@@ -81,7 +82,7 @@ if __name__ == "__main__":
 
     range_frame(
         ax,
-        np.array([-0.5, 1.5]),
+        np.array([-0.4, 1.4]),
         np.array(
             [
                 all_aligned["flesch_kincaid_reading_ease"].min(),

### 2024-03-20 src/tex/appendix.tex
use nice fonts if they are available

@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/performance_per_topic_tiny.pdf}
+%     \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
+%         We see that the performance of the models varies significantly between the different data sources.
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic_tiny}
-    \script{analyze_performance_per_source.py}
-\end{figure}
+%     }
+%     \label{fig:performance_per_topic_tiny}
+%     \script{analyze_performance_per_source.py}
+% \end{figure}
 
@@
 
-\begin{figure} 
+\begin{figure}[htb] 
     \centering

### 2024-03-20 src/tex/appendix.tex
remove tiny for compilation

--- 2024-03-20 src/tex/appendix.tex
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic_tiny}
-    \script{analyze_performance_per_source.py}
-\end{figure}
-
 \begin{figure}

### 2024-03-20 src/tex/authors.tex
update authors hopefully the last time
one more author
update manuscript

@@
 \credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
 \credit{Lea Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
@@
 \credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Nicole Roesner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
 \credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
@@
 \author[]{Benedict~Emoekabu}
-\author[qpi]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
-\author[nccrcat,liac]{Tanya~Gupta}
+\author[4]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
+\author[5,6]{Tanya~Gupta}
 \author[1]{Mara~Willhelmi}
@@
 
-\author[cambridge]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
+\author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
 \author[1]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
-\author[lsmo]{Amir~Elahi}
+\author[8]{Amir~Elahi}
 \author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
 \author[1]{Maximilian~Greiner}
+\author[1]{Tim~Hoffmann}
 \author[1]{Lea~Klepsch~\orcidlink{0000-0003-4978-4670}}
 \author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
-\author[ituc, 2]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[2, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
 \author[1]{Jakob~Meyer}
@@
 \author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
+\author[1]{Nicole~Roesner}
 
@@
 \author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
-\author[dresden]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
+\author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
 
 % Senior authors
-\author[open,stability]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
-\author[nccrcat,liac]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-\author[1,2, 3]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+\author[11,12]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[5,6]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
+\author[1,2, 3, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
 
@@
 \affil[2]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[3]{Helmholtz Institute for Polymers in Energy Applications (HIPOLE), Philosophenweg 7a, 07743 Jena, Germany}
+\affil[3]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Philosophenweg 7a, 07743 Jena, Germany}
 
-\affil[liac]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland.}
-\affil[nccrcat]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland.}
+\affil[5]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[6]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
 
-\affil[lsmo]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Valais, Switzerland.}
-\affil[dresden]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
-\affil[open]{OpenBioML.org}
-\affil[stability]{Stability.AI}
-\affil[ituc]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[qpi]{QpiVolta Technologies Pvt Ltd}
-\affil[cambridge]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, UK}
+\affil[8]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
+\affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
+\affil[11]{OpenBioML.org}
+\affil[12]{Stability.AI}
+\affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[4]{QpiVolta Technologies Pvt Ltd}
+\affil[7]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
 
-\affil[]{\texttt{mail@kjablonka.com}}
+\affil[\Letter]{\texttt{mail@kjablonka.com}}
 \affil[$\star$]{These authors contributed equally.}

### 2024-03-20 src/tex/ms.tex
we need to create the heatmap on subsets
update manuscript

@@
 
-\title{\textsf{Are frontier models superhuman chemists?}}
+\title{\textsf{Are large language models superhuman chemists?}}
 
@@
 
+\clearpage
 \begin{abstract}
@@
     \includegraphics{figures/human_subset_performance.pdf}
-    \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly.
+    \caption{\textbf{Performance of models and humans on \chembench.} The figure shows the percentage of questions that the models answered completely correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
@@
 % when it comes to reasoning tasks (NMR, point groups) the humans tend to beat LLMs
+
+%%%%% IMPORTANT emphasize the importance of breadth as performance on tasks does not necessarily correlate with performance on other tasks
 
@@
 M.A.
-expresses gratitude to the European Research Council (ERC) for evaluating the project with the reference number 101106377 titled ‘‘CLARIFIER’’ and accepting it for funding under the HORIZON TMA MSCA Postdoctoral Fellowships - European Fellowships.
+expresses gratitude to the European Research Council (ERC) for evaluating the project with the reference number 101106377 titled \enquote{CLARIFIER} and accepting it for funding under the HORIZON TMA MSCA Postdoctoral Fellowships - European Fellowships.
 Furthermore, M.A.
@@
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI. M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.AI.
+K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.AI.
 
@@
 \input{appendix.tex}
-
+\clearpage
 \printbibliography[heading=subbibintoc]

### 2024-03-20 src/tex/parse_check_desc.tex
feat: description of parsing checks

\paragraph*{Manual validation of parsing performance}

As stated, we tried to account for the variation in outputs with custom regular expressions.
We selected a large, diverse subset of 791 questions and manually investigated where the parsed output does not match the actual answer intended by the model.
We found that for \glstext{mcq} questions, the parsing was accurate in 99.76\% of the cases, while for floating point questions, the parsing was accurate in 99.17\% of the cases.
The models generating errors are \texttt{pplx-7b-chat} and \texttt{Mixtral-8x7b}, which in the specified amount of cases fail to follow the prompt.

### 2024-03-20 src/tex/ms.tex
update text

@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of the models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
 While some see in them \enquote{sparks of \gls{agi}},\autocite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
 In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
-This led to tremendous economic interest and investment in those generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
+This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
 
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober} or generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
-But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of autonomous systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
 For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
 Most of these insights are in form of text in scientific publications. 
-Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
-This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
+Thus, operating on such text might be our best way of unlocking these insights and to learn from them.
+This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
 Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
-
-It is important to realize that the user base of such models is not only limited to experts in chemistry and materials science that reflect every output such models produce. 
+To some extent, this is not surprising as any technology that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
+Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect every output such models produce. 
 For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
-Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
+Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
 
-To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry (for a total of number of more than \variable{output/total_hours.txt}\unskip) on a subset of the benchmark corpus to be able to compare the performance of current frontier models with the one of humans.
+To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with the one of humans.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, included safety-related ones, might be very misleading.
-The results indicate the limitations that current models still need to overcome to find use in autonomous systems for chemists and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.
+We find that there are plenty limitations that current models still need to overcome to find use in autonomous systems for chemists and that carefully created broad benchmarks are an important stepping stone for progress in this field.
+%and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.
 
@@
 To compile our benchmark corpus we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator.
+For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator as well as automatic checks.
 
@@
 In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as large distinct clusters.
+It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster.
 
@@
     \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model\cite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
 We then project the embeddings into a two-dimensional space using \gls{pca}.
-We color the points based on a classification into topics.}
+We color the points based on a classification into topics.
+Safety related aspects cover a large part of the figure that is not covered by questions from other topics.}
     \label{fig:question_diversity}
@@
 It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
-For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the \gls{helm} benchmark on a model. 
+For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the app we used for the human baseline study.
+We also used this subset to seed the web application we used for the human baseline study.
 
@@
 Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
-For instance, Galactica\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
-To support this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
+For instance, Galactica\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
+Current benchmarking suites, however, do not account for such special treatment of scientific information.
+To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
 For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.
@@
 Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
-This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
+This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
+Such systems can augment the capabilities of \glspl{llm} through the use external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).
+The text completions, however, are the final outputs of the system that would also be used in a real-world application. Hence, we use them for our evaluations.
 
-To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the chembench corpus, including systems augmented with external tools.
+To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
 An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
@@
 We use horizontal bars to indicate the performance of various models and highlight statistics (e.g., the average) of the human performance with vertical lines.
-Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly exceeds the average performance of our human scores.
+Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
 Many other models also outperform the average human performance.
@@
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web).
-This observation highlights the importance of transparency not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
+This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
 
@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of the \chembench corpus.
+    \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly).
+A larger colored area indicates a better performance.
     This figure shows the performance on the subset of questions that was answered by humans, too. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
@@
 One can observe that this performance varies widely across models and topics. 
-While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4)
-While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound.
+While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs.
 \variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
-These findings also shine an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
-On those questions the models tend to perform better than on some our semi-automatically constructed tasks.
+These findings also shine an interesting light on the value of textbook-inspired questions. 
+A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+On those questions the models tend to perform better than on some our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
+We suppose that this is due to the fact that some of the textbook questions require less reasoning and more memorization and combination of facts.
 
@@
 We then plot the correctness of the answers [which can be either completely correct (1) or completely incorrect (0)] against the confidence score.
-        The stripplots show the individual data points, the lines show the mean and the errorbars show the standard error of the mean.
+        The stripplots show the individual data points, a strong density indicating a higher number of points.
+The lines show the mean and the errorbars show the standard error of the mean.
 The figure shows that the reliability of the confidence estimates varies widely across models.
@@
 For our dataset we curated questions from existing exams, but also programmatically created new questions.
-Questions were added via Pull Requests on our GitHub repository and only merged into the main collection after passing manual review.
+Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review.
 
@@
 
-\subparagraph{Chemical reactivity}
-\variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
-
 \subparagraph{Textbook questions}
@@
 Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals.
-Furthermore, it includes inquiries about the Federal/State Working Group on Chemical Safety (BLAC), consisting of \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions. 
-These questions are based on the question bank used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
+Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
 Moreover, this category addresses questions lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
@@
 \subparagraph{NMR spectroscopy}
-Questions are based on images shared via the Twitter account \url{https://twitter.com/NMRspectroscopy}, focusing choosing appropriate \gls{nmr} experiments.
+Questions are based on images shared via the Twitter account \url{https://twitter.com/NMRspectroscopy}, focusing on the choice and design of  appropriate \gls{nmr} experiments.
 
 \paragraph{Semi-programatically generated questions}
+
+\subparagraph{Chemical reactivity}
+\variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
 
@@
 \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} questions pertaining to the total electron count of molecules are based on data from the website \url{https://www.cheminfo.org/}.
-These inquiries involve determining the electron count of atoms in various chemical contexts. The calculations for these tasks are facilitated using the RDKit package \autocite{rdkit}.
 
 \subparagraph{Number of isomers}
-We used MAYGEN\autocite{Yirik_2021} to compute the number of isomers for a set of randomly sampled \gls{smiles} from the ZINC dataset.\autocite{Irwin_2012}
-To obtain the answers for false options we sampled from the set of number of isomers. In total, we generated \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} questions.
+We used MAYGEN\autocite{Yirik_2021} to compute the number of isomers for a set of  \gls{smiles} randomly sampled from the ZINC dataset.\autocite{Irwin_2012}
+To obtain the answers for false options we sampled from the set of all observed isomer counts.
+In total, we generated \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} questions.
 
@@
 \subparagraph{SMILES-IUPAC}
-We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \autocite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions involve converting \gls{smiles} notation to \gls{iupac} nomenclature.
+We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \autocite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions around the task of converting \gls{smiles} notation to \gls{iupac} nomenclature.
 \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt} questions involve the reverse task, i.e., converting \gls{iupac} nomenclature to \gls{smiles} notation.
@@
 To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
-We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
-Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
-For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
-Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
+We impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
+Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
+For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
+Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
+This step can be easily customized in our codebase.
 
@@
 In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
-Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
+%Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
 
@@
 Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree and for \variable{output/num_human_highschool.txt} the highest awarded degree is from high school.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 
@@
 To allow for analysis of the model performance as a function of the topic, we leverage this information together with sequence classification models.
-For questions which can easily be assigned to a topic based on the source (e.g., number of NMR peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
+For questions which can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
 For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.

### 2024-03-20 src/tex/parse_check_desc.tex
Update parse_check_desc.tex

@@
 As stated, we tried to account for the variation in outputs with custom regular expressions.
-We selected a large, diverse subset of 791 questions and manually investigated where the parsed output does not match the actual answer intended by the model.
+We selected a large, diverse subset of 791 questions (10 questions per topic for all model reports) and manually investigated where the parsed output does not match the actual answer intended by the model.
 We found that for \glstext{mcq} questions, the parsing was accurate in 99.76\% of the cases, while for floating point questions, the parsing was accurate in 99.17\% of the cases.

### 2024-03-20 src/tex/appendix.tex
update manuscript
update text
move details to appendix

--- 2024-03-20 src/tex/appendix.tex
@@
 \subsection{Human baseline}
+\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
+This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
+We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
+The application is hosted on the Vercel web hosting platform.
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
@@
 
-\printnoidxglossary[type=\acronymtype]  % https://github.com/tectonic-typesetting/tectonic/issues/704
+\clearpage
+
+\printnoidxglossary[type=\acronymtype, nonumberlist]  % https://github.com/tectonic-typesetting/tectonic/issues/704

### 2024-03-20 src/tex/ms.tex
move details to appendix

@@
 \subparagraph{Completion models}
-We used \texttt{Galactica-120b}\autocite{taylor2022galactica} with the default settings.
+We used Galactica (120b)\autocite{taylor2022galactica} with the default settings.
 
-\subparagraph{Instruction-tuned models} In addition, we used \texttt{Claude 2}, \texttt{Claude3 (Opus)},\autocite{anthropicClaudeModelFamily2024} \texttt{GPT-4},\autocite{openai2024gpt4} \texttt{GPT-3.5-turbo},\autocite{brown2020language}
-\texttt{Gemini Pro},\autocite{gemini}
-\texttt{Mixtral-8x7b}\autocite{jiang2024mixtral}
-\texttt{Llama2-70b},\autocite{touvron2023llama}
-\texttt{pplx-7b-chat}
+\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} Llama2 (70b),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
 
@@
 To estimate the models' confidence we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the right answer on a scale from 1 to 5. 
+We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases compared to other prompting strategies, which might be more suitable when implemented in systems.
 
 \subsection{Human baseline}
-
-\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
-This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
-The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
-The application is hosted on the Vercel web hosting platform.
 

### 2024-03-20 src/tex/ms.tex
clean references

@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this we, on the one hand, investigated the \texttt{online} model of Perplexity.AI and, on the other hand, \texttt{gpt-3.5-turbo} as well as \texttt{claude-2}, with ReAct-style tool augmentation.\autocite{yao2023react}
+For this we, on the one hand, investigated the 7B parameter \enquote{online} model of Perplexity.AI and, on the other hand, gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
 The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
@@
 This work was supported by the Carl Zeiss Foundation and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We also want to acknowledge access to the HPC cluster of Stability.AI.
+We also want to acknowledge access to the HPC cluster of Stability AI.
 M.A.
@@
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability.ai and A.M.\ and N.A.\ are paid contractors of Stability.AI.
+K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability AI and A.M.\ and N.A.\ are paid contractors of Stability AI.
 

### 2024-03-20 src/tex/ms.tex
start working on discussion

@@
 \section{Discussion and Conclusions}
-Our findings also lay the groundwork for discussions for the chemical education in the future. 
-Given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
+One the one hand, our findings show that it is impressive how far we got. Leading models outperform domain experts in chemistry questions on many topics. 
+On the other hand, there are still striking limitations. On very relevant topics the answers models provide are wrong. And models are not able to estimate their own limitations.
+However, the success in our benchmarks perhaps also reveals more about the limitations about the exams we use to evaluate models---and chemists---than the models themselves.
+For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning.
+
+Yet, given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
 Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
+
+Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks.
 

### 2024-03-20 src/tex/parse_check_desc.tex
Remove #

--- 2024-03-20 src/tex/parse_check_desc.tex
@@
 As stated, we tried to account for the variation in outputs with custom regular expressions.
-We selected a large, diverse subset of 791 questions and manually investigated where the parsed output does not match the actual answer intended by the model.
+We selected a large, diverse subset of questions (10 questions per topic for all model reports) and manually investigated where the parsed output does not match the actual answer intended by the model.
 We found that for \glstext{mcq} questions, the parsing was accurate in 99.76\% of the cases, while for floating point questions, the parsing was accurate in 99.17\% of the cases.

### 2024-03-20 src/tex/ms.tex
need to think a bit about how to phrase conclusions

@@
 
+The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific tasks. 
+They cannot easily be used to evaluate systems, or models built for scientific applications. Thus, we had on little understanding of the capabilities of \glspl{llm} in the chemical sciences.
+Our work,  shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
+It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
+
 % given that LLMs have parsed almost the entire internet, it is perhaps not surprising to seem them perform well on many of the task which often require 

### 2024-03-20 src/tex/appendix.tex
cite -> autocite

@@
     \item \emph{End-to-end automation}. For model development, the evaluations will need to be run many times (e.g., on regular intervals of a training run).
-    Approaches that rely on humans scoring the answers of a system\cite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
-    \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\cite{northcutt2021pervasive}
-    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\cite{Frye_2023, Awg}
+    Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
+    \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
+    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
     \item \emph{Usable with models that support special treatment of molecules}.
-Some models such as Galactica\cite{taylor2022galactica} use special tokenization or encoding procedures for molecules or equations.
+Some models such as Galactica\autocite{taylor2022galactica} use special tokenization or encoding procedures for molecules or equations.
     To support this, the benchmark system must encode the semantic meaning of various parts of the question or answer.
     \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or even just the raw logits.
-    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
+    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
     Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
@@
     \item \emph{Cover a diverse set of topics}.
-Chemistry, as the \enquote{central science}, bridges multiple disciplines.\cite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
+Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
 \end{itemize}
@@
 \subsection{Related work}
-Existing benchmarks such as those from \textcite{guo2023large}, \cite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
+Existing benchmarks such as those from \textcite{guo2023large}, \autocite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
 While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
@@
 \subsection{Benchmark corpus}
-To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\cite{srivastava2022beyond}
+To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\autocite{srivastava2022beyond}
 This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
@@
     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\cite{flesch1948new} is a measure of how easy a text is to read.
+    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} is a measure of how easy a text is to read.
 It is calculated based on the average number of syllables per word and the average number of words per sentence.

### 2024-03-20 src/tex/ms.tex
some nuggets for conclusion, need now to reshape
cite -> autocite

@@
     \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model\cite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
 We then project the embeddings into a two-dimensional space using \gls{pca}.
@@
 
-Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks.
+Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks. % link to the missing correlation
 
@@
 It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
-
-% given that LLMs have parsed almost the entire internet, it is perhaps not surprising to seem them perform well on many of the task which often require 
-% specific knowledge humans might not have at hand without consulting an additional knowledge bank 
-% when it comes to reasoning tasks (NMR, point groups) the humans tend to beat LLMs
-
-%%%%% IMPORTANT emphasize the importance of breadth as performance on tasks does not necessarily correlate with performance on other tasks
-
-% \subsection{Future work}
-% - Advanced prompting techniques such as CoT and test time strategies
-
-% - More advanced systems with chemnistry specific tools
-
-% - More systematic human baseline: Ask people to come in, do not allow to skip questions
-
-% clearly defined metrics important for deep learning https://towardsdatascience.com/the-road-to-biology-2-0-will-pass-through-black-box-data-bbd00fabf959
+Clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision. 
+We hope that our open source \chembench will be a stepping stone for the development of such breakthrough for the chemical sciences.
 

### 2024-03-20 src/tex/broader_implications.tex
feat: initial thoughts

\paragraph*{Broader implications}

Benchmarking a \glspl{llm}s is not trivial due to many factors.
What does one really investigate in these benchmarks? And what would the ultimate goal be?
If we focus on automated machine-machine interaction where there is a closed-feedback loop,
then focusing on the appropriate prompts and answer quality is a good start. However, if we focus on
the human-machine interaction, we should investigate the simplest prompts, mimicking a real-life use case
for LLMs (e.g. safety profile of a specific molecule in the lab).

Typically, chemists specialize in 1-3 fields, and any human lecturer would be preferred to a \glspl{llm}
at the current stage. However, \glspl{llm} can be used to assist in the teaching process, since its general 
knowledge is vast and sufficient for trivial questions.

### 2024-03-21 src/tex/parse_check_desc.tex

--- 2024-03-20 src/tex/parse_check_desc.tex
@@
 As stated, we tried to account for the variation in outputs with custom regular expressions.
-We selected a large, diverse subset of 791 questions and manually investigated where the parsed output does not match the actual answer intended by the model.
+We selected a large, diverse subset of questions (10 per topic for all model reports) and manually investigated where the parsed output does not match the actual answer intended by the model.
 We found that for \glstext{mcq} questions, the parsing was accurate in 99.76\% of the cases, while for floating point questions, the parsing was accurate in 99.17\% of the cases.

### 2024-03-21 src/tex/ms.tex
parallel coordinates and updated plot

@@
 % Radar chart
-To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
+\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
 For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics based and hand-crafted rules as well as classifier models.
@@
 Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
-For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.
+For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
 
 % Estimate of difficulty
-One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
+\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
@@
 Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks. % link to the missing correlation 
+% also need for better ways of how humans actually use LLMs. How do they get the confidence?  
+% an LLM is not a chemist and at least the current systems seem to be far from reasoning like a chemist 
+% for fact like questions, that requires knowledge they might have seen on the internet, they work well, but for reasoning they are not good yet.
 

### 2024-03-21 src/tex/ms.tex
make folder for sections
update text for question selection

@@
 \maketitle
+
+% make the app somewhat more prominent
 
@@
 
+% emphasize that the topics are more balanced in this subset
+
 \subsection{Model evaluation}
@@
     \includegraphics{figures/human_subset_performance.pdf}
-    \caption{\textbf{Performance of models and humans on \chembench.} The figure shows the percentage of questions that the models answered completely correctly.
+    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
@@
     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
@@
 % in contrast to LLM the human performance is also less spiky
+% emphasise that the exact ranking depends on the subset of questions we look at
 
@@
 
+\input{sections/parse_check_desc.tex}
+
 \paragraph{Models}
@@
 \end{itemize}
-To answer the first question we selected X questions that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-From those we picked X diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART (see below).
+To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
+In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
 

### 2024-03-21 src/scripts/plotutils.py

@@ -302,15 +302,13 @@ def parallel_coordinates_plot(
     legend_handles = []
     for j in range(N):
         # to just draw straight lines between the axes:
-        lines = host.plot(
-            range(ys.shape[1]), zs[j, :], c=colors[(category[j] - 1) % len(colors)]
-        )
+        lines = host.plot(range(ys.shape[1]), zs[j, :], c=colors[j])
         legend_handles.append(lines[0])
 
     host.legend(
         legend_handles,
         category_names,
-        ncol=5,
-        bbox_to_anchor=(1, 2.2),
+        ncol=4,
+        bbox_to_anchor=(1.1, -0.1),
         # loc="upper right",
     )

### 2024-03-21 src/scripts/collect_model_scores_human_subset.py
human score distribution rule

@@ -5,6 +5,8 @@ import subprocess
 from utils import obtain_chembench_repo
 from paths import output
 
+
+
 if __name__ == "__main__":
     chembench_repo = obtain_chembench_repo()
     reports = glob(

### 2024-03-21 src/tex/appendix.tex
parallel coordinates and updated plot
discuss the appendix figures

@@
 
+\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.  
+We see that the questions are generally complex to read.
+
 \begin{figure}[htb]
@@
 
+\Cref{fig:question_count_barplot_mcq_vs_general} shows that most questions in our corpus are \gls{mcq}.
+A substantial fraction, in contrast to other benchmarks, is open-ended. 
 \begin{figure}[htb]
@@
 \Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
+Note that this ranking is different from the one on the \enquote{tiny} subset.
 
@@
 \Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
+The general pattern that the performance varies significantly between the different topics is also observed when the models are evaluated on the entire corpus. 
+However, since some subjects are composed of questions from different sources the ranking of the models is in some instances different from the one on the \enquote{tiny} subset.
 
@@
 \end{figure}
+
+\Cref{fig:parallel_coordinates_overall} shows this data in the form of a parallel coordinates plot. 
+This visualization highlights the important observation that the ranking of the models highly depends on the questions they are evaluated on.
+Only very broad benchmark have hence the potential to provide a comprehensive view of the capabilities of a model. However, even in those cases, the weighting of the different topics is crucial.
+Hence, we believe that fine-grained analysis of model performance is crucial for the development of future benchmarks.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_overall}
+    \script{make_parallel_coordinates_plot.py}
+\end{figure}
+
+To further investigate the performance of the models, we also compared the performance on different data sources.
+Compared to topics, this is a more fine-grained analysis as topics can be composed of questions from different sources.
+In \cref{fig:performance_per_topic} we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to better for our models than some of the semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
 
@@
     \centering
+    \includegraphics{figures/human_score_distribution.pdf}
+    \script{plot_human_score_distribution.py}
+    \label{fig:human_score_distribution}
+    \caption{\textbf{Distribution of human scores.} The histogram shows the fraction of questions that have been answered completely correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
     \includegraphics{figures/human_timing.pdf}

### 2024-03-21 src/tex/appendix.tex
update manuscript

@@
 
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_human_aligned.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_human_aligned}
+    \script{make_parallel_coordinates_plot.py}
+\end{figure}
+
 To further investigate the performance of the models, we also compared the performance on different data sources.
@@
 
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.
+
 \begin{figure}[htb]
@@
 \end{figure}
+
+One might wonder if questions that are more difficult to parse lead to worse performance of the models.
+\Cref{fig:reading_ease_vs_model_performance} shows that there is no clear correlation between the reading ease of the questions and the performance of the models.
 
@@
 
+\paragraph{Statistics}
+\Cref{fig:human_score_distribution} shows the distribution of scores our human scorers achieved.
+
 \begin{figure}[htb]
@@
     \label{fig:human_score_distribution}
-    \caption{\textbf{Distribution of human scores.} The histogram shows the fraction of questions that have been answered completely correctly.
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions that have been answered completely correctly.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
 \end{figure}
+
+We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+Additionally, we prompted users to provide additional information about their experience in chemistry. 
+While we recorded very fine-grained information, e.g., including their specialization, we focussed our analysis on the number of years since the first university-level chemistry course.
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 

### 2024-03-21 src/tex/ms.tex
human score distribution rule
update manuscript

@@
 For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
-To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
+To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
 We also used this subset to seed the web application we used for the human baseline study.
@@
 In this figure, we show the percentage of questions that the models answered completely correctly.
-We use horizontal bars to indicate the performance of various models and highlight statistics (e.g., the average) of the human performance with vertical lines.
+Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
@@
 In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
+% https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
 

### 2024-03-21 src/tex/appendix.tex
typesetting

@@
     \centering
-    \includegraphics{figures/parallel_coordinates_human_aligned.pdf}
+    \includegraphics{figures/parallel_coordinates_tiny.pdf}
     \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
@@
 }
-    \label{fig:parallel_coordinates_human_aligned}
+    \label{fig:parallel_coordinates_tiny}
     \script{make_parallel_coordinates_plot.py}
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 While we recorded very fine-grained information, e.g., including their specialization, we focussed our analysis on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 Since it is important to understand if models can provide an indication of whether their answer might likely be incorrect, we prompted some of our top performing \glspl{llm} to return the confidence in providing a correct answer on an ordinal scale.
-
+This is similar to the verbalized confidence scores reported by \textcite{xiong2023llms}.
 \Cref{fig:confidence_score_distributions} plots the distribution of those scores.
+We find that the models show different distributions of confidence scores, which for some are skewed to the extremes.
 

### 2024-03-21 src/tex/ms.tex
typesetting

@@
 
-    We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4, and found that the best models, on average, outperform the best human chemists in our study.
+    We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4 and Claude 3, and found that the best models, on average, outperform the best human chemists in our study.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g., about the safety profile of chemicals.
@@
 The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
-Due to the lack of widely excepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
+Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
 In these cases, the models are  evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, or reactivity) or on predicting the outcome of specific chemical reactions.
@@
 
-\begin{figure}
+\begin{figure}[!htb]
     \centering
     \includegraphics{figures/question_count_barplot.pdf}
-    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question.
+    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions.
     The figure shows that not all aspects of chemistry are equally represented in our corpus.
@@
 
-\begin{figure}
+\begin{figure}[!htb]
     \centering
     \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
 We then project the embeddings into a two-dimensional space using \gls{pca}.
@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
-%As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\autocite{kincaid1975derivation}
 
@@
 
-% emphasize that the topics are more balanced in this subset
-
 \subsection{Model evaluation}
 
-Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
+\paragraph{Benchmark suite design} Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
 For instance, Galactica\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
@@
 To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
-For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.
+For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
+This allows the model to treat the \gls{smiles} string different from other text. 
+\chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.
 
@@
 
-To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
-An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
-In this figure, we show the percentage of questions that the models answered completely correctly.
-Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
-Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
-Many other models also outperform the average human performance.
-It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
-% Bar chart
-
-\begin{figure}[htb]
+\begin{figure}[!h]
     \centering
@@
     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
+    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
     \Cref{fig:barplot_all_correct_all_questions} gives an overview of the performance of various models on the entire corpus.
+    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
+    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
+    In this case, we consider the answer as incorrect. 
     }
@@
 
-Given the large interest in tool-augmented systems, the bad performance of these systems in our benchmark is striking. 
+\paragraph{System performance} 
+To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
+An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
+In this figure, we show the percentage of questions that the models answered completely correctly.
+Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
+Many other models also outperform the average human performance.
+It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
+
+Given the large interest in tool-augmented systems, the bad performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
@@
 
-%% Note also while the performance of the systems is better than what humans do, it is still far from perfect
-
 % Radar chart
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
-For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified the questions into these topics based and hand-crafted rules as well as classifier models.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
 We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
-
-\begin{figure}[htb]
+\begin{figure}[!h]
     \centering
@@
     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+The performance is measured as the fraction of questions that were answered completely correct by the models.
     The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly).
 A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that was answered by humans, too. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
     }
@@
 \end{figure}
-
 In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
@@
 Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
-This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs.
-\variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
 These findings also shine an interesting light on the value of textbook-inspired questions. 
@@
 On those questions the models tend to perform better than on some our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
-We suppose that this is due to the fact that some of the textbook questions require less reasoning and more memorization and combination of facts.
-
-% GFK performance -> would the models obtain a chemical license 
-% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
-% in contrast to LLM the human performance is also less spiky
-% emphasise that the exact ranking depends on the subset of questions we look at
+For instance, while the overall performance in the chemical safety topic is low, the models would pass the examination for a chemical license based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
 
@@
 For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
+
+\clearpage
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[!h]
     \centering
@@
 
-Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks. % link to the missing correlation 
-% also need for better ways of how humans actually use LLMs. How do they get the confidence?  
-% an LLM is not a chemist and at least the current systems seem to be far from reasoning like a chemist 
-% for fact like questions, that requires knowledge they might have seen on the internet, they work well, but for reasoning they are not good yet.
+Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks. 
+The analysis of model performance on different topics shows that the performance of models varies widely across topics. 
+But even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
 
@@
 It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
-Clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision. 
-We hope that our open source \chembench will be a stepping stone for the development of such breakthrough for the chemical sciences.
+Our findings also illustrate that more focus is needed in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
+
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision.
+We thus believe that although current systems are far from reasoning like a chemist our open source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
+
+\clearpage
 
@@
 
-To ensure that the questions do not enter a training dataset we use the canary string of the Big Bench project.
+To ensure that the questions do not enter a training dataset we use the same canary string as the Big Bench project.
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
 
-\begin{figure}[htb]
+\begin{figure}[!h]
     \includegraphics[width = \textwidth]{figures/chem-bench.pdf}
@@
 \subparagraph{Organic reactivity}
-\variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt}  questions are primarily sourced from tutorials and exam papers at for the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), exploring reactions and mechanisms in organic chemistry.
+\variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt}  questions are primarily sourced from tutorials and exam papers for the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), exploring reactions and mechanisms in organic chemistry.
 
@@
 Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
-Moreover, this category addresses questions lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
+Moreover, this category addresses questions about lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
 The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
@@
 
-To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
+To maintain consistency with the training, we employ distinct prompt templates tailored for completion and instruction-tuned models. 
 We impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
@@
 \paragraph{Parsing}
-Our parsing workflow is, by default, multistep and primarily based on regular expressions.
+Our parsing workflow is multistep and primarily based on regular expressions.
 In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this we, on the one hand, investigated the 7B parameter \enquote{online} model of Perplexity.AI and, on the other hand, gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
+For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
+In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
 The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
@@
 \variable{output/num_humans_with_204_scores.txt} humans answered more than 200 questions.
+For the analysis, we treated each human as a model. For analyses grouped by topic, we computed the topic aggregated averages per human and then averaged over all humans.
 
 \subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset we systematically recorded keywords and sources.
-To allow for analysis of the model performance as a function of the topic, we leverage this information together with sequence classification models.
+To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
 For questions which can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
@@
 This work was supported by the Carl Zeiss Foundation and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We also want to acknowledge access to the HPC cluster of Stability AI.
+We thank Stability.AI for the use of its HPC cluster and for its generous donations.
+We also thank the OpenBioML.org community and their ChemNLP project team for useful discussions.
+
 M.A.
@@
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability AI and A.M.\ and N.A.\ are paid contractors of Stability AI.
+K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability.AI and A.M.\ and N.A.\ are paid contractors of Stability.AI.
 
@@
 
-\footnotesize
+\scriptsize
 \insertcredits

### 2024-03-21 src/scripts/make_parallel_coordinates_plot.py

@@ -8,11 +8,9 @@ import pandas as pd
 
 import numpy as np
 
-from plotutils import range_frame, parallel_coordinates_plot, model_color_map
+from plotutils import parallel_coordinates_plot, model_color_map
 from utils import (
-    ONE_COL_WIDTH_INCH,
     ONE_COL_GOLDEN_RATIO_HEIGHT_INCH,
-    TWO_COL_GOLDEN_RATIO_HEIGHT_INCH,
     TWO_COL_WIDTH_INCH,
 )
 import pickle
@@ -90,7 +88,7 @@ def prepare_data_for_parallel_coordinates(model_score_dict, human_data=None):
 
 def plot_parallel_coordinates(parallel_coordinates_data, suffix=""):
     fig, ax = plt.subplots(
-        1, 1, figsize=(TWO_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
+        1, 1, figsize=(TWO_COL_WIDTH_INCH * 0.8, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
     )
 
     parallel_coordinates_data_raw = [
@@ -132,4 +130,4 @@ if __name__ == "__main__":
         model_scores["human_aligned"], human_data=human_scores
     )
 
-    plot_parallel_coordinates(parallel_coordinates_data, suffix="human_aligned")
+    plot_parallel_coordinates(parallel_coordinates_data, suffix="tiny")

### 2024-03-21 src/scripts/reading_ease_vs_model_performance.py

@@ -39,7 +39,7 @@ if __name__ == "__main__":
     all_aligned = pd.concat(all_aligned)
 
     fig, ax = plt.subplots(
-        1, 1, figsize=(TWO_COL_WIDTH_INCH, TWO_COL_GOLDEN_RATIO_HEIGHT_INCH)
+        1, 1, figsize=(TWO_COL_WIDTH_INCH * 0.8, TWO_COL_GOLDEN_RATIO_HEIGHT_INCH)
     )
 
     relevant_models = [

### 2024-03-21 src/scripts/plot_statistics.py
typographic updates

@@ -26,7 +26,7 @@ def overall_question_count_barplot(df):
     ax.set_xscale("log")
     range_frame(ax, counts, np.arange(len(topics)))
 
-    ax.set_xlabel("Number of Questions")
+    ax.set_xlabel("number of questions")
 
     # fig.tight_layout()
 
@@ -75,7 +75,7 @@ def question_count_barplot(df):
     ax[1].set_xscale("log")
     range_frame(ax[1], all_counts, np.arange(len(topics_general)))
 
-    ax[1].set_xlabel("Question Count")
+    ax[1].set_xlabel("question count")
 
     # add "MCQ" and "General" labels
     ax[0].text(0.5, 1.05, "MCQ", transform=ax[0].transAxes, ha="center")

### 2024-03-22 src/tex/authors.tex
need to think a bit about how to phrase conclusions
parallel coordinates and updated plot
typesetting
affiliations will need reordering

@@
 \credit{Mara~Wilhemli}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Macjonathan~Oreke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Macjonathan~Okereke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
 
@@
 \credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Caroline Holick}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
 \credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Lea Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
 \credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
@@
 \credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
+\credit{Leanne M. Stafast}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
 \credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
@@
 % senior authors 
-\credit{Michael~Pieler}{1,0,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Philippe~Schwaller}{1,1,0,1,1,0,0,0,0,1,0,0,0,0}
+\credit{Michael~Pieler}{1,0,0,1,0,1,0,1,0,0,0,0,0,1}
+\credit{Philippe~Schwaller}{1,1,0,1,1,1,0,0,0,1,0,0,0,0}
 \credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}
@@
 \author[1,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
-\author[1,$\star$]{Nawaf~Alampara~\orcidlink{}}
+\author[1,$\star$]{Nawaf~Alampara~\orcidlink{0009-0001-7697-7315}}
 \author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
@@
 \author[1]{Mara~Willhelmi}
-\author[1]{Macjonathan~Oreke}
+\author[1]{Macjonathan~Okereke~\orcidlink{0009-0007-1013-0502}}
 
 \author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
-\author[1]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
+\author[13]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
 \author[8]{Amir~Elahi}
@@
 \author[1]{Maximilian~Greiner}
+\author[1]{Caroline~Holick}
 \author[1]{Tim~Hoffmann}
-\author[1]{Lea~Klepsch~\orcidlink{0000-0003-4978-4670}}
+\author[1]{Lea~C.~Klepsch~\orcidlink{0009-0009-3849-1670}}
 \author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
@@
 \author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[1]{Leanne~M.~Stafast}
 \author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
@@
 \affil[2]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[3]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Philosophenweg 7a, 07743 Jena, Germany}
+\affil[3]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
 
@@
 \affil[7]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
-
+\affil[13]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
 \affil[\Letter]{\texttt{mail@kjablonka.com}}

### 2024-03-23 src/tex/ms.tex
typographic updates
details about inference temperature and calculation impact start
review until discussion

@@
 \begin{abstract}
-    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly been trained on. 
+    Large language models (LLMs) have drawn great interest as they can perform tasks they have not been explicitly been trained on. 
     This is relevant for the chemical sciences, which face the problem of learning from many small and diverse datasets that are frequently in the form of text.
@@
 
-    We curated more than 6000 question answer pairs about a wide array of subfields, evaluated leading models such as GPT-4 and Claude 3, and found that the best models, on average, outperform the best human chemists in our study.
+    We curated more than 6000 question-answer pairs about a wide array of subfields, evaluated leading models such as GPT-4 and Claude 3, and found that the best models, on average, outperform the best human chemists in our study.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g., about the safety profile of chemicals.
@@
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a great untapped potential in \glspl{llm} for chemistry and materials science.
 For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
@@
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).
+To some extent, this is not surprising as any technology that can be used to predict the toxicity of a compound for drug discovery, can be used in a contrastive way in chemical warfare.
 Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect every output such models produce. 
@@
 Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
-However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from anecdotal reports there is little evidence on how \glspl{llm} perform compared to experts.
+However, even for experts, the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
+Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to experts.
 
@@
 While some benchmark based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
-This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
+Likely, this is because they cannot be used automatically with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
 On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights in how the models compare relative to experts.
@@
 In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs: manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
 Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
@@
 For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
-This allows the model to treat the \gls{smiles} string different from other text. 
+This allows the model to treat the \gls{smiles} strings different from other text. 
 \chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.
@@
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in a real-world application. Hence, we use them for our evaluations.
+The text completions, however, are the final outputs of the system that would also be used in real-world applications.
 
@@
 In this figure, we show the percentage of questions that the models answered completely correctly.
-Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the highest, average, and lowest performances of the human experts, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
 Many other models also outperform the average human performance.
-It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
+However, it is interesting to point out that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
 
 Given the large interest in tool-augmented systems, the bad performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
-Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
+Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of 10 \gls{llm} calls.
 With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
-This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+This observation highlights the importance of communicating and measuring not only the predictive performance but also the computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
 
@@
     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correct by the models.
@@
 While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4), while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
 Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 These findings also shine an interesting light on the value of textbook-inspired questions. 
-A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+A subset of the questions in \chembench are based on textbooks targeted at undergraduate students. 
 On those questions the models tend to perform better than on some our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
 For instance, while the overall performance in the chemical safety topic is low, the models would pass the examination for a chemical license based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
-While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such a question set or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
 
@@
 % Estimate of difficulty
-\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
+\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they could be able to answer a question correctly. 
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
-To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate the on an ordinal scale their confidence in their ability to answer the question correctly.
+To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate (on an ordinal scale) their confidence to answer the question correctly.
 
@@
 In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
+For applications in which humans might rely on the models to provide answers, this is a concerning observation which highlights the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered in correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
@@
 \section{Discussion and Conclusions}
-One the one hand, our findings show that it is impressive how far we got. Leading models outperform domain experts in chemistry questions on many topics. 
+On one hand, our findings show why \glspl{llm} are impressive to the scietists, but also to the general public. Leading models outperform domain experts in chemistry questions on many topics. 
 On the other hand, there are still striking limitations. On very relevant topics the answers models provide are wrong. And models are not able to estimate their own limitations.
@@
 \paragraph{Models}
+We find that the performance of models depends on the sampling temperature. Typically we find the best temperature with greedy argmax decoding ($T=0$).
 \subparagraph{Completion models}

### 2024-03-24 src/tex/ms.tex
Update ms.tex

@@
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
+    Here, we introduce \enquote{\chembench}, an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
 
@@
 Yet, given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
-Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
+Critical reasoning is increasingly important and robotic solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
 

### 2024-03-24 src/tex/ms.tex
discussion reviewed

@@
 The analysis of model performance on different topics shows that the performance of models varies widely across topics. 
-But even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
+But even within a topic, the performance of models can vary significantly depending on the type of question and the reasoning required to answer it.
 
 The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific tasks. 
-They cannot easily be used to evaluate systems, or models built for scientific applications. Thus, we had on little understanding of the capabilities of \glspl{llm} in the chemical sciences.
-Our work,  shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
+They cannot easily be used to evaluate systems, or models built for scientific applications. As a result, we only had limited understanding of the capabilities of \glspl{llm} in the chemical sciences.
+Our work shows that carefully curated benchmarks can provide a deeper understanding of the capabilities of \glspl{llm} in the chemical sciences.
 It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
@@
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision.
-We thus believe that although current systems are far from reasoning like a chemist our open source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml}, such as computer vision.
+We believe that while current systems are far from reasoning like a chemist, our open-source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
 

### 2024-03-24 src/tex/ms.tex
review curation workflow
reviewed manually curated categories

@@
 
-To ensure that the questions do not enter a training dataset we use the same canary string as the Big Bench project.
+To ensure that the questions do not enter a training dataset we use the same canary string as the BIG-bench project.
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
@@
 \subparagraph{Combustion engineering}
-\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} questions are based on master student examination papers at the University of Magdeburg (Germany). 
+\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question are based on master student examination papers at the University of Magdeburg (Germany). 
 They explore topics related to combustion processes and engineering principles.
@@
 \subparagraph{Functional materials and nanomaterials}
-\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
+\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises from a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
 
@@
 
-\subparagraph{International Chemical Olympiad}
-\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
+\subparagraph{Chemistry Olympiads}
+\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various local and international chemistry competitions, including those held in the USA, UK, and Moldova.
 
@@
 Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
-Moreover, this category addresses questions about lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
+Moreover, this category includes questions about lab safety, pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
 The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
@@
 For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
-In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
-The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
+In addition, we used gpt-3.5-turbo and Claude 2 with ReAct-style tool augmentation.\autocite{yao2023react}
+The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (through DuckDuckGo).
 We implemented the system using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.

### 2024-03-24 src/tex/ms.tex
more review

@@
 For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+That is, while users were prompted to not use external tools other than a calculator and to not consult other humans, we do not have any way to verify that the participants complied with those rules. 
 Note that users were also allowed to skip questions.
 
-Another aspect of requiring unsupervised question answering is that in real life humans have tools and are able to use them for answering any question of interest. 
-In our current study, we prompted users to not use those tools.
+Another aspect of requiring unsupervised question answering is that in the real world humans also have tools and are able to use them for answering any question of interest.
 

### 2024-03-24 src/tex/ms.tex
more updates

@@
 For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted to not use external tools other than a calculator and to not consult other humans, we do not have any way to verify that the participants complied with those rules. 
-Note that users were also allowed to skip questions.
-
-Another aspect of requiring unsupervised question answering is that in the real world humans also have tools and are able to use them for answering any question of interest.
+That is, while users were prompted to not use external tools other than a calculator and to not consult other humans, we do not have any way to verify that the participants complied with those rules (note that users were also allowed to skip questions).
+Another aspect of requiring unsupervised question answering is the real world equivalent of our benchmark, where humans have unencumbered access to tools and are able to answer any question of interest.
 
@@
 Overall, \variable{output/num_users_with_education_info.txt} did so. 
-Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a doctoral degree.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoctorate, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 

### 2024-03-24 src/tex/ms.tex
Update ms.tex

@@
 To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
+In addition, we picked 100 diverse ones using the greedy MaxMin sampling on the embeddings of the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
 % https://github.com/lamalab-org/chem-bench-app/issues/92 has the details

### 2024-03-25 src/tex/ms.tex
rephrased messy paragraph

@@
 \section{Discussion and Conclusions}
-On one hand, our findings show why \glspl{llm} are impressive to the scietists, but also to the general public. Leading models outperform domain experts in chemistry questions on many topics. 
-On the other hand, there are still striking limitations. On very relevant topics the answers models provide are wrong. And models are not able to estimate their own limitations.
-However, the success in our benchmarks perhaps also reveals more about the limitations about the exams we use to evaluate models---and chemists---than the models themselves.
-For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning.
+Our findings show why \glspl{llm} are catching the interest of scientists across domains, but also the attention of the general public.
+As observed, leading models outperform domain experts in chemistry questions on a majority of topics.
+Unsurprinsigly, there are still striking limitations. 
+First, it is important to note that the models are not able to estimate their own limitations.
+Moreover, on very relevant topics (such as GHS classification) the answers provided by the models are mostly inaccurate.
+Looking in more detail at the relationship between the model performance and the question type/source revealed that the studied models perform well on textbook questions, while underperforming on questions that require more reasoning (for example questions that require calculations).
 

### 2024-03-25 src/tex/ms.tex
for example -> e.g.

@@
 Moreover, on very relevant topics (such as GHS classification) the answers provided by the models are mostly inaccurate.
-Looking in more detail at the relationship between the model performance and the question type/source revealed that the studied models perform well on textbook questions, while underperforming on questions that require more reasoning (for example questions that require calculations).
+Looking in more detail at the relationship between the model performance and the question type/source revealed that the studied models perform well on textbook questions, while underperforming on questions that require more reasoning (e.g.
+questions that require calculations).
 

### 2024-03-25 src/tex/ms.tex
contrastive -> divergent

@@
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology that can be used to predict the toxicity of a compound for drug discovery, can be used in a contrastive way in chemical warfare.
+To some extent, this is not surprising as any technology that can be used to predict the toxicity of a compound for drug discovery, can be used in a divergent manner in chemical warfare.
 Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect every output such models produce. 

### 2024-03-25 src/tex/ms.tex
significantly -> notably
grammar of combustion questions

@@
 The analysis of model performance on different topics shows that the performance of models varies widely across topics. 
-But even within a topic, the performance of models can vary significantly depending on the type of question and the reasoning required to answer it.
+But even within a topic, the performance of models can vary notably depending on the type of question and the reasoning required to answer it.
 
@@
 \subparagraph{Combustion engineering}
-\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question are based on master student examination papers at the University of Magdeburg (Germany). 
-They explore topics related to combustion processes and engineering principles.
+\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question is based on master student examination papers at the University of Magdeburg (Germany).
+It explores topics related to combustion processes and engineering principles.
 

### 2024-03-25 src/tex/ms.tex
Update ms.tex

@@
 \section{Discussion and Conclusions}
-Our findings show why \glspl{llm} are catching the interest of scientists across domains, but also the attention of the general public.
+Our findings show why \glspl{llm} are catching the interest of scientists across domains, but also the attention of the general public, as such models seemed inconceivable just a few years ago.
 As observed, leading models outperform domain experts in chemistry questions on a majority of topics.
@@
 questions that require calculations).
+The success of ChemBench also revealed limitations about the exams we use to test the models (but also the chemists), rather than the limitations of the models themselves.
 

### 2024-03-25 src/tex/ms.tex

@@
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench}, an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
 

### 2024-03-25 src/tex/appendix.tex
feat: first addition of chemical analysis
co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>

@@
 
+\subsubsection{Model performance as a function of molecular structure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/correlation_is_number_nmr_peaks_num_atoms.pdf}
+    \caption{\textbf{Model performance as a function of molecular structure.} The violinplots show the distribution of the number of atoms in the molecules for questions that were answered completely correctly and those that were not.
+We do not observe a clear correlation between the number of atoms in the molecules and the performance of the models.
+}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:molecular_structure_vs_model_performance}
+\end{figure}
+
 \subsection{Human baseline}

### 2024-03-26 src/scripts/utils.py
update plots

@@ -26,7 +26,7 @@ def obtain_chembench_repo():
         os.system(
             f"wget --header 'Authorization: token {GH_PAT}' https://github.com/lamalab-org/chem-bench/archive/refs/heads/main.zip"
         )
-        os.system("unzip chem-bench-main.zip")
+        os.system("unzip main.zip")
 
     else:
         os.chdir(os.path.join(data, "chem-bench-main"))

### 2024-03-26 src/scripts/analyze_model_reports.py
radarplot opacity

@@ -208,7 +208,10 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
     ax.set_yticks([])
 
     # Improve legend placement
-    ax.legend(loc=(-0.1, 1.2), ncols=3)
+    leg = ax.legend(loc=(-0.1, 1.2), ncols=3)
+
+    for lh in leg.legend_handles:
+        lh.set_alpha(1)
 
     fig.tight_layout()
 
@@ -501,7 +504,9 @@ def radarplot_requires_calculation(df_dict, human_dicts, suffix):
     ax.set_yticklabels([])
     ax.set_yticks([])
     # Improve legend placement
-    ax.legend(loc=(-0.1, 1.2), ncols=4)
+    leg = ax.legend(loc=(-0.1, 1.2), ncols=4)
+    for lh in leg.legend_handles:
+        lh.set_alpha(1)
 
     fig.tight_layout()
 

### 2024-03-26 src/tex/ms.tex
radarplot opacity
update dependencies for end-to-end-run

@@
 
-% Calculation vs. no calculation
-Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
-For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
+% % Calculation vs. no calculation
+% Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
+% For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
 
@@
 \subparagraph{Number of NMR peaks} 
-To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
+To generate the \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
 We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.

### 2024-03-26 src/scripts/count_json_files.py

@@ -2,7 +2,6 @@ import os
 from utils import obtain_chembench_repo
 from paths import output as output
 from loguru import logger
-import os
 from glob import glob
 
 output_path = output / "question_count_per_dir"

### 2024-03-26 src/tex/appendix.tex
update plots
[showyourwork]

--- 2024-03-21 src/tex/appendix.tex
@@
 This visualization highlights the important observation that the ranking of the models highly depends on the questions they are evaluated on.
-Only very broad benchmark have hence the potential to provide a comprehensive view of the capabilities of a model. However, even in those cases, the weighting of the different topics is crucial.
-Hence, we believe that fine-grained analysis of model performance is crucial for the development of future benchmarks.
+Only very broad benchmarks have hence the potential to provide a comprehensive view of the capabilities of a model. However, even in those cases, the weighting of the different topics is crucial.
+Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
 
@@
         We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.

### 2024-03-26 src/tex/ms.tex
[showyourwork]
[showyourwork] overleaf sync

--- 2024-03-25 src/tex/ms.tex
@@
 \begin{abstract}
-    Large language models (LLMs) have drawn great interest as they can perform tasks they have not been explicitly been trained on. 
-    This is relevant for the chemical sciences, which face the problem of learning from many small and diverse datasets that are frequently in the form of text.
-    LLMs are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
+    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly trained on. 
+    This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
+    LLMs have shown promise to address these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
 
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
 
-    We curated more than 6000 question-answer pairs about a wide array of subfields, evaluated leading models such as GPT-4 and Claude 3, and found that the best models, on average, outperform the best human chemists in our study.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and give overconfident misleading predictions, e.g., about the safety profile of chemicals.
+    We curated more than 7000 question answer pairs about a wide array of subfields, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.
 
@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass the BAR exam or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
-While some see in them \enquote{sparks of \gls{agi}},\autocite{bubeck2023sparks} others consider them as \enquote{stochasitc parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination\cite{kung2023performance} or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
+In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
 This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
@@
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of autonomous systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of autonomous systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a great untapped potential in \glspl{llm} for chemistry and materials science.
-For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
-Most of these insights are in form of text in scientific publications. 
-Thus, operating on such text might be our best way of unlocking these insights and to learn from them.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
+Many of these insights are in form of text in scientific publications. 
+Thus, operating on such texts might be our best way of unlocking these insights and to learn from them.
 This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
 
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology that can be used to predict the toxicity of a compound for drug discovery, can be used in a divergent manner in chemical warfare.
-Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect every output such models produce. 
-For example, many students frequently consolidate those tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
-This also implies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
+To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used in the inverse way to predict toxic ones.
+Yet, it is important to realize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
+However, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
+For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
+This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
 Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
-However, even for experts, the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
+However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
 Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to experts.
@@
 For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmarks suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
-The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related} whereas the latter contains no specific chemistry tasks.
+The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
 Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
-In these cases, the models are evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, or reactivity) or on predicting the outcome of specific chemical reactions.
+In these cases, the models are  evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, melting temperature or reactivity) or on predicting the outcome of specific chemical reactions.
 This, however, only gives a very limited view of the general chemical capabilities of the models.
 
-While some benchmark based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
-Likely, this is because they cannot be used automatically with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
-On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights in how the models compare relative to experts.
+While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
+This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
+On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.
 
-In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs: manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
 Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
 
-To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with the one of humans.
-Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, included safety-related ones, might be very misleading.
-We find that there are plenty limitations that current models still need to overcome to find use in autonomous systems for chemists and that carefully created broad benchmarks are an important stepping stone for progress in this field.
+To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specialization.
+Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, including safety-related ones, might be very misleading.
+We find that there are still plenty of limitations for current models to overcome, to find use in autonomous systems for chemists.
+Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.
 %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.
@@
 While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
+For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
 
@@
 For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
-To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the web application we used for the human baseline study.
+To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
+We also used this subset to seed the web application for the human baseline study.
 
@@
 To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
-For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
-This allows the model to treat the \gls{smiles} strings different from other text. 
+For instance, molecules represented in \gls{smiles} are enclosed in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
+This allows the model to treat the \gls{smiles} string different from other text. 
 \chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.
@@
 This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
-Such systems can augment the capabilities of \glspl{llm} through the use external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
+Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in real-world applications.
+The text completions, however, are the final outputs of the system that would also be used in a real-world application. Hence, we use them for our evaluations.
 
@@
     The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
-    \Cref{fig:barplot_all_correct_all_questions} gives an overview of the performance of various models on the entire corpus.
+    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
@@
 An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
-In this figure, we show the percentage of questions that the models answered completely correctly.
-Moreover, we show the highest, average, and lowest performances of the human experts, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+In this figure, we show the percentage of questions that the models answered correctly.
+Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
 Many other models also outperform the average human performance.
-However, it is interesting to point out that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
+It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
 
 Given the large interest in tool-augmented systems, the bad performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
-Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of 10 \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to find the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
-This observation highlights the importance of communicating and measuring not only the predictive performance but also the computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
+With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
+This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
 
@@
     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.
+    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correct by the models.
@@
 While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4), while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
 Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 These findings also shine an interesting light on the value of textbook-inspired questions. 
-A subset of the questions in \chembench are based on textbooks targeted at undergraduate students. 
-On those questions the models tend to perform better than on some our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
+A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
 For instance, while the overall performance in the chemical safety topic is low, the models would pass the examination for a chemical license based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
-While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such a question set or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
 
@@
 Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
-For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
+For this purpose, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
 
@@
 % Estimate of difficulty
-\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they could be able to answer a question correctly. 
+\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
-To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate (on an ordinal scale) their confidence to answer the question correctly.
+To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.
 
@@
 To obtain those estimates, we prompted the models to return a confidence score on an ordinal scale.
-We then plot the correctness of the answers [which can be either completely correct (1) or completely incorrect (0)] against the confidence score.
+We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
         The stripplots show the individual data points, a strong density indicating a higher number of points.
@@
 In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation which highlights the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
-For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered in correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
+For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
@@
 \section{Discussion and Conclusions}
-Our findings show why \glspl{llm} are catching the interest of scientists across domains, but also the attention of the general public, as such models seemed inconceivable just a few years ago.
-As observed, leading models outperform domain experts in chemistry questions on a majority of topics.
-Unsurprinsigly, there are still striking limitations.
-First, it is important to note that the models are not able to estimate their own limitations.
-Moreover, on very relevant topics (such as GHS classification) the answers provided by the models are mostly inaccurate.
-Looking in more detail at the relationship between the model performance and the question type/source revealed that the studied models perform well on textbook questions, while underperforming on questions that require more reasoning (e.g.
-questions that require calculations).
-The success of ChemBench also revealed limitations about the exams we use to test the models (but also the chemists), rather than the limitations of the models themselves.
+On one hand, our findings show that it is impressive how far we got. Leading models outperform domain experts in chemistry questions on many topics. 
+On the other hand, there are still striking limitations. On very relevant topics the answers models provide are wrong. And models are not able to estimate their own limitations.
+However, the success of our benchmark perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
+For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning.
 
 Yet, given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
-Critical reasoning is increasingly important and robotic solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
+Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
 
-Our findings also highlight the nuanced tradeoff between breadth and depth of evaluation frameworks. 
+Our findings also highlight the nuanced trade-off between breadth and depth of evaluation frameworks. 
 The analysis of model performance on different topics shows that the performance of models varies widely across topics. 
-But even within a topic, the performance of models can vary notably depending on the type of question and the reasoning required to answer it.
+But even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
 
 The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific tasks. 
-They cannot easily be used to evaluate systems, or models built for scientific applications. As a result, we only had limited understanding of the capabilities of \glspl{llm} in the chemical sciences.
-Our work shows that carefully curated benchmarks can provide a deeper understanding of the capabilities of \glspl{llm} in the chemical sciences.
+They cannot easily be used to evaluate systems, or models built for scientific applications. Thus, we had only little understanding of the capabilities of \glspl{llm} in the chemical sciences.
+Our work, shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
 It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
-Our findings also illustrate that more focus is needed in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
+Our findings also illustrate that more focus is required in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml}, such as computer vision.
-We believe that while current systems are far from reasoning like a chemist, our open-source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision.
+We thus believe that although current systems are far from reasoning like a chemist our open source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
 
@@
 
-To ensure that the questions do not enter a training dataset we use the same canary string as the BIG-bench project.
+To ensure that the questions do not enter a training dataset we use the same canary string as the Big Bench project.
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
@@
 \subparagraph{Analytical chemistry}
-\variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the University of Jena (Germany). 
+\variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the Friedrich Schiller University Jena (Germany). 
 These questions focus on the principles and techniques employed in analytical chemistry.
@@
 \subparagraph{Combustion engineering}
-\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question is based on master student examination papers at the University of Magdeburg (Germany).
+\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question is based on a master student examination paper at the University of Magdeburg (Germany). 
 It explores topics related to combustion processes and engineering principles.
@@
 \subparagraph{Functional materials and nanomaterials}
-\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises from a seminar conducted at the University of Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
+\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the Friedrich Schiller University Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
 
@@
 
-\subparagraph{Chemistry Olympiads}
-\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various local and international chemistry competitions, including those held in the USA, UK, and Moldova.
+\subparagraph{International Chemical Olympiad}
+\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
 
@@
 This category comprises a total of \variable{output/question_count_per_dir/json_file_counts_safety.txt} questions, including semi-programmatically created questions covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \autocite{pubchem} (see below for details).
-Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals.
+Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals, via semi-programmatically created questions.
 Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
-Moreover, this category includes questions about lab safety, pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
+Moreover, this category addresses questions about lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
 The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
@@
 To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
-We then sampled from the set of number of peaks (excluding the correct answer) to obtain the false answer options.
+We then sampled from the set of all observed peak counts (excluding the correct answer) to obtain the false answer options.
 
@@
 Our parsing workflow is multistep and primarily based on regular expressions.
-In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
+In the case of instruction-tuned models, we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
 In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
@@
 In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
-%Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
+%Manual verification indicates that this \gls{llm}-based parsing represent not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
 
@@
 \paragraph{Models}
-We find that the performance of models depends on the sampling temperature. Typically we find the best temperature with greedy argmax decoding ($T=0$).
 \subparagraph{Completion models}
@@
 For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
-In addition, we used gpt-3.5-turbo and Claude 2 with ReAct-style tool augmentation.\autocite{yao2023react}
-The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (through DuckDuckGo).
-We implemented the system using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
+In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
+The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
+We implemented the systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
 
@@
 To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-In addition, we picked 100 diverse ones using the greedy MaxMin sampling on the embeddings of the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
+In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
 % https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
@@
 For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted to not use external tools other than a calculator and to not consult other humans, we do not have any way to verify that the participants complied with those rules (note that users were also allowed to skip questions).
-Another aspect of requiring unsupervised question answering is the real world equivalent of our benchmark, where humans have unencumbered access to tools and are able to answer any question of interest.
+That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+Note that users were also allowed to skip questions.
+
+Another aspect of requiring unsupervised question answering is that in real life humans have tools and are able to use them for answering any question of interest. 
+In our current study, we prompted users to not use those tools.
 
@@
 Overall, \variable{output/num_users_with_education_info.txt} did so. 
-Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a doctoral degree.
-\variable{output/num_human_postdoc.txt} are beyond a first postdoctorate, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 
@@
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
-To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
+To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706 }.
 
@@
 acknowledges the funding provided by UK Research and Innovation (UKRI) under the UK government’s Horizon Europe funding guarantee (Grant Reference: EP/Y023447/1; Organization Reference: 101106377).
+
+M.R.
+and U.S.S.
+thanks the “Deutsche Forschungsgemeinschaft” for funding under the regime of the priority programme SPP 2363 “Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning” (SCHU 1229/63-1; project number 497115849).
 

### 2024-03-26 src/tex/ms.tex
update authors and acknowledgements

@@
 \section*{Acknowledgements}
-This work was supported by the Carl Zeiss Foundation and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We thank Stability.AI for the use of its HPC cluster and for its generous donations.
-We also thank the OpenBioML.org community and their ChemNLP project team for useful discussions.
+This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena and the FAIRmat consortium.
+We also thank Stability.AI for the use of its HPC cluster and for its generous donations.
+In addition, we thank the OpenBioML.org community and their ChemNLP project team for useful discussions.
+Moreover, we thank Pepe Márquez, and Santiago Miret for discussion and support and Julian Kimmig for feedback on the web app.
 
@@
 and U.S.S.
-thanks the “Deutsche Forschungsgemeinschaft” for funding under the regime of the priority programme SPP 2363 “Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning” (SCHU 1229/63-1; project number 497115849).
-
-The authors thank Julian Kimmig for feedback on the web app.
+thanks the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
 

### 2024-03-26 src/tex/appendix.tex
nice mae vs. number atoms for NMR

@@
 
+\subsection{Performance as a function of molecular features}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
+    \caption{\textbf{Impact of the mean absolute error in the prediction of number of NMR peaks on the number of atoms.}}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
+\end{figure}
+
 \subsection{Human baseline}

### 2024-03-26 src/tex/showyourwork-metadata.tex
nice mae vs. number atoms for NMR

@@
     \def\syw@slug{lamalab-org/chembench-paper}
-    \def\syw@sha{9e3e1d5a5f3dc81ac1ab792b79e58cce26f6bbda}
+    \def\syw@sha{e9da465b586df3b66d049cbdbf915503a23b2cb2}
     \def\syw@runid{}
@@
         \addvalue{fig:topic_barplot_script}{src/scripts/plot_statistics.py}
-        \addvalue{fig:wordcloud_script}{src/scripts/wordcloud.py}
+        \addvalue{fig:question_diversity_script}{src/scripts/plot_question_diversity.py}
+        \addvalue{fig:human_vs_models_bar_script}{src/scripts/plot_overview_performance_plot.py}
+        \addvalue{fig:all_questions_models_completely_correct_radar_human_script}{src/scripts/analyze_model_reports.py}
+        \addvalue{fig:all_questions_models_completely_correct_radar_human_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:confidence_vs_performance_script}{src/scripts/joint_analysis_confidence_performance.py}
+        \addvalue{fig:flesch_kincaid_reading_ease_script}{src/scripts/wordcloud.py}
+        \addvalue{fig:question_count_barplot_mcq_vs_general_script}{src/scripts/plot_statistics.py}
+        \addvalue{fig:barplot_all_correct_all_questions_script}{src/scripts/plot_overview_performance_plot.py}
+        \addvalue{fig:all_questions_models_completely_correct_radar_overall_script}{src/scripts/analyze_model_reports.py}
+        \addvalue{fig:all_questions_models_completely_correct_radar_overall_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:parallel_coordinates_overall_script}{src/scripts/make_parallel_coordinates_plot.py}
+        \addvalue{fig:parallel_coordinates_tiny_script}{src/scripts/make_parallel_coordinates_plot.py}
+        \addvalue{fig:performance_per_topic_script}{src/scripts/analyze_performance_per_source.py}
+        \addvalue{fig:performance_per_topic_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:performance_per_topic_tiny_script}{src/scripts/analyze_performance_per_source.py}
+        \addvalue{fig:performance_per_topic_tiny_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:reading_ease_vs_model_performance_script}{src/scripts/reading_ease_vs_model_performance.py}
+        \addvalue{fig:reading_ease_vs_model_performance_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:correlation_plot_is_number_nmr_peaks_num_atoms_script}{src/scripts/correlate_with_molecule_features.py}
+        \addvalue{fig:human_score_distribution_script}{src/scripts/plot_human_score_distribution.py}
         \addvalue{fig:human_timing_script}{src/scripts/analyze_human_data.py}
         \addvalue{fig:experience_vs_correctness_script}{src/scripts/analyze_human_data.py}
+        \addvalue{fig:confidence_score_distributions_script}{src/scripts/plot_confidence_score_distributions.py}
 
-        \addvalue{output/total_number_of_questions.txt_rule}{Snakefile\#L1}
-        \addvalue{output/manually_generated.txt_rule}{Snakefile\#L1}
-        \addvalue{output/automatically_generated.txt_rule}{Snakefile\#L1}
-        \addvalue{output/number_experts.txt_rule}{Snakefile\#L20}
-        \addvalue{output/total_hours.txt_rule}{Snakefile\#L20}
-        \addvalue{output/flesch_kincaid_reading_ease.txt_rule}{Snakefile\#L30}
-        \addvalue{output/num_h_statements.txt_rule}{Snakefile\#L1}
-        \addvalue{output/num_pictograms.txt_rule}{Snakefile\#L1}
-        \addvalue{output/num_dai.txt_rule}{Snakefile\#L1}
-        \addvalue{output/spearman_experience_score.txt_rule}{Snakefile\#L20}
-        \addvalue{output/spearman_experience_score_p.txt_rule}{Snakefile\#L20}
+        \addvalue{output/total_number_of_questions.txt_rule}{Snakefile\#L20}
+        \addvalue{output/manually_generated.txt_rule}{Snakefile\#L20}
+        \addvalue{output/automatically_generated.txt_rule}{Snakefile\#L20}
+        \addvalue{output/number_experts.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_topics.txt_rule}{Snakefile\#L75}
+        \addvalue{output/mcq_questions.txt_rule}{Snakefile\#L20}
+        \addvalue{output/non_mcq_questions.txt_rule}{Snakefile\#L20}
+        \addvalue{output/num_tiny_questions.txt_rule}{Snakefile\#L20}
+        \addvalue{output/subset_scores/is_number_nmr_peaks_gpt4.txt_rule}{Snakefile\#L184}
+        \addvalue{output/human_subset_scores/is_number_nmr_peaks.txt_rule}{Snakefile\#L184}
+        \addvalue{output/subset_scores/is_gfk_gpt4.txt_rule}{Snakefile\#L184}
+        \addvalue{output/subset_scores/is_gfk_claude3.txt_rule}{Snakefile\#L184}
+        \addvalue{output/human_subset_scores/is_gfk.txt_rule}{Snakefile\#L184}
+        \addvalue{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt_rule}{Snakefile\#L273}
+        \addvalue{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_combustion_engineering.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_ac_faessler_tum.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_pum_tum.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_icho.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_materials_synthesis.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_organic_reactivity.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_oup.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_safety.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_materials_compatibility.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_blac_gfk.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_reactive_groups.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_oxidation_states.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_electron_counts.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_number_of_isomers.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_point_group.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_smiles_to_name.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_name_to_smiles.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_pictograms.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_h_statements.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_dai.txt_rule}{Snakefile\#L235}
+        \addvalue{output/num_users_with_education_info.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_human_phd.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_human_postdoc.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_human_master.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_human_bachelor.txt_rule}{Snakefile\#L89}
+        \addvalue{output/num_humans_with_more_than_100_scores.txt_rule}{Snakefile\#L39}
+        \addvalue{output/num_humans_with_204_scores.txt_rule}{Snakefile\#L39}
+        \addvalue{output/spearman_experience_score.txt_rule}{Snakefile\#L89}
+        \addvalue{output/spearman_experience_score_p.txt_rule}{Snakefile\#L89}

### 2024-03-26 src/tex/sections/semi_programatically_sources_table.tex
Create semi_programatically_sources_table.tex

\begin{table}[h]
\begin{tabular}{lp{7 cm}r}
\hline
\textbf{Subset} & \textbf{Source description} & \textbf{\#} \\
\hline
Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset. & 24 \\
\hline
\begin{tabular}[c]{@{}l@{}}Total electron count of \\ molecules\end{tabular} & Electron counts based on the data from https://www.cheminfo.org/. & 25 \\

\hline
Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & 156 \\

\hline
Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity). & 652 \\
\hline
Number of NMR peaks & Molecules are sampled from the ZINC database and the number of NMR peaks is computed with OpenChemLib & 520 \\
\hline
Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases. & 16 \\
IUPAC-SMILES pairs & Sampled from the PubChem database & 175 + 175 \\
\hline
\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & 101 \\
 & Definitions of hazard statements & 88 \\
 & GHS classification of chemicals mined through the API & 787 \\
\hline
\multirow{2}{*}{Safety}
& Materials' compatibility & 3196 \\
 & Chemical compatibility & 297 \\
\hline
\end{tabular}
\end{table}

### 2024-03-26 src/tex/sections/manually_sources_table.tex
Create manually_sources_table.tex

\hline
NMR spectroscopy & Questions are based on images shared via the Twitter account (https://twitter.com/NMRspectroscopy). & 5 \\
\hline
Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany. & 22 \\
\hline
\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria. & 6 \\
 & Questions originate from the Bachelor of Science courses  at the Technical University of Munich, Germany. & 8 \\
 \hline
\begin{tabular}[c]{@{}l@{}}Functional materials and \\ nanomaterials\end{tabular} & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany. & 8 \\
\hline
Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & 1 \\
\hline
Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany. & 11 \\

\hline
Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & 193 \\
Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & 34 \\
\hline
Periodic table & Manually created & 15 \\
\hline
Polymer chemistry & Based on examinations at the University of Hamburg, Germany & 17 \\
\hline
Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography. & 202 \\
\hline
\multirow{1}{*}{Safety} &
 Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & 90 \\
\hline
\end{tabular}
\end{table}

### 2024-03-26 src/tex/appendix.tex
update correlation plot and main text

@@
 \subsection{Related work}
-Existing benchmarks such as those from \textcite{guo2023large}, \autocite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
+Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
 While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
@@
 Compared to topics, this is a more fine-grained analysis as topics can be composed of questions from different sources.
-In \cref{fig:performance_per_topic} we see that the performance of the models varies significantly between the different data sources.
-Interestingly, the performance of the models on questions sourced based on textbooks seems to better for our models than some of the semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
+In \Cref{fig:performance_per_topic} we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to better for our models than some of the semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
 
@@
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
     }
@@
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
     }
@@
     \centering 
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
+    \caption{\textbf{Dependence of the mean absolute error in the prediction of number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information theoretic additive measure of compound complexity that follows chemical intuitions.
+    The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
+For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
+\end{figure}
+
+\begin{figure}
+    \centering
     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Impact of the mean absolute error in the prediction of number of NMR peaks on the number of atoms.}}
+    \caption{\textbf{Dependence of the mean absolute error in the prediction of number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
+    For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
     \script{correlate_with_molecule_features.py}
     \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
+    \caption{\textbf{Dependence of the mean absolute error in the prediction of total electron counts on the number of atoms.} The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_electron_counts_num_atoms}
 \end{figure}

### 2024-03-26 src/tex/ms.tex
update correlation plot and main text

@@
     However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, against the expertise of human chemists.
 
-    We curated more than 7000 question answer pairs about a wide array of subfields, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
+    We curated more than 7000 question answer-pairs about a wide array of subfields, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.
@@
 While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
 Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
@@
 
-\subparagraph{Number of NMR peaks} 
+\subparagraph{Number of NMR signals} 
 To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
@@
 To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
+In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} signals and the point group of molecules.
 % https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
@@
 To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
-For questions which can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} peaks, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
+For questions which can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
 For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.

### 2024-03-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \section{Results}
+
+\begin{figure}
+    \includegraphics{figures/overview_figure.pdf}
+    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework. 
+    The foundation of the framework is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
+    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework.
+To provide a baseline, we built a web application that we used to survey experts in chemistry.
+    The results of the evaluations are then compiled in publicly accessible leaderboards which we propose as a foundation for the evaluation of future models.
+    }
+    \label{fig:overview_figure}
+\end{figure}
 

### 2024-03-27 src/scripts/complexity.py

@@ -165,7 +165,7 @@ def GetBottcherComplexity(themol, debug=False):
     return complexity
 
 
-@lru_cache()
+@lru_cache(maxsize=1024)
 def complexity_from_smiles(smiles):
     try:
         mol = Chem.MolFromSmiles(smiles)

### 2024-03-27 src/tex/authors.tex
Update authors.tex
update authors with latest feedback
update authors
sync with overleaf
[showyourwork]
update authors and acknowledgements
update correlation plot and main text
[showyourwork] overleaf sync
add JCSM

@@
 % Question contributors
-\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,0}
-\credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Mara~Wilhemli}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
-\credit{Macjonathan~Okereke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
+\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
+\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,1}
+\credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
+\credit{Mara~Wilhelmi}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
+\credit{Macjonathan~Okereke}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
 
 % Question answerers
-\credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Caroline Holick}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Jan Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Nicole Roesner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
-\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
-\credit{Leanne M. Stafast}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
-\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
+\credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Caroline T. Holick}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Jan Matthias Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Nicole Roesner}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,1}
+\credit{Leanne M. Stafast}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
 
@@
 \credit{Michael~Pieler}{1,0,0,1,0,1,0,1,0,0,0,0,0,1}
-\credit{Philippe~Schwaller}{1,1,0,1,1,1,0,0,0,1,0,0,0,0}
+\credit{Philippe~Schwaller}{1,1,0,1,1,1,0,0,0,1,0,0,0,1}
 \credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}
@@
 \author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
-\author[]{Benedict~Emoekabu}
-\author[4]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
-\author[5,6]{Tanya~Gupta}
-\author[1]{Mara~Willhelmi}
+\author[ ]{Benedict~Emoekabu}
+\author[2]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
+\author[3,4]{Tanya~Gupta~\orcidlink{0009-0001-9523-3290}}
+\author[1]{Mara~Wilhelmi}
 \author[1]{Macjonathan~Okereke~\orcidlink{0009-0007-1013-0502}}
 
-\author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
-\author[13]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
-\author[8]{Amir~Elahi}
+\author[5]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
+\author[6]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
+\author[7]{Amir~Elahi}
 \author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
 \author[1]{Maximilian~Greiner}
-\author[1]{Caroline~Holick}
-\author[1]{Tim~Hoffmann}
+\author[1]{Caroline~T.~Holick\orcidlink{0009-0000-1724-2725}}
+\author[1]{Tim~Hoffmann~\orcidlink{0009-0004-0230-6115}}
 \author[1]{Lea~C.~Klepsch~\orcidlink{0009-0009-3849-1670}}
 \author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
-\author[2, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[8, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
 \author[1]{Jakob~Meyer}
-\author[1]{Jan~Peschel~\orcidlink{0009-0002-4787-2757}}
+\author[1]{Jan~Matthias~Peschel~\orcidlink{0009-0002-4787-2757}}
 \author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
-\author[1]{Nicole~Roesner}
+\author[1, 10]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
 
-\author[1]{Johanna Schreiber}
-\author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[1, 10]{Johanna Schreiber~\orcidlink{0009-0000-0991-8967}}
+\author[1,8,10, 11]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
 \author[1]{Leanne~M.~Stafast}
@@
 % Senior authors
-\author[11,12]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
-\author[5,6]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-\author[1,2, 3, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+\author[12,13]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[3,4]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
+\author[1,8,10, 11, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
 
 \affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
-\affil[2]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[3]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
+\affil[8]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[11]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
+\affil[10]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
 
-\affil[5]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
-\affil[6]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[3]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[4]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
 
-\affil[8]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
+\affil[7]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
 \affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
-\affil[11]{OpenBioML.org}
-\affil[12]{Stability.AI}
+\affil[12]{OpenBioML.org}
+\affil[13]{Stability.AI}
 \affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[4]{QpiVolta Technologies Pvt Ltd}
-\affil[7]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
-\affil[13]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
+\affil[2]{QpiVolta Technologies Pvt Ltd}
+\affil[5]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
+\affil[6]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
 \affil[\Letter]{\texttt{mail@kjablonka.com}}

### 2024-03-27 src/tex/ms.tex
add JCSM

@@
 
-In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
+In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
@@
 On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
-For instance, while the overall performance in the chemical safety topic is low, the models would pass the examination for a chemical license based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
 While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
 
-% Calculation vs. no calculation
-Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
-For this purpose, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here
+It is important to note that the model performance for some topics, however, is underestimated in the current evaluation.
+This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
+For instance, models might refuse to provide answers about cyanides.
+To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
 

### 2024-03-27 src/tex/sections/semi_programatically_sources_table.tex
Update semi_programatically_sources_table.tex

@@
 \hline
-Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset. & 24 \\
+Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset.
+& \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
 \hline
-\begin{tabular}[c]{@{}l@{}}Total electron count of \\ molecules\end{tabular} & Electron counts based on the data from https://www.cheminfo.org/. & 25 \\
-
+\begin{tabular}[c]{@{}l@{}}Total electron count of \\ molecules\end{tabular} & Electron counts based on the data from https://www.cheminfo.org/.
+& \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
 \hline
-Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & 156 \\
-
+Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
 \hline
-Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity). & 652 \\
+Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity).
+& \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
 \hline
-Number of NMR peaks & Molecules are sampled from the ZINC database and the number of NMR peaks is computed with OpenChemLib & 520 \\
+Number of NMR peaks & Molecules are sampled from the ZINC database and the number of NMR peaks is computed with OpenChemLib & \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
 \hline
-Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases. & 16 \\
-IUPAC-SMILES pairs & Sampled from the PubChem database & 175 + 175 \\
+Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases.
+& \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
+IUPAC-SMILES pairs & Sampled from the PubChem database & \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
 \hline
-\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & 101 \\
- & Definitions of hazard statements & 88 \\
- & GHS classification of chemicals mined through the API & 787 \\
+\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
+ & Definitions of hazard statements &  \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} \\
+ & GHS classification of chemicals mined through the API & \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} \\
 \hline
 \multirow{2}{*}{Safety}
-& Materials' compatibility & 3196 \\
- & Chemical compatibility & 297 \\
+& Materials' compatibility & \variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} \\
+ & Chemical compatibility & \variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} \\
 \hline

### 2024-03-27 src/tex/appendix.tex
add note about leaderboard

@@
 
+\subsection{Leaderboard}
+Our leaderboard is based on the toolchain developed for matbench.
+Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
+The Markdown tables and plots are then automatically generated and updated on the \chembench website.
+
 \clearpage

### 2024-03-27 src/tex/appendix.tex
add matbench reference

--- 2024-03-26 src/tex/appendix.tex
@@
 
+\subsection{Leaderboard}
+Our leaderboard is based on the tool chain developed for matbench.\autocite{Dunn_2020} Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
+The Markdown tables and plots are then automatically generated and updated on the \chembench website.
+
 \clearpage

### 2024-03-27 src/tex/ms.tex
add note about leaderboard
add matbench reference

@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination\cite{kung2023performance} or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination\autocite{kung2023performance} or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
@@
 To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
+This is facilitated with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.
 
@@
 The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific tasks. 
-They cannot easily be used to evaluate systems, or models built for scientific applications. Thus, we had only little understanding of the capabilities of \glspl{llm} in the chemical sciences.
+They cannot easily be used to evaluate systems, or models built for scientific applications. 
+Thus, we had only little understanding of the capabilities of \glspl{llm} in the chemical sciences.
 Our work, shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.

### 2024-03-27 src/tex/ms.tex
update authors
adrian feedback

@@
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of autonomous systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
@@
 
-To ensure that the questions do not enter a training dataset we use the same canary string as the Big Bench project.
+To ensure that the questions do not enter a training dataset we use the same canary string as the BigBench project.
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
@@
 Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals, via semi-programmatically created questions.
-Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).
+Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).\autocite{blac}
 Moreover, this category addresses questions about lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 

### 2024-03-27 src/scripts/correlate_with_molecule_features.py

@@ -142,20 +142,20 @@ def prepare_data(
         )
         questions_ = questions_.dropna(subset=["smiles"])
         questions_["molecular_weight"] = questions_["smiles"].map(smiles_to_mass)
-        questions_["num_atoms"] = questions_["smiles"].apply(
+        questions_["num_atoms"] = questions_["smiles"].map(
             lambda x: (Chem.MolFromSmiles(x).GetNumAtoms() if x is not None else None)
         )
-        questions_["complexity"] = questions_["smiles"].apply(
+        questions_["complexity"] = questions_["smiles"].map(
             lambda x: complexity_from_smiles(x) if x is not None else None
         )
     elif type_mol == "molecular_formula":
         questions_["molecular_formula"] = questions_["question"].str.extract(
             r"\\ce\{(.*)\}"
         )
-        questions_["molecular_formula_weight"] = questions_["molecular_formula"].apply(
+        questions_["molecular_formula_weight"] = questions_["molecular_formula"].map(
             lambda x: compute_molecular_mass(x) if x is not None else None
         )
-        questions_["num_atoms"] = questions_["molecular_formula"].apply(
+        questions_["num_atoms"] = questions_["molecular_formula"].map(
             lambda x: composition_to_num_atoms(x) if x is not None else None
         )
 
@@ -265,7 +265,7 @@ def plot_correlations_num_atoms(questions: dict):
                 questions_ = prepare_data(questions, model, topic, type_mol=type_mol)
                 questions_["model"] = model
                 all_data.append(questions_)
-            all_data = pd.concat(all_data).dropna().reset_index(drop=True)
+            all_data = pd.concat(all_data).reset_index(drop=True)
             covariate_range = all_data[covariate]
 
             print("Plotting models")

### 2024-03-27 src/scripts/plot_question_diversity.py

@@ -43,8 +43,22 @@ def plot_pca_map():
     # make legend in two columns
 
     plt.axis("off")
-    sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=10, alpha=0.7, ax=ax)
-    ax.legend(loc="center left", ncol=3, bbox_to_anchor=(0, 1.2))
+    plot = sns.scatterplot(x=pca_1, y=pca_2, hue=labels, s=10, alpha=0.7, ax=ax)
+
+    legend_handles_with_increased_size = plot.legend_.legendHandles
+    for handle in legend_handles_with_increased_size:
+        handle.set_sizes([10])
+
+    legend_labels = plot.legend_.texts
+
+    ax.legend(
+        legend_handles_with_increased_size,
+        legend_labels,
+        loc="center left",
+        ncol=3,
+        bbox_to_anchor=(0, 1.2),
+    )
+
     f.tight_layout()
     f.savefig(figures / "question_diversity.pdf", bbox_inches="tight")
 

### 2024-03-27 src/tex/ms.tex
reading pass

@@
 
-    We curated more than 7000 question answer-pairs about a wide array of subfields, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
+    We curated more than 7000 question-answer pairs about a wide array of subfields of the chemical sciences, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.
@@
     These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, there is a critical need for further research to enhance their safety and utility in chemical sciences.
-    Our findings also indicate a need for adaptation to chemistry curricula and highlight the importance of continued development of evaluation frameworks for systematic improvement of safe and useful LLMs.
+    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continued development of evaluation frameworks for systematic improvement of safe and useful LLMs.
 \end{abstract}
@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination\autocite{kung2023performance} or identify and autonomously buy chemicals when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4}
+Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination.\autocite{kung2023performance}
+They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow}
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-In any case, the promise of so-called foundation models (the leading ones being called frontier models) is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+In any case, the promise of these models is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
 This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
@@
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build first prototypes of systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
 For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
@@
 Yet, it is important to realize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
-However, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
+Still, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to experts.
+Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.
 
-Thus, to better understand what \glspl{llm} can do for chemistry and materials science, and where they might be improved with further developments, evaluation frameworks are needed to allow us to systematically measure progress and mitigate potential harms.
-For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmarks suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
+Thus, to better understand what \glspl{llm} can do for the chemical sciences, and where they might be improved with further developments, evaluation frameworks are needed to allow us to systematically measure progress and mitigate potential harms.
+For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmark suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
 The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
@@
 
-\section{Results}
+\section{Results and Discussion}
 
@@
 To compile our benchmark corpus we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator as well as automatic checks.
+For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator as well as automated checks.
 
@@
 It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
-For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
+For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
@@
 \paragraph{Benchmark suite design} Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
-For instance, Galactica\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
+For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
 Current benchmarking suites, however, do not account for such special treatment of scientific information.
@@
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in a real-world application. Hence, we use them for our evaluations.
+The text completions, however, are the final outputs of the system that would also be used in a real-world application. 
+Hence, we use them for our evaluations.
 
@@
 
-Given the large interest in tool-augmented systems, the bad performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
+Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
@@
 
-% Radar chart
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
@@
 \end{figure}
-In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
+In this spider chart, the worst score for every dimension is zero (no question answered correctly) and the best score is one (all questions answered correctly). 
 Thus, a larger colored area indicates a better performance. 
@@
 On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
-For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
 While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+
+We also gain insight in the struggles the models have with chemical reasoning tasks when we look at the performance as a function of molecular descriptors.
+If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
+We, however, find that the performance of the models does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{fig:performance_per_molecular_size}).
+This indicates that the models might not (yet)  be able to reason about the structures of the molecules (in the way one might expect) but rather rely on retrieval of fragments of the training data.
+Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
 
@@
 The figure shows that the reliability of the confidence estimates varies widely across models.
-While the ones for GPT-4 and Claude-2 are not particularly reliable, the ones for Claude-3 follow the expected trend.}
+While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend.}
     \label{fig:confidence_vs_performance}
@@
 
-\section{Discussion and Conclusions}
-On one hand, our findings show that it is impressive how far we got. Leading models outperform domain experts in chemistry questions on many topics. 
-On the other hand, there are still striking limitations. On very relevant topics the answers models provide are wrong. And models are not able to estimate their own limitations.
-However, the success of our benchmark perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
+\section{Conclusions}
+On one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in certain chemistry questions on many topics. 
+On the other hand, there are still striking limitations. 
+For very relevant topics the answers models provide are wrong. 
+On top of that, many models are not able to reliably estimate their own limitations.
+Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
 For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning. 
-
-Yet, given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
+Given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
 Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
@@
 Our findings also highlight the nuanced trade-off between breadth and depth of evaluation frameworks. 
-The analysis of model performance on different topics shows that the performance of models varies widely across topics. 
+The analysis of model performance on different topics shows that the performance of models varies widely across subfields they are tested on. 
 But even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
 
-The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific tasks. 
-They cannot easily be used to evaluate systems, or models built for scientific applications. 
+The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific property prediction tasks. 
+They cannot be used to evaluate reasoning or systems built for scientific applications. 
 Thus, we had only little understanding of the capabilities of \glspl{llm} in the chemical sciences.
-Our work, shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
-It lays the groundwork for the development of more advanced systems that can be used in the chemical sciences and highlights the importance of continued development of evaluation frameworks for systematic improvement of safe and useful \glspl{llm}.
-Our findings also illustrate that more focus is required in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
+Our work shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
+Importantly, our findings also illustrate that more focus is required in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
 
 While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision.
-We thus believe that although current systems are far from reasoning like a chemist our open source \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
+We thus believe that although current systems are far from reasoning like a chemist, our \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
 
@@
 For our dataset we curated questions from existing exams, but also programmatically created new questions.
-Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review.
+Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review (\Cref{fig:curation_workflow}).
 
@@
 \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
-In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
+In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude 2, to parse the completion.
 %Manual verification indicates that this \gls{llm}-based parsing represent not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
@@
 
-M.A.
-expresses gratitude to the European Research Council (ERC) for evaluating the project with the reference number 101106377 titled \enquote{CLARIFIER} and accepting it for funding under the HORIZON TMA MSCA Postdoctoral Fellowships - European Fellowships.
-Furthermore, M.A.
-acknowledges the funding provided by UK Research and Innovation (UKRI) under the UK government’s Horizon Europe funding guarantee (Grant Reference: EP/Y023447/1; Organization Reference: 101106377).
+M.A.\ expresses gratitude to the European Research Council (ERC) for evaluating the project with the reference number 101106377 titled \enquote{CLARIFIER} and accepting it for funding under the HORIZON TMA MSCA Postdoctoral Fellowships - European Fellowships.
+Furthermore, M.A.\ acknowledges the funding provided by UK Research and Innovation (UKRI) under the UK government’s Horizon Europe funding guarantee (Grant Reference: EP/Y023447/1; Organization Reference: 101106377).
 
-M.R.
-and U.S.S.
-thanks the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
+M.R.\ and U.S.S.\ thanks the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
 

### 2024-03-28 src/tex/sections/manually_sources_table.tex

@@
 \hline
-Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany. & 22 \\
+Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany.
+&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
 \hline
-\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria. & 6 \\
+\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria.
+& \variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} \\
  & Questions originate from the Bachelor of Science courses  at the Technical University of Munich, Germany. & 8 \\
  \hline
-\begin{tabular}[c]{@{}l@{}}Functional materials and \\ nanomaterials\end{tabular} & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany. & 8 \\
+\begin{tabular}[c]{@{}l@{}}Functional materials and \\ nanomaterials\end{tabular} & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany.
+& \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
 \hline
-Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & 1 \\
+Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} \\
 \hline
-Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany. & 11 \\
-
+Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany.
+& \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
 \hline
-Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & 193 \\
-Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & 34 \\
+Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
+Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
 \hline
-Periodic table & Manually created & 15 \\
+Periodic table & Manually created & \variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} \\
 \hline
-Polymer chemistry & Based on examinations at the University of Hamburg, Germany & 17 \\
+Polymer chemistry & Based on examinations at the University of Hamburg, Germany & \variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} \\
 \hline
-Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography. & 202 \\
+Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography.
+& \variable{output/question_count_per_dir/json_file_counts_oup.txt} \\
 \hline
 \multirow{1}{*}{Safety} &
- Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & 90 \\
+ Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
 \hline

### 2024-03-28 src/tex/appendix.tex
update authors
include tables of dataset sizes

--- 2024-03-26 src/tex/appendix.tex
@@
 
+\subsection{Leaderboard}
+Our leaderboard is based on the tool chain developed for matbench.\autocite{dunn2020benchmarking} 
+Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
+The Markdown tables and interactive plots are then automatically generated and updated on the \chembench website.
+
 \clearpage

### 2024-03-28 src/tex/authors.tex
update authors
adrian feedback
include tables of dataset sizes

@@
 \credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
-\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Amir~Mohammad~Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
 \credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
@@
 \author[6]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
-\author[7]{Amir~Elahi}
+\author[7]{Amir~Mohammad~Elahi~\orcidlink{0009-0001-5907-101X}}
 \author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
@@
 \author[1]{Jakob~Meyer}
+\author[10]{Santiago~Miret~\orcidlink{0000-0002-4787-2757}}
 \author[1]{Jan~Matthias~Peschel~\orcidlink{0009-0002-4787-2757}}
 \author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
-\author[1, 10]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
+\author[1, 11]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
 
 \author[1, 10]{Johanna Schreiber~\orcidlink{0009-0000-0991-8967}}
-\author[1,8,10, 11]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
-\author[1]{Leanne~M.~Stafast}
-\author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
+\author[1,8,11, 12]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[1, 11]{Leanne~M.~Stafast~\orcidlink{0009-0008-5604-261X}}
+\author[13]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
 
 % Senior authors
-\author[12,13]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[14,15]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
 \author[3,4]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-\author[1,8,10, 11, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+\author[1,8,11, 12, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
 
@@
 \affil[8]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[11]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
-\affil[10]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
+\affil[12]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
+\affil[11]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
 
@@
 \affil[7]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
-\affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
-\affil[12]{OpenBioML.org}
-\affil[13]{Stability.AI}
+\affil[13]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
+\affil[14]{OpenBioML.org}
+\affil[15]{Stability.AI}
 \affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
@@
 \affil[6]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
+\affil[10]{Intel Labs}
 \affil[\Letter]{\texttt{mail@kjablonka.com}}

### 2024-03-28 src/tex/ms.tex
include tables of dataset sizes

@@
 If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
-We, however, find that the performance of the models does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{fig:performance_per_molecular_size}).
+We, however, find that the performance of the models does not correlate with complexity indicators but rather trivially with the size of the compounds (see XX).
 This indicates that the models might not (yet)  be able to reason about the structures of the molecules (in the way one might expect) but rather rely on retrieval of fragments of the training data.
@@
 
-\subparagraph{Analytical chemistry}
-\variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} questions are based on master student examination papers at the Friedrich Schiller University Jena (Germany). 
-These questions focus on the principles and techniques employed in analytical chemistry.
-
-\subparagraph{Combustion engineering}
-\variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} question is based on a master student examination paper at the University of Magdeburg (Germany). 
-It explores topics related to combustion processes and engineering principles.
-
-\subparagraph{Functional materials and nanomaterials}
-\variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt}  questions are based on exercises in a seminar conducted at the Friedrich Schiller University Jena (Germany), delving into topics concerning functional materials and nanomaterials, exploring their properties and applications.
-
-\subparagraph{General chemistry}
-\variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} questions originate from examinations for grade 10 at Marist Comprehensive Academy Uturu (Nigeria). 
-They cover fundamental concepts in chemistry suitable for students at the secondary school level.
-In addition, we included questions based on the general chemistry courses of the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), focusing on inorganic chemistry (\variable{output/question_count_per_dir/json_file_counts_ac_faessler_tum.txt}) and principles of chemistry (\variable{output/question_count_per_dir/json_file_counts_pum_tum.txt}).
-
-\subparagraph{International Chemical Olympiad}
-\variable{output/question_count_per_dir/json_file_counts_icho.txt} questions are based on various countries' Chemical Olympiads, including those held in the USA, UK, and Moldova.
-
-\subparagraph{Material synthesis}
-\variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} questions are based on seminars on material synthesis at the University of Jena (Germany), focusing on the methods and techniques employed in the synthesis of various materials.
-
-\subparagraph{Organic reactivity}
-\variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt}  questions are primarily sourced from tutorials and exam papers for the Bachelor of Science in Chemistry at the Technical University of Munich (Germany), exploring reactions and mechanisms in organic chemistry.
-
-\subparagraph{Periodic table}
-\variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} questions are manually created to cover various aspects and trends of the periodic table.
-
-\subparagraph{Polymer chemistry}
-\variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} questions are based on examinations held at the University of Hamburg (Germany), covering topics related to polymer chemistry and its applications.
-
-\subparagraph{Textbook questions}
-\variable{output/question_count_per_dir/json_file_counts_oup.txt} questions are based on questions from various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography.
-
-\subparagraph{Chemical safety} 
-This category comprises a total of \variable{output/question_count_per_dir/json_file_counts_safety.txt} questions, including semi-programmatically created questions covering areas such as GHS classification, hazard statements, \gls{dai}, and pictograms from the PubChem database \autocite{pubchem} (see below for details).
-Additionally, it explores crucial topics like materials' compatibility (\variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} questions) and chemical compatibility (\variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} questions), essential for understanding the hazardous impact of mixing certain materials and chemicals, via semi-programmatically created questions.
-Furthermore, it includes \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} questions based on the question bank published by the Federal/State Working Group on Chemical Safety (BLAC), which is used for the expertise examination (\enquote{Sachkundeprüfung}) according to §11 of the Chemical Prohibition Ordinance (\enquote{Chemikalien-Verbotsverordnung}).\autocite{blac}
-Moreover, this category addresses questions about lab safety and topics such as pharmacology, toxicology, material safety data sheets, and chemical safety data sheets. 
-The chemical safety category stands out as one of the primary focus areas within the \chembench corpus.
-
-\subparagraph{NMR spectroscopy}
-Questions are based on images shared via the Twitter account \url{https://twitter.com/NMRspectroscopy}, focusing on the choice and design of  appropriate \gls{nmr} experiments.
+\input{sections/manually_sources_table.tex}
 
 \paragraph{Semi-programatically generated questions}
+
+\input{sections/semi_programatically_sources_table.tex}
 
@@
 \section*{Data and code availability}
-The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
+The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench}.
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
@@
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
-To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706 }.
+To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
 
@@
 In addition, we thank the OpenBioML.org community and their ChemNLP project team for useful discussions.
-Moreover, we thank Pepe Márquez, and Santiago Miret for discussion and support and Julian Kimmig for feedback on the web app.
+Moreover, we thank Pepe Márquez for discussion and support and Julian Kimmig for feedback on the web app.
 

### 2024-03-28 src/tex/sections/manually_sources_table.tex
include tables of dataset sizes

--- 2024-03-26 src/tex/sections/manually_sources_table.tex
@@
-\hline
+\begin{table}[h]
+    \begin{tabularx}{\textwidth}{Xp{7 cm}X}
+    \toprule
+    source & description & question count \\
+    \midrule
 NMR spectroscopy & Questions are based on images shared via the Twitter account (https://twitter.com/NMRspectroscopy). & 5 \\
-\hline
-Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany. & 22 \\
-\hline
-\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria. & 6 \\
- & Questions originate from the Bachelor of Science courses  at the Technical University of Munich, Germany. & 8 \\
- \hline
-\begin{tabular}[c]{@{}l@{}}Functional materials and \\ nanomaterials\end{tabular} & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany. & 8 \\
-\hline
-Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & 1 \\
-\hline
-Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany. & 11 \\
-
-\hline
-Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & 193 \\
-Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & 34 \\
-\hline
-Periodic table & Manually created & 15 \\
-\hline
-Polymer chemistry & Based on examinations at the University of Hamburg, Germany & 17 \\
-\hline
-Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography. & 202 \\
-\hline
-\multirow{1}{*}{Safety} &
- Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & 90 \\
-\hline
-\end{tabular}
+\
+Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany.
+&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
+\
+\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria.
+& \variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} \\
+ & Questions originate from the Bachelor of Science (for chemistry) courses at the Technical University of Munich, Germany. & 8 \\
+Functional materials and  nanomaterials & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany.
+& \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
+Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} \\
+Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany.
+& \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
+Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
+Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
+Periodic table & Manually created & \variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} \\
+Polymer chemistry & Based on examinations at the University of Hamburg, Germany & \variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} \\
+Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography.
+& \variable{output/question_count_per_dir/json_file_counts_oup.txt} \\
+Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
+\bottomrule
+\end{tabularx}
 \end{table}

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
Update src/tex/sections/semi_programatically_sources_table.tex
include tables of dataset sizes

@@
 \begin{table}[h]
-\begin{tabular}{lp{7 cm}r}
-\hline
-\textbf{Subset} & \textbf{Source description} & \textbf{\#} \\
-\hline
+\begin{tabularx}{\textwidth}{Xp{7 cm}X}
+    \toprule
+    source & description & question count \\
+\midrule
 Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset.
 & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
-\hline
-\begin{tabular}[c]{@{}l@{}}Total electron count of \\ molecules\end{tabular} & Electron counts based on the data from https://www.cheminfo.org/.
-& \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
-\hline
+Total electron count of molecules & Electron counts based on the data from https://www.cheminfo.org/. & \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
 Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
-\hline
 Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity).
 & \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
-\hline
-Number of NMR peaks & Molecules are sampled from the ZINC database and the number of NMR peaks is computed with OpenChemLib & \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
-\hline
+Number of NMR signals & Molecules are sampled from the ZINC database and the number of NMR signals is computed with OpenChemLib & \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
 Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases.
@@
 IUPAC-SMILES pairs & Sampled from the PubChem database & \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
-\hline
 \multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
@@
  & GHS classification of chemicals mined through the API & \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} \\
-\hline
 \multirow{2}{*}{Safety}
@@
  & Chemical compatibility & \variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} \\
-\hline
-\end{tabular}
+\bottomrule
+\end{tabularx}
 \end{table}

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
sorted affil

--- 2024-03-26 src/tex/sections/semi_programatically_sources_table.tex
@@
 \begin{table}[h]
-\begin{tabular}{lp{7 cm}r}
-\hline
-\textbf{Subset} & \textbf{Source description} & \textbf{\#} \\
-\hline
-Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset. & 24 \\
-\hline
-\begin{tabular}[c]{@{}l@{}}Total electron count of \\ molecules\end{tabular} & Electron counts based on the data from https://www.cheminfo.org/. & 25 \\
-
-\hline
-Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & 156 \\
-
-\hline
-Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity). & 652 \\
-\hline
-Number of NMR peaks & Molecules are sampled from the ZINC database and the number of NMR peaks is computed with OpenChemLib & 520 \\
-\hline
-Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases. & 16 \\
-IUPAC-SMILES pairs & Sampled from the PubChem database & 175 + 175 \\
-\hline
-\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & 101 \\
- & Definitions of hazard statements & 88 \\
- & GHS classification of chemicals mined through the API & 787 \\
-\hline
+\begin{tabularx}{\textwidth}{p{4 cm}p{6 cm}p{.5cm}X}
+    \toprule
+    source & description & question count \\
+\midrule
+Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset.
+& & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
+Total electron count of molecules & Electron counts based on the data from https://www.cheminfo.org/. & &\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
+Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. && \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
+Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity).
+&& \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
+Number of NMR signals & Molecules are sampled from the ZINC database and the number of NMR signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
+Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well-defined cases.
+&& \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
+IUPAC-SMILES pairs & Sampled from the PubChem database && \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
+\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization && \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
+ & Definitions of hazard statements &&  \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} \\
+ & GHS classification of chemicals mined through the API& & \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} \\
 \multirow{2}{*}{Safety}
-& Materials' compatibility & 3196 \\
- & Chemical compatibility & 297 \\
-\hline
-\end{tabular}
+& Materials' compatibility && \variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} \\
+ & Chemical compatibility && \variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} \\
+\bottomrule
+\end{tabularx}
 \end{table}

### 2024-03-28 src/tex/ms.tex
update overviedw
edits mar 28

--- 2024-03-27 src/tex/ms.tex
@@
 \begin{abstract}
-    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly trained on. 
-    This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
+    Large language models (LLMs) have gathered widespread interest due to their abilities to process human language and perform tasks they have not been explicitly trained on. 
+    This is particularly relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in diverse forms of text.
     LLMs have shown promise to address these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
 
-    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be helpful in understanding limitations, create more performant models and mitigate potential harms.
     Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, against the expertise of human chemists.
 
-    We curated more than 7000 question-answer pairs about a wide array of subfields of the chemical sciences, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.
+    We curated more than 7000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs and found that the best models, on average, outperform the best human chemists in our study.
+    The best-performing models, however, struggle with certain chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, such as describing the safety profile of chemicals.
 
@@
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-In any case, the promise of these models is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
-This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
+Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
 
 Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science \autocite{miret2024llms}.
 For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
@@
 This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous, sim2023chemos}
 
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used in the inverse way to predict toxic ones. 
-Yet, it is important to realize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
-Still, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
+Yet, it is important to recognize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
+Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 
-While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
+While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely adopted.
 This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
@@
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, including safety-related ones, might be very misleading.
-We find that there are still plenty of limitations for current models to overcome, to find use in autonomous systems for chemists. 
+We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists. 
 Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.
@@
 In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster.
+It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster across the embedding space.
 
@@
     \includegraphics{figures/question_count_barplot.pdf}
-    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions.
+    \caption{\textbf{Number of questions for different chemistry topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions.
     The figure shows that not all aspects of chemistry are equally represented in our corpus.
 The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}.
-This figure represents the combined count of \glstext{mcq} and open-ended questions.}
+This figure represents the combined count of multiple choice questions (\glstext{mcq}) and open-ended questions.}
     \label{fig:topic_barplot}
@@
     \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
-We then project the embeddings into a two-dimensional space using \gls{pca}.
-We color the points based on a classification into topics.
-Safety related aspects cover a large part of the figure that is not covered by questions from other topics.}
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} This figure represents a projection of text embeddings from \chembench questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+We project the embeddings into a two-dimensional space using \gls{pca} and color the points based on a classification into topics.
+Safety related aspects, the primary focus of \chembench, cover a large part of the figure that is not covered by questions from other topics.}
     \label{fig:question_diversity}
@@
 
-While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
+While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of today's chemistry education and research practices.
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
@@
 To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the web application for the human baseline study.
+This \enquote{tiny} also provided the seed of the web application for our human baseline study.
 
@@
 
-\paragraph{Benchmark suite design} Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
+\paragraph{Benchmark suite design} Given that the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a targeted manner.
 For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
@@
 
-Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
-This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
+Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is also designed to operate on text completions.
+This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
 Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in a real-world application. 
-Hence, we use them for our evaluations.
+The text completions, however, are the final outputs of the system that would also be used in a real-world application prompting us to them for our evaluations.
 
@@
     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly.
-We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
-    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
+Horizontal bars indicate the performance of various models and highlight statistics of the human performance.
+    Since many of the humans did not answer all the questions, we created this plot based on the common subset of questions that most humans answered.
+    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, with partially correct answers being considered as incorrect.
     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
-    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
-    In this case, we consider the answer as incorrect. 
+    However, we limit those systems to a maximum of ten calls to the \gls{llm}.
+This constraint led the systems to often fail to find the correct answer within the specified number of calls.
+If the \gls{llm} exceeds ten calls, we consider the answer as incorrect.
     }
@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
-Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
-Many other models also outperform the average human performance. 
-It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
+Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) used to survey the experts.
+Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study according to the overall performance metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
+Furthermore, many commercial models also outperform the average human performance. 
+Interestingly, the Galactica model, which was trained specifically for scientific applications, underperforms compared to many advanced commerical and some open-source models and barely exceeds the random baseline.
 
-Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
+Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is particuarly salient. 
 Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
+Using the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), these systems were often unable to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
 This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
@@
     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correct by the models.
+The performance is measured as the fraction of questions that were answered completely correct by each of the models.
     The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly).
@@
 
-It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. 
+It is important to note that the model performance for some topics, however, is underestimated in the current evaluation procedure. 
 This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
 For instance, models might refuse to provide answers about cyanides. 
-To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
-This is facilitated with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.
+To overcome this, direct access to the model weights would be required, and as such we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
+We aim to faciliate this with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.
 
@@
 % Estimate of difficulty
-\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
+\paragraph{Confidence estimates} One interesting question to investigate is whether the models are capable of estimating if they will be able to answer a question correctly. 
 If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
@@
 In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
@@
 On the other hand, there are still striking limitations. 
-For very relevant topics the answers models provide are wrong. 
+For very relevant topics the answers all models provide are wrong. 
 On top of that, many models are not able to reliably estimate their own limitations.
 Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
-For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning. 
+For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning conforming findings of previous studies of LLM reasoning outside of the chemistry domain \autocite{behnamghader-etal-2023-retriever}.
 Given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
@@
 For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
-In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
+Additionally, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
 The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
-We implemented the systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
+We implemented all systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
 
@@
 Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
+Our results confirmed our assumption as only XXX percent of questions received multiple answers.
 For selecting the subset, we decided on addressing two questions:
@@
 Users were open to report about their experience in chemistry. 
-Overall, \variable{output/num_users_with_education_info.txt} did so. 
+Overall, \variable{output/num_users_with_education_info.txt} out of YYYY did so. 
 Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
polishing tables

@@
 \begin{table}[h]
-\begin{tabularx}{\textwidth}{p{4 cm}p{6 cm}p{.5cm}X}
+\begin{tabularx}{\textwidth}{p{3.5 cm}p{6.5 cm}p{.5cm}X}
     \toprule
@@
 & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
+\midrule
 Total electron count of molecules & Electron counts based on the data from https://www.cheminfo.org/. & &\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
+\midrule
 Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. && \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
+\midrule
 Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity).
 && \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
+\midrule
 Number of NMR signals & Molecules are sampled from the ZINC database and the number of NMR signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
+\midrule
 Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well-defined cases.
 && \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
+\midrule
 IUPAC-SMILES pairs & Sampled from the PubChem database && \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
+\midrule
 \multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization && \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
@@
  & GHS classification of chemicals mined through the API& & \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} \\
+ \midrule
 \multirow{2}{*}{Safety}

### 2024-03-28 src/tex/showyourwork-metadata.tex
polishing tables

@@
     \def\syw@slug{lamalab-org/chembench-paper}
-    \def\syw@sha{e9da465b586df3b66d049cbdbf915503a23b2cb2}
+    \def\syw@sha{829e14e5ec3be9b46f822ed113d6c6d6fa5fb029}
     \def\syw@runid{}
@@
         \addvalue{fig:reading_ease_vs_model_performance_cache}{https://sandbox.zenodo.org/record/34705}
+        \addvalue{fig:correlation_plot_is_number_nmr_peaks_complexity_script}{src/scripts/correlate_with_molecule_features.py}
         \addvalue{fig:correlation_plot_is_number_nmr_peaks_num_atoms_script}{src/scripts/correlate_with_molecule_features.py}
+        \addvalue{fig:correlation_plot_is_electron_counts_num_atoms_script}{src/scripts/correlate_with_molecule_features.py}
         \addvalue{fig:human_score_distribution_script}{src/scripts/plot_human_score_distribution.py}
@@
         \addvalue{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_combustion_engineering.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_ac_faessler_tum.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_pum_tum.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_materials_synthesis.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_icho.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_materials_synthesis.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_organic_reactivity.txt_rule}{Snakefile\#L235}
@@
         \addvalue{output/question_count_per_dir/json_file_counts_oup.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_safety.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_blac_gfk.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_number_of_isomers.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_electron_counts.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_oxidation_states.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_reactive_groups.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_point_group.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_smiles_to_name.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_dai.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_h_statements.txt_rule}{Snakefile\#L235}
+        \addvalue{output/question_count_per_dir/json_file_counts_pictograms.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_materials_compatibility.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_blac_gfk.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_reactive_groups.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_oxidation_states.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_electron_counts.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_number_of_isomers.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_point_group.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_smiles_to_name.txt_rule}{Snakefile\#L235}
         \addvalue{output/question_count_per_dir/json_file_counts_name_to_smiles.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_pictograms.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_h_statements.txt_rule}{Snakefile\#L235}
-        \addvalue{output/question_count_per_dir/json_file_counts_dai.txt_rule}{Snakefile\#L235}
         \addvalue{output/num_users_with_education_info.txt_rule}{Snakefile\#L89}

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
polish table

@@
 \midrule
-Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset.
-& & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
+Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
 \midrule
-Total electron count of molecules & Electron counts based on the data from https://www.cheminfo.org/. & &\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
+Total electron count of molecules & Electron counts based on the data from  \url{https://www.cheminfo.org/} & &\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
 \midrule
-Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. && \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
+Oxidation states & Oxidation states questions based on the data from \url{https://www.cheminfo.org/} && \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
 \midrule
-Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity).
-&& \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
+Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}) && \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
 \midrule
-Number of NMR signals & Molecules are sampled from the ZINC database and the number of NMR signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
+Number of \gls[nmr] signals & Molecules are sampled from the ZINC database and the number of \gls{nmr} signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
 \midrule
-Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well-defined cases.
-&& \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
+Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well-defined cases && \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
 \midrule

### 2024-03-28 src/tex/sections/manually_sources_table.tex
sorted affil
polish table

--- 2024-03-26 src/tex/sections/manually_sources_table.tex
@@
-\hline
-NMR spectroscopy & Questions are based on images shared via the Twitter account (https://twitter.com/NMRspectroscopy). & 5 \\
-\hline
-Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany. & 22 \\
-\hline
-\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria. & 6 \\
- & Questions originate from the Bachelor of Science courses  at the Technical University of Munich, Germany. & 8 \\
- \hline
-\begin{tabular}[c]{@{}l@{}}Functional materials and \\ nanomaterials\end{tabular} & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany. & 8 \\
-\hline
-Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany & 1 \\
-\hline
-Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany. & 11 \\
-
-\hline
-Chemistry olympiad & National and International Olympiads held in US, UK and Moldova & 193 \\
-Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany & 34 \\
-\hline
-Periodic table & Manually created & 15 \\
-\hline
-Polymer chemistry & Based on examinations at the University of Hamburg, Germany & 17 \\
-\hline
-Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography. & 202 \\
-\hline
-\multirow{1}{*}{Safety} &
- Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) & 90 \\
-\hline
-\end{tabular}
+\begin{table}[h]
+    \begin{tabularx}{\textwidth}{p{3.5 cm}p{6.5 cm}p{.5cm}X}
+    \toprule
+    source & description & question count \\
+    \midrule
+NMR spectroscopy & Questions are based on images shared via the Twitter account \href{https://twitter.com/NMRspectroscopy}{NMRspectroscopy} && 5 \\
+Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany &&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
+\
+\multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria && \variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} \\
+ & Questions originate from the Bachelor of Science (for chemistry) courses at the Technical University of Munich, Germany && 8 \\
+Functional materials and  nanomaterials & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
+Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany && \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} \\
+Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
+Chemistry olympiad & National and International Olympiads held in US, UK and Moldova && \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
+Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
+Periodic table & Manually created && \variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} \\
+Polymer chemistry & Based on examinations at the University of Hamburg, Germany && \variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} \\
+Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography && \variable{output/question_count_per_dir/json_file_counts_oup.txt} \\
+Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) && \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
+\bottomrule
+\end{tabularx}
 \end{table}

### 2024-03-28 src/tex/sections/manually_sources_table.tex
update credits

@@
 NMR spectroscopy & Questions are based on images shared via the Twitter account \href{https://twitter.com/NMRspectroscopy}{NMRspectroscopy} && 5 \\
+\midrule
 Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany &&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
-\
+\midrule
 \multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria && \variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} \\
  & Questions originate from the Bachelor of Science (for chemistry) courses at the Technical University of Munich, Germany && 8 \\
+ \midrule
 Functional materials and  nanomaterials & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
+\midrule
 Combustion engineering & Based on a Master of Science examination paper at the University of Magdeburg, Germany && \variable{output/question_count_per_dir/json_file_counts_combustion_engineering.txt} \\
+\midrule
 Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
+\midrule
 Chemistry olympiad & National and International Olympiads held in US, UK and Moldova && \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
-Organic reactivity & Based on Bachelor of Science in Chemistry at the Technical University of Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
+\midrule
+Organic reactivity & Based on exams in the Bachelor of Science in Chemistry at the Technical University of Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
+\midrule
 Periodic table & Manually created && \variable{output/question_count_per_dir/json_file_counts_periodic_table_properties.txt} \\
+\midrule
 Polymer chemistry & Based on examinations at the University of Hamburg, Germany && \variable{output/question_count_per_dir/json_file_counts_polymer_chemistry.txt} \\
+\midrule
 Textbook questions & Various textbooks covering biomolecular science, drug synthesis, molecular structure, organic chemistry, and X-ray crystallography && \variable{output/question_count_per_dir/json_file_counts_oup.txt} \\
+\midrule
 Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) && \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\

### 2024-03-28 src/tex/sections/manually_sources_table.tex
make safety table more complete

@@
 \midrule
-Analytical chemistry & Questions are based on exercises in master student examination at the Friedrich Schiller University, Jena, Germany &&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
+Analytical chemistry & Questions are based on exercises in master student examination at the FSU Jena, Germany &&  \variable{output/question_count_per_dir/json_file_counts_analytical_chemistry.txt} \\
 \midrule
 \multirow{2}{*}{General chemistry} & Questions originate from examinations for grade 10 at the Marist Comprenhensive Academy Uturu, Nigeria && \variable{output/question_count_per_dir/json_file_counts_Gen_Chem_MCA.txt} \\
- & Questions originate from the Bachelor of Science (for chemistry) courses at the Technical University of Munich, Germany && 8 \\
+ & Questions originate from the Bachelor of Science (for chemistry) courses at the TUM, Germany && \variable{output/question_count_per_dir/json_file_counts_ac_faessler_tum.txt} + \variable{output/question_count_per_dir/json_file_counts_pum_tum.txt} \\
  \midrule
-Functional materials and  nanomaterials & Questions are based on exercises in a seminar conducted at the Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
+Functional materials and  nanomaterials & Questions are based on exercises in a seminar conducted at the FSU Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_func_mats_and_nanomats.txt} \\
 \midrule
@@
 \midrule
-Materials' synthesis & Questions are based on seminars on material synthesis from Friedrich Schiller University, Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
+Materials' synthesis & Questions are based on seminars on material synthesis from FSU Jena, Germany && \variable{output/question_count_per_dir/json_file_counts_materials_synthesis.txt} \\
 \midrule
@@
 \midrule
-Organic reactivity & Based on exams in the Bachelor of Science in Chemistry at the Technical University of Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
+Organic reactivity & Based on exams in the Bachelor of Science in Chemistry at the TUM, Germany && \variable{output/question_count_per_dir/json_file_counts_organic_reactivity.txt} \\
 \midrule
@@
 Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) && \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
+       & Based on toxicology exams at the LMU Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_LMU_tox.txt} \\
+       & Based on toxicology exams at the University of Vienna, Austria && \variable{output/question_count_per_dir/json_file_counts_tox_pharma_vienna.txt} \\
+       & Based on toxicology exams at WWU Munster, Germany && \variable{output/question_count_per_dir/json_file_counts_tox_wwu.txt} \\
+       & Based on exercises for lectures in high-energy density materials at LMU Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_hedm_munich.txt} \\
+       & Based on pharmacology exams at the University of Vienna, Austria && \variable{output/question_count_per_dir/json_file_counts_pharmacology_vienna.txt} \\
 \bottomrule

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
make safety table more complete

@@
 \midrule
-Number of \gls[nmr] signals & Molecules are sampled from the ZINC database and the number of \gls{nmr} signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
+Number of \gls{nmr} signals & Molecules are sampled from the ZINC database and the number of \gls{nmr} signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
 \midrule

### 2024-03-28 src/tex/sections/manually_sources_table.tex
update tables

@@
 \midrule
-Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundeprufung”) && \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
+Safety & Based on the question bank published by the Federal/State Working Group  on Chemical Safety (BLAC), which is used for the expertise examination (“Sachkundepr\"ufung”) && \variable{output/question_count_per_dir/json_file_counts_blac_gfk.txt} \\
        & Based on toxicology exams at the LMU Munich, Germany && \variable{output/question_count_per_dir/json_file_counts_LMU_tox.txt} \\

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
update tables

@@
 \midrule
-Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
+Number of isomers & MAYGEN was used to compute the number of isomers for a set of \gls{smiles} extracted from the ZINC dataset & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
 \midrule
@@
 \midrule
-Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}) && \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
+Chemical reactivity & Questions are framed based on the information from the \href{https://cameochemicals.noaa.gov/reactivity}{Cameo Chemicals website} && \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
 \midrule

### 2024-03-28 src/tex/ms.tex
update credits
review

--- 2024-03-28 src/tex/ms.tex
@@
     This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
-    LLMs have shown promise to address these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
+    LLMs have shown promise in addressing these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
 
-    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, against the expertise of human chemists.
+    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
 
-    We curated more than 7000 question-answer pairs about a wide array of subfields of the chemical sciences, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.
+    We curated more than 7,000 question-answer pairs about a wide array of chemical sciences subfields, evaluated leading models, and found that the best models outperformed the best human chemists in our study on average.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
 
-    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, there is a critical need for further research to enhance their safety and utility in chemical sciences.
-    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continued development of evaluation frameworks for systematic improvement of safe and useful LLMs.
+    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
+    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to systematically improve safe and useful LLMs.
 \end{abstract}
@@
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text, there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
 For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
-Many of these insights are in form of text in scientific publications. 
-Thus, operating on such texts might be our best way of unlocking these insights and to learn from them.
+Many of these insights are in the form of text in scientific publications. 
+Thus, operating on such texts might be our best way of unlocking these insights and learning from them.
 This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
+Such a usage mode is, in particular, interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
 
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used in the inverse way to predict toxic ones. 
-Yet, it is important to realize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
-Still, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
+To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones. 
+Yet, it is important to realize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to special facilities such as laboratories.
+Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
-However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
+However, even for experts, chemical understanding and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
 Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.
 
-Thus, to better understand what \glspl{llm} can do for the chemical sciences, and where they might be improved with further developments, evaluation frameworks are needed to allow us to systematically measure progress and mitigate potential harms.
-For the development of \glspl{llm}, evaluation is currently mostly performed via standardized benchmark suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
-The former contains, among 204 tasks (such as linguistic puzzles), only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
+Thus, to better understand what \glspl{llm} can do for the chemical sciences and where they might be improved with further developments, evaluation frameworks are needed to allow us to measure progress and mitigate potential harms systematically.
+For the development of \glspl{llm}, evaluation is currently primarily performed via standardized benchmark suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
+Among 204 tasks (such as linguistic puzzles), the former contains only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
 Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
@@
 In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
-Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
+It can be used to evaluate any system that can return text (i.e., including tool-augmented systems).
 
-To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specialization.
-Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, including safety-related ones, might be very misleading.
+To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
 We find that there are still plenty of limitations for current models to overcome, to find use in autonomous systems for chemists. 
@@
 \begin{figure}
-    \includegraphics{figures/overview_figure.pdf}
+    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
     \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework. 
-    The foundation of the framework is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
+    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
     We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework.
 To provide a baseline, we built a web application that we used to survey experts in chemistry.
-    The results of the evaluations are then compiled in publicly accessible leaderboards which we propose as a foundation for the evaluation of future models.
+    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
     }
@@
 
-To compile our benchmark corpus we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator as well as automated checks.
+To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
+For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks.
 
@@
 The distribution of topics is also evident from \Cref{fig:question_diversity} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
-In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster.
+In this representation, semantically similar questions are close to each other, and we color the points based on classification into \variable{output/num_topics.txt} topics. 
+It is clear that a focus of \chembench (by design) lies on safety-related aspects, which in \Cref{fig:question_diversity} appear as a large distinct cluster.
 
@@
     \includegraphics{figures/question_count_barplot.pdf}
-    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions.
+    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) and a classifier operating on word embeddings of the questions.
     The figure shows that not all aspects of chemistry are equally represented in our corpus.
@@
     \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results).
+    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model, leads to qualitatively similar results).
 We then project the embeddings into a two-dimensional space using \gls{pca}.
 We color the points based on a classification into topics.
-Safety related aspects cover a large part of the figure that is not covered by questions from other topics.}
+Safety-related aspects cover a large part of the figure that is not covered by questions from other topics.}
     \label{fig:question_diversity}
@@
 
-While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
+While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
@@
 \paragraph{\enquote{Tiny} subset}
-It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
+It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
@@
 
-\paragraph{Benchmark suite design} Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
+\paragraph{Benchmark suite design} Because the text used in scientific settings differs from typical natural language, many models have been developed that deal with such text in a particular way.
 For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
 Current benchmarking suites, however, do not account for such special treatment of scientific information.
-To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
+To address this, \chembench encodes the semantic meaning of various parts of the question or answer.  
 For instance, molecules represented in \gls{smiles} are enclosed in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
-This allows the model to treat the \gls{smiles} string different from other text. 
+This allows the model to treat the \gls{smiles} string differently from other text. 
 \chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.
 
-Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
+Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
 This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
@@
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
-The text completions, however, are the final outputs of the system that would also be used in a real-world application. 
+The text completions, however, are the system's final outputs, which would also be used in a real-world application. 
 Hence, we use them for our evaluations.
@@
     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
+    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
 Many other models also outperform the average human performance. 
-It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
+It is, however, interesting to note that the Galactica model, which was trained specifically for scientific applications, is outperformed by more general models.
 
-Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
-Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
+Given the considerable interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
+Their lack of performance, however, is partially because we limited the system to a maximum of ten \gls{llm} calls.
 With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
-This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+This observation highlights the importance of communicating and measuring not only predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
 
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
-For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
-We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}).
+We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
+We then computed the percentage of questions the models or humans answered correctly for each topic.
 \begin{figure}[!h]
@@
     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correct by the models.
-    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly).
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly).
 A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
     }
@@
 \end{figure}
-In this spider chart, the worst score for every dimension is zero (no question answered correctly) and the best score is one (all questions answered correctly). 
+In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
 Thus, a larger colored area indicates a better performance. 
@@
 A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
-On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
+On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
 For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
@@
 
-We also gain insight in the struggles the models have with chemical reasoning tasks when we look at the performance as a function of molecular descriptors.
+We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
 If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
-We, however, find that the performance of the models does not correlate with complexity indicators but rather trivially with the size of the compounds (see XX).
-This indicates that the models might not (yet)  be able to reason about the structures of the molecules (in the way one might expect) but rather rely on retrieval of fragments of the training data.
+However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see XX).
+This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
 Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
@@
 It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. 
-This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
+This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
 For instance, models might refuse to provide answers about cyanides. 
 To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
-This is facilitated with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.
+This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
 
@@
 % Estimate of difficulty
-\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
-If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
+\paragraph{Confidence estimates} One might wonder whether the models can estimate if they can answer a question correctly. 
+If they could do so, incorrect answers would be less problematic as one could detect when an answer is incorrect.
 To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.
@@
     \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model.
-To obtain those estimates, we prompted the models to return a confidence score on an ordinal scale.
+We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
 We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
         The stripplots show the individual data points, a strong density indicating a higher number of points.
-The lines show the mean and the errorbars show the standard error of the mean.
+The lines show the mean, and the error bars show the standard error of the mean.
 The figure shows that the reliability of the confidence estimates varies widely across models.
@@
 
-In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
+In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023}
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
@@
 \section{Conclusions}
-On one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in certain chemistry questions on many topics. 
+On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics. 
 On the other hand, there are still striking limitations. 
@@
 For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning. 
-Given that the models outperform the average human in our study, it is clear that we need to rethink the way we teach and examine chemistry.
-Critical reasoning is increasingly important and rote solving of problems or memorization of facts is clearly a domain in which \glspl{llm} will continue to outperform humans.
+Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
+Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans.
 
 Our findings also highlight the nuanced trade-off between breadth and depth of evaluation frameworks. 
-The analysis of model performance on different topics shows that the performance of models varies widely across subfields they are tested on. 
-But even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
+The analysis of model performance on different topics shows that models' performance varies widely across the subfields they are tested on. 
+However, even within a topic, the performance of models can vary widely depending on the type of question and the reasoning required to answer it.
 
-The current evaluation frameworks for \glspl{llm} are mostly designed to measure the performance of the models on specific property prediction tasks. 
+The current evaluation frameworks for chemical \glspl{llm} are primarily designed to measure the performance of the models on specific property prediction tasks. 
 They cannot be used to evaluate reasoning or systems built for scientific applications. 
-Thus, we had only little understanding of the capabilities of \glspl{llm} in the chemical sciences.
+Thus, we had little understanding of the capabilities of \glspl{llm} in the chemical sciences.
 Our work shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
-Importantly, our findings also illustrate that more focus is required in the development of better human-model interaction frameworks given that models are not able to estimate their own limitations.
+Importantly, our findings also illustrate that more focus is required in the development of better human-model interaction frameworks, given that models are not able to estimate their limitations.
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key for the progress of many fields of \gls{ml} such computer vision.
-We thus believe that although current systems are far from reasoning like a chemist, our \chembench framework will be a stepping stone for the development of systems that might come closer to this goal.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
+Although current systems are far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.
 
@@
 \subsection{Curation workflow}\label{sec:curation}
-For our dataset we curated questions from existing exams, but also programmatically created new questions.
+For our dataset, we curated questions from existing exams but also programmatically created new questions.
 Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review (\Cref{fig:curation_workflow}).
 
-To ensure that the questions do not enter a training dataset we use the same canary string as the BigBench project.
+To ensure that the questions do not enter a training dataset, we use the same canary string as the BigBench project.
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
@@
 Some tasks were manually curated, others semi-programmatically.
-For all questions we added semantic annotations to make them compatible with systems that use special processing for modalities that are not conventional natural text.
-Before adding the questions to the corpus, we reviewed with manual as well as automatic methods.}
+We added semantic annotations for all questions to make them compatible with systems that use special processing for modalities that are not conventional natural text.
+We reviewed the questions using manual and automatic methods before adding them to the corpus.}
     \label{fig:curation_workflow}
@@
 \paragraph{Manually curated questions}
-
+Manually curated questions were sourced from various sources, including university exams, textbooks, and question banks.
+\Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
+\begin{table}
 \input{sections/manually_sources_table.tex}
+\label{tab:manually_sources}
+\end{table}
 
@@
 
+\begin{table}
 \input{sections/semi_programatically_sources_table.tex}
-
-\subparagraph{Chemical reactivity}
-\variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} questions are framed based on information from the Cameo Chemicals website (\url{https://cameochemicals.noaa.gov/reactivity}), exploring the reactivity of various chemicals and compounds.
-
-\subparagraph{Oxidation states}
-\variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} questions regarding oxidation states are based on data from the website \url{https://www.cheminfo.org/}.
-
-\subparagraph{Total electron count of molecules}
-\variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} questions pertaining to the total electron count of molecules are based on data from the website \url{https://www.cheminfo.org/}.
-
-\subparagraph{Number of isomers}
-We used MAYGEN\autocite{Yirik_2021} to compute the number of isomers for a set of  \gls{smiles} randomly sampled from the ZINC dataset.\autocite{Irwin_2012}
-To obtain the answers for false options we sampled from the set of all observed isomer counts.
-In total, we generated \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} questions.
-
-\subparagraph{Point group for molecules}
-\variable{output/question_count_per_dir/json_file_counts_point_group.txt} questions involve predicting the point group for molecules.
-Our ChemCaption tool (\url{https://github.com/lamalab-org/chem-caption}) is utilized for data generation.
-Given a molecule, it uses spglib\autocite{spglib} to identify the point group. 
-Assigned point groups were manually verified to only include well-defined cases.
-
-\subparagraph{SMILES-IUPAC}
-We scraped pairs of \gls{smiles} and \gls{iupac} names from the PubChem database \autocite{pubchem} and then framed \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} questions around the task of converting \gls{smiles} notation to \gls{iupac} nomenclature.
-\variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt} questions involve the reverse task, i.e., converting \gls{iupac} nomenclature to \gls{smiles} notation.
-
-\subparagraph{Number of NMR signals} 
-To generate the \variable{output/question_count_per_dir/json_file_counts_name_to_smiles.txt}  tasks about the number of \gls{nmr} peaks, we randomly sampled \gls{smiles} from the ZINC database\autocite{Irwin_2012} and then used OpenChemLib\autocite{openchemlib} to compute the number of diasterotopically distinct hydrogen atoms.
-We then sampled from the set of all observed peak counts (excluding the correct answer) to obtain the false answer options.
-
-\subparagraph{GHS, hazard statements and DAI}
-The \gls{ghs} classification, hazard statements and \gls{dai} data have been extracted from PubChem.\autocite{pubchem}
-The former two have been mined via the PubChem \gls{api}, while the latter has been manually compiled. 
-In total, we generated \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} questions about the \gls{ghs} classification of chemicals, \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} questions about definition of hazard statements, and \variable{output/question_count_per_dir/json_file_counts_dai.txt} questions about \glspl{dai}.
-The \glspl{dai} have been curated to contain only records approved by the \gls{who}.
-The chemicals in this class of questions belong to one of the three classes: pesticides (e.g., calcium arsenate), insecticides (e.g., cyfluthrin) or herbicides (e.g. 2,4-D).
+\end{table}
 
@@
 
-To maintain consistency with the training, we employ distinct prompt templates tailored for completion and instruction-tuned models. 
-We impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
+We employ distinct prompt templates tailored for completion and instruction-tuned models to maintain consistency with the training. 
+We impose constraints on the models within these templates to receive responses in a specific format so that robust, fair, and consistent parsing can be performed, as explained below.
 Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
 For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
-Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).
+Our prompting strategy consistently adheres to these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations, and physical units (by either adding or removing wrappers).
 This step can be easily customized in our codebase.
@@
 In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
-In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
-As initial tests indicated that models sometimes return integers in the form of words, e.g.
-\enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
-In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude 2, to parse the completion.
+In the case of numbers, our regular expression was engineered to deal with various forms of scientific notation.
+As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
+If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.
 %Manual verification indicates that this \gls{llm}-based parsing represent not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
+For this, we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
 In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
-The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
+The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, and web search (using DuckDuckGo).
 We implemented the systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
@@
 \subsection{Confidence estimate}
-To estimate the models' confidence we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the right answer on a scale from 1 to 5. 
-We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases compared to other prompting strategies, which might be more suitable when implemented in systems.
+To estimate the models' confidence, we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the correct answer on a scale from 1 to 5. 
+We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases than other prompting strategies, which might be more suitable when implemented in systems.
 
@@
 \paragraph{Question selection} \label{sec:subset-selection}
-Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
-For selecting the subset, we decided on addressing two questions:
+Since we anticipated that we would not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided to show a relevant subset of all questions to the human scorers.
+For selecting the subset, we decided to address two questions:
 \begin{itemize}
@@
 \end{itemize}
-To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
-In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} signals and the point group of molecules.
+To answer the first question, we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
+In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART.
+We added five random questions about the number of \gls{nmr} signals and the point group of molecules.
 % https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
@@
 \paragraph{Study design}
-For our initial study we wanted to maximize the response rate given our available resources. 
+For our initial study, we wanted to maximize the response rate given our available resources. 
 For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted to not use external tools other than a calculator and to not consult with other humans, we do not have any way to verify that the participants complied with those rules.
+That is, while users were prompted not to use external tools other than a calculator and not consult with other humans, we do not have any way to verify that the participants complied with those rules.
 Note that users were also allowed to skip questions.
 
-Another aspect of requiring unsupervised question answering is that in real life humans have tools and are able to use them for answering any question of interest. 
-In our current study, we prompted users to not use those tools.
+Another aspect of requiring unsupervised question answering is that in real life, humans have tools and can use them to answer any question of interest. 
+In our current study, we prompted users not to use those tools.
 
 \paragraph{Participants}
-Users were open to report about their experience in chemistry. 
+Users were open to reporting about their experience in chemistry. 
 Overall, \variable{output/num_users_with_education_info.txt} did so. 
@@
 \paragraph{Comparison with models}
-To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions which at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
-The latter threshold was chosen to limit it to humans who made serious attempts at systematically answering a part of the questions. 
-This analysis might lead to potential biases, most likely in favor of humans as they were allowed to skip questions.
+To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions that at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
+The latter threshold was chosen to limit it to humans who seriously attempted to answer a part of the questions systematically. 
+This analysis might lead to potential biases, most likely in favor of humans, as they were allowed to skip questions.
 \variable{output/num_humans_with_204_scores.txt} humans answered more than 200 questions.
-For the analysis, we treated each human as a model. For analyses grouped by topic, we computed the topic aggregated averages per human and then averaged over all humans.
+For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans.
 
-\subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset we systematically recorded keywords and sources.
+\subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset, we systematically recorded keywords and sources.
 To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
-For questions which can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions) we use this information to make the assignment.
+We use this information to make the assignment for questions that can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions).
 For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
@@
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
-To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
+To facilitate reproduction, some intermediate analysis results are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
 
@@
 This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena and the FAIRmat consortium.
-We also thank Stability.AI for the use of its HPC cluster and for its generous donations.
-In addition, we thank the OpenBioML.org community and their ChemNLP project team for useful discussions.
-Moreover, we thank Pepe Márquez for discussion and support and Julian Kimmig for feedback on the web app.
+We also thank Stability.AI for the access to its HPC cluster and donations.
 
@@
 
+In addition, we thank the OpenBioML.org community and their ChemNLP project team for valuable discussions.
+Moreover, we thank Pepe Márquez for discussions and support and Julian Kimmig for feedback on the web app. 
+We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}).
+
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is employee of Stability.AI and A.M.\ and N.A.\ are paid contractors of Stability.AI.
+K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is an employee of Stability.AI, and A.M.\ and N.A.\ are paid contractors of Stability.AI.
 

### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex
review

@@
 \midrule
-Number of isomers & MAYGEN was used to compute the number of isomers for a set of \gls{smiles} extracted from the ZINC dataset & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
+Number of isomers & MAYGEN\autocite{Yirik_2021} was used to compute the number of isomers for a set of \gls{smiles} extracted from the ZINC dataset\autocite{Irwin_2012} & & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
 \midrule
@@
 \midrule
-Number of \gls{nmr} signals & Molecules are sampled from the ZINC database and the number of \gls{nmr} signals is computed with OpenChemLib && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
+Number of \gls{nmr} signals & Molecules are sampled from the ZINC database\autocite{Irwin_2012}, OpenChemLib\autocite{openchemlib} is used to compute the number of diasterotopically distinct hydrogen atoms && \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
 \midrule
-Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well-defined cases && \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
+Point group of molecules & Our \href{https://github.com/lamalab-org/chem-caption}{ChemCaption} tool is used to assign the point group using spglib,\autocite{spglib}  and then each case was manually checked to select well-defined cases && \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
 \midrule
-IUPAC-SMILES pairs & Sampled from the PubChem database && \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
+IUPAC-SMILES pairs & Sampled from the PubChem \autocite{pubchem}  database && \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
 \midrule
-\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization && \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
+\multirow{3}{*}{PubChem \autocite{pubchem}  safety data} & Daily allowable intakes according to the World Health Organization && \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
  & Definitions of hazard statements &&  \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} \\

### 2024-03-28 src/tex/appendix.tex
review

--- 2024-03-27 src/tex/appendix.tex
@@
 \begin{itemize}
-    \item \emph{End-to-end automation}. For model development, the evaluations will need to be run many times (e.g., on regular intervals of a training run).
+    \item \emph{End-to-end automation}. For model development, the evaluations must be run many times (e.g., on regular intervals of a training run).
     Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
-    \item \emph{Careful validation by experts}. Manual curation is needed to minimize number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
+    \item \emph{Careful validation by experts}. Manual curation is needed to minimize the number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
     This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
     \item \emph{Usable with models that support special treatment of molecules}.
-Some models such as Galactica\autocite{taylor2022galactica} use special tokenization or encoding procedures for molecules or equations.
-    To support this, the benchmark system must encode the semantic meaning of various parts of the question or answer.
-    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or even just the raw logits.
+Some models, such as Galactica\autocite{taylor2022galactica}, use special tokenization or encoding procedures for molecules or equations.
+    The benchmark system must encode the semantic meaning of various parts of the question or answer to support this.
+    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or raw logits.
     This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
     Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
-    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real world chemistry as well as higher-level university education multiple choice question are seldom utilized.
-    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond the answering of \gls{mcq}.
+    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real-world chemistry, as well as higher-level university education, multiple-choice questions are seldom utilized.
+    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond answering \gls{mcq}.
     \item \emph{Cover a diverse set of topics}.
@@
 Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
-While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
-\chembench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols.
-Moreover, we believe that our robust framework, including the platform \url{chembench.org}, will engage the community in open-source contributions.
+While these benchmarks could provide valuable insights in the short term, they cannot follow the rapid additions to the \gls{llm} space. 
+\chembench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, a particular focus on chemical safety, employment of diverse prompting strategies, and specialized notation for molecules and mathematical symbols.
+Moreover, our robust framework, including the platform \url{chembench.org}, will engage the community in open-source contributions.
 
@@
 To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\autocite{srivastava2022beyond}
-This also implies that future baselines can be built on top of our infrastructure as long as they are saved in the same format.
+This also implies that future baselines can be built on top of our infrastructure if saved in the same format.
 
@@
     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} is a measure of how easy a text is to read.
-It is calculated based on the average number of syllables per word and the average number of words per sentence.
+    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+It is calculated based on the average number of syllables per word and words per sentence.
 The higher the score, the easier the text is to read.
-The distribution of the scores of the questions is shown in the histogram.
+The distribution of the questions' scores is shown in the histogram.
 }
@@
     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of multiple choice questions and general questions per topic.}
+    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
     \label{fig:question_count_barplot_mcq_vs_general}
@@
 \Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
-Note that this ranking is different from the one on the \enquote{tiny} subset.
+Note that this ranking differs from the one on the \enquote{tiny} subset.
 
@@
 \Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
-The general pattern that the performance varies significantly between the different topics is also observed when the models are evaluated on the entire corpus. 
-However, since some subjects are composed of questions from different sources the ranking of the models is in some instances different from the one on the \enquote{tiny} subset.
+The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus. 
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
 
@@
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
-    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none of the questions were answered completely correctly.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
     }
@@
 
-\Cref{fig:parallel_coordinates_overall} shows this data in the form of a parallel coordinates plot. 
-This visualization highlights the important observation that the ranking of the models highly depends on the questions they are evaluated on.
-Only very broad benchmarks have hence the potential to provide a comprehensive view of the capabilities of a model. However, even in those cases, the weighting of the different topics is crucial.
+\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
+This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
+Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
 Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
@@
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+The performance is measured as the fraction of questions answered completely correctly by the models.
 }
@@
     \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+The performance is measured as the fraction of questions answered completely correctly by the models.
 }
@@
 To further investigate the performance of the models, we also compared the performance on different data sources.
-Compared to topics, this is a more fine-grained analysis as topics can be composed of questions from different sources.
-In \Cref{fig:performance_per_topic} we see that the performance of the models varies significantly between the different data sources.
-Interestingly, the performance of the models on questions sourced based on textbooks seems to better for our models than some of the semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
+In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some of the semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
 
@@
     \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions that were answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
         We see that the performance of the models varies significantly between the different data sources.
@@
 One might wonder if questions that are more difficult to parse lead to worse performance of the models.
-\Cref{fig:reading_ease_vs_model_performance} shows that there is no clear correlation between the reading ease of the questions and the performance of the models.
+\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.
 
@@
     \includegraphics{figures/reading_ease_vs_model_performance.pdf}
-    \caption{\textbf{Model performance as a function of reading ease.} The violinplots show the distribution of reading ease scores for questions that were answered completely correctly and those that were not.
+    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
 We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
@@
     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in the prediction of number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information theoretic additive measure of compound complexity that follows chemical intuitions.
-    The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
+    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
@@
     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in the prediction of number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
     For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
@@
     \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in the prediction of total electron counts on the number of atoms.} The plot shows that for the \glspl{llm} the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
+    \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
     \script{correlate_with_molecule_features.py}
@@
 \paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
-This application handles serving the user interface as well as exposes various \gls{rest} \glspl{api} for relevant operations.
+This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
 We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
@@
     \label{fig:human_score_distribution}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions that have been answered completely correctly.
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
@@
     \label{fig:human_timing}
-    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
 \end{figure}
@@
 Additionally, we prompted users to provide additional information about their experience in chemistry. 
-While we recorded very fine-grained information, e.g., including their specialization, we focussed our analysis on the number of years since the first university-level chemistry course.
+While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
 \Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
@@
     \label{fig:experience_vs_correctness}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
 \end{figure}
@@
 \Cref{fig:confidence_score_distributions} plots the distribution of those scores.
-We find that the models show different distributions of confidence scores, which for some are skewed to the extremes.
+We find that the models show different distributions of confidence scores, which, for some, are skewed to the extremes.
 
@@
 The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
-The barplots show how many questions were answered with each confidence score.}
+The bar plots show how many questions were answered with each confidence score.}
     \label{fig:confidence_score_distributions}
@@
 \subsection{Leaderboard}
-Our leaderboard is based on the toolchain developed for matbench.
+Our leaderboard is based on the toolchain developed for Matbench.\autocite{dunn2020benchmarking} 
 Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
-The Markdown tables and plots are then automatically generated and updated on the \chembench website.
+The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.
 

### 2024-03-28 src/tex/ms.tex
update tables

@@
 \Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
-\begin{table}
+\begin{table}[h]
+\caption{\textbf{Sources of manually curated questions.} The table shows the sources and a brief description  as well as the number of the manually curated questions.}
 \input{sections/manually_sources_table.tex}
@@
 \paragraph{Semi-programatically generated questions}
+In addition to the manually curated questions, we also generated questions programmatically.
+An overview of the sources of the semi-programatically generated questions is provided in \Cref{tab:semi_programatically_sources}.
 
-\begin{table}
+\begin{table}[h]
+\caption{\textbf{Sources of semi-programatically generated questions.} The table shows the sources and a brief description as well as the number of the semi-programatically generated questions.}
 \input{sections/semi_programatically_sources_table.tex}
+\label{tab:semi_programatically_sources}
 \end{table}
+
+\clearpage
 

### 2024-03-28 src/tex/ms.tex
review

--- 2024-03-28 src/tex/ms.tex
@@
     These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
-    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to systematically improve safe and useful LLMs.
+    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs systematically.
 \end{abstract}
@@
 
-Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
+Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
 This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones. 
-Yet, it is important to realize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to special facilities such as laboratories.
+Yet, it is important to realize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to controlled facilities such as laboratories.
 Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
@@
 \Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
-\begin{table}
+
 \input{sections/manually_sources_table.tex}
-\label{tab:manually_sources}
-\end{table}
 
 \paragraph{Semi-programatically generated questions}
+In addition to the manually curated questions, we also generated questions programmatically.
+An overview of the sources of the semi-programatically generated questions is provided in \Cref{tab:semi_programatically_sources}.
 
-\begin{table}
 \input{sections/semi_programatically_sources_table.tex}
-\end{table}
+
+\clearpage
 

### 2024-03-28 src/tex/sections/manually_sources_table.tex
update tables
review
cleanup tables

@@
-\begin{table}[h]
-    \begin{tabularx}{\textwidth}{p{3.5 cm}p{6.5 cm}p{.5cm}X}
+    \begin{xltabular}{\textwidth}{p{3.5 cm}p{6.5 cm}p{.5cm}X}
+        \caption{\textbf{Sources of manually curated questions.} The table shows the sources and a brief description  as well as the number of the manually curated questions.} \label{tab:manually_sources} \\
+
     \toprule
@@
     \midrule
-NMR spectroscopy & Questions are based on images shared via the Twitter account \href{https://twitter.com/NMRspectroscopy}{NMRspectroscopy} && 5 \\
+NMR spectroscopy & Questions are based on images shared via the Twitter account \href{https://twitter.com/NMRspectroscopy}{NMRspectroscopy} && \variable{output/question_count_per_dir/json_file_counts_nmr.txt} \\
 \midrule
@@
 \bottomrule
-\end{tabularx}
-\end{table}
+\end{xltabular}

### 2024-03-28 src/tex/ms.tex
smiret edits

@@
 \begin{abstract}
-    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly trained on. 
+    Large language models (LLMs) have found widespread interest due to their abilities to process human language and to perform tasks they have not been explicitly trained on. 
     This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
@@
 
-    We curated more than 7,000 question-answer pairs about a wide array of chemical sciences subfields, evaluated leading models, and found that the best models outperformed the best human chemists in our study on average.
+    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
@@
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-In any case, the promise of these models is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
-This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
+Nevertheless, the promise of these models is that they have show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
 
 Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
-In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text, there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
+In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text, there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.\autocite{miret2024llms}
 For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones. 
-Yet, it is important to realize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to controlled facilities such as laboratories.
+Yet, it is important to recognize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to controlled facilities such as laboratories.
 Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
@@
 
-While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
+While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
 This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
@@
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
-We find that there are still plenty of limitations for current models to overcome, to find use in autonomous systems for chemists. 
+We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
 Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.
@@
 In this representation, semantically similar questions are close to each other, and we color the points based on classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects, which in \Cref{fig:question_diversity} appear as a large distinct cluster.
+It is clear that a focus of \chembench (by design) lies on safety-related aspects, which in \Cref{fig:question_diversity} appear as a large distinct clusters across the embedding space.
 
@@
 Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
-This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
+This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems.
 Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
@@
 Many other models also outperform the average human performance. 
-It is, however, interesting to note that the Galactica model, which was trained specifically for scientific applications, is outperformed by more general models.
+Interestingly, the Galactica model, which was trained specifically for scientific applications, underperformed compared to many advanced commercial and some open-source models and barely exceeds the random baseline.
 
@@
 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023}
+For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.

### 2024-03-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-    Large language models (LLMs) have found widespread interest due to their abilities to process human language and to perform tasks they have not been explicitly trained on. 
+    Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks they have not been explicitly trained on. 
     This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
@@
 
-    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.

### 2024-03-28 src/tex/ms.tex
acknowledge sandeep
manual curation table

@@
 \subsection{Curation workflow}\label{sec:curation}
-For our dataset, we curated questions from existing exams but also programmatically created new questions.
-Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review (\Cref{fig:curation_workflow}).
+For our dataset, we curated questions from existing exams or exercise sheets but also programmatically created new questions.
+Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review (\Cref{fig:curation_workflow}) as well as automated checks (e.g., for compliance with a standardized schema).
 
@@
 \paragraph{Manually curated questions}
-Manually curated questions were sourced from various sources, including university exams, textbooks, and question banks.
+Manually curated questions were sourced from various sources, including university exams, exercises, and question banks.
 \Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
@@
 Moreover, we thank Pepe Márquez for discussions and support and Julian Kimmig for feedback on the web app. 
+In addition, we acknowledge support from Sandeep Kumar with an initial prototype of the web app.
 We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}).

### 2024-03-28 src/tex/appendix.tex
prepare addition of scaling analysis

@@
 
+\subsection{Influence of scale}
+To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
+Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
+
+\Cref{fig:llama_performance} shows the performance of the LLaMA (chat) models with different parameter counts on the \chembench corpus.
+
+\begin{figure}
+    \centering 
+    \includegraphics{figures/llama_performance.pdf}
+    \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
+    \label{fig:llama_performance}
+    \script{analyze_llama_performance.py}
+\end{figure}
+
 \subsection{Human baseline}

### 2024-03-28 src/tex/appendix.tex
add some correlation analysis

@@
 \subsection{Performance as a function of molecular features}
+To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
+\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}).
+The corresponding Spearman correlation coefficients are listed in \Cref{}.
 
@@
 \end{figure}
+
+\begin{table}
+    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation}
+    \begin{tabularx}{\textwidth}{XXXX XXXX}
+        \toprule
+        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+        \midrule
+        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude_3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude_3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+        \midrule 
+        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude_3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+        \midrule 
+        \gls{smiles}/\gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude_3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude_3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+        \bottomrule
+    \end{tabularx}
+\end{table}
 

### 2024-03-28 src/tex/appendix.tex
prepare complete rerun

--- 2024-03-28 src/tex/appendix.tex
@@
 \subsection{Performance as a function of molecular features}
+To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
+\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}).
+The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
 
@@
 
+\begin{table}
+    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation}
+    \begin{tabularx}{\textwidth}{XXXX XXXX}
+        \toprule
+        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+        \midrule
+        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+        \midrule 
+        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+        \midrule 
+        \gls{smiles}/\gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+        \bottomrule
+    \end{tabularx}
+    \label{tab:correlation_coefficients}
+\end{table}
+
 \subsection{Influence of scale}
@@
     \centering 
-    \includegraphics{figures/llama_performance.pdf}
+    %\includegraphics{figures/llama_performance.pdf}
     \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
     \label{fig:llama_performance}
-    \script{analyze_llama_performance.py}
+    %\script{analyze_llama_performance.py}
 \end{figure}

### 2024-03-28 src/scripts/analyze_llama_performance.py

@@ -1,0 +1,13 @@
+import matplotlib.pyplot as plt
+from paths import scripts, figures
+from utils import ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH
+
+plt.style.use(scripts / "lamalab.org")
+
+# make a dummy plots
+fig, ax = plt.subplots(
+    figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH))
+
+ax.plot([1, 2, 3], [1, 2, 3])
+
+fig.savefig(figures / "llama_performance.pdf", bbox_inches="tight")

### 2024-03-29 src/tex/ms.tex
prepare addition of scaling analysis

@@
 
-% make the app somewhat more prominent
-
 \clearpage
@@
 
-    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve model performance and mitigate potential harms.
     Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
@@
     We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as predictions related to chemicals' safety profiles.
 
@@
 This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems.
-Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
+Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react, buehler2023generative}
 In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
@@
     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
-    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
-    In this case, we consider the answer as incorrect. 
+    However, we limit those systems to a maximum of ten calls to the \gls{llm}.
+This constraint led the systems to often fail to find the correct answer within the specified number of calls.
+If the \gls{llm} exceeds ten calls, we consider the answer as incorrect.
     }
@@
 In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
-Thus, a larger colored area indicates a better performance. 
+Thus, a larger colored area indicates a better overall performance across different substasks. 
 One can observe that this performance varies widely across models and topics. 
@@
 On the other hand, there are still striking limitations. 
-For very relevant topics the answers models provide are wrong. 
+For very relevant topics the answers all models provide are wrong. 
 On top of that, many models are not able to reliably estimate their own limitations.

### 2024-03-29 src/tex/appendix.tex
include scale plot
trying to compile

--- 2024-03-28 src/tex/appendix.tex
@@
 
+% \begin{table}
+%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation}
+%     \begin{tabularx}{\textwidth}{XXXX XXXX}
+%         \toprule
+%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+%         \midrule
+%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+%         \midrule 
+%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+%         \midrule 
+%         \gls{smiles}/\gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{tab:correlation_coefficients}
+% \end{table}
+
 \subsection{Influence of scale}
@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering 
@@
 \subsection{Leaderboard}
-Our leaderboard is based on the toolchain developed for Matbench.\autocite{dunn2020benchmarking} 
+Our leaderboard is based on the toolchain developed for Matbench.\autocite{Dunn_2020} 
 Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.

### 2024-03-29 src/tex/ms.tex
trying to compile

--- 2024-03-28 src/tex/ms.tex
@@
 \maketitle
-
-% make the app somewhat more prominent
 
@@
 Among 204 tasks (such as linguistic puzzles), the former contains only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
-Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
+Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{Dunn_2020}
 In these cases, the models are evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, melting temperature or reactivity) or on predicting the outcome of specific chemical reactions.
@@
 Our work shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
-Importantly, our findings also illustrate that more focus is required in the development of better human-model interaction frameworks, given that models are not able to estimate their limitations.
+Importantly, our findings also illustrate that more focus is required in developing better human-model interaction frameworks, given that models cannot estimate their limitations.
 
@@
 If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.
-%Manual verification indicates that this \gls{llm}-based parsing represent not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
 

### 2024-03-29 src/scripts/analyze_llama_performance.py

@@ -4,7 +4,7 @@ from utils import ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH
 import pickle
 
 
-plt.style.use(scripts / "lamalab.org")
+plt.style.use(scripts / "lamalab.mplstyle")
 
 
 def plot_performance(model_score_dict):
@@ -25,7 +25,7 @@ def plot_performance(model_score_dict):
     )
 
     ax.set_xlabel("number of parameters in billions")
-    ax.set_ylabel("fraction of correct predictions")
+    ax.set_ylabel("fraction ")
 
     fig.savefig(figures / "llama_performance.pdf", bbox_inches="tight")
 

### 2024-03-30 src/tex/appendix.tex
fix compilation bugs introduced via bibtex files added in the last iterations

--- 2024-03-28 src/tex/appendix.tex
@@
         \midrule 
-        \gls{smiles}/\gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
@@
 
-\begin{figure}
+\begin{figure}[htb]
     \centering 
-    %\includegraphics{figures/llama_performance.pdf}
+    \includegraphics{figures/llama_performance.pdf}
     \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
     \label{fig:llama_performance}
-    %\script{analyze_llama_performance.py}
+    \script{analyze_llama_performance.py}
 \end{figure}

### 2024-03-30 src/tex/ms.tex
update
attempt to recompile again
fix compilation bugs introduced via bibtex files added in the last iterations

@@
 \begin{abstract}
-    Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks they have not been explicitly trained on. 
+    Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks on which they have not been explicitly trained. 
     This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
@@
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-Nevertheless, the promise of these models is that they have show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
 This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
@@
 Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
@@
 Their lack of performance, however, is partially because we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
+With the default tool augmented setup (using the so-called ReAct method\autocite{yao2022react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
 This observation highlights the importance of communicating and measuring not only predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
@@
 
-\paragraph{Semi-programatically generated questions}
+\paragraph{Semi-programmatically generated questions}
 In addition to the manually curated questions, we also generated questions programmatically.
-An overview of the sources of the semi-programatically generated questions is provided in \Cref{tab:semi_programatically_sources}.
+An overview of the sources of the semi-programmatically generated questions is provided in \Cref{tab:semi_programatically_sources}.
 
@@
 We employ distinct prompt templates tailored for completion and instruction-tuned models to maintain consistency with the training. 
-We impose constraints on the models within these templates to receive responses in a specific format so that robust, fair, and consistent parsing can be performed, as explained below.
+As explained below, we impose constraints on the models within these templates to receive responses in a specific format so that robust, fair, and consistent parsing can be performed.
 Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
@@
 For this, we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
-In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
+In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2022react}
 The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, and web search (using DuckDuckGo).

### 2024-03-30 src/tex/appendix.tex
update

@@
 \begin{table}
-    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation}
+    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+}
     \begin{tabularx}{\textwidth}{XXXX XXXX}

### 2024-03-30 src/tex/appendix.tex
remove scaling plot and instead just add numbers in the text

@@
 In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
-Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some of the semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
 
@@
 To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
-\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}).
+\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
 The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
@@
 
-\subsection{Influence of scale}
+\subsection{Influence of model scale}
 To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
+Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance ().
+
 Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
-
-\Cref{fig:llama_performance} shows the performance of the LLaMA (chat) models with different parameter counts on the \chembench corpus.
-
-\begin{figure}[htb]
-    \centering 
-    \includegraphics{figures/llama_performance.pdf}
-    \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
-    \label{fig:llama_performance}
-    \script{analyze_llama_performance.py}
-\end{figure}
 
@@
 \subsection{Leaderboard}
-Our leaderboard is based on the toolchain developed for Matbench.\autocite{Dunn_2020} 
+Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020} 
 Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.

### 2024-03-30 src/tex/ms.tex
remove scaling plot and instead just add numbers in the text

@@
 If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
-However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see XX).
+However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
 This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.

### 2024-03-30 src/tex/ms.tex
done (?)

@@
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
-To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\\showyourwork} framework.\autocite{Luger2021}
+To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\showyourwork} framework.\autocite{Luger2021}
 The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 

### 2024-03-30 src/tex/appendix.tex
style updates

@@
 
+\clearpage
+
 \subsection{Performance as a function of molecular features}
@@
 To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance ().
+Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
 
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 

### 2024-03-30 src/tex/ms.tex
style updates

@@
 
-    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+    We curated more than 7,000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
     The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
@@
     These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
-    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs systematically.
+    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs.
 \end{abstract}
@@
 However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
-To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones. 
-Yet, it is important to recognize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to controlled facilities such as laboratories.
+To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
 Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
@@
 
-To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
@@
 We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
-If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
+If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
 However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
@@
 The figure shows that the reliability of the confidence estimates varies widely across models.
-While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend.}
+While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
     \label{fig:confidence_vs_performance}
@@
 For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
-While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases. 
+For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 Our parsing workflow is multistep and primarily based on regular expressions.
-In the case of instruction-tuned models, we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
+In the case of instruction-tuned models, we first identify the \texttt{[ANSWER]}\texttt{[\textbackslash ANSWER]} environment we prompt the model to report the answer in.
 In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
@@
 
-\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} Llama2 (70b),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
+\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
 
@@
 \end{itemize}
-To answer the first question, we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
+To answer the first question, we selected up to 13 questions per source that \glspl{llm}did not answer correctly in an initial scoring round.
 In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART.

### 2024-03-30 src/scripts/analyze_llama_performance.py

@@ -15,7 +15,6 @@ plt.style.use(scripts / "lamalab.mplstyle")
 
 
 def plot_performance(model_score_dict):
-    # make a dummy plots
     fig, ax = plt.subplots(
         figsize=(ONE_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH)
     )

### 2024-03-30 src/scripts/analyze_model_reports.py

@@ -149,13 +149,13 @@ def make_overall_performance_radar_plot(df_dict, suffix, human_dicts=None):
                     ],
                     [
                         "GPT-4",
-                        "Claude2",
+                        "Claude 2",
                         # "Claude2-ReAct",
-                        "Claude3",
+                        "Claude 3",
                         "Gemini-Pro",
                         "GPT-3.5-Turbo",
                         #   "GPT-3.5-Turbo-ReAct",
-                        "LLAMA-2-70B",
+                        "LLaMA 2 (70B)",
                         "Galactica-120B",
                         "MixTRAL-8x7B",
                         # "PPLX-7B-Chat",
@@ -390,14 +390,14 @@ def radarplot_requires_calculation(df_dict, human_dicts, suffix):
                 ],
                 [
                     "GPT-4",
-                    "Claude2",
+                    "Claude 2",
                     # "Claude2-ReAct",
-                    "Claude3",
+                    "Claude 3",
                     "Gemini-Pro",
                     "GPT-3.5-Turbo",
                     "GPT-3.5-Turbo-ReAct",
-                    "LLAMA-2-70B",
-                    "Galactica-120B",
+                    "LLaMA (70B)",
+                    "Galactica (120B)",
                     "MixTRAL-8x7B",
                     "PPLX-7B-Chat",
                     "PPLX-7B-Online",

### 2024-03-30 src/scripts/analyze_performance_per_source.py

@@ -75,9 +75,9 @@ model_rename_dict = {
     "gpt4": "GPT-4",
     "claude2": "Claude 2",
     "claude3": "Claude 3",
-    "llama70b": "Llama 70B",
+    "llama70b": "LLaMA (70B)",
     "gemini_pro": "Gemini Pro",
-    "galactica_120b": "Galactica 120B",
+    "galactica_120b": "Galactica (120B)",
     "mixtral": "Mixtral",
     "pplx7b_chat": "PPLX7B Chat",
     "pplx7b_online": "PPLX7B Online",
@@ -85,8 +85,8 @@ model_rename_dict = {
     "claude2_react": "Claude 2 ReAct",
     "claude2_zero_t": "Claude 2 Zero T",
     "gemini_pro_zero_t": "Gemini Pro Zero T",
-    "gpt35turbo": "GPT-35 Turbo",
-    "gpt35turbo_react": "GPT-35 Turbo ReAct",
+    "gpt35turbo": "GPT-3.5 Turbo",
+    "gpt35turbo_react": "GPT-3.5 Turbo ReAct",
 }
 
 

### 2024-03-30 src/scripts/plot_statistics.py

@@ -68,12 +68,12 @@ def question_count_barplot(df):
     ax[0].hlines(topics_mcq, xmin=0, xmax=counts_mcq, linewidth=5, alpha=0.2)
     ax[0].plot(counts_mcq, topics_mcq, "o", markersize=5, alpha=0.6)
     ax[0].set_xscale("log")
-    range_frame(ax[0], all_counts, np.arange(len(topics_general)))
+    range_frame(ax[0], all_counts, np.arange(len(topics_general)), pad=0.15)
 
     ax[1].hlines(topics_general, xmin=0, xmax=counts_general, linewidth=5, alpha=0.2)
     ax[1].plot(counts_general, topics_general, "o", markersize=5, alpha=0.6)
     ax[1].set_xscale("log")
-    range_frame(ax[1], all_counts, np.arange(len(topics_general)))
+    range_frame(ax[1], all_counts, np.arange(len(topics_general)), pad=0.15)
 
     ax[1].set_xlabel("question count")
 

### 2024-03-31 src/tex/sections/manually_sources_table.tex
[showyourwork]

@@
        & Based on pharmacology exams at the University of Vienna, Austria && \variable{output/question_count_per_dir/json_file_counts_pharmacology_vienna.txt} \\
+       & Lab safety quizzes based on various sources && \variable{output/question_count_per_dir/json_file_counts_sci_lab_safety_test.txt} + \variable{output/question_count_per_dir/json_file_counts_lab_safety.txt} + \variable{output/question_count_per_dir/json_file_counts_stolaf.txt} + \variable{output/question_count_per_dir/json_file_counts_chemical_safety_mcq_exam.txt} + \variable{output/question_count_per_dir/json_file_counts_anderson.txt}\\
 \bottomrule

### 2024-04-01 src/tex/appendix.tex
[showyourwork]
feat: add requires-calc analysis

@@
 
+In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
+We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/model_overall_cal.pdf}
+    \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+    \script{requires_calculation_plot.py}
+    \label{fig:requires_cal}
+\end{figure}
+
 \clearpage

### 2024-04-01 src/scripts/plot_overview_performance_plot.py

@@ -15,7 +15,7 @@ model_file_name_to_label = {
     "claude2-react": "Claude 2 + ReAct",
     "claude3": "Claude 3",
     "galatica_120b": "Galactica 120B",
-    "gemin-pro-zero-T": "Gemini Pro",
+    "gemini-pro-zero-T": "Gemini Pro",
     # "gemini-pro": "Gemini Pro",
     "gpt-3.5-turbo": "GPT-3.5 Turbo",
     # "gpt-4": "GPT-4",

### 2024-08-08 src/tex/appendix.tex
small fixes (typo, abbreviations) + fix incomplete citations

@@
 To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
+Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
 

### 2024-08-08 src/tex/ms.tex
[showyourwork]
small fixes (typo, abbreviations) + fix incomplete citations

@@
 The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}.
-This figure represents the combined count of \glstext{mcq} and open-ended questions.}
+This figure represents the combined count of multiple-choice questions (\gls{mcq}) and open-ended questions.}
     \label{fig:topic_barplot}
@@
 
-\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
+\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b,\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
 

### 2024-08-19 src/tex/ms.tex
Update src/tex/ms.tex

@@
 
-\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b,\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
+\subparagraph{Instruction-tuned models} In addition, we tested Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b,\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} and LLaMa3 (8B, 70B, 405B).
 

### 2024-08-21 src/tex/ms.tex
feat: shorten abstract

--- 2024-03-30 src/tex/ms.tex
@@
 \begin{abstract}
-    Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks on which they have not been explicitly trained. 
-    This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
-    LLMs have shown promise in addressing these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
+    Large language models (LLMs) have gained widespread interest due to their ability to process human language. 
+    Also, LLMs have shown promise in addressing the limited data problem in chemistry and are increasingly being harnessed to improve and extend existing workflows.
 
-    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
+    However, we possess only a limited systematic understanding of the reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    Here, we introduce \enquote{\chembench,} an automated framework for evaluating the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of chemists.
 
-    We curated more than 7,000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
-    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
-
-    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
-    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs.
+    We curated more than 7,500 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
+    The models, however, struggle with some chemical reasoning tasks that are easy for the human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles. 
+    Although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
 \end{abstract}

### 2024-08-21 src/tex/ms.tex
feat: shorten abstract

@@
 
-\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.
+\subparagraph{Instruction-tuned models} In addition, we tested Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b,\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} and LLaMa3 (8B, 70B, 405B).
 

### 2024-08-21 src/tex/ms.tex
debug appendix

@@
 For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks. 
-
-Importantly, our large pool of questions encompasses a wide range of topics.
-This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
-The distribution of topics is also evident from \Cref{fig:question_diversity} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
-In this representation, semantically similar questions are close to each other, and we color the points based on classification into \variable{output/num_topics.txt} topics. 
-It is clear that a focus of \chembench (by design) lies on safety-related aspects, which in \Cref{fig:question_diversity} appear as a large distinct clusters across the embedding space.
+Importantly, our large pool of questions encompasses a wide range of topics and question types.
+The topics range from general chemistry to more specialized fields such as inorganic, analytical or technical chemistry.
+We also classify the questions based on what techniques are required to answer them.
+Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
+Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty.
 
@@
     \script{plot_statistics.py}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
-    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model, leads to qualitatively similar results).
-We then project the embeddings into a two-dimensional space using \gls{pca}.
-We color the points based on a classification into topics.
-Safety-related aspects cover a large part of the figure that is not covered by questions from other topics.}
-    \label{fig:question_diversity}
-    \script{plot_question_diversity.py}
 \end{figure}
@@
 \section{Appendix}
-\input{appendix.tex}
+% \input{appendix.tex}
 \clearpage

### 2024-08-26 src/tex/ms.tex
Update ms.tex
feat: add reference

--- 2024-08-21 src/tex/ms.tex
@@
 \begin{abstract}
-    Large language models (LLMs) have gained widespread interest due to their ability to process human language. 
-    Also, LLMs have shown promise in addressing the limited data problem in chemistry and are increasingly being harnessed to improve and extend existing workflows.
+    Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks on which they have not been explicitly trained. 
+    This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
+    LLMs have shown promise in addressing these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.
 
-    However, we possess only a limited systematic understanding of the reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
-    Here, we introduce \enquote{\chembench,} an automated framework for evaluating the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of chemists.
+    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
+    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.
 
-    We curated more than 7,500 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
-    The models, however, struggle with some chemical reasoning tasks that are easy for the human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles. 
-    Although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
+    We curated more than 7,000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
+
+    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
+    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs.
 \end{abstract}
@@
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
-Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
+Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier, ai4science2023impact}
 This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}

### 2024-08-26 src/tex/ms.tex
7500 -> 2500

--- 2024-08-21 src/tex/ms.tex
@@
 
-    We curated more than 7,500 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
+    We curated more than 2,500 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
     The models, however, struggle with some chemical reasoning tasks that are easy for the human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles. 
@@
 \section{Appendix}
-% \input{appendix.tex}
+\input{appendix.tex}
 \clearpage

### 2024-08-31 src/tex/ms.tex
Update ms.tex

@@
 
-    We curated more than 2,500 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
+    We curated more than 2,800 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists. 
     The models, however, struggle with some chemical reasoning tasks that are easy for the human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles. 

### 2024-10-06 src/tex/appendix.tex
small fixes (typo, abbreviations) + fix incomplete citations (#25)
Co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>
Co-authored-by: Martiño Rios-Garcia <147555961+MrtinoRG@users.noreply.github.com
feat: start rewrite

--- 2024-03-30 src/tex/appendix.tex
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
 It is calculated based on the average number of syllables per word and words per sentence.
@@
 }
-    \label{fig:flesch_kincaid_reading_ease}
-    \script{wordcloud.py}
-\end{figure}
+%     \label{fig:flesch_kincaid_reading_ease}
+%     \script{wordcloud.py}
+% \end{figure}
 
@@
 A substantial fraction, in contrast to other benchmarks, is open-ended. 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-    \label{fig:question_count_barplot_mcq_vs_general}
-    \script{plot_statistics.py}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+%     \label{fig:question_count_barplot_mcq_vs_general}
+%     \script{plot_statistics.py}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/overall_performance.pdf}
-    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/overall_performance.pdf}
+%     \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
 Scores computed on the entire \chembench corpus.}
-    \label{fig:barplot_all_correct_all_questions}
-    \script{plot_overview_performance_plot.py}
-\end{figure}
+%     \label{fig:barplot_all_correct_all_questions}
+%     \script{plot_overview_performance_plot.py}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
+%     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
 The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
-    }
-    \label{fig:all_questions_models_completely_correct_radar_overall}
-    \script{analyze_model_reports.py}
-\end{figure}
+%     A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+%     }
+%     \label{fig:all_questions_models_completely_correct_radar_overall}
+%     \script{analyze_model_reports.py}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/parallel_coordinates_overall.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/parallel_coordinates_overall.pdf}
+%     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 }
-    \label{fig:parallel_coordinates_overall}
-    \script{make_parallel_coordinates_plot.py}
-\end{figure}
+%     \label{fig:parallel_coordinates_overall}
+%     \script{make_parallel_coordinates_plot.py}
+% \end{figure}
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/parallel_coordinates_tiny.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/parallel_coordinates_tiny.pdf}
+%     \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 }
-    \label{fig:parallel_coordinates_tiny}
-    \script{make_parallel_coordinates_plot.py}
-\end{figure}
+%     \label{fig:parallel_coordinates_tiny}
+%     \script{make_parallel_coordinates_plot.py}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/performance_per_topic.pdf}
+%     \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
+%         We see that the performance of the models varies significantly between the different data sources.
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic}
-    \script{analyze_performance_per_source.py}
-\end{figure}
+%     }
+%     \label{fig:performance_per_topic}
+%     \script{analyze_performance_per_source.py}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/performance_per_topic_tiny.pdf}
+%     \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
+%         We see that the performance of the models varies significantly between the different data sources.
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic_tiny}
-    \script{analyze_performance_per_source.py}
-\end{figure}
+%     }
+%     \label{fig:performance_per_topic_tiny}
+%     \script{analyze_performance_per_source.py}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \hspace*{-1cm}
-    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
-    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
+% \begin{figure}
+%     \centering
+%     \hspace*{-1cm}
+%     \includegraphics{figures/reading_ease_vs_model_performance.pdf}
+%     \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
 We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
 }
-    \script{reading_ease_vs_model_performance.py}
-    \label{fig:reading_ease_vs_model_performance}
-\end{figure}
+%     \script{reading_ease_vs_model_performance.py}
+%     \label{fig:reading_ease_vs_model_performance}
+% \end{figure}
+
+% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
+% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/model_overall_cal.pdf}
+%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+%     \script{requires_calculation_plot.py}
+%     \label{fig:requires_cal}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \centering 
-    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
-    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
+% \begin{figure}
+%     \centering 
+%     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
+%     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
+%     The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
-    \script{correlate_with_molecule_features.py}
-    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
-\end{figure}
+%     \script{correlate_with_molecule_features.py}
+%     \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
-    For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
-    \script{correlate_with_molecule_features.py}
-    \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
+%     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
+%     For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
+%     \script{correlate_with_molecule_features.py}
+%     \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
-    \script{correlate_with_molecule_features.py}
-    \label{fig:correlation_plot_is_electron_counts_num_atoms}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
+%     \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
+%     \script{correlate_with_molecule_features.py}
+%     \label{fig:correlation_plot_is_electron_counts_num_atoms}
+% \end{figure}
 
-\begin{table}
-    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+% \begin{table}
+%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
 }
-    \begin{tabularx}{\textwidth}{XXXX XXXX}
-        \toprule
-        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
-        \midrule
-        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
-                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
-        \midrule 
-        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
-        \midrule 
-        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
-                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
-        \bottomrule
-    \end{tabularx}
-    \label{tab:correlation_coefficients}
-\end{table}
+%     \begin{tabularx}{\textwidth}{XXXX XXXX}
+%         \toprule
+%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+%         \midrule
+%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+%         \midrule 
+%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+%         \midrule 
+%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{tab:correlation_coefficients}
+% \end{table}
 
 \subsection{Influence of model scale}
-To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
+% To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
+% Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
 
-Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
+% Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
+REWRITE.
 
@@
 This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
-The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
+We utilize a Postgresql.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
 The application is hosted on the Vercel web hosting platform.
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/human_score_distribution.pdf}
-    \script{plot_human_score_distribution.py}
-    \label{fig:human_score_distribution}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/human_score_distribution.pdf}
+%     \script{plot_human_score_distribution.py}
+%     \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
+%     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
+%     \label{fig:human_score_distribution}
+% \end{figure}
 
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/human_timing.pdf}
-    \script{analyze_human_data.py}
-    \label{fig:human_timing}
-    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/human_timing.pdf}
+%     \script{analyze_human_data.py}
+%     \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
+%     \label{fig:human_timing}
+% \end{figure}
 
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+%\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/experience_vs_correctness.pdf}
-    \script{analyze_human_data.py}
-    \label{fig:experience_vs_correctness}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/experience_vs_correctness.pdf}
+%     \script{analyze_human_data.py}
+%     \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
+%     \label{fig:experience_vs_correctness}
+% \end{figure}
 
@@
 
-\begin{figure}[htb] 
-    \centering
-    \includegraphics{figures/confidence_score_distributions.pdf}
-    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
+% \begin{figure}[htb] 
+%     \centering
+%     \includegraphics{figures/confidence_score_distributions.pdf}
+%     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
 The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
 The bar plots show how many questions were answered with each confidence score.}
-    \label{fig:confidence_score_distributions}
-    \script{plot_confidence_score_distributions.py}
-\end{figure}
+%     \label{fig:confidence_score_distributions}
+%     \script{plot_confidence_score_distributions.py}
+% \end{figure}
 

### 2024-10-06 src/tex/ms.tex
feat: start rewrite

@@
 In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+%Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
 Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
@@
 
-To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+%To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
@@
 
-\begin{figure}[!htb]
-    \centering
-    \includegraphics{figures/question_count_barplot.pdf}
-    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) and a classifier operating on word embeddings of the questions.
-    The figure shows that not all aspects of chemistry are equally represented in our corpus.
+% \begin{figure}[!htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot.pdf}
+%     \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) and a classifier operating on word embeddings of the questions.
+%     The figure shows that not all aspects of chemistry are equally represented in our corpus.
 The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}.
 This figure represents the combined count of multiple-choice questions (\gls{mcq}) and open-ended questions.}
-    \label{fig:topic_barplot}
-    \script{plot_statistics.py}
-\end{figure}
+%     \label{fig:topic_barplot}
+%     \script{plot_statistics.py}
+% \end{figure}
 
 While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
+%For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
 
@@
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
-To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
+%To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
 We also used this subset to seed the web application for the human baseline study.
@@
 
-\begin{figure}[!h]
-    \centering
-    \includegraphics{figures/human_subset_performance.pdf}
-    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly.
+% \begin{figure}[!h]
+%     \centering
+%     \includegraphics{figures/human_subset_performance.pdf}
+%     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
-    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
-    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
-    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
-    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
-    In this case, we consider the answer as incorrect. 
-    }
-    \label{fig:human_vs_models_bar}
-    \script{plot_overview_performance_plot.py}
-\end{figure}
+%     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
+%     The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
+%     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
+%     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
+%     However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
+%     In this case, we consider the answer as incorrect. 
+%     }
+%     \label{fig:human_vs_models_bar}
+%     \script{plot_overview_performance_plot.py}
+% \end{figure}
 
@@
 We then computed the percentage of questions the models or humans answered correctly for each topic.
-\begin{figure}[!h]
-    \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+% \begin{figure}[!h]
+%     \centering
+%     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
+%     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
-    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly).
+%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly).
 A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
-    }
-    \label{fig:all_questions_models_completely_correct_radar_human}
-    \script{analyze_model_reports.py}
-\end{figure}
-In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
-Thus, a larger colored area indicates a better performance. 
-One can observe that this performance varies widely across models and topics. 
-While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
-In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
-Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
-These findings also shine an interesting light on the value of textbook-inspired questions. 
-A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
-On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
-For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
-While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+%     This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+%     }
+%     \label{fig:all_questions_models_completely_correct_radar_human}
+%     \script{analyze_model_reports.py}
+% \end{figure}
 
-We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
-If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
-However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
-This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
-Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
+% In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
+% Thus, a larger colored area indicates a better performance. 
+% One can observe that this performance varies widely across models and topics. 
+% While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
+% In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+% Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+% These findings also shine an interesting light on the value of textbook-inspired questions. 
+% A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
+% On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
+% For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+% While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
 
-It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. 
-This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
-For instance, models might refuse to provide answers about cyanides. 
-To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
-This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
+% We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
+% If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
+% However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
+% This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
+% Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
+
+% It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. 
+% This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
+% For instance, models might refuse to provide answers about cyanides. 
+% To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
+% This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
 
@@
 If they could do so, incorrect answers would be less problematic as one could detect when an answer is incorrect.
-To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.
 
-\begin{figure}[!h]
-    \centering
-    \includegraphics{figures/confidence_vs_performance_overall.pdf}
-    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model.
+REWRITE
+
+%To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.
+
+% \begin{figure}[!h]
+%     \centering
+%     \includegraphics{figures/confidence_vs_performance_overall.pdf}
+%     \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model.
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
 We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
-        The stripplots show the individual data points, a strong density indicating a higher number of points.
+%         The stripplots show the individual data points, a strong density indicating a higher number of points.
 The lines show the mean, and the error bars show the standard error of the mean.
@@
 While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
-    \label{fig:confidence_vs_performance}
-    \script{joint_analysis_confidence_performance.py}
-\end{figure}
+%     \label{fig:confidence_vs_performance}
+%     \script{joint_analysis_confidence_performance.py}
+% \end{figure}
 
-In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
-For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
-While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases. 
-For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+% In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
+% For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
+% For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+% While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases. 
+% For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 
-\input{sections/manually_sources_table.tex}
+%\input{sections/manually_sources_table.tex}
 
@@
 
-\input{sections/semi_programatically_sources_table.tex}
+%\input{sections/semi_programatically_sources_table.tex}
 
@@
 \subsection{Model evaluation workflow}
+EXPAND THIS section.
 
@@
 
-\input{sections/parse_check_desc.tex}
-
 \paragraph{Models}
-\subparagraph{Completion models}
-We used Galactica (120b)\autocite{taylor2022galactica} with the default settings.
-
-\subparagraph{Instruction-tuned models} In addition, we tested Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b,\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} and LLaMa3 (8B, 70B, 405B).
-
-\subparagraph{Tool augmented models}
-In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this, we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
-In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2022react}
-The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, and web search (using DuckDuckGo).
-We implemented the systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.
+Rewrite - as table. The table can then also contain if it has been completion or instruction-prompted.
 
@@
 \paragraph{Question selection} \label{sec:subset-selection}
-Since we anticipated that we would not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided to show a relevant subset of all questions to the human scorers.
-For selecting the subset, we decided to address two questions:
-\begin{itemize}
-    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
-    \item Are there areas in which the performance of humans is very different from the ones of the models?
-\end{itemize}
-To answer the first question, we selected up to 13 questions per source that \glspl{llm}did not answer correctly in an initial scoring round.
-In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART.
-We added five random questions about the number of \gls{nmr} signals and the point group of molecules.
-% https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
+Rewrite.
 
 \paragraph{Study design}
-For our initial study, we wanted to maximize the response rate given our available resources. 
-For this reason, we did not opt for a highly controlled study setting. 
-That is, while users were prompted not to use external tools other than a calculator and not consult with other humans, we do not have any way to verify that the participants complied with those rules.
-Note that users were also allowed to skip questions.
-
-Another aspect of requiring unsupervised question answering is that in real life, humans have tools and can use them to answer any question of interest. 
-In our current study, we prompted users not to use those tools.
+Rewrite.
 
 \paragraph{Participants}
-Users were open to reporting about their experience in chemistry. 
-Overall, \variable{output/num_users_with_education_info.txt} did so. 
-Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+% Users were open to reporting about their experience in chemistry. 
+% Overall, \variable{output/num_users_with_education_info.txt} did so. 
+% Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
+% \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 
 \paragraph{Comparison with models}
-To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions that at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
-The latter threshold was chosen to limit it to humans who seriously attempted to answer a part of the questions systematically. 
-This analysis might lead to potential biases, most likely in favor of humans, as they were allowed to skip questions.
+%To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions that at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
+%The latter threshold was chosen to limit it to humans who seriously attempted to answer a part of the questions systematically. 
+%This analysis might lead to potential biases, most likely in favor of humans, as they were allowed to skip questions.
 \variable{output/num_humans_with_204_scores.txt} humans answered more than 200 questions.
@@
 
-\subsection{Classification of questions into topics}\label{sec:meth-topic} When curating our dataset, we systematically recorded keywords and sources.
-To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
-We use this information to make the assignment for questions that can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions).
-For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
+\subsection{Classification of questions into topics}\label{sec:meth-topic} 
+% When curating our dataset, we systematically recorded keywords and sources.
+% To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
+% We use this information to make the assignment for questions that can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions).
+% For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
+
+REWRITE.
 
@@
 \section*{Acknowledgements}
-This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena and the FAIRmat consortium.
-We also thank Stability.AI for the access to its HPC cluster and donations.
+This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
+We also thank Stability.AI for the access to its HPC cluster and donations. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme. 
+K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
 
@@
 In addition, we acknowledge support from Sandeep Kumar with an initial prototype of the web app.
-We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}).
+We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}) and thank Berend Smit and Pepe Márquez for feedback on an early version of the manuscript.
 
 \section*{Conflicts of interest}
-K.M.J.\ is a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is an employee of Stability.AI, and A.M.\ and N.A.\ are paid contractors of Stability.AI.
+K.M.J.\ was a paid consultant for OpenAI (as part of the red teaming network). M.P.\ is an employee of Stability.AI, and A.M.\ and N.A.\ were paid contractors of Stability.AI.
 

### 2024-10-06 src/scripts/classify_questions.py
add statistics about number of questions

@@ -2,7 +2,9 @@ from utils import obtain_chembench_repo
 from paths import output, data
 import os
 import pandas as pd 
-
+import json
+from chembench.task import Task
+from chembench.analysis import add_group_info
 def classify():
     # Load the data
     chembench_repo = obtain_chembench_repo()
@@ -15,13 +17,29 @@ def classify():
     def update_path(path):
         # the dataframe contains paths like '../data/questions/00000000-0000-0000-0000-000000000000.json'
         # we want to get the path like chembench_repo/data/questions/00000000-0000-0000-0000-000000000000.json
-        path_without_dots = path.replace("../data", "")
+        
+        path_without_dots = path.replace("../data/", "")
         return os.path.join(data_, path_without_dots)
     
-    df_categories['path'] = df_categories['path'].apply(update_path)
-    
+    df_categories['name'] = df_categories['index']
+    df_categories['path'] = df_categories['question'].apply(update_path)
+    df_categories = df_categories[df_categories['path'].apply(lambda x: os.path.exists(x))].copy()
+
+    #import pdb; pdb.set_trace()
+    is_mcq = []
+    keywords = []
+    for path in df_categories['path']:
+        task = Task.from_json(path)
+        is_mcq.append('multiple_choice_grade' in task._metrics)
+        keywords.append(task._keywords)
+
+    df_categories['is_classification'] = is_mcq
+    df_categories['keywords'] = keywords
+
     # now, keep those rows where the path exists
-    df = df_categories[df_categories['path'].apply(lambda x: os.path.exists(x))].copy()
+    
+    df = add_group_info(df_categories)
+    
 
     df.to_csv(data / "questions.csv")
     df.to_pickle(data / "questions.pkl")

### 2024-10-06 src/scripts/compute_basic_statistics.py

@@ -14,15 +14,6 @@ def load_data():
     with open(output / "total_number_of_questions.txt", "w") as f:
         f.write(str(total_number_of_questions) + "\endinput")
 
-    automatically_generated = df[df["is_semiautomatically_generated"]]
-
-    with open(output / "automatically_generated.txt", "w") as f:
-        f.write(str(len(automatically_generated)) + "\endinput")
-
-    manually_generated = df[~df["is_semiautomatically_generated"]]
-    with open(output / "manually_generated.txt", "w") as f:
-        f.write(str(len(manually_generated)) + "\endinput")
-
     mcq_questions = df[df["is_classification"]]
     with open(output / "mcq_questions.txt", "w") as f:
         f.write(str(len(mcq_questions)) + "\endinput")
@@ -52,11 +43,11 @@ def load_data():
     with open(output / "num_pictograms.txt", "w") as f:
         f.write(str(len(pictograms)) + "\endinput")
 
-    # read number of questions in "tiny" subset
-    df = pd.read_csv(os.path.join(chembench_repo, "reports", "humans", "questions.csv"))
+    # # read number of questions in "tiny" subset
+    # df = pd.read_csv(os.path.join(chembench_repo, "reports", "humans", "questions.csv"))
 
-    with open(output / "num_tiny_questions.txt", "w") as f:
-        f.write(str(len(df)) + "\endinput")
+    # with open(output / "num_tiny_questions.txt", "w") as f:
+    #     f.write(str(len(df)) + "\endinput")
 
 
 if __name__ == "__main__":

### 2024-10-06 src/scripts/compute_basic_statistics.py
automatically vs. manually

@@ -22,6 +22,16 @@ def load_data():
     with open(output / "non_mcq_questions.txt", "w") as f:
         f.write(str(len(non_mcq_questions)) + "\endinput")
 
+
+    automatically_generated = df[df["is_semiautomatically_generated"]] 
+
+    with open(output / "automatically_generated.txt", "w") as f:    
+        f.write(str(len(automatically_generated)) + "\endinput")
+
+    manually_generated = df[~df["is_semiautomatically_generated"]]  
+    with open(output / "manually_generated.txt", "w") as f: 
+        f.write(str(len(manually_generated)) + "\endinput")
+
     dai_data = glob(
         os.path.join(chembench_repo, "data", "safety", "pubchem_data", "DAI*.json")
     )

### 2024-10-06 src/tex/appendix.tex
feat: add corpus table

--- 2024-08-08 src/tex/appendix.tex
@@
 \end{figure}
+
+The corpus of the questions in Chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong, or which type of knowledge is required in order to solve them correctly.
+
+\begin{table*}[!t]
+    \centering
+        \caption{Summary of ChemBench question types.
+    }
+    \begin{tabular}{p{2.7cm}cp{10cm}}
+        \toprule
+        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
+        \midrule
+        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
+        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
+        \midrule
+        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
+        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+        \midrule
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
+        & & A. $E_1$ \\
+        & & B. $E_{cb}$ \\
+        & & C. $S_N1$ \\
+        & & D. $S_N2$ \\
+        \midrule
+        Toxicity/Safety & 24.2 & What details pertaining to a hazardous substance can be located on a Material Safety Data Sheet (MSDS)? \\
+        & & A. All hazards regarding the material \\
+        & & B. Physical and chemical properties \\
+        & & C. Precautions for safe handling and storage \\
+        \midrule
+        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+        & & A. X-ray fluorescence analysis \\
+        & & B. Differential pulse polarography \\
+        & & C. Flame-atomic absorption spectroscopy \\
+        & & D. Gas chromatography with flame ionization detector \\
+        & & E. Hydride generation atomic absorption spectroscopy \\
+        \midrule
+        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
+Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
+\\
+        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
+        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
+        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
+        \midrule
+        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
+        \midrule
+        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+\\
+        & & A. \\ce{C60} \\
+        & & B. \\ce{CH4} \\
+        & & C. \\ce{Fe(CO)5} \\
+        & & D. \\ce{H2+} \\
+        & & E. \\ce{NaCl} \\
+        \midrule
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
+        \toprule
+        \textbf{Reasoning Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        \midrule
+        Knowledge & 770 & Which of the following salts is an acidic salt? \\
+        & & A. \\
+        \midrule
+        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        & & A. By measuring atomic force microscopy \\
+        & & B. By measuring fourier transform infrared spectroscopy \\
+        & & C. By measuring photoluminescence spectrum \\
+        & & D. By measuring the absorption spectroscopy \\
+        \midrule
+        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
+        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
+        \midrule
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
+The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
+What is the average number of monomer units in each polymer chain of this sample?
+\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:chembench_corpus}
+\end{table*}
 
@@
     \script{plot_human_score_distribution.py}
-    \label{fig:human_score_distribution}
     \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
+    \label{fig:human_score_distribution}
 \end{figure}
@@
     \script{analyze_human_data.py}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
     \label{fig:human_timing}
-    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
 \end{figure}
@@
     \script{analyze_human_data.py}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
     \label{fig:experience_vs_correctness}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
 \end{figure}

### 2024-10-07 src/tex/ms.tex
add statistics about number of questions
automatically vs. manually
rebuild from scratch

@@
 In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
-%Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs compiled from diverse sources (\variable{output/manually_generated.txt}\xspace manually generated, and \variable{output/automatically_generated.txt}\xspace automatically generated).
 Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
@@
 
-\clearpage
+\paragraph{Judging chemical preference}
+
+One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman}
+If models could do so for chemical compounds, this would open opportunities for novel optimization approaches. 
+Such open-ended tasks, however, depend on an external observer defining what interestingness is.\autocite{hughes2024openendednessessentialartificialsuperhuman}
+Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting).
+While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench.
+This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.
 
@@
 In addition, we acknowledge support from Sandeep Kumar with an initial prototype of the web app.
-We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}) and thank Berend Smit and Pepe Márquez for feedback on an early version of the manuscript.
+We thank Bastian Rieck for developing the \LaTeX-credit package (\url{https://github.com/Pseudomanifold/latex-credits}) and thank Berend Smit for feedback on an early version of the manuscript.
 

### 2024-10-07 src/tex/appendix.tex
update human plots

--- 2024-10-06 src/tex/appendix.tex
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/human_timing.pdf}
-%     \script{analyze_human_data.py}
-%     \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
-%     \label{fig:human_timing}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/human_timing.pdf}
+    \script{analyze_human_data.py}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.
+However, we see that human scorers typically took longer to correctly answer questions with tool use.}
+    \label{fig:human_timing}
+\end{figure}
 
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-%\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/experience_vs_correctness.pdf}
-%     \script{analyze_human_data.py}
-%     \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
-%     \label{fig:experience_vs_correctness}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/experience_vs_correctness.pdf}
+    \script{analyze_human_data.py}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers wasp correlated with the correctness of their answers.}
+    \label{fig:experience_vs_correctness}
+\end{figure}
 

### 2024-10-14 src/tex/appendix.tex
work on subset scores

@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/performance_per_topic.pdf}
-%     \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic.pdf}
+    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
-%         We see that the performance of the models varies significantly between the different data sources.
+        We see that the performance of the models varies significantly between the different data sources.
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-%     }
-%     \label{fig:performance_per_topic}
-%     \script{analyze_performance_per_source.py}
-% \end{figure}
+    }
+    \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
+\end{figure}
 

### 2024-10-14 src/scripts/analyze_performance_per_source.py

@@ -23,7 +23,6 @@ subsets = [
     "is_dai",
     "is_pictograms",
     "is_name",
-    "is_smiles_name",
     "is_organic_reactivity",
     "is_electron_counts",
     "is_chemical_compatibility",
@@ -42,7 +41,6 @@ subset_clean_names = [
     "DAI (daily allowed intake)",
     "GHS pictograms",
     "name to SMILES",
-    "SMILES to name",
     "organic reactivity",
     "electron counts",
     "chemical compatibility",
@@ -53,41 +51,6 @@ subset_clean_names = [
     "polymer chemistry",
 ]
 
-relevant_models = [
-    "claude2",
-    "claude2_react",
-    "claude2_zero_t",
-    "claude3",
-    "galactica_120b",
-    "gemini_pro",
-    "gemini_pro_zero_t",
-    "gpt35turbo",
-    "gpt35turbo_react",
-    "gpt4",
-    "lama70b",
-    "mixtral",
-    "pplx7b_chat",
-    "pplx7b_online",
-    "random_baseline",
-]
-
-model_rename_dict = {
-    "gpt4": "GPT-4",
-    "claude2": "Claude 2",
-    "claude3": "Claude 3",
-    "llama70b": "LLaMA (70B)",
-    "gemini_pro": "Gemini Pro",
-    "galactica_120b": "Galactica (120B)",
-    "mixtral": "Mixtral",
-    "pplx7b_chat": "PPLX7B Chat",
-    "pplx7b_online": "PPLX7B Online",
-    "random_baseline": "Random Baseline",
-    "claude2_react": "Claude 2 ReAct",
-    "claude2_zero_t": "Claude 2 Zero T",
-    "gemini_pro_zero_t": "Gemini Pro Zero T",
-    "gpt35turbo": "GPT-3.5 Turbo",
-    "gpt35turbo_react": "GPT-3.5 Turbo ReAct",
-}
 
 
 rename_dict = dict(zip(subsets, subset_clean_names))
@@ -152,8 +115,10 @@ if __name__ == "__main__":
     # human aligned
     model_scores = obtain_subset_scores(model_scores["human_aligned"], outdir)
 
-    with open(os.path.join(data, "humans_as_models_scores.pkl"), "rb") as handle:
+    with open(os.path.join(data, "humans_as_models_scores_no_tools.pkl"), "rb") as handle:
         human_scores = pickle.load(handle)
+
+    
 
[…]

### 2024-10-14 src/scripts/get_model_performance_dicts.py

@@ -49,7 +49,7 @@ def load_reports(topic_frame: pd.DataFrame, human_aligned: bool = False) -> Dict
     chembench = obtain_chembench_repo()
     datafolder = os.path.join(chembench, "data")
     
-    with open(data / "name_dir_map.pkl", "rb") as handle:
+    with open(data / "name_to_dir_map.pkl", "rb") as handle:
         name_dir_map = pickle.load(handle)
     
     human_baseline_folder = os.path.join(chembench, "reports/humans") if human_aligned else None

### 2024-10-14 src/scripts/make_name_dir_map.py

@@ -41,8 +41,8 @@ def build_name_dir_map(report_base_dir=None):
             model_name = yaml_data["model_name"]
 
         name_dir_map[model_name] = os.path.abspath(report_subdir)
-        
-    with open(data / "name_dir_map.pkl", "wb") as f:
+
+    with open(data / "name_to_dir_map.pkl", "wb") as f:
         pickle.dump(name_dir_map, f)
 
     return name_dir_map

### 2024-10-16 src/tex/appendix.tex
reading ease

@@
 Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
+    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition. 
+    \item \emph{Cover a range of difficulty levels}.
+To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
+    \item \emph{Impossible to completely solve with current models}.
+A benchmark should contain questions that are impossible to solve with current models.
+If current models can solve all questions, the benchmark provides no useful signal.
 \end{itemize}
@@
 We also evaluated the model performance on the entire \chembench corpus. 
-\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
-Note that this ranking differs from the one on the \enquote{tiny} subset.
+\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models. 
+Note that this ranking differs from the one on the \enquote{tiny} subset
 
@@
     \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
@@
 
-% \begin{figure}
-%     \centering
-%     \hspace*{-1cm}
-%     \includegraphics{figures/reading_ease_vs_model_performance.pdf}
-%     \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
+\begin{figure}
+    \centering
+    \hspace*{-1cm}
+    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
+    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
 We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
 }
-%     \script{reading_ease_vs_model_performance.py}
-%     \label{fig:reading_ease_vs_model_performance}
-% \end{figure}
+    \script{reading_ease_vs_model_performance.py}
+    \label{fig:reading_ease_vs_model_performance}
+\end{figure}
 

### 2024-10-16 src/tex/ms.tex
implement some human statistics
reading ease

@@
 
-%To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
@@
 
-% \begin{figure}[!h]
-%     \centering
-%     \includegraphics{figures/human_subset_performance.pdf}
-%     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly.
+\begin{figure}[!h]
+    \centering
+    \includegraphics{figures/human_subset_performance.pdf}
+    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
-%     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
-%     The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
-%     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
-%     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
-%     However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
-%     In this case, we consider the answer as incorrect. 
-%     }
-%     \label{fig:human_vs_models_bar}
-%     \script{plot_overview_performance_plot.py}
-% \end{figure}
+    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
+    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
+    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
+    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
+    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
+    In this case, we consider the answer as incorrect. 
+    }
+    \label{fig:human_vs_models_bar}
+    \script{plot_overview_performance_plot.py}
+\end{figure}
 
@@
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
-%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly).
-A larger colored area indicates a better performance.
+%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
 %     This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
@@
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
-We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
+We then plot the correctness of the answers (which is calculated as the mean value of all answers being either correct (1) or incorrect (0)) against the confidence score.
 %         The stripplots show the individual data points, a strong density indicating a higher number of points.

### 2024-10-16 src/tex/appendix.tex
add radar plots back in

@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/overall_performance.pdf}
-%     \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/overall_performance.pdf}
+    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
 Scores computed on the entire \chembench corpus.}
-%     \label{fig:barplot_all_correct_all_questions}
-%     \script{plot_overview_performance_plot.py}
-% \end{figure}
+    \label{fig:barplot_all_correct_all_questions}
+    \script{plot_overview_performance_plot.py}
+\end{figure}
 
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
-%     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
 The performance is measured as the fraction of questions answered completely correctly by the models.
-%     A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
-%     }
-%     \label{fig:all_questions_models_completely_correct_radar_overall}
-%     \script{analyze_model_reports.py}
-% \end{figure}
-
-\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
-This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
-Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
-Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/parallel_coordinates_overall.pdf}
-%     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered completely correctly by the models.
+    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
     }
-%     \label{fig:parallel_coordinates_overall}
-%     \script{make_parallel_coordinates_plot.py}
-% \end{figure}
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/parallel_coordinates_tiny.pdf}
-%     \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-}
-%     \label{fig:parallel_coordinates_tiny}
-%     \script{make_parallel_coordinates_plot.py}
-% \end{figure}
+    \label{fig:all_questions_models_completely_correct_radar_overall}
+    \script{analyze_model_reports.py}
+\end{figure}
 
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/performance_per_topic_tiny.pdf}
-%     \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic_tiny.pdf}
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
 A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
-%         We see that the performance of the models varies significantly between the different data sources.
+        We see that the performance of the models varies significantly between the different data sources.
 For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
 However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-%     }
-%     \label{fig:performance_per_topic_tiny}
-%     \script{analyze_performance_per_source.py}
-% \end{figure}
-
-One might wonder if questions that are more difficult to parse lead to worse performance of the models.
-\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.
-
-\begin{figure}
-    \centering
-    \hspace*{-1cm}
-    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
-    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
-We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
     }
-    \script{reading_ease_vs_model_performance.py}
-    \label{fig:reading_ease_vs_model_performance}
+    \label{fig:performance_per_topic_tiny}
+    \script{analyze_performance_per_source.py}
 \end{figure}
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/human_score_distribution.pdf}
-%     \script{plot_human_score_distribution.py}
-%     \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
-%     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
-%     \label{fig:human_score_distribution}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/human_score_distribution.pdf}
+    \script{plot_human_score_distribution.py}
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
+    \label{fig:human_score_distribution}
+\end{figure}
 

### 2024-10-16 src/scripts/collect_human_scores.py

@@ -1,33 +1,42 @@
 from glob import glob
-import os
-from pathlib import Path
-import subprocess
-from utils import obtain_chembench_repo
 from paths import output, data
+from utils import obtain_chembench_repo
+import numpy as np
+import json
+import os
+import pickle
 
-if __name__ == "__main__":
+
+def obtain_human_scores():
     chembench_repo = obtain_chembench_repo()
-    human_files = os.listdir(os.path.join(chembench_repo, "reports", "humans"))
-    human_files = [
-        os.path.join(chembench_repo, "reports", "humans", p)
-        for p in human_files
-    ]
 
-    outpath = os.path.join(output, "human_scores")
-    if not os.path.exists(outpath):
-        os.mkdir(outpath)
+    human_reports_dir = os.path.join(chembench_repo, "reports", "humans")
 
-    datafolder = os.path.join(chembench_repo, "data")
-    human_baseline_folder = os.path.join(chembench_repo, "reports", "humans")
-    scriptpath = os.path.join(chembench_repo, "scripts", "collect_scores.py")
+    human_scores_with_tools_files = glob(
+        os.path.join(human_reports_dir, "*_h_tool*.json")
+    )
 
-    for file in human_files:
-        try:
-            p = Path(file).parts[-1]
-            outfile = os.path.join(outpath, p + ".json")
-            subprocess.run(
-                f"python {scriptpath} {file} {outfile} --datafolder={datafolder} --human_baseline_folder={human_baseline_folder}",
-                shell=True,
-            )
-        except Exception:
-            pass
+    human_scores_without_tools_files = glob(
+        os.path.join(human_reports_dir, "*_h_notool*.json")
+    )
+
+    human_scores_with_tools, human_scores_without_tools = [], []
+
+    for file in human_scores_with_tools_files:
+        with open(file, "r") as handle:
+            d = json.load(handle)
+            human_scores_with_tools.append(d["fraction_correct"])
+
+    for file in human_scores_without_tools_files:
+        with open(file, "r") as handle:
+            d = json.load(handle)
+            human_scores_without_tools.append(d["fraction_correct"])
+
+    human_scores_with_tools = np.array(human_scores_with_tools)
+    human_scores_without_tools = np.array(human_scores_without_tools)
+
+    all_human_scores = np.concatenate(
+        [human_scores_with_tools, human_scores_without_tools]
+    )
+
+    return human_scores_with_tools, human_scores_without_tools, all_human_scores

### 2024-10-16 src/tex/ms.tex

@@
 We then computed the percentage of questions the models or humans answered correctly for each topic.
-% \begin{figure}[!h]
-%     \centering
-%     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
-%     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+\begin{figure}[!h]
+    \centering
+    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
+    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
-%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-%     This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
-%     }
-%     \label{fig:all_questions_models_completely_correct_radar_human}
-%     \script{analyze_model_reports.py}
-% \end{figure}
+    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
+    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    }
+    \label{fig:all_questions_models_completely_correct_radar_human}
+    \script{analyze_model_reports.py}
+\end{figure}
 

### 2024-10-16 src/tex/ms.tex
feat: add some lines about preference data

@@
 
+\subparagraph{Chemical Preference data}
+
+These questions assess the ability to establish a \enquote{preference}, such as favoring a specific molecule.
+Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that lasts several years within a chemist's career.
+Our data corpus is adapted from the published dataset by \textcite{Choung2023}, which consists of more than 5000 question-answer pairs about chemical intuition.
+To build the dataset, they presented 35 medicinal chemists with two different molecules, asking them what molecule they would like to continue with when imaging an early virtual screening campaign setting.
+The question was designed so the scientists do not spend much time answering it, relying only on their feelings or \enquote{chemical preference}.
+
+To understand whether the capabilities of the leading models align with the preferences of professional chemists, we selected 1000 data points from the original dataset to create a meaningful evaluation set, where molecules are represented as SMILES.
+We only considered molecules for which we could obtain IUAPC names to ablate the effect of different molecular representations.
+
 %\input{sections/semi_programatically_sources_table.tex}

### 2024-10-17 src/tex/appendix.tex

@@
 
-% \begin{figure}
-%     \centering 
-%     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
-%     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
-%     The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
+\begin{figure}
+    \centering 
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
+    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
-%     \script{correlate_with_molecule_features.py}
-%     \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
-% \end{figure}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
+\end{figure}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-%     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
-%     For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
-%     \script{correlate_with_molecule_features.py}
-%     \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
-% \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
+    For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
+\end{figure}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
-%     \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
-%     \script{correlate_with_molecule_features.py}
-%     \label{fig:correlation_plot_is_electron_counts_num_atoms}
-% \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
+    \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
+    \script{correlate_with_molecule_features.py}
+    \label{fig:correlation_plot_is_electron_counts_num_atoms}
+\end{figure}
 

### 2024-10-17 src/scripts/collect_performance_table_scores.py
feat: add performance tables

@@ -1,0 +1,250 @@
+import pickle
+from paths import output, scripts, figures
+
+import json
+import os
+
+from loguru import logger
+
+from chembench.analysis import (
+    all_correct,
+    load_all_reports,
+)
+from chembench.types import PathType
+
+from collections import defaultdict
+
+from glob import glob
+import os
+from pathlib import Path
+from utils import obtain_chembench_repo
+from paths import output
+from loguru import logger
+
+import json
+import numpy as np
+import matplotlib.pyplot as plt
+from plotutils import range_frame
+from paths import output, scripts, figures
+from loguru import logger
+import pandas as pd
+
+model_file_name_to_label = {
+    # "claude2": "Claude 2",
+    # "claude3": "Claude 3",
+    "claude3.5": "Claude 3.5 Sonnet",
+    # "command-r+": "Command R+",
+    "galatica_120b": "Galactica 120B",
+    # "gemini-pro": "Gemini Pro",
+    # "gemma-1-1-7b-it": "Gemma 1.1 7B Instruct",
+    # "gemma-1-1-7b-it-T-one": "Gemma 1.1 7B Instruct T-One",
+    # "gemma-2-9b-it": "Gemma 2 9B Instruct",
+    # "gemma-2-9b-it-T-one": "Gemma 2 9B Instruct T-One",
+    "gpt-3.5-turbo": "GPT-3.5-Turbo",
+    # "gpt-4": "GPT-4",
+    "gpt-4o": "GPT-4o",
+    "o1": "OpenAI o1",
+    # "llama2-70b-chat": "Llama 2 70B Chat",
+    # "llama3-70b-instruct": "Llama 3 70B Instruct",
+    # "llama3-70b-instruct-T-one": "Llama 3 70B Instruct T-One",
+    # "llama3-8b-instruct": "Llama 3 8B Instruct",
+    # "llama3-8b-instruct-T-one": "Llama 3 8B Instruct T-One",
+    "llama3.1-405b-instruct": "Llama 3.1 405B Instruct",
+    "llama3.1-70b-instruct": "Llama 3.1 70B Instruct",
+    # "llama3.1-70b-instruct-T-one": "Llama 3.1 70B Instruct T-One",
+    "llama3.1-8b-instruct": "Llama 3.1 8B Instruct",
+    # "llama3.1-8b-instruct-T-one": "Llama 3.1 8B Instruct T-One",
+    # "mixtral-8x7b-instruct": "Mixtral 8x7B Instruct",
[…]

### 2024-10-17 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 
+%\input{sections/semi_programatically_sources_table.tex}
+
 \subparagraph{Chemical Preference data}
@@
 
-To understand whether the capabilities of the leading models align with the preferences of professional chemists, we selected 1000 data points from the original dataset to create a meaningful evaluation set, where molecules are represented as SMILES.
-We only considered molecules for which we could obtain IUAPC names to ablate the effect of different molecular representations.
-
-%\input{sections/semi_programatically_sources_table.tex}
+To understand whether the capabilities of the leading models align with the preferences of professional chemists, we randomly selected 1000 data points from the original dataset to create a meaningful evaluation set, where molecules are represented as SMILES.
+To ablate the effect of different molecular representations, we only considered questions for which we could obtain IUPAC names for both molecules present.
 

### 2024-10-17 src/tex/appendix.tex
add distribution of topics image

@@
     \centering 
-    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
+    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
@@
     \centering
-    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
+    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
@@
 \subsection{Influence of model scale}
-% To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-% Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
+\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
+We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.
 
-% Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
-REWRITE.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/model_size_plot.pdf}
+    \caption{\textbf{Performance of models as a function of model size.} The plot shows the performance of the models as a function of the parameter count.
+The performance is measured as the fraction of questions answered correctly by the models.
+We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.}
+    \script{performance_vs_model_size.py}
+    \label{fig:model_size_plot}
+\end{figure}
 
@@
 The application is hosted on the Vercel web hosting platform.
+
+In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
+\LaTeX\xspace equations and chemical equations were rendered using MathJax\autocite{mathjax}.
 

### 2024-10-17 src/tex/ms.tex
humanset pie chart

--- 2024-10-16 src/tex/ms.tex
@@
 
-% \begin{figure}[!htb]
-%     \centering
-%     \includegraphics{figures/question_count_barplot.pdf}
-%     \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) and a classifier operating on word embeddings of the questions.
-%     The figure shows that not all aspects of chemistry are equally represented in our corpus.
-The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}.
-This figure represents the combined count of multiple-choice questions (\gls{mcq}) and open-ended questions.}
-%     \label{fig:topic_barplot}
-%     \script{plot_statistics.py}
-% \end{figure}
+\begin{figure}[!htb]
+    \centering
+    \includegraphics{figures/cb_corpus_v1.pdf}
+    \caption{\textbf{Distribution of topics and required skills.} TThis circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
+The topics were manually classified, showing a varied representation across different aspects of chemistry.
+Each topic is associated with a combination of three key skills: Calculation, Reasoning, and Knowledge, as indicated by the colored bars.
+\chembench samples diverse topics and diverse skills, setting a high bar for \glspl{llm} to demonstrate human-competitive performance across a wide range of chemistry tasks.}
+    \label{fig:corpus}
+\end{figure}
 
@@
 We also used this subset to seed the web application for the human baseline study.
+
+\begin{figure}
+    \centering
+    \includegraphics[]{figures/cb_humanset_v1.pdf}
+    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
+\end{figure}
 

### 2024-10-17 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-17 src/tex/appendix.tex
@@
 Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition. 
-    \item \emph{Cover a range of difficulty levels}.
-To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
-    \item \emph{Impossible to completely solve with current models}.
-A benchmark should contain questions that are impossible to solve with current models.
-If current models can solve all questions, the benchmark provides no useful signal.
 \end{itemize}
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
 It is calculated based on the average number of syllables per word and words per sentence.
@@
 }
-%     \label{fig:flesch_kincaid_reading_ease}
-%     \script{wordcloud.py}
-% \end{figure}
+    \label{fig:flesch_kincaid_reading_ease}
+    \script{wordcloud.py}
+\end{figure}
 
@@
 A substantial fraction, in contrast to other benchmarks, is open-ended. 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-%     \label{fig:question_count_barplot_mcq_vs_general}
-%     \script{plot_statistics.py}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+    \label{fig:question_count_barplot_mcq_vs_general}
+    \script{plot_statistics.py}
+\end{figure}
+
+The corpus of the questions in Chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong, or which type of knowledge is required in order to solve them correctly.
+
+\begin{table*}[!t]
+    \centering
+        \caption{Summary of \chembench question types.}
+    }
+    \begin{tabular}{p{2.7cm}cp{10cm}}
+        \toprule
+        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
+        \midrule
+        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
+        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
+        \midrule
+        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
+        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+        \midrule
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
+        & & A. $E_1$ \\
+        & & B. $E_{cb}$ \\
+        & & C. $S_N1$ \\
+        & & D. $S_N2$ \\
+        \midrule
+        Toxicity/Safety & 24.2 & What details pertaining to a hazardous substance can be located on a Material Safety Data Sheet (MSDS)? \\
+        & & A. All hazards regarding the material \\
+        & & B. Physical and chemical properties \\
+        & & C. Precautions for safe handling and storage \\
+        \midrule
+        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+        & & A. X-ray fluorescence analysis \\
+        & & B. Differential pulse polarography \\
+        & & C. Flame-atomic absorption spectroscopy \\
+        & & D. Gas chromatography with flame ionization detector \\
+        & & E. Hydride generation atomic absorption spectroscopy \\
+        \midrule
+        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
+Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
+\\
+        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
+        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
+        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
+        \midrule
+        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
+        \midrule
+        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+\\
+        & & A. \\ce{C60} \\
+        & & B. \\ce{CH4} \\
+        & & C. \\ce{Fe(CO)5} \\
+        & & D. \\ce{H2+} \\
+        & & E. \\ce{NaCl} \\
+        \midrule
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
+        \toprule
+        \textbf{Reasoning Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        \midrule
+        Knowledge & 770 & Which of the following salts is an acidic salt? \\
+        & & A. \\
+        \midrule
+        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        & & A. By measuring atomic force microscopy \\
+        & & B. By measuring fourier transform infrared spectroscopy \\
+        & & C. By measuring photoluminescence spectrum \\
+        & & D. By measuring the absorption spectroscopy \\
+        \midrule
+        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
+        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
+        \midrule
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
+The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
+What is the average number of monomer units in each polymer chain of this sample?
+\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:chembench_corpus}
+\end{table*}
 
@@
 We also evaluated the model performance on the entire \chembench corpus. 
-\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models. 
-Note that this ranking differs from the one on the \enquote{tiny} subset
+\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
+Note that this ranking differs from the one on the \enquote{tiny} subset.
 
@@
 
+\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
+This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
+Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
+Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_overall}
+    \script{make_parallel_coordinates_plot.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_tiny.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_tiny}
+    \script{make_parallel_coordinates_plot.py}
+\end{figure}
+
 To further investigate the performance of the models, we also compared the performance on different data sources.
@@
     \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
@@
 
-% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
-% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+One might wonder if questions that are more difficult to parse lead to worse performance of the models.
+\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/model_overall_cal.pdf}
-%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-%     \script{requires_calculation_plot.py}
-%     \label{fig:requires_cal}
-% \end{figure}
+\begin{figure}
+    \centering
+    \hspace*{-1cm}
+    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
+    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
+We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
+}
+    \script{reading_ease_vs_model_performance.py}
+    \label{fig:reading_ease_vs_model_performance}
+\end{figure}
+
+In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
+We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/model_overall_cal.pdf}
+    \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+    \script{requires_calculation_plot.py}
+    \label{fig:requires_cal}
+\end{figure}
 
@@
 
-% \begin{table}
-%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+\begin{table}
+    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
 }
-%     \begin{tabularx}{\textwidth}{XXXX XXXX}
-%         \toprule
-%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
-%         \midrule
-%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
-%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
-%         \midrule 
-%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
-%         \midrule 
-%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
-%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
-%         \bottomrule
-%     \end{tabularx}
-%     \label{tab:correlation_coefficients}
-% \end{table}
+    \begin{tabularx}{\textwidth}{XXXX XXXX}
+        \toprule
+        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+        \midrule
+        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+        \midrule 
+        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+        \midrule 
+        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+        \bottomrule
+    \end{tabularx}
+    \label{tab:correlation_coefficients}
+\end{table}
 
 \subsection{Influence of model scale}
-% To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-% Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
+To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
+Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
 
-% Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
-REWRITE.
+Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
 
@@
 This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a Postgresql.
-The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
+We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
 The application is hosted on the Vercel web hosting platform.
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
     \script{analyze_human_data.py}
-    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.
-However, we see that human scorers typically took longer to correctly answer questions with tool use.}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
     \label{fig:human_timing}
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
     \script{analyze_human_data.py}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers wasp correlated with the correctness of their answers.}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
     \label{fig:experience_vs_correctness}
@@
 
-% \begin{figure}[htb] 
-%     \centering
-%     \includegraphics{figures/confidence_score_distributions.pdf}
-%     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
+\begin{figure}[htb] 
+    \centering
+    \includegraphics{figures/confidence_score_distributions.pdf}
+    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
 The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
 The bar plots show how many questions were answered with each confidence score.}
-%     \label{fig:confidence_score_distributions}
-%     \script{plot_confidence_score_distributions.py}
-% \end{figure}
+    \label{fig:confidence_score_distributions}
+    \script{plot_confidence_score_distributions.py}
+\end{figure}
 

### 2024-10-17 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 
-The corpus of the questions in Chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong, or which type of knowledge is required in order to solve them correctly.
+The corpus of the questions in  \chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong to or which type of knowledge is required to solve them correctly.
 

### 2024-10-17 src/tex/appendix.tex
fix: applied suggestions from code review

@@
 
-The corpus of the questions in  \chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong to or which type of knowledge is required to solve them correctly.
+The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
 
@@
     \centering
-        \caption{Summary of \chembench question types.}
-    }
+        \caption{Examples for each of the topics considered in the evaluation of the \chembench corpus.}
     \begin{tabular}{p{2.7cm}cp{10cm}}
@@
         \midrule
-        Toxicity/Safety & 24.2 & What details pertaining to a hazardous substance can be located on a Material Safety Data Sheet (MSDS)? \\
-        & & A. All hazards regarding the material \\
-        & & B. Physical and chemical properties \\
-        & & C. Precautions for safe handling and storage \\
+        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
+What follows from this?
+\\
+        & & A. Pindolol has a greater therapeutic range than propranolol \\
+        & & B. Pindolol has a longer half-life than propranolol \\
+        & & C. Pindolol has intrinsic activity \\
+        & & D. Pindolol is more lipophilic than propranolol \\
+        & & E. Pindolol is more potent than propranolol \\
         \midrule
@@
         Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
+    \end{tabular}
+    \label{tab:chembench_corpus_topic}
+\end{table*}
+
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different mental approaches needed to solve the questions.
+
+\begin{table*}[!t]
+    \centering
+        \caption{Examples for each of the mental approaches considered in the \chembench corpus.}
+    \begin{tabular}{p{2.7cm}cp{10cm}}
         \toprule
-        \textbf{Reasoning Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        \textbf{Cognitive Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
         \midrule
         Knowledge & 770 & Which of the following salts is an acidic salt? \\
-        & & A. \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
         \midrule
@@
     \end{tabular}
-    \label{tab:chembench_corpus}
+    \label{tab:chembench_corpus_cognitive}
 \end{table*}

### 2024-10-18 src/tex/appendix.tex
add process figure

--- 2024-10-17 src/tex/appendix.tex
@@
 A substantial fraction, in contrast to other benchmarks, is open-ended. 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-    \label{fig:question_count_barplot_mcq_vs_general}
-    \script{plot_statistics.py}
-\end{figure}
-
-The corpus of the questions in  \chembench, as shown in \Cref{tab:chembench_corpus}, can be divided according to which topic they belong to or which type of knowledge is required to solve them correctly.
-
-\begin{table*}[!t]
-    \centering
-        \caption{Summary of \chembench question types.}
-    }
-    \begin{tabular}{p{2.7cm}cp{10cm}}
-        \toprule
-        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
-        \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
-        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
-        \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-        \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
-        \midrule
-        Toxicity/Safety & 24.2 & What details pertaining to a hazardous substance can be located on a Material Safety Data Sheet (MSDS)? \\
-        & & A. All hazards regarding the material \\
-        & & B. Physical and chemical properties \\
-        & & C. Precautions for safe handling and storage \\
-        \midrule
-        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-        & & A. X-ray fluorescence analysis \\
-        & & B. Differential pulse polarography \\
-        & & C. Flame-atomic absorption spectroscopy \\
-        & & D. Gas chromatography with flame ionization detector \\
-        & & E. Hydride generation atomic absorption spectroscopy \\
-        \midrule
-        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
-\\
-        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
-        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
-        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
-        \midrule
-        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \\ce{NH4Cl} \\
-        & & B. \\ce{Na2CO3} \\
-        & & C. \\ce{NaH2PO4} \\
-        & & D. \\ce{Zn(OH)Cl} \\
-        \midrule
-        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
-\\
-        & & A. \\ce{C60} \\
-        & & B. \\ce{CH4} \\
-        & & C. \\ce{Fe(CO)5} \\
-        & & D. \\ce{H2+} \\
-        & & E. \\ce{NaCl} \\
-        \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
-        \toprule
-        \textbf{Reasoning Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
-        \midrule
-        Knowledge & 770 & Which of the following salts is an acidic salt? \\
-        & & A. \\
-        \midrule
-        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By measuring atomic force microscopy \\
-        & & B. By measuring fourier transform infrared spectroscopy \\
-        & & C. By measuring photoluminescence spectrum \\
-        & & D. By measuring the absorption spectroscopy \\
-        \midrule
-        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
-        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
-        \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
-The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
-What is the average number of monomer units in each polymer chain of this sample?
-\\
-        \bottomrule
-    \end{tabular}
-    \label{tab:chembench_corpus}
-\end{table*}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+%     \label{fig:question_count_barplot_mcq_vs_general}
+%     \script{plot_statistics.py}
+% \end{figure}
 
@@
 
-\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
-This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
-Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
-Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/parallel_coordinates_overall.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-}
-    \label{fig:parallel_coordinates_overall}
-    \script{make_parallel_coordinates_plot.py}
-\end{figure}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/parallel_coordinates_tiny.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-}
-    \label{fig:parallel_coordinates_tiny}
-    \script{make_parallel_coordinates_plot.py}
-\end{figure}
-
 To further investigate the performance of the models, we also compared the performance on different data sources.
@@
 
-One might wonder if questions that are more difficult to parse lead to worse performance of the models.
-\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.
+% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
+% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/model_overall_cal.pdf}
+%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+%     \script{requires_calculation_plot.py}
+%     \label{fig:requires_cal}
+% \end{figure}
+
+\clearpage
+
+\subsection{Implementation}
 
@@
     \centering 
-    \hspace*{-1cm}
-    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
-    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
-We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
-}
-    \script{reading_ease_vs_model_performance.py}
-    \label{fig:reading_ease_vs_model_performance}
+    \includegraphics{figures/process.pdf}
+    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
+\texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
+    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types.
+For each task type, appropriate prompts are generated and passed to the \gls{llm}.
+The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
+    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
+    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback.
+The parsed or extracted answers then undergo metric computation and refusal detection.}
+    \label{fig:process}
 \end{figure}
-
-In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
-We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
-
-\begin{figure}
-    \centering
-    \includegraphics{figures/model_overall_cal.pdf}
-    \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-    \script{requires_calculation_plot.py}
-    \label{fig:requires_cal}
-\end{figure}
-
-\clearpage
 
@@
 
-\begin{table}
-    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+% \begin{table}
+%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
 }
-    \begin{tabularx}{\textwidth}{XXXX XXXX}
-        \toprule
-        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
-        \midrule
-        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
-                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
-        \midrule 
-        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
-        \midrule 
-        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
-                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
-        \bottomrule
-    \end{tabularx}
-    \label{tab:correlation_coefficients}
-\end{table}
+%     \begin{tabularx}{\textwidth}{XXXX XXXX}
+%         \toprule
+%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+%         \midrule
+%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+%         \midrule 
+%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+%         \midrule 
+%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{tab:correlation_coefficients}
+% \end{table}
 
 \subsection{Influence of model scale}
-To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
-Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
+\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
+We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.
 
-Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/model_size_plot.pdf}
+    \caption{\textbf{Performance of models as a function of model size.} The plot shows the performance of the models as a function of the parameter count.
+The performance is measured as the fraction of questions answered correctly by the models.
+We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.}
+    \script{performance_vs_model_size.py}
+    \label{fig:model_size_plot}
+\end{figure}
 
@@
 This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
-The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
+We utilize a Postgresql.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
 The application is hosted on the Vercel web hosting platform.
+
+In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
+\LaTeX\xspace equations and chemical equations were rendered using MathJax\autocite{mathjax}.
 
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
     \script{analyze_human_data.py}
-    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.}
+    \caption{\textbf{Time taken by human scorers to answer questions vs.\ correctness of their answers.} From the plot, it is clear that there is no clear dependence of the correctness of the answers on the time taken by the human scorers to answer the questions.
+However, we see that human scorers typically took longer to correctly answer questions with tool use.}
     \label{fig:human_timing}
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
     \script{analyze_human_data.py}
-    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers was not significantly correlated with the correctness of their answers.}
+    \caption{\textbf{Experience of human  scorers vs.\ correctness of their answers.} The experience (in the number of years since the first university-level chemistry course) of the human scorers wasp correlated with the correctness of their answers.}
     \label{fig:experience_vs_correctness}

### 2024-10-18 src/tex/ms.tex
updated arxiv references

@@
 Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination.\autocite{kung2023performance}
-They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow}
+They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow, darvish2024organa}
 While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
@@
 Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 

### 2024-10-18 src/tex/ms.tex
rewrite intro and corpus

@@
 \Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
-Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination.\autocite{kung2023performance}
+Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass the United States Medical Licensing Examination\autocite{kung2023performance} or other professional licensing exams.
 They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow, darvish2024organa}
-While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
+While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on\autocite{bender2021dangers} and that show inherent limitations due to the way they are trained.\autocite{mccoy2023embersautoregressionunderstandinglarge}
 Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier, ai4science2023impact}
-This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}
 
@@
 This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
-Such a usage mode is, in particular, interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
 
@@
 However, even for experts, chemical understanding and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.
+Unfortunately, apart from anecdotal reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.
 
@@
 While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
-This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
+This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
 On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.
@@
 Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs compiled from diverse sources (\variable{output/manually_generated.txt}\xspace manually generated, and \variable{output/automatically_generated.txt}\xspace automatically generated).
-Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
+Our corpus measures reasoning, knowledge and intuition across a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
 It can be used to evaluate any system that can return text (i.e., including tool-augmented systems).
@@
 To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
 We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
-Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.
-%and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.
 
@@
     The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
+    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
     We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework.
@@
 
-To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
+To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel manually crafted questions, over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
 For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks. 
@@
 Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
-Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty.
+Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty by the annotator.
 
@@
     \centering
-    \includegraphics{figures/cb_corpus_v1.pdf}
-    \caption{\textbf{Distribution of topics and required skills.} TThis circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
+    \includegraphics[width=\textwidth]{figures/cb_corpus_v1.pdf}
+    \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
 The topics were manually classified, showing a varied representation across different aspects of chemistry.
@@
 While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
-%For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
+For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
+In addition, \chembench samples skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
 
@@
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
-%To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset to seed the web application for the human baseline study.
+To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
+While it is impossible to fully represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
+We also used this subset for the human baseline study.
 

### 2024-10-18 src/scripts/analyze_human_data.py

@@ -108,6 +108,9 @@ def make_human_performance_plots():
         for q in human_answered_questions:
             f.write(f"{q}" + "\n")
 
+    with open(data / "num_human_answered_questions.txt", "w") as f:
+        f.write(f"{str(len(human_answered_questions))}" + "\endinput")
+
     with open(data / "human_tool_answered_questions.txt", "w") as f:
         human_tool_answered_questions = long_df[long_df["tool_allowed"] == True][
             ("name", 0)

### 2024-10-18 src/tex/ms.tex
add macros for model names

@@
 Current benchmarking suites, however, do not account for such special treatment of scientific information.
-To address this, \chembench encodes the semantic meaning of various parts of the question or answer.  
+To address this, \chembench encodes the semantic meaning of various parts (e.g., chemicals, units, equations) of the question or answer.  
 For instance, molecules represented in \gls{smiles} are enclosed in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
@@
 The text completions, however, are the system's final outputs, which would also be used in a real-world application. 
-Hence, we use them for our evaluations.
+Hence, we use them for our evaluations.\autocite{xiong2023llms}
 
@@
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
-    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
+    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
     The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
-    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
-    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
-    In this case, we consider the answer as incorrect. 
+    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer.
+We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
     }

### 2024-10-18 src/tex/ms.tex
update results

@@
 
+% ToDO: Link to full table of results in the SI
 \paragraph{System performance} 
@@
 Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
-Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
+Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
 Many other models also outperform the average human performance. 
-Interestingly, the Galactica model, which was trained specifically for scientific applications, underperformed compared to many advanced commercial and some open-source models and barely exceeds the random baseline.
+Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
 

### 2024-10-18 src/tex/ms.tex
add more details

@@
     \centering
-    \includegraphics{figures/human_subset_performance.pdf}
+    \includegraphics[width=\textwidth]{figures/human_subset_performance.pdf}
     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly.
@@
     \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
+    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
@@
 \paragraph{Participants}
-% Users were open to reporting about their experience in chemistry. 
-% Overall, \variable{output/num_users_with_education_info.txt} did so. 
-% Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
-% \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+Users were open to reporting about their experience in chemistry. 
+Overall, \variable{output/num_users_with_education_info.txt} did so. 
+Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 

### 2024-10-18 src/tex/ms.tex
update text

@@
 
+% Importance of scale 
+% Still lacks in knowledge-intensive questions
+
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
@@
 \subsection{Model evaluation workflow}
-EXPAND THIS section.
+An overview of the pipeline is shown in \Cref{fig:evaluation_workflow}.
 
@@
 As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
-If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.
+If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., \ClaudeThreeFiveSonnet, to parse the completion.
 
 \paragraph{Models}
-Rewrite - as table. The table can then also contain if it has been completion or instruction-prompted.
+For all models we performed inference using greedy decoding (i.e., temperature 0). We used the \gls{api} endpoints provided by the model developers and the ones provided by Groq.
 
@@
 \paragraph{Comparison with models}
-%To compare the performance of humans (who might have answered only some questions) with the performance of models (which answered all questions), we focussed on questions that at least four humans answered and limited the pool of human scorers to those who answered at least 100 questions (i.e., \variable{output/num_humans_with_more_than_100_scores.txt} humans).
-%The latter threshold was chosen to limit it to humans who seriously attempted to answer a part of the questions systematically. 
-%This analysis might lead to potential biases, most likely in favor of humans, as they were allowed to skip questions.
-\variable{output/num_humans_with_204_scores.txt} humans answered more than 200 questions.
 For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans.
 
-\subsection{Classification of questions into topics}\label{sec:meth-topic} 
+\subsection{Data annotation}\label{sec:meth-topic} 
 % When curating our dataset, we systematically recorded keywords and sources.

### 2024-10-18 src/tex/sections/human_subset_selection.tex
feat: describe the human subset selection procedure

\paragraph{Human subset selection}

For selecting a human subset, several design choices have been made. Firstly, from the full dataset, we keep all the questions that are labeled as advanced. In this way, we can obtain a deeper insight about the capabilities of LLMs on advanced task
when compared to actual chemists. Secondly, we sample a maximum of 3 questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry). Thirdly, we do not include any intuition questions
in this subset. In total, 248 questions have been sampled for the human study. Then, this subset is divided into 2 subsets based on the aforementioned combinations. One of the question subsets allows tool use, and the other does not.

### 2024-10-18 src/tex/appendix.tex
updated arxiv references
rewrite intro and corpus
fix: change into required skills

--- 2024-10-17 src/tex/appendix.tex
@@
 
-In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different mental approaches needed to solve the questions.
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 
@@
     \centering
-        \caption{Examples for each of the mental approaches considered in the \chembench corpus.}
+        \caption{Examples for each of the required skills considered in the \chembench corpus.}
     \begin{tabular}{p{2.7cm}cp{10cm}}
         \toprule
-        \textbf{Cognitive Type} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        \textbf{Required skills} & \textbf{Number of Questions} & \textbf{Example Question} \\
         \midrule

### 2024-10-18 src/scripts/collect_performance_table_scores.py
include the human result in table

@@ -1,258 +1,0 @@
-import pickle
-from paths import output, scripts, figures
-
-import json
-import os
-
-from loguru import logger
-
-from chembench.analysis import (
-    all_correct,
-    load_all_reports,
-)
-from chembench.types import PathType
-
-from collections import defaultdict
-
-from glob import glob
-import os
-from pathlib import Path
-from utils import obtain_chembench_repo
-from paths import output
-from loguru import logger
-
-import json
-import numpy as np
-import matplotlib.pyplot as plt
-from plotutils import range_frame
-from paths import output, scripts, figures
-from loguru import logger
-import pandas as pd
-
-model_file_name_to_label = {
-    # "claude2": "Claude 2",
-    # "claude3": "Claude 3",
-    "claude3.5": "Claude 3.5 Sonnet",
-    # "command-r+": "Command R+",
-    "galatica_120b": "Galactica 120B",
-    # "gemini-pro": "Gemini Pro",
-    # "gemma-1-1-7b-it": "Gemma 1.1 7B Instruct",
-    # "gemma-1-1-7b-it-T-one": "Gemma 1.1 7B Instruct T-One",
-    # "gemma-2-9b-it": "Gemma 2 9B Instruct",
-    # "gemma-2-9b-it-T-one": "Gemma 2 9B Instruct T-One",
-    "gpt-3.5-turbo": "GPT-3.5-Turbo",
-    # "gpt-4": "GPT-4",
-    "gpt-4o": "GPT-4o",
-    "o1": "OpenAI o1",
-    # "llama2-70b-chat": "Llama 2 70B Chat",
-    # "llama3-70b-instruct": "Llama 3 70B Instruct",
-    # "llama3-70b-instruct-T-one": "Llama 3 70B Instruct T-One",
-    # "llama3-8b-instruct": "Llama 3 8B Instruct",
-    # "llama3-8b-instruct-T-one": "Llama 3 8B Instruct T-One",
-    "llama3.1-405b-instruct": "Llama 3.1 405B Instruct",
-    "llama3.1-70b-instruct": "Llama 3.1 70B Instruct",
-    # "llama3.1-70b-instruct-T-one": "Llama 3.1 70B Instruct T-One",
-    "llama3.1-8b-instruct": "Llama 3.1 8B Instruct",
-    # "llama3.1-8b-instruct-T-one": "Llama 3.1 8B Instruct T-One",
-    # "mixtral-8x7b-instruct": "Mixtral 8x7B Instruct",
[…]

### 2024-10-18 src/tex/appendix.tex
add variables

--- 2024-10-18 src/tex/appendix.tex
@@
 
+% \begin{figure}
+%     \centering
+%     \includegraphics[]{figures/cb_humanset_v1.pdf}
+%     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
+% \end{figure}
+
 \subsection{Model performance}
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \variable{output/performance_table.tex}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \variable{output/performance_table_human_subset.tex}
+\end{table}
 
@@
     \centering 
-    \includegraphics{figures/process.pdf}
+    \includegraphics[height=\textheight]{figures/process.pdf}
     \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
@@
     \centering
-    \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
+    \includegraphics[width=\textwidth]{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}

### 2024-10-18 src/tex/appendix.tex

@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[]{figures/cb_humanset_v1.pdf}
-%     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/cb_humanset_v2.pdf}
+    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
 The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
-% \end{figure}
+\end{figure}
 
@@
     \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
     \variable{output/performance_table.tex}
+    }
 \end{table}
@@
     \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
     \variable{output/performance_table_human_subset.tex}
+    }
 \end{table}
@@
     \centering
-    \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.

### 2024-10-18 src/tex/appendix.tex

@@
 % \end{figure}
+
+The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
+
+\begin{table*}[!t]
+    \centering
+        \caption{Examples for each of the topics considered in the evaluation of the \chembench corpus.}
+    \begin{tabular}{p{2.7cm}cp{10cm}}
+        \toprule
+        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
+        \midrule
+        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
+        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
+        \midrule
+        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
+        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+        \midrule
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
+        & & A. $E_1$ \\
+        & & B. $E_{cb}$ \\
+        & & C. $S_N1$ \\
+        & & D. $S_N2$ \\
+        \midrule
+        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
+What follows from this?
+\\
+        & & A. Pindolol has a greater therapeutic range than propranolol \\
+        & & B. Pindolol has a longer half-life than propranolol \\
+        & & C. Pindolol has intrinsic activity \\
+        & & D. Pindolol is more lipophilic than propranolol \\
+        & & E. Pindolol is more potent than propranolol \\
+        \midrule
+        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+        & & A. X-ray fluorescence analysis \\
+        & & B. Differential pulse polarography \\
+        & & C. Flame-atomic absorption spectroscopy \\
+        & & D. Gas chromatography with flame ionization detector \\
+        & & E. Hydride generation atomic absorption spectroscopy \\
+        \midrule
+        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
+Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
+\\
+        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
+        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
+        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
+        \midrule
+        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
+        \midrule
+        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+\\
+        & & A. \\ce{C60} \\
+        & & B. \\ce{CH4} \\
+        & & C. \\ce{Fe(CO)5} \\
+        & & D. \\ce{H2+} \\
+        & & E. \\ce{NaCl} \\
+        \midrule
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
+    \end{tabular}
+    \label{tab:chembench_corpus_topic}
+\end{table*}
+
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+
+\begin{table*}[!t]
+    \centering
+        \caption{Examples for each of the required skills considered in the \chembench corpus.}
+    \begin{tabular}{p{2.7cm}cp{10cm}}
+        \toprule
+        \textbf{Required skills} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        \midrule
+        Knowledge & 770 & Which of the following salts is an acidic salt? \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
+        \midrule
+        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        & & A. By measuring atomic force microscopy \\
+        & & B. By measuring fourier transform infrared spectroscopy \\
+        & & C. By measuring photoluminescence spectrum \\
+        & & D. By measuring the absorption spectroscopy \\
+        \midrule
+        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
+        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
+        \midrule
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
+The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
+What is the average number of monomer units in each polymer chain of this sample?
+\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:chembench_corpus_cognitive}
+\end{table*}
 

### 2024-10-18 src/tex/appendix.tex
escape special things

@@
 \\
-        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
-        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
+        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
+        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
         \midrule

### 2024-10-18 src/tex/appendix.tex
escape more

@@
         \toprule
-        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
+        Question Type & \% & Example Questions  \\ 
         \midrule
@@
         \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END\_RXNSMILES]?
+\\
         & & A. $E_1$ \\
@@
         Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
+Why would one choose \ce{HF} over \ce{HCl} for this purpose?
 \\
-        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
-        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
-        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
+        & & A. \ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
+        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
         \midrule
         General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \\ce{NH4Cl} \\
-        & & B. \\ce{Na2CO3} \\
-        & & C. \\ce{NaH2PO4} \\
-        & & D. \\ce{Zn(OH)Cl} \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
         \midrule
@@
 \\
-        & & A. \\ce{C60} \\
-        & & B. \\ce{CH4} \\
-        & & C. \\ce{Fe(CO)5} \\
-        & & D. \\ce{H2+} \\
-        & & E. \\ce{NaCl} \\
+        & & A. \ce{C60} \\
+        & & B. \ce{CH4} \\
+        & & C. \ce{Fe(CO)5} \\
+        & & D. \ce{H2+} \\
+        & & E. \ce{NaCl} \\
         \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
     \end{tabular}
@@
         \toprule
-        \textbf{Required skills} & \textbf{Number of Questions} & \textbf{Example Question} \\
+        Required Skills & Number of Questions & Example Question \\
         \midrule
         Knowledge & 770 & Which of the following salts is an acidic salt? \\
-        & & A. \\ce{NH4Cl} \\
-        & & B. \\ce{Na2CO3} \\
-        & & C. \\ce{NaH2PO4} \\
-        & & D. \\ce{Zn(OH)Cl} \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
         \midrule
@@
 \\
-        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
-        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
+        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
         \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
-The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \pu{595 g mol^{-1}} using end-group analysis, where \pu{0.1619 g} of PLA was dissolved in \pu{25 cm^3} of benzyl alcohol and titrated with \pu{0.0400 mol dm^{-3}} \ce{KOH} solution.
+The volume of \ce{KOH} solution required to reach the endpoint was \pu{6.81 cm^3}.
 What is the average number of monomer units in each polymer chain of this sample?

### 2024-10-18 src/tex/appendix.tex
escape

@@
         \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \pu{595 g mol^{-1}} using end-group analysis, where \pu{0.1619 g} of PLA was dissolved in \pu{25 cm^3} of benzyl alcohol and titrated with \pu{0.0400 mol dm^{-3}} \ce{KOH} solution.
-The volume of \ce{KOH} solution required to reach the endpoint was \pu{6.81 cm^3}.
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
 What is the average number of monomer units in each polymer chain of this sample?

### 2024-10-18 src/tex/appendix.tex
table size

@@
         \caption{Examples for each of the topics considered in the evaluation of the \chembench corpus.}
-    \begin{tabular}{p{2.7cm}cp{10cm}}
+    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}
         \toprule
@@
         Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-    \end{tabular}
+    \end{tabularx}
     \label{tab:chembench_corpus_topic}
@@
     \centering
-        \caption{Examples for each of the required skills considered in the \chembench corpus.}
-    \begin{tabular}{p{2.7cm}cp{10cm}}
+        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.}}
+    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}
         \toprule
@@
         \bottomrule
-    \end{tabular}
+    \end{tabularx}
     \label{tab:chembench_corpus_cognitive}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/cb_humanset_v2.pdf}
-    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\textwidth]{figures/cb_humanset_v2.pdf}
+%     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
 The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
-\end{figure}
+% \end{figure}
 

### 2024-10-18 src/tex/sections/human_subset_selection.tex
Update human_subset_selection.tex
Update src/tex/sections/human_subset_selection.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
-\paragraph{Human subset selection}
+\paragraph{Question selection}
 
-For selecting a human subset, several design choices have been made.
+Several design choices were made when selecting a human subset.
 Firstly, from the full dataset, we keep all the questions that are labeled as advanced.
-In this way, we can obtain a deeper insight about the capabilities of LLMs on advanced task
+In this way, we can obtain a deeper insight about the capabilities of \glspl{llm} on advanced tasks
 when compared to actual chemists.

### 2024-10-18 src/tex/sections/human_subset_selection.tex
Update src/tex/sections/human_subset_selection.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 in this subset.
-In total, 248 questions have been sampled for the human study.
-Then, this subset is divided into 2 subsets based on the aforementioned combinations.
+In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for the human study.
+Then, this subset is divided into two subsets based on the aforementioned combinations.
 One of the question subsets allows tool use, and the other does not.

### 2024-10-18 src/tex/sections/human_subset_selection.tex
Update src/tex/sections/human_subset_selection.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 when compared to actual chemists.
-Secondly, we sample a maximum of 3 questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry).
+Secondly, we sample a maximum of three questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry).
 Thirdly, we do not include any intuition questions

### 2024-10-18 src/tex/appendix.tex
more cleanup

@@
 
+\clearpage
 \subsection{Benchmark corpus}
@@
 
-\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
-We see that the questions are generally complex to read.
+%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
+%We see that the questions are generally complex to read.
 
@@
     \centering
-        \caption{Examples for each of the topics considered in the evaluation of the \chembench corpus.}
-    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}
+        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
+    \scriptsize
+    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}
         \toprule
@@
 
+\normalsize
+
 In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
@@
     \centering
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.}}
-    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}
+        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.}
+    \scriptsize
+    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}
         \toprule
@@
 
-% \subsection{Parsing verification}\label{sec:manually-verified-parsing}
-% For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
+\normalsize
 
@@
     }
+    \label{tab:performance_table}
 \end{table}
@@
     }
+    \label{tab:performance_table_human_subset}
 \end{table}
@@
 
-\subsection{Implementation}
-
-\begin{figure}
-    \centering
-    \includegraphics[height=\textheight]{figures/process.pdf}
-    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
-\texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
-    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types.
-For each task type, appropriate prompts are generated and passed to the \gls{llm}.
-The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
-    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
-    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback.
-The parsed or extracted answers then undergo metric computation and refusal detection.}
-    \label{fig:process}
-\end{figure}
-
 \subsection{Performance as a function of molecular features}
@@
 \Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
-The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
+%The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
 
@@
 
+\clearpage
 \subsection{Influence of model scale}
 \Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
-We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.
+We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.
 
@@
 
-\subsection{Human baseline}
+\clearpage
+\subsection{Implementation}
+An overview of the benchmarking pipeline implemented in \chembench is shown in \cref{fig:process}.
+More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
+\begin{figure}
+    \centering
+    \includegraphics[height=.8\textheight]{figures/process.pdf}
+    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
+\texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
+    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types.
+For each task type, appropriate prompts are generated and passed to the \gls{llm}.
+The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
+    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
+    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback.
+The parsed or extracted answers then undergo metric computation and refusal detection.}
+    \label{fig:process}
+\end{figure}
+
+\clearpage
+\subsection{Human baseline} \label{sec:human_baseline}
 \paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
-%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 
+\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}
@@
 
+\clearpage
 \subsection{Leaderboard}

### 2024-10-18 src/tex/ms.tex
Update ms.tex
more cleanup

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/cb_corpus_v1.pdf}
+    \includegraphics[width=\textwidth]{figures/cb_corpus_v2.pdf}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/human_subset_performance.pdf}
+    \includegraphics{figures/human_subset_performance.pdf}
     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly.
@@
 
-% ToDO: Link to full table of results in the SI
-\paragraph{System performance} 
+\paragraph{Overall system performance}
 To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
-An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
+An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
 In this figure, we show the percentage of questions that the models answered correctly.
@@
 
-Given the considerable interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
-Their lack of performance, however, is partially because we limited the system to a maximum of ten \gls{llm} calls.
-With the default tool augmented setup (using the so-called ReAct method\autocite{yao2022react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
-This observation highlights the importance of communicating and measuring not only predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.
+Interestingly, we find that models are still limited in their ability to answer knowledge-intensive questions.
+Importantly, this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo.
+This is likely because the required knowledge cannot easily be accessed via papers but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions.
+This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
 
-% Importance of scale 
-% Still lacks in knowledge-intensive questions
+In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
+This is in line with observations in other domains but also indicates that chemical \gls{llm} could be further improved by scaling them up.
 
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
-For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}).
-We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
-We then computed the percentage of questions the models or humans answered correctly for each topic.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics.
+We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
+In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly).
+Thus, a larger colored area indicates a better performance.
+
 \begin{figure}[!h]
@@
 
-% In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
-% Thus, a larger colored area indicates a better performance. 
-% One can observe that this performance varies widely across models and topics. 
-% While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
-% In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
-% Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+One can observe that this performance varies across models and topics.
+While general and technical chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as toxicity and safety or analytical chemistry.
+
+In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
+ Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 % These findings also shine an interesting light on the value of textbook-inspired questions.
@@
 
-REWRITE
-
-%To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.
+To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly (see \Cref{sec:confidence_estimates} for details on methodology and comparison to logit-based approaches).
 
@@
 
-%\input{sections/semi_programatically_sources_table.tex}
+\input{sections/semi_programatically_sources_table.tex}
 
-\subparagraph{Chemical Preference data}
+\subparagraph{Chemical preference data}
 
@@
 We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases than other prompting strategies, which might be more suitable when implemented in systems.
+In addition, this approach captures semantic uncertainty which is not the same as the probability of a token given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).
 
@@
 \paragraph{Study design}
-Rewrite.
+Human participants were asked the questions in a custom-built web interface (\Cref{sec:human_baseline}) which rendered chemicals and equations.
+Questions were shown in random order and participants were not allowed to skip questions.
+For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
+Prior to answering questions, participants were asked to provide information about their education and experience in chemistry.
+The study was conducted in English.
 
@@
 Overall, \variable{output/num_users_with_education_info.txt} did so.
-Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+Out of those,
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
 
@@
 \subsection{Data annotation}\label{sec:meth-topic}
-% When curating our dataset, we systematically recorded keywords and sources.
-% To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
-% We use this information to make the assignment for questions that can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions).
-% For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.
-
-REWRITE.
+In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question.
+We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
+the H/P statements of chemicals).
+Reasoning is required if answering a question requires multiple reasoning steps.
+Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.
 
@@
 
+\section{Statement of ethical compliance}
+The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe).
+Informed consent was obtained from all participants.
+
 \section*{Conflicts of interest}
@@
 
+\resizebox{\textwidth}{!}{%
 \scriptsize
 \insertcredits
+}
 \normalsize

### 2024-10-18 src/tex/appendix.tex
chore: add plot in appendix

--- 2024-10-18 src/tex/appendix.tex
@@
         \toprule
-        Question Type & \% & Example Questions  \\ 
+        \textbf{Question Type} & \textbf{\%} & \textbf{Example Questions}  \\ 
         \midrule
@@
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
+        & & A. [START_SMILES]N#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END_SMILES] \\
+        & & B. [START_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END_SMILES] \\
         \midrule
@@
         \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END\_RXNSMILES]?
-\\
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END_RXNSMILES]? \\
         & & A. $E_1$ \\
@@
         Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \ce{HF} over \ce{HCl} for this purpose?
+Why would one choose \\ce{HF} over \\ce{HCl} for this purpose?
 \\
-        & & A. \ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
-        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
+        & & A. \\ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \\ce{HF} has a better water solubility than \\ce{HCl} \\
+        & & C. \\ce{HF} has a higher boiling point than \\ce{HCl} \\
+        & & D. \\ce{HF} is a weaker acid than \\ce{HCl} \\
         \midrule
         General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
         \midrule
@@
 \\
-        & & A. \ce{C60} \\
-        & & B. \ce{CH4} \\
-        & & C. \ce{Fe(CO)5} \\
-        & & D. \ce{H2+} \\
-        & & E. \ce{NaCl} \\
+        & & A. \\ce{C60} \\
+        & & B. \\ce{CH4} \\
+        & & C. \\ce{Fe(CO)5} \\
+        & & D. \\ce{H2+} \\
+        & & E. \\ce{NaCl} \\
         \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \\ce{[ZrF7]^{3-}}\\
     \end{tabular}
@@
         \toprule
-        Required Skills & Number of Questions & Example Question \\
+        \textbf{Required skills} & \textbf{Number of Questions} & \textbf{Example Question} \\
         \midrule
         Knowledge & 770 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
+        & & A. \\ce{NH4Cl} \\
+        & & B. \\ce{Na2CO3} \\
+        & & C. \\ce{NaH2PO4} \\
+        & & D. \\ce{Zn(OH)Cl} \\
         \midrule
@@
 \\
-        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
+        & & A. [START_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C@@H]1O[END_SMILES] \\
+        & & B. [START_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END_SMILES] \\
         \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
-The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
+        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \\pu{595 g mol^{-1}} using end-group analysis, where \\pu{0.1619 g} of PLA was dissolved in \\pu{25 cm^3} of benzyl alcohol and titrated with \\pu{0.0400 mol dm^{-3}} \\ce{KOH} solution.
+The volume of \\ce{KOH} solution required to reach the endpoint was \\pu{6.81 cm^3}.
 What is the average number of monomer units in each polymer chain of this sample?
@@
 
+Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models. 
+GPT-4o demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers. 
+However, when GPT-4o displays high confidence, it accurately predicts correct answers approximately 80\% of the time.
+In contrast, Llama-3.1-Instruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5 probability.
+High-confidence predictions from Llama are less frequent compared to GPT-4o, and unlike GPT-4o, high confidence does not necessarily correlate with a higher chance of correct answers.
+\begin{figure}[htb] 
+    \centering
+    \includegraphics{figures/calibration_plot_overall_filtered.pdf}
+    \caption{\textbf{Calibration between linear probability of \glspl{llm} and correctness.} For this analysis we obtained the linear probability from the logprobs of the models.
+Only logprobs of the tokens corresponding to the answers were considered.
+Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
+The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities.
+The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness.
+The ECE (Expected Calibration Error) value quantifies the overall calibration performance, with a lower ECE indicating better calibration.}
+    \label{fig:confidence_score_distributions}
+    \script{plot_logprobs.py}
+\end{figure}
+
 % \begin{figure}[htb] 

### 2024-10-18 src/tex/appendix.tex
additional sections

--- 2024-10-18 src/tex/appendix.tex
@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\textwidth]{figures/cb_humanset_v2.pdf}
-%     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/cb_humanset_v3.pdf}
+    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
 The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
-% \end{figure}
+\end{figure}
 
@@
 \clearpage
+\subsection{Refusal detection}
+
+To automatically detect refusals, we use a modified regular expression reported by LLM Guard\autocite{llmguard} to detect commonly used refusal phrases.
+
+\Cref{tab:refusal_counts} lists how many refusals were detected for the reponses of different models on \chembench.
+Overall, we find that refusals do not majorly affect the performance measured by \chembench.
+
+\clearpage
+\subsection{LLM Parsing}
+
+In our parsing workflow, we use pipelines based on regular expressions to extract the answers.
+In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions.
+In these cases, we use a fallback mechanism to extract the answers.
+The fallback mechanism uses an \gls{llm} to extract the answers from the responses.
+The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer.
+\Cref{tab:llm_parsing_count} tabulates the number of times the fallback mechanism was used for each model.
+
+\clearpage
 \subsection{Implementation}

### 2024-10-18 src/tex/appendix.tex

@@
 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/confidence_score_distributions.pdf}
-%     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
-The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
-The bar plots show how many questions were answered with each confidence score.}
-%     \label{fig:confidence_score_distributions}
-%     \script{plot_confidence_score_distributions.py}
-% \end{figure}
+Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models. 
+GPT-4o demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers. 
+However, when GPT-4o displays high confidence, it accurately predicts correct answers approximately 80\% of the time.
+In contrast, Llama-3.1-Instruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5 probability.
+High-confidence predictions from Llama are less frequent compared to GPT-4o, and unlike GPT-4o, high confidence does not necessarily correlate with a higher chance of correct answers.
+\begin{figure}[htb] 
+    \centering
+    \includegraphics{figures/calibration_plot_overall_filtered.pdf}
+    \caption{\textbf{Calibration between linear probability of \glspl{llm} and correctness.} For this analysis we obtained the linear probability from the logprobs of the models.
+Only logprobs of the tokens corresponding to the answers were considered.
+Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
+The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities.
+The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness.
+The ECE (Expected Calibration Error) value quantifies the overall calibration performance, with a lower ECE indicating better calibration.}
+    \label{fig:confidence_score_distributions}
+    \script{plot_logprobs.py}
+\end{figure}
 

### 2024-10-18 src/tex/ms.tex

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/cb_corpus_v2.pdf}
+    \includegraphics[width=\textwidth]{figures/cb_corpus_v3.pdf}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
 
-% We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
-% If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
-% However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
-% This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
-% Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
+ We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
+ If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
+ However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
+ This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on proximity of the molecules to the training data.\autocite{mccoy2023embersautoregressionunderstandinglarge}
 
-% It is important to note that the model performance for some topics, however, is underestimated in the current evaluation.
-% This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
-% For instance, models might refuse to provide answers about cyanides.
-% To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
-% This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
+It is important to note that the model performance for some topics, however, is slightly underestimated in the current evaluation.
+ This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
+ For instance, models might refuse to provide answers about cyanides. Statistics of the frequency of such refusals are shown in \Cref{tab:refusals}.
+ To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
+ This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
 

### 2024-10-18 src/tex/appendix.tex
cleanup
try to make ready for logprobs figure

--- 2024-10-18 src/tex/appendix.tex
@@
 
+\Cref{fig:cb_humanset} shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/cb_humanset_v3.pdf}
+    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
+    \label{fig:cb_humanset}
+\end{figure}
+
 %\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
@@
 
-\begin{table*}[!t]
-    \centering
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
         \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
-    \scriptsize
-    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}
+
         \toprule
@@
         Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-    \end{tabularx}
     \label{tab:chembench_corpus_topic}
-\end{table*}
+\end{xltabular}
 
@@
 
-\begin{table*}[!t]
-    \centering
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
         \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.}
-    \scriptsize
-    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}
+
         \toprule
@@
         \bottomrule
-    \end{tabularx}
+
     \label{tab:chembench_corpus_cognitive}
-\end{table*}
+\end{xltabular}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/cb_humanset_v3.pdf}
-    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
-The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
-\end{figure}
-
+\clearpage
 \subsection{Model performance}
@@
 \subsection{Implementation}
-An overview of the benchmarking pipeline implemented in \chembench is shown in \cref{fig:process}.
+An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}.
 More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
@@
 
+\paragraph{Tool use}
+In our study, humans were allowed to use tools for answering some questions.
+They could also report what tools they used for answering questions.
+As \Cref{fig:tool_use} shows, the most commonly tool was some form of web search (which, according to the freetext responses, often was a multi-step process).
+
 \clearpage
@@
 This is similar to the verbalized confidence scores reported by \textcite{xiong2023llms}.
-\Cref{fig:confidence_score_distributions} plots the distribution of those scores.
 We find that the models show different distributions of confidence scores, which, for some, are skewed to the extremes.
 
-% \begin{figure}[htb]
+In addition, we also analyzed the confidence estimated via the log probabilities of the answer tokens.
+This probability of a token given the context is not necessarily the same as the confidence in the correctness of the answer.
+However, it is still often used as a proxy.
+
+Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models.
+\GPTFourO demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers.
+However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately 80\% of the time.
+In contrast, \LlamaThreeOneEightBInstruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5.
+High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.
+%\begin{figure}[htb]
 %    \centering
-%     \includegraphics{figures/confidence_score_distributions.pdf}
-%     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
-The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
-The bar plots show how many questions were answered with each confidence score.}
+%    \includegraphics{figures/calibration_plot_overall_filtered.pdf}
+%    \caption{\textbf{Calibration between linear probability of \glspl{llm} and correctness.} For this analysis we obtained the linear probability from the logprobs of the models.
+Only logprobs of the tokens corresponding to the answers were considered.
+Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
+The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities.
+The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness.
+The ECE (Expected Calibration Error) value quantifies the overall calibration performance, with a lower ECE indicating better calibration.}
 %    \label{fig:confidence_score_distributions}
-%     \script{plot_confidence_score_distributions.py}
-% \end{figure}
+%    \script{plot_logprobs.py}
+%\end{figure}
 

### 2024-10-18 src/tex/sections/human_subset_selection.tex
Update human_subset_selection.tex

@@
 Thirdly, we do not include any intuition questions
-in this subset.
+in this subset, because the correct answers are subjective, and would hinder an objective evaluation of the chemists against \glspl{llm}.
 In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for the human study.
-Then, this subset is divided into two subsets based on the aforementioned combinations.
+Then, this set is divided into two subsets based on the aforementioned combinations.
 One of the question subsets allows tool use, and the other does not.

### 2024-10-19 src/tex/sections/manually_sources_table.tex
[showyourwork]
fix overleaf sync messup
first additions

@@
        & Lab safety quizzes based on various sources && \variable{output/question_count_per_dir/json_file_counts_sci_lab_safety_test.txt} + \variable{output/question_count_per_dir/json_file_counts_lab_safety.txt} + \variable{output/question_count_per_dir/json_file_counts_stolaf.txt} + \variable{output/question_count_per_dir/json_file_counts_chemical_safety_mcq_exam.txt} + \variable{output/question_count_per_dir/json_file_counts_anderson.txt}\\
+\midrule
+Spectroscopy & Based on exams at the University of Hamburg, Germany && \variable{output/question_count_per_dir/json_file_counts_spectroscopy.txt} \\
+\midrule
+Solid state chemistry & Based on lectures published on the YouTube channels "Material Universe", "Advanced Cyberinfrastructure Training at RPI", "Phase Diagrams in Materials Science and Engineering" && 9\\
+                    & Based on the website \href{Wayback Machine}{https://archive.org/details/AshcroftSolidState/page/128/mode/2up} && 4\\
+                    & Based on the \href{https://ocw.mit.edu/courses/3-091sc-introduction-to-solid-state-chemistry-fall-2010/pages/organic-materials/29-polymers-synthesis-properties-applications/}{lecture notes} published by the Massachusetts Institute of Technology (MIT) && 4\\
+Technical chemistry & Based on notes from lectures at school at the University of Oviedo && 12\\
+                    & Based on the chemical engineering textbook from Octave Levenspiel && 8\\
+                    & Based on a lecture at the IIT Kharagpur, India && 1\\
+\midrule
+Hybridization questions &  Based on exams at the Ecole Polytechnique Federale de Lausanne && 9\\
+\midrule
+Biochemistry & Based on exams at the FSU Jena, Germany && 7\\
+\midrule
+Physical Chemistry & Based on a textbook & 6\\
 \bottomrule

### 2024-10-19 src/tex/sections/manually_sources_table.tex
first additions

@@
 \midrule
-Solid state chemistry & Based on lectures published on the YouTube channels "Material Universe", "Advanced Cyberinfrastructure Training at RPI", "Phase Diagrams in Materials Science and Engineering" && 9\\
+Solid state chemistry & Based on lectures published on the YouTube channels Materials Universe, Advanced Cyberinfrastructure Training at RPI, Phase Diagrams in Materials Science and Engineering" && 9\\
                     & Based on the website \href{Wayback Machine}{https://archive.org/details/AshcroftSolidState/page/128/mode/2up} && 4\\

### 2024-10-19 src/tex/sections/manually_sources_table.tex
more additions

@@
 \midrule
-Chemistry olympiad & National and International Olympiads held in US, UK and Moldova && \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
+Chemistry olympiad & National and International Olympiads held in US, UK, Romania and Moldova && \variable{output/question_count_per_dir/json_file_counts_icho.txt} \\
 \midrule
@@
 Physical Chemistry & Based on a textbook & 6\\
+\midrule Inorganic Chemistry & Based on questions from the national examination of the Russian Federation && 3\\
 \bottomrule

### 2024-10-19 src/tex/sections/human_subset_selection.tex
Update human_subset_selection.tex

@@
 Secondly, we sample a maximum of three questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry).
-Thirdly, we do not include any intuition questions
-in this subset, because the correct answers are subjective, and would hinder an objective evaluation of the chemists against \glspl{llm}.
+Thirdly, we do not include any intuition questions in this subset, because the intended used of the human subset is to provide a fast and fair evaluation of \glspl{llm} independent of any human baseline.
 In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for the human study.

### 2024-10-19 src/tex/sections/manually_sources_table.tex
Update manually_sources_table.tex
add 2 more sources

@@
                     & Based on the \href{https://ocw.mit.edu/courses/3-091sc-introduction-to-solid-state-chemistry-fall-2010/pages/organic-materials/29-polymers-synthesis-properties-applications/}{lecture notes} published by the Massachusetts Institute of Technology (MIT) && 4\\
-Technical chemistry & Based on notes from lectures at school at the University of Oviedo && 12\\
+Technical chemistry & Based on notes from lectures at the University of Oviedo && 12\\
                     & Based on the chemical engineering textbook from Octave Levenspiel && 8\\
@@
 \midrule
-Physical Chemistry & Based on a textbook & 6\\
+Physical Chemistry & Based on a textbook && 6\\
 \midrule Inorganic Chemistry & Based on questions from the national examination of the Russian Federation && 3\\
+Analytical Chemistry & Based from exams at IIT Bombay, India && 5\\
+\midrule
+Polymer Chemistry & Based on lecture notes from FSU, Jena && 15\\
+\midrule
 \bottomrule

### 2024-10-19 src/tex/appendix.tex
add in logprobs figure

@@
 \begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
+        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
 
@@
 \begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.}
+        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
 
@@
 High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.
-%\begin{figure}[htb]
-%    \centering
-%    \includegraphics{figures/calibration_plot_overall_filtered.pdf}
-%    \caption{\textbf{Calibration between linear probability of \glspl{llm} and correctness.} For this analysis we obtained the linear probability from the logprobs of the models.
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
+    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models.
 Only logprobs of the tokens corresponding to the answers were considered.
@@
 The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness.
-The ECE (Expected Calibration Error) value quantifies the overall calibration performance, with a lower ECE indicating better calibration.}
-%    \label{fig:confidence_score_distributions}
-%    \script{plot_logprobs.py}
-%\end{figure}
+The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
+    \label{fig:confidence_score_distributions}
+    \script{plot_logprobs.py}
+\end{figure}
 

### 2024-10-19 src/scripts/plot_logprobs.py

@@ -7,8 +7,9 @@ import pandas as pd
 from loguru import logger
 from paths import figures, scripts
 from utils import (
-    ONE_COL_WIDTH_INCH,
+    TWO_COL_WIDTH_INCH,
     TWO_COL_GOLDEN_RATIO_HEIGHT_INCH,
+    ONE_COL_GOLDEN_RATIO_HEIGHT_INCH,
     obtain_chembench_repo,
 )
 from plotutils import model_color_map
@@ -65,11 +66,12 @@ def create_calibration_plot(df, num_bins):
 
 def make_plot_of_calibration(merged_dicts, num_bins, suffix: str = ""):
     fig, ax = plt.subplots(
-        2,
-        1,  # Two subplots for two models
-        figsize=(ONE_COL_WIDTH_INCH, TWO_COL_GOLDEN_RATIO_HEIGHT_INCH),
+        1,
+        2,  # Two subplots for two models
+        figsize=(TWO_COL_WIDTH_INCH, ONE_COL_GOLDEN_RATIO_HEIGHT_INCH),
         sharex=True,
     )
+    plt.subplots_adjust(wspace=0.5)
 
     for i, (model, df) in enumerate(merged_dicts.items()):
         mean_predicted_probs, std_dev_probs, fraction_positives, bin_counts = (
@@ -100,7 +102,8 @@ def make_plot_of_calibration(merged_dicts, num_bins, suffix: str = ""):
         ax[i].plot([0, 1], [0, 1], linestyle="--", color="gray")
 
         ax[i].set_title(rename_dict[model])
-        ax[i].set_ylabel("fraction of positives")
+        if i == 0:
+            ax[i].set_ylabel("fraction of positives")
         ax[i].set_ylim(0, 1)
         ax[i].set_xlim(0, 1)
 
@@ -115,7 +118,8 @@ def make_plot_of_calibration(merged_dicts, num_bins, suffix: str = ""):
             color=model_color_map[rename_dict[model]],
             align="edge",
         )
-        ax_twin.set_ylabel("Count")
+        if i == 1:
+            ax_twin.set_ylabel("Count")
 
         # Calculate and display ECE
         ece = np.mean(
@@ -129,7 +133,8 @@ def make_plot_of_calibration(merged_dicts, num_bins, suffix: str = ""):
             verticalalignment="top",
         )
 
-    ax[-1].set_xlabel("predicted probability")
+    # add xlabel
[…]

### 2024-10-19 src/tex/appendix.tex
add in logprobs figure

@@
 \clearpage
+\subsection{Impact of sampling temperature}
+We also investigated the impact of sampling temperature on the performance of the models.
+\Cref{fig:temperature_impact} shows that, generally, the performance of models tends to degrease with increasing temperature.
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/swarm_plot_combined.pdf}
+    \caption{\textbf{Impact of sampling temperature on the performance of the models.} The plot shows the performance of the models at zero temperature (i.e., greedy decoding) and temperature of one.
+The performance is measured in terms of the fraction of questions answered correctly.}
+    \label{fig:temperature_impact}
+    \script{plot_temperature_diffs.py}
+\end{figure}
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-19 src/data/chem-bench-main/scripts/build_leaderboard.py
it compiles
Former-commit-id: aad21b06db862d3491ade91a2dec79ff349b347e

@@ -1,0 +1,384 @@
+import json
+import os
+from functools import partial
+from operator import itemgetter
+from pathlib import Path
+from typing import Any, Dict, List
+
+import pandas as pd
+import plotly.express as px
+import plotly.graph_objects as go
+import yaml
+
+THIS_DIR = os.path.dirname(os.path.abspath(__file__))
+DOCS_DIR = os.path.join(THIS_DIR, "../docs")
+REPORTS_DIR = os.path.join(THIS_DIR, "../reports")
+BENCHMARKS_DIR = os.path.join(DOCS_DIR, "Benchmarks")
+LEADERBOARDS_DIR = os.path.join(DOCS_DIR, "task_leaderboards")
+
+
+def load_yaml(file: str) -> Any:
+    """Load YAML file and return its contents."""
+    with open(file) as f:
+        data = yaml.load(f, Loader=yaml.FullLoader)
+    return data
+
+
+def load_json_as_dict(file: str) -> Dict:
+    """Load JSON file as dictionary and return it."""
+    with open(file) as f:
+        data = json.load(f)
+    return data
+
+
+def write_main_leaderboard(
+    main_leaderboard_results: List[Dict[str, Any]],
+    docs_folder: str,
+    get_name_from_yaml_: partial,
+) -> None:
+    """Write main leaderboard results to markdown file."""
+    main_leaderboard_results = sorted(main_leaderboard_results, key=lambda x: x["fraction_correct"], reverse=True)
+    names = [get_name_from_yaml_(result["model"]) for result in main_leaderboard_results]
+    scores = [result["fraction_correct"] for result in main_leaderboard_results]
+    with open(f"{docs_folder}/leaderboard.md", "w") as f:
+        f.write("# :crown: ChemBenchmark\n")
+        f.write("| Rank | Model | Fraction Correct |\n")
+        f.write("| --- | --- | --- |\n")
+        for i, result in enumerate(main_leaderboard_results, start=1):
+            fraction_correct = round(result["fraction_correct"], 3)
+            model = result["model"]
+            md_path = os.path.join("Benchmarks", model)
+            f.write(f"|  {i} |  [{get_name_from_yaml_(model)}]({md_path}.md)  | {fraction_correct} |\n")
+
+    fig = px.bar(x=names, y=scores, labels={"x": "Model", "y": "Fraction Correct"})
+    # fig.update_traces(marker_line_color="#7AA2E3", marker_line_width=0.5)
+
[…]

### 2024-10-19 src/data/chem-bench-main/src/chembench/analysis.py

@@ -1,0 +1,510 @@
+import json
+import math
+import os
+from concurrent.futures import ProcessPoolExecutor, as_completed
+from glob import glob
+from pathlib import Path
+from typing import Dict, List, Optional, Union
+
+import pandas as pd
+from loguru import logger
+
+from chembench.metrics import _AGGREGATION_MAP, _METRIC_MAP, classification_scores
+from chembench.task import Task
+from chembench.types import PathType
+
+
+def is_classification(report: dict):
+    return not isinstance(report["targets_"], (float, int))
+
+
+def is_number(obj: Union[str, int, float]) -> bool:
+    try:
+        float(obj)
+        return True
+    except ValueError:
+        return False
+
+
+def are_parsed_values_syntactically_correct(report: Union[dict, list]):
+    """
+    Check if the parsed values in a report are syntactically correct.
+
+    Args:
+        report (Union[dict, list]): The report to check. If a list, the first item is used.
+
+    Returns:
+        bool: True if the parsed values are syntactically correct, False otherwise.
+    """
+    if isinstance(report, list):
+        report = report[0]
+
+    # check if report["targets_"] can be converted to float
+    try:
+        is_classification = not (is_number(report["targets_"]))
+    except Exception:
+        is_classification = True
+
+    try:
+        if "exact_str_match" in report["parsed_output"]:
+            parsed = report["parsed_output"]["exact_str_match"]
+        else:
+            logger.warning(f"parsed_output not found in {report['name']}")
+            return False
+    except Exception as e:
+        logger.warning("Exception in are_parsed_values_syntactically_correct", e, report)
+        return False
+
+    # if none or nan, return false, or list that only contains none or nan
+    if not parsed or all([v is None for v in parsed]):
+        logger.warning(f"Parsed data for {report['name']} is None or nan")
+        return False
+
+    if is_classification:
+        num_target = list(report["targets_"].keys())
+        alphabet = [chr(i) for i in range(ord("A"), ord("Z") + 1)]
+        allowed_values = [alphabet[i] for i in range(len(num_target))]
+        # make sure there is no value that is not in allowed_values
+
+        print(parsed, allowed_values)
[…]

### 2024-10-19 src/data/chem-bench-main/src/chembench/baseline.py

@@ -1,0 +1,73 @@
+import random
+import re
+from dataclasses import dataclass
+from typing import List
+
+from loguru import logger
+
+
+@dataclass
+class Generation:
+    text: str
+
+
+@dataclass
+class Generations:
+    generations: List[List[Generation]]
+
+    def __getitem__(self, index):
+        return self.generations[index]
+
+
+class RandomBaseline:
+    """Class for generating random baseline answers."""
+
+    def __init__(
+        self,
+        mcq_strategy: str = "random",
+        general_constant: float = 0,
+        instruct: bool = False,
+        mcq_constant: float = "A",
+    ) -> None:
+        self.mcq_strategy = mcq_strategy
+        self.general_constant = general_constant
+        self.instruct = instruct
+        self.mcq_constant = mcq_constant
+
+    def generate(self, prompt: List[str]) -> Generations:
+        """
+        Generates random baseline answers based on the given prompt.
+
+        Args:
+            prompt (List[str]): List of prompts.
+
+        Returns:
+            Generations: List of generated answers.
+        """
+        generations = []
+        for prompt_ in prompt:
+            if "Options:" in prompt_:
+                option_letters = re.findall(r"[\n][A-Z]\.", prompt_)
+                # remove the dot
+                if self.mcq_strategy == "random":
+                    option_letters = [letter[:-1].strip() for letter in option_letters]
+                    answer = random.choice(option_letters)
+
+                elif self.mcq_strategy == "constant":
+                    answer = self.mcq_constant
+
+                elif self.mcq_strategy == "first":
+                    answer = option_letters[0].strip()
+                else:
+                    raise ValueError("mcq_strategy must be random, constant or first")
+            # otherwise, we return 0
+            else:
+                answer = f"{self.general_constant}"
+            if self.instruct:
+                answer = f"[ANSWER]{answer}[/ANSWER]"
+            logger.info(f"Answer: {answer}")
+            generation = [Generation(text=answer)]
+            generations.append(generation)
+
+        generations = Generations(generations=generations)
+        return generations

### 2024-10-19 src/data/chem-bench-main/src/chembench/constant.py

@@ -1,0 +1,139 @@
+# ruff: noqa: E501
+# prompt templates for instruction finetuned models and completion models
+INSTRUCT_TEMPLATE_1 = """The following is a multiple choice question about chemistry.
+Please answer by responding with the letter of the correct answer.{cot}
+
+Question: {question}
+
+Options:
+{answers}
+
+You MUST include the letter(s) of the correct answer (separated by comma if there are many) within the following tags: [ANSWER] and [/ANSWER].
+For example, '[ANSWER]<answer>[/ANSWER]', where <answer> is comma- or space-separated list of the correct letters. Always answer in exactly this format of comma-separated letters between the two tags, even if you are unsure. We require this because we use automatic parsing."""
+
+
+COMPLETION_TEMPLATE_1 = """The following is a question about chemistry.
+Please answer by responding with the letter(s) of the correct answer(s). Separate multiple answers with a comma.{cot}
+
+Question: {question}
+
+Options:
+{answers}
+
+Constraints: You must answer by returning ONLY the LETTER(S) of the correct answers(s) and NOTHING ELSE. Do not use other characters, words, or numbers, only the LETTER(S) of the correct answer(s). You must also return with the correct letter(s) if you are unsure. You can choose from {letters}.
+
+Answer:"""
+
+
+INSTRUCT_TEMPLATE_2 = """The following is a question about chemistry.{cot}
+
+Question: {question}
+
+You MUST include the final answer within the following tags: [ANSWER] and [/ANSWER].
+For example, '[ANSWER]<answer>[/ANSWER]', where <answer> is only one number. Always answer in exactly this format, with ONE NUMBER between the two tags, even if you are unsure. Use dots as decimal separator. We require this because we use automatic parsing."""
+
+
+COMPLETION_TEMPLATE_2 = """The following is a question about chemistry. Please return the answer as a number.{cot}
+
+Question: {question}
+
[…]

### 2024-10-19 src/data/chem-bench-main/src/chembench/metrics.py

@@ -1,0 +1,192 @@
+import math
+from typing import Dict, Union
+
+import numpy as np
+from loguru import logger
+
+
+def _is_number_like(x):
+    """Check if the input is float-like."""
+    try:
+        float(x)
+        return True
+    except Exception:
+        return False
+
+
+def all_correct(found: Union[str, list, tuple, float, int], expected: Union[dict, float, int, str]) -> bool:
+    """
+    Check if the found value(s) is/are exactly the same as the expected value(s).
+
+    If the expected value is a dictionary, the function checks if the found value(s) match the keys in the dictionary
+    where the corresponding value is 1.
+
+    If the expected value is a number, the function checks if the mean absolute error between the found and expected
+    values is 0.
+
+    If the expected value is a string, the function checks if the found and expected strings are exactly the same.
+
+    Args:
+        found: The found value(s), can be a string, list, tuple, float, or integer.
+        expected: The expected value, can be a dictionary, float, integer, or string.
+
+    Returns:
+        A boolean value indicating whether the found value(s) is/are exactly the same as the expected value(s).
+    """
+    if isinstance(expected, dict):
+        if not isinstance(found, (list, tuple)):
+            found = [found]
+        all_correct = set(found) == set([k for k in expected.keys() if expected[k] == 1])
+    elif _is_number_like(expected):
+        expected = float(expected)
+        found = float(found)
+        all_correct = bool(mae(found, expected) == 0)
+    else:
+        all_correct = bool(exact_str_match(found, expected))
+    return all_correct
+
+
+def mae(found: Union[float, int, str], expected: Union[float, int, str]) -> float:
+    """Compute the mean absolute error between two values.
+
+    Casts the values to float before computing the error.
+    If the values cannot be cast to float, returns np.nan.
+
+    Args:
+        found: The found value.
+        expected: The expected value.
+
+    Returns:
+        The mean absolute error.
+    """
+    try:
+        found = float(found)
+        expected = float(expected)
+        return abs(found - expected)
+    except Exception:
+        return np.nan
+
+
+def mse(found: Union[float, int, str], expected: Union[float, int, str]) -> float:
+    """Compute the mean squared error between two values.
+
+    Casts the values to float before computing the error.
[…]

### 2024-10-19 src/data/chem-bench-main/src/chembench/prompter.py

@@ -1,0 +1,631 @@
+import os
+from abc import ABC, abstractmethod
+from collections import defaultdict
+from typing import Any, Dict, List, Literal, Optional, Tuple
+
+import numpy as np
+from fastcore.basics import basic_repr
+from loguru import logger
+
+from chembench.constant import (
+    FLOATQ_REGEX_TEMPLATE_1,
+    MCQ_REGEX_TEMPLATE_1,
+    MCQ_REGEX_TEMPLATE_2,
+    NUM_REGEX,
+)
+from chembench.extractor import LLMExtractor
+from chembench.metrics import _AGGREGATION_MAP, _METRIC_MAP, classification_scores
+from chembench.report import Report
+from chembench.settings import MODEL_POST_PROCESS_ENV
+from chembench.task import Task
+from chembench.utils import (
+    LangChainChatModelWrapper,
+    create_multiple_choice_confidence_prompt,
+    create_multiple_choice_prompt,
+    create_multiple_choice_prompt_completion,
+    create_multiple_choice_regex,
+    detect_refusal,
+    general_prompt,
+    general_prompt_completion,
+    general_prompt_confidence,
+    passthrough,
+    post_process_prompts,
+    prepare_general_answer,
+    prepare_mcq_answer,
+    remove_ce,
+    remove_math,
+    remove_pu,
+    remove_rxnsmiles,
+    remove_smiles,
+    run_regex,
+)
+from chembench.watermark import get_watermark
+
+
+class BasePrompter(ABC):
+    """Prompters take a task and build prompts for the LLM.
+
+    For this reason, Prompters have specific methods to build prompts
+    that might be different for different LLMs.
+
+    The Prompters extract information from the tasks and use
+    it to also implement the `report` method that will prompt the LLM
+    and compute the metrics of interest.
+
+    For this, we will parse the output in a comparatively robust way,
+    for example, taking into account various ways of writing scientific
+    numbers.
+
+    Classes that inherit from this class should implement the following
+    methods:
+
+    - `_prompts_with_choices`
+    - `_prompts_general`
+
+    And also set the following attributes:
+
+    - `_mcq_regex_template`
+    - `_floatq_regex_template`
+
+
+    Args:
+        model: LLM to prompt.
+        post_process_ce: Function to post process chemical equations.
+        post_process_math: Function to post process math equations.
+        post_process_pu: Function to post process physical units.
+        post_process_smiles: Function to post process SMILES.
[…]

### 2024-10-19 src/data/chem-bench-main/tests/test_analysis.py

@@ -1,0 +1,43 @@
+import math
+import os
+
+from chembench.analysis import (
+    are_parsed_values_syntactically_correct,
+    construct_name_to_path_dict,
+    load_all_reports,
+    load_json,
+    score_report,
+)
+
+THIS_DIR = os.path.dirname(os.path.abspath(__file__))
+
+
+def test_are_parsed_values_syntactically_correct():
+    d = {
+        "targets_": {"A": 1, "B": 0},
+        "parsed_output": {"exact_str_match": ["A", "B", "A"]},
+    }
+    assert are_parsed_values_syntactically_correct(d)
+
+
+def test_load_all_reports():
+    df = load_all_reports(
+        os.path.join(THIS_DIR, "test_files", "to_analyze"),
+        os.path.join(THIS_DIR, "..", "data"),
+    )
+    row = df[df[("name", 0)] == "2019-1b1-icho_uk_2019_1b1"]
+    assert math.isclose(row[("metrics", "exact_str_match")].values[0], 0.0, rel_tol=1e-5)
+
+
+def test_construct_name_to_path_dict():
+    d = construct_name_to_path_dict(os.path.join(THIS_DIR, "..", "data"))
+    assert isinstance(d, dict)
+
+    assert "chem_chem_comp-chemComp_177-ChemComp" in d
+
+
+def test_score_report(get_report_file):
+    d = load_json(get_report_file)
+    name_to_path = construct_name_to_path_dict(os.path.join(THIS_DIR, "..", "data"))
+    report = score_report(d, name_to_path)
+    assert isinstance(report, dict)

### 2024-10-19 src/data/chem-bench-main/tests/test_baseline.py

@@ -1,0 +1,29 @@
+from chembench.baseline import RandomBaseline
+
+
+def test_random_baseline():
+    example_prompt = """The following is a multiple choice question about chemistry.
+Answer by responding with the letter of the correct answer.
+
+Question: What is the capital of France?
+
+Options:
+A. Paris
+B. London
+C. Berlin
+
+Make sure to include the letter(s) of the correct answer (separated by comma if there are many) within the following tags: [ANSWER] and [/ANSWER]."""
+
+    baseline = RandomBaseline()
+    assert baseline.generate([example_prompt]).generations[0][0].text.replace("[ANSWER]", "").replace("[/ANSWER]", "") in ["A", "B", "C"]
+
+    # non mcq prompt
+
+    example_prompt = """The following is a question about chemistry.
+
+Question: What the the height of the Eiffel Tower?
+
+Answer by including the final answer within the following tags: [ANSWER] and [/ANSWER]."""
+
+    baseline = RandomBaseline()
+    assert baseline.generate([example_prompt]).generations[0][0].text.replace("[ANSWER]", "").replace("[/ANSWER]", "") == "0"

### 2024-10-19 src/data/chem-bench-main/tests/test_regex.py

@@ -1,0 +1,395 @@
+import os
+
+import pytest
+
+from chembench.constant import MCQ_REGEX_TEMPLATE_2, NUM_REGEX, NUM_WORDS_REGEX
+from chembench.helper import fix_unicode
+from chembench.utils import (
+    create_multiple_choice_regex,
+    find_numbers,
+    passthrough,
+    post_process_prompts,
+    prepare_general_answer,
+    prepare_mcq_answer,
+    remove_ce,
+    remove_smiles,
+    run_regex,
+)
+
+
+def pytest_runtest_setup(item):
+    if "CI" in os.environ:
+        if "skip_in_ci" in item.keywords:
+            pytest.skip("Skipping test in CI environment")
+
+
+test_cases = [
+    ("The speed of light is 3.00 x 10^8 m/s.", 3e8),
+    ("The temperature is 25.5°C.", 25.5),
+    ("Avogadro's number is 6.022e23.", 6.022e23),
+    ("Pi is approximately 3.14159265359.", 3.14159265359),
+    ("There are 1000 millimeters in a meter.", 1000),
+    ("Hello .1", 0.1),
+    ("Hello 1.", 1),
+    ("None is none", None),
+    ("e-10", 1e-10),
+    ("-1.1", -1.1),
+    ("-1e-10", -1e-10),
+    ("10^10", 10**10),
+    ("1.77E-6", 1.77e-6),  #
+    ("1E10", 1e10),
+    ("-.2E-4", -0.2e-4),
+    ("10x10^{10}", 10 * 10**10),
+]
+
+
+# Use pytest.mark.parametrize to create parametrized tests
+@pytest.mark.parametrize("input_string, expected_numbers", test_cases)
+def test_find_numbers(input_string, expected_numbers):
+    result = find_numbers(input_string)
+    assert result == expected_numbers
+
+
+unicode_test_cases = [
+    (
+        "Assume during fermentation at 21 °C,",
+        "Assume during fermentation at 21 degree C,",
+    ),
+    ("The √ of 4 is 2.", "The square root of 4 is 2."),
+    ("π is approximately 3.14159.", "pi is approximately 3.14159."),
+    ("This is a text with no symbols.", "This is a text with no symbols."),
+    ("∑ Summation ∑", "summation Summation summation"),
+    ("The ∠ of a triangle is 90°.", "The angle of a triangle is 90degree ."),
+]
+
+
[…]

### 2024-10-19 src/data/chem-bench-main/tests/test_utils.py

@@ -1,0 +1,175 @@
+import os
+import random
+import re
+
+import pytest
+
+from chembench.constant import NUM_REGEX
+from chembench.task import Task
+from chembench.utils import (
+    LangChainChatModelWrapper,
+    create_multiple_choice_prompt,
+    detect_refusal,
+    load_json,
+    permute_options,
+    remove_ce,
+    remove_pu,
+    run_regex,
+    run_regex_iterator,
+    text2int,
+)
+
+_THIS_DIR = os.path.dirname(os.path.realpath(__file__))
+
+
+@pytest.fixture
+def get_multiple_choice_json():
+    json_file = random.choice(["multiple_choice.json", "multiple_choice_2.json"])
+    return os.path.join(_THIS_DIR, "test_files", json_file)
+
+
+def test_permute_prompter_and_extract(get_multiple_choice_json):
+    random.seed()
+    seed = random.randint(1, 100)
+    path = get_multiple_choice_json
+    task = Task.from_json(path)
+    example = task._examples[0]
+    prompt, answer_to_score = create_multiple_choice_prompt(example, permute=True, random_seed=seed)
+    answers_scores = example["target_scores"].values()
+    answer = permute_options(answers_scores, seed)
+    assert answer == list(answer_to_score.values())
+
+
+def test_load_json(get_one_float_json):
+    result = load_json(get_one_float_json)
+    assert isinstance(result, dict)
+
+
+def test_run_regex():
+    assert run_regex("[-+]?\\d+", "34") == "34"
+    assert run_regex("[-+]?\\d+", "-34") == "-34"
+    assert run_regex("[-+]?\\d+", "tgbf") is None
+
+
+def test_run_regex_iterator():
+    num_regex = NUM_REGEX
+    pattern = re.compile(num_regex)
+    assert run_regex_iterator(pattern, "1.77E-6") == "1.77E-6"
+    assert run_regex_iterator(pattern, "3.27e22") == "3.27e22"
+    assert run_regex_iterator(pattern, "0.001") == "0.001"
+    assert run_regex_iterator(pattern, "1E10") == "1E10"
+    assert run_regex_iterator(pattern, "1e10") == "1e10"
+    assert run_regex_iterator(pattern, "1exp10") == "1exp10"
+    assert run_regex_iterator(pattern, "-.2E-4") == "-.2E-4"
[…]

### 2024-10-19 src/scripts/plot_temperature_diffs.py

@@ -8,7 +8,7 @@ import pandas as pd
 import seaborn as sns
 import numpy as np
 from itertools import cycle
-from plotutils import range_frame
+from plotutils import range_frame, model_color_map
 
 plt.style.use(scripts / "lamalab.mplstyle")
 
@@ -52,8 +52,7 @@ def compute_scores_subset(data, title):
     df_melted = df.melt(id_vars=["Model"], value_vars=["T=0", "T=1"],
                         var_name="Temperature", value_name="Score")
 
-    with open("color_palette.json", "r") as f:
-        model_color_map = json.load(f)
+
 
     unique_models = df_melted["Model"].unique()
     color_palette = {model: color for model, color in model_color_map.items() if model in unique_models}

### 2024-10-19 src/tex/appendix.tex
trying to make it compile again
add sreekanth's figures
feat: add refusal table + some lines about refusal in the appendix

--- 2024-10-19 src/tex/appendix.tex
@@
 Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition.
-    \item \emph{Cover a range of difficulty levels}.
-To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
-    \item \emph{Impossible to completely solve with current models}.
-A benchmark should contain questions that are impossible to solve with current models.
-If current models can solve all questions, the benchmark provides no useful signal.
 \end{itemize}
@@
 
-\clearpage
 \subsection{Benchmark corpus}
@@
 
-\Cref{fig:cb_humanset} shows the distribution of topics and required skills in the human subset of the \chembench corpus.
+\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.  
+We see that the questions are generally complex to read.
 
-\begin{figure}
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=\textwidth]{figures/cb_humanset_v3.pdf}
-    \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
-The human subset is a representative subset of the full corpus, with a balanced distribution of topics and skills.}
-    \label{fig:cb_humanset}
-\end{figure}
-
-%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
-%We see that the questions are generally complex to read.
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+    \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+    \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
 It is calculated based on the average number of syllables per word and words per sentence.
@@
 A substantial fraction, in contrast to other benchmarks, is open-ended. 
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-%     \label{fig:question_count_barplot_mcq_vs_general}
-%     \script{plot_statistics.py}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+    \label{fig:question_count_barplot_mcq_vs_general}
+    \script{plot_statistics.py}
+\end{figure}
 
-The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
+% \subsection{Parsing verification}\label{sec:manually-verified-parsing}
+% For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-
-        \toprule
-        Question Type & \% & Example Questions  \\
-        \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
-        \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-        \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END\_RXNSMILES]?
-\\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
-        \midrule
-        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
-However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
-What follows from this?
-\\
-        & & A. Pindolol has a greater therapeutic range than propranolol \\
-        & & B. Pindolol has a longer half-life than propranolol \\
-        & & C. Pindolol has intrinsic activity \\
-        & & D. Pindolol is more lipophilic than propranolol \\
-        & & E. Pindolol is more potent than propranolol \\
-        \midrule
-        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-        & & A. X-ray fluorescence analysis \\
-        & & B. Differential pulse polarography \\
-        & & C. Flame-atomic absorption spectroscopy \\
-        & & D. Gas chromatography with flame ionization detector \\
-        & & E. Hydride generation atomic absorption spectroscopy \\
-        \midrule
-        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \ce{HF} over \ce{HCl} for this purpose?
-\\
-        & & A. \ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
-        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
-        \midrule
-        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
-        \midrule
-        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
-\\
-        & & A. \ce{C60} \\
-        & & B. \ce{CH4} \\
-        & & C. \ce{Fe(CO)5} \\
-        & & D. \ce{H2+} \\
-        & & E. \ce{NaCl} \\
-        \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-    \label{tab:chembench_corpus_topic}
-\end{xltabular}
-
-\normalsize
-
-In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
-
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
-
-        \toprule
-        Required Skills & Number of Questions & Example Question \\
-        \midrule
-        Knowledge & 770 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
-        \midrule
-        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By measuring atomic force microscopy \\
-        & & B. By measuring fourier transform infrared spectroscopy \\
-        & & C. By measuring photoluminescence spectrum \\
-        & & D. By measuring the absorption spectroscopy \\
-        \midrule
-        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
-        \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
-The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
-What is the average number of monomer units in each polymer chain of this sample?
-\\
-        \bottomrule
-
-    \label{tab:chembench_corpus_cognitive}
-\end{xltabular}
-
-\normalsize
-
-\clearpage
 \subsection{Model performance}
 We also evaluated the model performance on the entire \chembench corpus. 
-\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.
-Note that this ranking differs from the one on the \enquote{tiny} subset
+\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
+Note that this ranking differs from the one on the \enquote{tiny} subset.
 
@@
 
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
-
 \begin{figure}[htb]
     \centering
-    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \includegraphics{figures/all_questions_models_completely_correct_radar_overall.pdf}
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
@@
     \script{analyze_model_reports.py}
+\end{figure}
+
+\Cref{fig:parallel_coordinates_overall} shows this data as a parallel coordinates plot. 
+This visualization highlights the critical observation that the ranking of the models highly depends on the questions they are evaluated on.
+Only very broad benchmarks have the potential to provide a comprehensive view of a model's capabilities. However, even in those cases, the weighting of the different topics is crucial.
+Hence, we believe that fine-grained analysis of model performance is vital for the development of future benchmarks.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_overall}
+    \script{make_parallel_coordinates_plot.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/parallel_coordinates_tiny.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+}
+    \label{fig:parallel_coordinates_tiny}
+    \script{make_parallel_coordinates_plot.py}
 \end{figure}
@@
 
-% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
-% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+One might wonder if questions that are more difficult to parse lead to worse performance of the models.
+\Cref{fig:reading_ease_vs_model_performance} shows no clear correlation between the reading ease of the questions and the performance of the models.
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/model_overall_cal.pdf}
-%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-%     \script{requires_calculation_plot.py}
-%     \label{fig:requires_cal}
-% \end{figure}
+\begin{figure}
+    \centering
+    \hspace*{-1cm}
+    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
+    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not.
+We do not observe a clear correlation between the reading ease of the questions and the performance of the models.
+}
+    \script{reading_ease_vs_model_performance.py}
+    \label{fig:reading_ease_vs_model_performance}
+\end{figure}
+
+In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps. 
+We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+\begin{figure}
+    \centering
+    \includegraphics{figures/model_overall_cal.pdf}
+    \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+    \script{requires_calculation_plot.py}
+    \label{fig:requires_cal}
+\end{figure}
 
@@
 \Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
-%The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
+The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
 
@@
     \centering 
-    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
+    \includegraphics{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
+    \includegraphics{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
     \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
@@
 
-% \begin{table}
-%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+\begin{table}
+    \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
 }
-%     \begin{tabularx}{\textwidth}{XXXX XXXX}
-%         \toprule
-%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\
-%         \midrule
-%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
-%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
-%         \midrule
-%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
-%         \midrule
-%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
-%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
-%         \bottomrule
-%     \end{tabularx}
-%     \label{tab:correlation_coefficients}
-% \end{table}
+    \begin{tabularx}{\textwidth}{XXXX XXXX}
+        \toprule
+        topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
+        \midrule
+        number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+                              & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+        \midrule 
+        total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+        \midrule 
+        \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+                & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+        \bottomrule
+    \end{tabularx}
+    \label{tab:correlation_coefficients}
+\end{table}
 
-\clearpage
 \subsection{Influence of model scale}
-\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
-We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.
+To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
+Interestingly, we find that the 7B and 70B models perform comparably, with the 13B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/model_size_plot.pdf}
-    \caption{\textbf{Performance of models as a function of model size.} The plot shows the performance of the models as a function of the parameter count.
-The performance is measured as the fraction of questions answered correctly by the models.
-We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.}
-    \script{performance_vs_model_size.py}
-    \label{fig:model_size_plot}
-\end{figure}
+Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}
 
-\clearpage
-\subsection{Refusal detection}
-
-To automatically detect refusals, we use a modified regular expression reported by LLM Guard\autocite{llmguard} to detect commonly used refusal phrases.
-
-\Cref{tab:refusal_counts} lists how many refusals were detected for the reponses of different models on \chembench.
-Overall, we find that refusals do not majorly affect the performance measured by \chembench.
-
-\clearpage
-\subsection{LLM Parsing}
-
-In our parsing workflow, we use pipelines based on regular expressions to extract the answers.
-In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions.
-In these cases, we use a fallback mechanism to extract the answers.
-The fallback mechanism uses an \gls{llm} to extract the answers from the responses.
-The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer.
-\Cref{tab:llm_parsing_count} tabulates the number of times the fallback mechanism was used for each model.
-
-\clearpage
-\subsection{Implementation}
-An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}.
-More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
-\begin{figure}
-    \centering
-    \includegraphics[height=.8\textheight]{figures/process.pdf}
-    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
-\texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
-    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types.
-For each task type, appropriate prompts are generated and passed to the \gls{llm}.
-The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
-    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
-    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback.
-The parsed or extracted answers then undergo metric computation and refusal detection.}
-    \label{fig:process}
-\end{figure}
-
-\clearpage
-\subsection{Human baseline} \label{sec:human_baseline}
+\subsection{Human baseline}
 \paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
 This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
-We utilize a Postgresql.
-The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
+We utilize a MySQL\autocite{mysql} database and Prisma \gls{orm}\autocite{prisma} for efficient database management.
+The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication and postMark for sending Emails.
 The application is hosted on the Vercel web hosting platform.
-
-In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
-\LaTeX\xspace equations and chemical equations were rendered using MathJax\autocite{mathjax}.
 
@@
 We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
+Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 
-\paragraph{Tool use}
-In our study, humans were allowed to use tools for answering some questions.
-They could also report what tools they used for answering questions.
-As \Cref{fig:tool_use} shows, the most commonly tool was some form of web search (which, according to the freetext responses, often was a multi-step process).
-
-\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}
@@
 This is similar to the verbalized confidence scores reported by \textcite{xiong2023llms}.
+\Cref{fig:confidence_score_distributions} plots the distribution of those scores.
 We find that the models show different distributions of confidence scores, which, for some, are skewed to the extremes.
 
-In addition, we also analyzed the confidence estimated via the log probabilities of the answer tokens.
-This probability of a token given the context is not necessarily the same as the confidence in the correctness of the answer.
-However, it is still often used as a proxy.
-
-Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models.
-\GPTFourO demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers.
-However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately 80\% of the time.
-In contrast, \LlamaThreeOneEightBInstruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5.
-High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.
 \begin{figure}[htb] 
     \centering
-    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
-    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models.
-Only logprobs of the tokens corresponding to the answers were considered.
-Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
-The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities.
-The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness.
-The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
+    \includegraphics{figures/confidence_score_distributions.pdf}
+    \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores.
+The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence.
+The bar plots show how many questions were answered with each confidence score.}
     \label{fig:confidence_score_distributions}
-    \script{plot_logprobs.py}
+    \script{plot_confidence_score_distributions.py}
 \end{figure}
 
-\clearpage
+\subsection{Question refusal and LLM fallback} \label{sec:refusal}
+
+\paragraph{Refusal}
+
+During the training of the \glspl{llm}, generally one of the steps is refusal training. 
+Thus, in this stage of the training, the models are trained to prevent them of generating harmful, undesirable or violent outputs.
+As a consequence, it can happen that the model refuses to answer some questions that they identify as potential adversarially crafted prompts.
+
+To have a measure of how much refusal happens in \chembench, we count the number of refused questions by each model with a rule-based approach. 
+This rule-based approach searches for the most common structures of the refusal completions made by \glspl{llm}.
+
+\paragraph{LLM-Fallback}
+
+As explained above, the parsing of the answers from the completions of the models are made by parsing an answer pattern imposed within the prompt.
+For the questions in which \glspl{llm} do not follow this pattern to answer, we implement the so-called \gls{llm}-fallback which consists in passing the question and the completion to another model asking to extract the question or return "None" otherwise.
+For the count, the model chosen for do this auxiliary extraction was Llama 3 70B accessed via the Groq API.
+
 \subsection{Leaderboard}

### 2024-10-20 src/tex/ms.tex

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/cb_corpus_v3.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V1.pdf}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
 
-% \begin{figure}[!h]
-%     \centering
-%     \includegraphics{figures/confidence_vs_performance_overall.pdf}
-%     \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model.
+ \begin{figure}[!h]
+     \centering
+     \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
+     \caption{\textbf{Calibration and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model.
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
-We then plot the correctness of the answers (which is calculated as the mean value of all answers being either correct (1) or incorrect (0)) against the confidence score.
-%         The stripplots show the individual data points, a strong density indicating a higher number of points.
-The lines show the mean, and the error bars show the standard error of the mean.
-The figure shows that the reliability of the confidence estimates varies widely across models.
-While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
-%     \label{fig:confidence_vs_performance}
-%     \script{joint_analysis_confidence_performance.py}
-% \end{figure}
+}
+     \label{fig:confidence_vs_performance}
+     \script{confidence_estimate.py}
+ \end{figure}
 
@@
 
-\paragraph{Question selection} \label{sec:subset-selection}
-Rewrite.
+\input{sections/human_subset_selection.tex}
 

### 2024-10-20 src/tex/appendix.tex

--- 2024-10-19 src/tex/appendix.tex
@@
     \centering
-    \includegraphics[height=.8\textheight]{figures/process.pdf}
+    \includegraphics[height=.7\textheight]{figures/process.pdf}
     \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}.
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/human_tool_usage_by_topic.pdf}
+    \script{human_tool_usage.py}
+    \caption{\textbf{Tool usage by human scorers.} The plot shows the most commonly used tools by human participants.}
+    \label{fig:tool_use}
+\end{figure}
+
 \clearpage
@@
 High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.
+
 \begin{figure}[htb]
@@
 
+\subsection{Question refusal and LLM fallback} \label{sec:refusal}
+
+\paragraph{Refusal}
+
+During the training of the \glspl{llm}, generally one of the steps is refusal training. 
+Thus, in this stage of the training, the models are trained to prevent them of generating harmful, undesirable or violent outputs.
+As a consequence, it can happen that the model refuses to answer some questions that they identify as potential adversarially crafted prompts.
+
+To have a measure of how much refusal happens in \chembench, we count the number of refused questions by each model with a rule-based approach. 
+This rule-based approach searches for the most common structures of the refusal completions made by \glspl{llm}.
+
+\paragraph{LLM-Fallback}
+
+As explained above, the parsing of the answers from the completions of the models are made by parsing an answer pattern imposed within the prompt.
+For the questions in which \glspl{llm} do not follow this pattern to answer, we implement the so-called \gls{llm}-fallback which consists in passing the question and the completion to another model asking to extract the question or return "None" otherwise.
+For the count, the model chosen for do this auxiliary extraction was Llama 3 70B accessed via the Groq API.
+
 \clearpage

### 2024-10-20 src/tex/appendix.tex
cleanup text for submission

--- 2024-10-19 src/tex/appendix.tex
@@
 \subsection{Refusal detection}
-
+\Glspl{llm} typically undergo refusal training to prevent harmful or undesirable outputs. As a result, models may decline to answer questions they perceive as potentially adversarial prompts.
 To automatically detect refusals, we use a modified regular expression reported by LLM Guard\autocite{llmguard} to detect commonly used refusal phrases.
@@
 The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer.
+We used \LlamaThreeSeventyBInstruct accessed via the Groq API.
 \Cref{tab:llm_parsing_count} tabulates the number of times the fallback mechanism was used for each model.

### 2024-10-20 src/tex/authors.tex
update overviedw
sorted affil
add mara orcid
[showyourwork]
fix overleaf sync messup
feat: start rewrite
cleanup text for submission

--- 2024-03-27 src/tex/authors.tex
@@
 \credit{Adrian~Mirza}{1,1,1,0,1,0,0,0,1,0,1,1,0,1}
-\credit{Nawaf~Alampara}{1,1,1,0,1,0,0,0,1,0,1,0,0,1}
-\credit{Sreekanth~Kunchapu}{1,1,0,0,1,0,0,0,0,0,1,0,0,1}
+\credit{Nawaf~Alampara}{1,1,1,0,1,0,0,0,1,0,1,1,0,1}
+\credit{Sreekanth~Kunchapu}{1,1,1,0,1,0,0,0,1,0,1,1,0,1}
+\credit{Martiño~Ríos-García}{1,1,1,0,1,0,0,0,1,0,1,1,0,1}
 
@@
 \credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
-\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Amir~Mohammad~Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Hani~M.~Elbeheiry~}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{María~Victoria~Gil}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
 \credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
@@
 \credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Abdelrahman~Ibrahim}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
 \credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
@@
 \credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
+\credit{Santiago~Miret}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
 \credit{Jan Matthias Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
@@
 
-\author[1,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
+\author[1,2,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
 \author[1,$\star$]{Nawaf~Alampara~\orcidlink{0009-0001-7697-7315}}
 \author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
+\author[1,2 $\star$]{Martiño~Ríos-García~\orcidlink{0000-0003-1507-4048}}
 \author[ ]{Benedict~Emoekabu}
-\author[2]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
-\author[3,4]{Tanya~Gupta~\orcidlink{0009-0001-9523-3290}}
-\author[1]{Mara~Wilhelmi}
+\author[4]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
+\author[5,6]{Tanya~Gupta~\orcidlink{0009-0001-9523-3290}}
+\author[1]{Mara~Wilhelmi~\orcidlink{0009-0007-4392-5918}}
 \author[1]{Macjonathan~Okereke~\orcidlink{0009-0007-1013-0502}}
 
-\author[5]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
-\author[6]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
-\author[7]{Amir~Elahi}
-\author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
+\author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
+\author[8]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
+\author[9]{Amir~Mohammad~Elahi~\orcidlink{0009-0001-5907-101X}}
+\author[1]{Hani~M.~Elbeheiry~\orcidlink{0000-0002-5205-2852}}
+\author[3]{María~Victoria~Gil~\orcidlink{0000-0003-4894-4660}}
+\author[ ]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
 \author[1]{Maximilian~Greiner}
-\author[1]{Caroline~T.~Holick\orcidlink{0009-0000-1724-2725}}
+\author[1]{Caroline~T.~Holick~\orcidlink{0009-0000-1724-2725}}
 \author[1]{Tim~Hoffmann~\orcidlink{0009-0004-0230-6115}}
+\author[1]{Abdelrahman~Ibrahim~\orcidlink{0009-0003-1460-4710}}
 \author[1]{Lea~C.~Klepsch~\orcidlink{0009-0009-3849-1670}}
 \author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
-\author[8, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[10, 11]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
 \author[1]{Jakob~Meyer}
+\author[12]{Santiago~Miret~\orcidlink{0000-0002-4787-2757}}
 \author[1]{Jan~Matthias~Peschel~\orcidlink{0009-0002-4787-2757}}
 \author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
-\author[1, 10]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
+\author[1, 13]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
 
-\author[1, 10]{Johanna Schreiber~\orcidlink{0009-0000-0991-8967}}
-\author[1,8,10, 11]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
-\author[1]{Leanne~M.~Stafast}
-\author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
+\author[1, 11]{Johanna~Schreiber~\orcidlink{0009-0000-0991-8967}}
+\author[1,2, 10, 13]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[1, 13]{Leanne~M.~Stafast~\orcidlink{0009-0008-5604-261X}}
+\author[14]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
 
 % Senior authors
-\author[12,13]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
-\author[3,4]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-\author[1,8,10, 11, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+\author[15,16]{Michael~Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[5, 6]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
+\author[1,2, 10, 13, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
 
 \affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
-\affil[8]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[11]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
-\affil[10]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
+\affil[2]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
 
-\affil[3]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
-\affil[4]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[3]{Institute of Carbon Science and Technology (INCAR), CSIC, Francisco Pintado Fe 26, 33011 Oviedo, Spain}
+\affil[4]{QpiVolta Technologies Pvt Ltd}
+\affil[5]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[6]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
+\affil[7]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
+\affil[8]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
+\affil[9]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
+\affil[10]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[11]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[12]{Intel Labs}
+\affil[13]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
 
-\affil[7]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
-\affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
-\affil[12]{OpenBioML.org}
-\affil[13]{Stability.AI}
-\affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[2]{QpiVolta Technologies Pvt Ltd}
-\affil[5]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
-\affil[6]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
+\affil[14]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
+\affil[15]{OpenBioML.org}
+\affil[16]{Stability.AI}
+
 \affil[\Letter]{\texttt{mail@kjablonka.com}}

### 2024-10-20 src/tex/ms.tex
cleanup text for submission

@@
 \begin{abstract}
-    Large language models (LLMs) have gained widespread interest due to their ability to process human language.
-    Also, LLMs have shown promise in addressing the limited data problem in chemistry and are increasingly being harnessed to improve and extend existing workflows.
+Large language models (LLMs) have gained widespread interest due to their ability to process human language and perform tasks on which they have not been explicitly trained.
 
-    However, we possess only a limited systematic understanding of the reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms.
+However, we possess only a limited systematic understanding of the chemical capabilities of LLMs, which would be required to improve models and mitigate potential harms.
 Here, we introduce \enquote{\chembench,} an automated framework for evaluating the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of chemists.
 
-    We curated more than 2,800 question-answer pairs, evaluated leading open and closed-source LLMs, and found that, on average, the best models outperformed the chemists.
-    The models, however, struggle with some chemical reasoning tasks that are easy for the human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.
-    Although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
+We curated more than 2,800 question-answer pairs, evaluated leading open—and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
+These findings reveal LLMs' impressive chemical capabilities while emphasizing the need for further research to improve their safety and usefulness.
+They also suggest adapting chemistry education and developing better evaluation methods for LLMs in chemistry.
 \end{abstract}
@@
 Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
-However, even for experts, chemical understanding and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from anecdotal reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.
+However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
+Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.
 
@@
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
-We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
+We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
 
@@
 For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks.
-Importantly, our large pool of questions encompasses a wide range of topics and question types.
+Importantly, our large pool of questions encompasses a wide range of topics and question types (\Cref{fig:corpus}).
 The topics range from general chemistry to more specialized fields such as inorganic, analytical or technical chemistry.
-We also classify the questions based on what techniques are required to answer them.
+We also classify the questions based on what skills are required to answer them.
 Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
@@
 \paragraph{Benchmark suite design} Because the text used in scientific settings differs from typical natural language, many models have been developed that deal with such text in a particular way.
-For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations.
+For instance, the Galactica model\autocite{taylor2022galactica} uses special encoding procedures for molecules and equations.
 Current benchmarking suites, however, do not account for such special treatment of scientific information.
@@
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
-    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
     The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
@@
 
-Interestingly, we find that models are still limited in their ability to answer knowledge-intensive questions.
-Importantly, this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo.
-This is likely because the required knowledge cannot easily be accessed via papers but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions.
+Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}).
+Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo.
+This is likely because the required knowledge cannot easily be accessed via papers but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
 This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
@@
 \paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
-For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics.
+For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
 We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
@@
 One can observe that this performance varies across models and topics.
-While general and technical chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as toxicity and safety or analytical chemistry.
+While general and technical receive relatively high scores for many models, this is not the case for topics such as toxicity and safety or analytical chemistry.
 
-In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
- Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
-% These findings also shine an interesting light on the value of textbook-inspired questions.
-% A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
-% On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
-% For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
-% While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+ These findings also shine an interesting light on the value of textbook-inspired questions.
+ A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
+ On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
+ For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5 (Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+ While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
+The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
 
@@
 \begin{refsection}
+\renewcommand{\thefigure}{A\arabic{figure}}
+\setcounter{figure}{0}
+
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{table}{0}
+
 \appendix

### 2024-10-20 src/scripts/analyze_human_data.py

@@ -64,6 +64,7 @@ def make_human_performance_plots():
     dirs_tool = list(set([os.path.dirname(p) for p in paths_tool]))
     dirs_no_tool = list(set([os.path.dirname(p) for p in paths_no_tool]))
     all_results = []
+    number_experts = 0
     for d in dirs_tool + dirs_no_tool:
         try:
             results = load_all_reports(d, os.path.join(chembench, "data"))
@@ -81,6 +82,7 @@ def make_human_performance_plots():
                 results["userid"] = userid
                 results["num_results"] = len(results)
                 results["experience"] = experience
+                number_experts += 1
                 results["highest_education"] = highest_education
                 if "tool-allowed" in d:
                     results["tool_allowed"] = True
@@ -94,10 +96,9 @@ def make_human_performance_plots():
             print(e)
             continue
 
-    number_humans = len(users)
 
     with open(output / "number_experts.txt", "w") as f:
-        f.write(f"{str(int(number_humans))}" + "\endinput")
+        f.write(f"{str(int(number_experts))}" + "\endinput")
 
     long_df = pd.concat(all_results).reset_index(drop=True)
     long_df["all_correct"] = long_df.apply(all_correct, axis=1)

### 2024-10-20 src/scripts/make_source_table.py
feat: add table about the sources

@@ -1,0 +1,114 @@
+
+import os
+import json
+import pickle
+import pandas as pd
+import numpy as np
+from loguru import logger
+import concurrent.futures
+from pandas import json_normalize
+
+from chembench.analysis import (
+    construct_name_to_path_dict,
+)
+from utils import obtain_chembench_repo
+
+from paths import tex, data
+
+def craft_table(df):
+    latex_table = df.to_latex(index=False)
+
+    latex_table = latex_table.replace(
+        r"\begin{tabular}{lr}",
+        r"\begin{tabular}{cc}"
+    )
+
+    with open(output_file, 'w') as f:
+        f.write(latex_table)
+
+    return
+
+def load_single_json(path):
+    with open(path, 'r') as file:
+        data = json.load(file)
+        return json_normalize(data)
+
+def load_data(data_paths):
+    results = []
+    for path in data_paths:
+        data = load_single_json(path)
+        results.append(load_single_json(path))
+    
+    combined_df = pd.concat(results, ignore_index=True, sort=False)
+    
+    return combined_df
+
+def collect_data(datafolder):
+    names = construct_name_to_path_dict(datafolder)
+    data_paths = list(names.values())
+
+    df = load_data(data_paths)
+
+    if df.empty:
+        raise ValueError("No data loaded from reports")
+    logger.info(f"Loaded {len(df)} rows of data")
+    
+    sources = {}
+    for i, row in df.iterrows():
+        try:
+            if not pd.isna(row["meta.exam.country"]):
+                if "Exam" in sources:
+                    sources["Exam"] += 1
+                else:
+                    sources["Exam"] = 0
+            elif not pd.isna(row["meta.urls"]):
+                if "URL" in sources:
+                    sources["URL"] += 1
+                else:
+                    sources["URL"] = 0
+            elif not pd.isna(row["meta.textbook.title"]):
+                if "Textbook" in sources:
+                    sources["Textbook"] += 1
+                else:
+                    sources["Textbook"] = 0
+            elif not pd.isna(row["meta.lecture.country"]):
+                if "Lectures" in sources:
+                    sources["Lectures"] += 1
+                else:
+                    sources["Lectures"] = 0 
+            elif not pd.isna(row["icho.country"]):
+                if "ICHO" in sources:
+                    sources["ICHO"] += 1
+                else:
+                    sources["ICHO"] = 0
+            else:
+                if "No source" in sources:
[…]

### 2024-10-20 src/tex/ms.tex
fix some references

@@
 
-One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman}
+One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman, Argyle_2023}
 If models could do so for chemical compounds, this would open opportunities for novel optimization approaches.
@@
 Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting).
-While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench.
+While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench (\Cref{tab:performance_table_human_subset})
 This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.
@@
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
+     The line plot shows the average fraction of correctly answered questions for each confidence level.
+The bar plot shows the distribution of confidence estimates.
+A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.
+     We find that the models are not well-calibrated and some provide misleading confidence estimates.
      }
@@
 
-% In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
-% For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
-% For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
-% While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
-% For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
+ For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
+ For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+ While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
+ For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
-For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning.
+For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
 Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
@@
 While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
-Although current systems are far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.
+Although current systems might be far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.
 
@@
 Manually curated questions were sourced from various sources, including university exams, exercises, and question banks.
-\Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
+%\Cref{tab:manually_sources} provides an overview of the sources of the manually curated questions.
 
@@
 Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that lasts several years within a chemist's career.
-Our data corpus is adapted from the published dataset by \textcite{Choung2023}, which consists of more than 5000 question-answer pairs about chemical intuition.
+Our data corpus is adapted from the published dataset by \textcite{Choung_2023}, which consists of more than 5000 question-answer pairs about chemical intuition.
 To build the dataset, they presented 35 medicinal chemists with two different molecules, asking them what molecule they would like to continue with when imaging an early virtual screening campaign setting.
@@
 
-\clearpage
-
 \subsection{Model evaluation workflow}
-An overview of the pipeline is shown in \Cref{fig:evaluation_workflow}.
+An overview of the pipeline is shown in \Cref{fig:process}.
 

### 2024-10-20 src/scripts/confidence_estimate.py

@@ -149,8 +149,45 @@ if __name__ == "__main__":
     try:
         merged_dicts = join_confidence_and_performance(json_file_paths)
         print(merged_dicts)
+        import pdb; pdb.set_trace()
         make_plot_of_confidence_vs_performance(merged_dicts)
         logger.info("Successfully completed analysis and saved plot as PDF")
     except Exception as e:
         logger.error(f"An error occurred during analysis: {e}")
         logger.exception("Error details:")
+
+
+
+
+
+    # for model, df in merged_dict.items():
+    #     relevant_model_performance = df[df[subset]]
+    #     correct = relevant_model_performance[
+    #         relevant_model_performance["all_correct_"].astype(bool)
+    #     ]
+    #     incorrect = relevant_model_performance[
+    #         ~relevant_model_performance["all_correct_"].astype(bool)
+    #     ]
+    #     num_correct = len(correct)
+    #     num_incorrect = len(incorrect)
+
+    #     average_confidence_correct = correct["estimate"].mean()
+    #     average_confidence_incorrect = incorrect["estimate"].mean()
+
+    #     with open(outpath / f"{model}_{subset}_num_correct_{suffix}.txt", "w") as f:
+    #         f.write(f"{num_correct}" + "\endinput")
+
+    #     with open(outpath / f"{model}_{subset}_num_incorrect_{suffix}.txt", "w") as f:
+    #         f.write(f"{num_incorrect}" + "\endinput")
+
+    #     with open(
+    #         outpath / f"{model}_{subset}_average_confidence_correct_{suffix}.txt", "w"
+    #     ) as f:
+    #         rounded = np.round(average_confidence_correct, 2)
+    #         f.write(f"{rounded}" + "\endinput")
+
+    #     with open(
+    #         outpath / f"{model}_{subset}_average_confidence_incorrect_{suffix}.txt", "w"
+    #     ) as f:
+    #         rounded = np.round(average_confidence_incorrect, 2)
+    #         f.write(f"{rounded}" + "\endinput")

### 2024-10-20 src/tex/ms.tex
more cleanup

@@
 As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
-If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., \ClaudeThreeFiveSonnet, to parse the completion.
+If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., \ClaudeThreeFiveSonnet, to parse the completion (\Cref{sec:llm-parsing} provides more details on this step).
 
@@
 In addition, this approach captures semantic uncertainty which is not the same as the probability of a token given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).
+In \Cref{sec:confidence_estimates} we provide more details and comparisons with a logit-based approach.
 
@@
 This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We also thank Stability.AI for the access to its HPC cluster and donations. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
+We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
 K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
@@
 
-\section{Statement of ethical compliance}
+\section*{Statement of ethical compliance}
 The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe).

### 2024-10-20 src/tex/appendix.tex
update references

@@
 
-\Cref{fig:question_count_barplot_mcq_vs_general} shows that most questions in our corpus are \gls{mcq}.
-A substantial fraction, in contrast to other benchmarks, is open-ended.
 % \begin{figure}[htb]
@@
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)(=O)c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
+        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
+        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
         \midrule
@@
         \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC[END\_RXNSMILES]?
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
 \\
@@
         & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
+        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
         \midrule
@@
 
-\subsection{Performance as a function of molecular features}
+\subsection{Performance as a function of molecular features} \label{sec:molecular_features}
 To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
@@
 
-\begin{figure}
+\begin{figure}[!h]
     \centering
@@
 \clearpage
-\subsection{LLM Parsing}
+\subsection{LLM Parsing} \label{sec:llm-parsing}
 
@@
 
-In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
-\LaTeX\xspace equations and chemical equations were rendered using MathJax\autocite{mathjax}.
+In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax.
 

### 2024-10-20 src/tex/ms.tex
update references

@@
  If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
- However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
+ However, we find that the models' performance does not correlate with complexity indicators (see \Cref{sec:molecular_features}).
  This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on proximity of the molecules to the training data.\autocite{mccoy2023embersautoregressionunderstandinglarge}
@@
 K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
+Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
 

### 2024-10-21 src/tex/ms.tex
it compiles now

@@
  On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
- For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5 (Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+ For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
  While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
@@
  For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
- For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+ For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
  While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.

### 2024-10-21 src/tex/appendix.tex
update for end-to-end run

@@
 \subsection{Refusal detection}
-\Glspl{llm} typically undergo refusal training to prevent harmful or undesirable outputs. As a result, models may decline to answer questions they perceive as potentially adversarial prompts.
+\Glspl{llm} typically undergo refusal training to prevent harmful or undesirable outputs. As a result, models may decline to answer questions perceived as potentially adversarial prompts.
 To automatically detect refusals, we use a modified regular expression reported by LLM Guard\autocite{llmguard} to detect commonly used refusal phrases.
 
-\Cref{tab:refusal_counts} lists how many refusals were detected for the reponses of different models on \chembench.
+\Cref{tab:refusal_counts_and_parsing} lists how many refusals were detected for the reponses of different models on \chembench.
 Overall, we find that refusals do not majorly affect the performance measured by \chembench.
@@
 We used \LlamaThreeSeventyBInstruct accessed via the Groq API.
-\Cref{tab:llm_parsing_count} tabulates the number of times the fallback mechanism was used for each model.
+\Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.
+
+\begin{table}
+    \centering
+    \caption{\textbf{Refusal counts and parsing.} The table shows the number of refusals detected and the number of times the fallback mechanism was used for each model.}
+    \variable{output/model_refusal_table.tex}
+    \label{tab:refusal_counts_and_parsing}
+\end{table}
 

### 2024-10-21 src/scripts/count_refusal.py

@@ -28,7 +28,7 @@ from chembench.constant import (
     NUM_REGEX,
 )
 from utils import obtain_chembench_repo
-
+from paths import output
 load_dotenv("/.env", override=True)
 
 def extract_examples(name, datafolder):
@@ -221,12 +221,10 @@ if __name__ == "__main__":
     logger.info(f"Found {len(models)} reports")
     logger.debug(f"First 5 reports: {models[:5]}")
 
-    outpath = os.path.join(chembench_repo, "output")
-    os.makedirs(outpath, exist_ok=True)
 
     datafolder = os.path.join(chembench_repo, "data")
-    json_dir_refusals = os.path.join(outpath, "refusals_files")
-    json_dir_extractions = os.path.join(outpath, "extractions_files")
+    json_dir_refusals = os.path.join(output, "refusals_files")
+    json_dir_extractions = os.path.join(output, "extractions_files")
     os.makedirs(json_dir_refusals, exist_ok=True)
     os.makedirs(json_dir_extractions, exist_ok=True)
 
@@ -242,7 +240,7 @@ if __name__ == "__main__":
             except Exception as exc:
                 logger.error(f'{model} generated an exception: {exc}')
 
-    with open(os.path.join(outpath, 'model_refusal_and_extraction_count.pkl'), 'wb') as f:
+    with open(os.path.join(output, 'model_refusal_and_extraction_count.pkl'), 'wb') as f:
         pickle.dump(results, f)
 
     logger.info("Processing completed successfully.")

### 2024-10-21 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Martiño Rios-Garcia <147555961+MrtinoRG@users.noreply.github.com>

@@
 
-We curated more than 2,800 question-answer pairs, evaluated leading open—and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
 However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.

### 2024-10-21 src/tex/appendix.tex
fix: correct required skill table + add script to automatically count required skills

@@
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1[END\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S(=O)(=O)N1[END\_SMILES] \\
+        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
         \midrule
@@
         Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+        \bottomrule
     \label{tab:chembench_corpus_topic}
@@
         \midrule
-        Knowledge & 770 & Which of the following salts is an acidic salt? \\
+        Knowledge & 919 & Which of the following salts is an acidic salt? \\
         & & A. \ce{NH4Cl} \\
@@
         \midrule
-        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
         & & A. By measuring atomic force microscopy \\
@@
         \midrule
-        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.

### 2024-10-21 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Martiño Rios-Garcia <147555961+MrtinoRG@users.noreply.github.com>

@@
 To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel manually crafted questions, over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
-For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks.
+For quality assurance, all questions have been reviewed by at least two scientists in addition to the original curator and automated checks.
 Importantly, our large pool of questions encompasses a wide range of topics and question types (\Cref{fig:corpus}).
@@
 Such open-ended tasks, however, depend on an external observer defining what interestingness is.\autocite{hughes2024openendednessessentialartificialsuperhuman}
-Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting).
+Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting) (see \Cref{tab:chembench_corpus_cognitive}).
 While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench (\Cref{tab:performance_table_human_subset})

### 2024-10-21 src/tex/ms.tex
minore cleanup

@@
 
-%\input{sections/manually_sources_table.tex}
+\begin{table}
+    \centering
+    \caption{\textbf{Overview of sources of the curated questions}.
+The table provides an overview of the types of sources the questions have been curated from.
+Detailed sources are available in the source data.
+Questions without source have been curated completely from scratch.
+Questions based on lecture notes or URL have been curated based on content presented in those resources.
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
+    \variable{output/sources_table.tex}
+    \label{tab:sources}
+\end{table}
 

### 2024-10-21 src/scripts/make_refusal_tables.py

@@ -20,7 +20,7 @@ model_file_name_to_label = {
     "gpt-3.5-turbo": "GPT-3.5 Turbo",
     "gpt-4": "GPT-4",
     "gpt-4o": "GPT-4o",
-    "o1": "OpenAI o1",
+    "o1": "o1",
     "llama2-70b-chat": "Llama 2 70B Chat",
     "llama3-70b-instruct": "Llama 3 70B",
     "llama3-70b-instruct-T-one": "Llama 3 70B Temp=1",
@@ -53,7 +53,7 @@ def list_of_dicts_to_latex_table(data, output_file):
 
     latex_table = latex_table.replace(
         r"\begin{tabular}{lrrrr}",
-        r"\begin{tabular}{ccccc}"
+        r"\begin{tabular}{lcccc}"
     )
     # Manually modify the LaTeX code for centering multi-level headers
     latex_table = latex_table.replace(

### 2024-10-21 src/scripts/make_source_table.py

@@ -14,14 +14,14 @@ from chembench.analysis import (
 )
 from utils import obtain_chembench_repo
 
-from paths import tex, data
+from paths import tex, data, output
 
 def craft_table(df):
     latex_table = df.to_latex(index=False)
 
     latex_table = latex_table.replace(
         r"\begin{tabular}{lr}",
-        r"\begin{tabular}{cc}"
+        r"\begin{tabular}{lc}"
     )
 
     with open(output_file, 'w') as f:
@@ -52,7 +52,7 @@ def collect_data(datafolder):
     if df.empty:
         raise ValueError("No data loaded from reports")
     logger.info(f"Loaded {len(df)} rows of data")
-    
+
     df["semiautomatically"] = df.apply(is_semiautomatically_generated, axis=1)
 
     sources = {}
@@ -120,7 +120,7 @@ if __name__ == "__main__":
     )
     output_file = (
         os.path.join(
-            tex,
+            output,
             "sources_table.tex"
         )
     )

### 2024-10-21 src/tex/ms.tex
add Bloom taxonomy ref

@@
 However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
+
 These findings reveal LLMs' impressive chemical capabilities while emphasizing the need for further research to improve their safety and usefulness.
@@
 
-Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language.}\autocite{White_2023}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions,\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
 In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text, there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.\autocite{miret2024llms}
-For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data.
+For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists interpreting the data.
 Many of these insights are in the form of text in scientific publications.
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
-Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce.
+Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
-Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs compiled from diverse sources (\variable{output/manually_generated.txt}\xspace manually generated, and \variable{output/automatically_generated.txt}\xspace automatically generated).
+Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question-answer pairs compiled from diverse sources (\variable{output/manually_generated.txt}\xspace manually generated, and \variable{output/automatically_generated.txt}\xspace semi-automatically generated).
 Our corpus measures reasoning, knowledge and intuition across a large fraction of the topics taught in undergraduate and graduate chemistry curricula.
@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}

### 2024-10-21 src/tex/ms.tex
minor tuning

@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
@@
 To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
-While it is impossible to fully represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
-We also used this subset for the human baseline study.
+While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
+Our human participants answered all the questions in this subset.
 
@@
 Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
-In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
+In those cases, the \gls{llm} that returns the probabilities for various tokens, i.e., text fragments, is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
 The text completions, however, are the system's final outputs, which would also be used in a real-world application.
@@
 
-Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}).
+Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts.
 Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo.
-This is likely because the required knowledge cannot easily be accessed via papers but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
+This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
 This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
@@
 In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
-This is in line with observations in other domains but also indicates that chemical \gls{llm} could be further improved by scaling them up.
+This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.
 
@@
 Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+
  These findings also shine an interesting light on the value of textbook-inspired questions.
@@
 \paragraph{Models}
-For all models we performed inference using greedy decoding (i.e., temperature 0). We used the \gls{api} endpoints provided by the model developers and the ones provided by Groq.
+For all models we performed inference using greedy decoding (i.e., temperature 0).
+We used the \gls{api} endpoints provided by the model developers and the ones provided by Groq.
+\PaperQATwo was used (in August 2024) via an \gls{api} provided by FutureHouse.
 

### 2024-10-21 src/tex/appendix.tex
add Bloom taxonomy ref
polishing

--- 2024-10-21 src/tex/appendix.tex
@@
         Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By measuring atomic force microscopy \\
-        & & B. By measuring fourier transform infrared spectroscopy \\
-        & & C. By measuring photoluminescence spectrum \\
-        & & D. By measuring the absorption spectroscopy \\
+        & & A. By atomic force microscopy \\
+        & & B. By transform infrared spectroscopy \\
+        & & C. By measuring a photoluminescence spectrum \\
+        & & D. By absorption spectroscopy \\
         \midrule
@@
 \Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.
-Note that this ranking differs from the one on the \enquote{tiny} subset
 
@@
 
-\begin{figure}
+\begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is correlated with the number of atoms in the molecule.
-    For reasoning-based inference, one would expect that the number of atoms in the molecules is not necessarily a good predictor, and certainly worse than complexity measures, of the difficulty of the question.}
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals).}
     \script{correlate_with_molecule_features.py}
@@
 
-\begin{figure}
+\begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/correlation_plot_is_electron_counts_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is correlated with the number of atoms in the molecule.}
+    \caption{\textbf{Dependence of the mean absolute error in predicting total electron counts on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the total electron counts) is sometimes correlated with the number of atoms in the molecule.}
     \script{correlate_with_molecule_features.py}
@@
 The performance is measured as the fraction of questions answered correctly by the models.
-We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.}
+We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.}
     \script{performance_vs_model_size.py}
@@
 
-\clearpage
 \subsection{LLM Parsing} \label{sec:llm-parsing}
@@
     \centering
-    \caption{\textbf{Refusal counts and parsing.} The table shows the number of refusals detected and the number of times the fallback mechanism was used for each model.}
+    \caption{\textbf{Refusal counts and parsing.} The table shows the number of refusals detected and the number of times the \gls{llm} fallback parsing mechanism was used for each model.}
     \variable{output/model_refusal_table.tex}
@@
 
-In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax.
+In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
+\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}.
+
+\begin{figure}
+    \subfloat[A physcial chemistry question.]{
+        \includegraphics[width=.45\textwidth]{figures/screenshot_a.png}
+    }
+    \subfloat[An organic chemistry question.]{
+        \includegraphics[width=.45\textwidth]{figures/screenshot_b.png}
+    }
+    \caption{\textbf{Examples of how questions were shown to the human participants.}}
+    \label{fig:screenshots}
+\end{figure}
 
@@
 While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
-\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
+\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was weakly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).
 
@@
 \GPTFourO demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers.
-However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately 80\% of the time.
+However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately \SI{80}{\percent} of the time.
 In contrast, \LlamaThreeOneEightBInstruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5.
@@
 The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.
+The leaderboard is available at \url{https://lamalab-org.github.io/chem-bench/leaderboard/}.
 

### 2024-10-21 src/tex/ms.tex
polishing

@@
  For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
- While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.
-The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
+ While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
 
@@
  This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
- For instance, models might refuse to provide answers about cyanides. Statistics of the frequency of such refusals are shown in \Cref{tab:refusals}.
+ For instance, models might refuse to provide answers about cyanides. Statistics of the frequency of such refusals are shown in \Cref{tab:refusal_counts_and_parsing}.
  To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
@@
 
-One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman, Argyle_2023}
+One interesting finding of recent research is that foundation models can judge interestingness or human preferences in some domains.\autocite{zhang2024omniopenendednessmodelshuman, Argyle_2023}
 If models could do so for chemical compounds, this would open opportunities for novel optimization approaches.
 Such open-ended tasks, however, depend on an external observer defining what interestingness is.\autocite{hughes2024openendednessessentialartificialsuperhuman}
-Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting) (see \Cref{tab:chembench_corpus_cognitive}).
-While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench (\Cref{tab:performance_table_human_subset})
+Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting, see \Cref{tab:chembench_corpus_cognitive} for an example).
+Despite chemists demonstrating a reasonable level of interrater agreement, our models largely fail to align with expert chemists' preferences.
+Their performance is often indistinguishable from random guessing, even though these same models excel in various other tasks in \chembench (\Cref{tab:performance_table_human_subset}).
 This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.
@@
 A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.
-     We find that the models are not well-calibrated and some provide misleading confidence estimates.
+     We find that most models are not well-calibrated and provide misleading confidence estimates.
      }
@@
  In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
- For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
+ For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
  For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
@@
 On top of that, many models are not able to reliably estimate their own limitations.
-Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
+Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
 For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
 Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
-Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans.
+Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans (when trained on the right training corpus).
 
@@
 \subsection{Curation workflow}\label{sec:curation}
-For our dataset, we curated questions from existing exams or exercise sheets but also programmatically created new questions.
+For our dataset, we curated questions from existing exams or exercise sheets but also programmatically created new questions (\Cref{tab:sources}).
 Questions were added via Pull Requests on our GitHub repository and only merged into the corpus after passing manual review (\Cref{fig:curation_workflow}) as well as automated checks (e.g., for compliance with a standardized schema).
@@
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
+
+\begin{table}
+    \centering
+    \caption{\textbf{Overview of sources of the curated questions}.
+The table provides an overview of the types of sources the questions have been curated from.
+Detailed sources are available in the source data on GitHub.
+Questions without source have been curated completely from scratch.
+Questions based on lecture notes or URL have been curated based on content presented in those resources.
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
+    \variable{output/sources_table.tex}
+    \label{tab:sources}
+\end{table}
 
@@
 
-\begin{table}
-    \centering
-    \caption{\textbf{Overview of sources of the curated questions}.
-The table provides an overview of the types of sources the questions have been curated from.
-Detailed sources are available in the source data.
-Questions without source have been curated completely from scratch.
-Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
-    \variable{output/sources_table.tex}
-    \label{tab:sources}
-\end{table}
-
 \paragraph{Semi-programmatically generated questions}
@@
 
-\subparagraph{Chemical preference data}
+\paragraph{Chemical preference data}
 
@@
 
-To understand whether the capabilities of the leading models align with the preferences of professional chemists, we randomly selected 1000 data points from the original dataset to create a meaningful evaluation set, where molecules are represented as SMILES.
-To ablate the effect of different molecular representations, we only considered questions for which we could obtain IUPAC names for both molecules present.
+To understand whether the capabilities of the leading models align with the preferences of professional chemists, we randomly selected 1000 data points from the original dataset to create a meaningful evaluation set, where molecules are represented as \gls{smiles}.
+To ablate the effect of different molecular representations, we only considered questions for which we could obtain \gls{iupac} names for both molecules present.
 
 \subsection{Model evaluation workflow}
-An overview of the pipeline is shown in \Cref{fig:process}.
+A graphical overview of the pipeline is shown in \Cref{fig:process}.
 
@@
 In addition, this approach captures semantic uncertainty which is not the same as the probability of a token given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).
+On top of that, many propietary models do not provide access to the logits, making this approach more general.
 In \Cref{sec:confidence_estimates} we provide more details and comparisons with a logit-based approach.
@@
 \paragraph{Study design}
-Human participants were asked the questions in a custom-built web interface (\Cref{sec:human_baseline}) which rendered chemicals and equations.
+Human participants were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
 Questions were shown in random order and participants were not allowed to skip questions.
@@
 \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.
 
 \paragraph{Comparison with models}
-For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans.
+For the analysis, we treated each human as a model.
+We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans.
+The performance metrics reported for models in the main text are computed on the same questions the humans answered.
+Metrics for the entire corpus are reported in the appendix.
 
@@
 K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
+K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
 Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.

### 2024-10-24 src/tex/ms.tex
done

@@
 
-\begin{table}
+\begin{table}[!h]
     \centering
@@
 Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
+A.A.\ gratefully acknowledges financial support for this (publication/project/research) by the Fulbright U.S.
+Student Program, which is sponsored by the U.S.
+Department of State and German-American Fulbright Commission.
+Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
 

### 2024-10-24 src/tex/appendix.tex
nawafs figure
feat: incorporate feedback from Macjonathan

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/cb_humanset_v3.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V1.pdf}
     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
 \Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
-We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.
+We see that the performance of the models correlates with their size for the models of the LLama-3 and Llama-3.1 herd of models.
 
@@
 In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
-\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}.
+\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).
 
@@
     \subfloat[A physcial chemistry question.]{
-        \includegraphics[width=.45\textwidth]{figures/screenshot_a.png}
+        \includegraphics[width=\textwidth]{figures/screenshot_a.png}
     }
+
     \subfloat[An organic chemistry question.]{
-        \includegraphics[width=.45\textwidth]{figures/screenshot_b.png}
+        \includegraphics[width=\textwidth]{figures/screenshot_b.png}
     }

### 2024-10-24 src/tex/ms.tex
feat: incorporate feedback from Macjonathan

@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
@@
 Many other models also outperform the average human performance.
-Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
+Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
 
@@
 In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
-Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string ofa a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 
@@
 \section*{Data and code availability}
-The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench}.
+The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench}.
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
@@
 Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
-A.A.\ gratefully acknowledges financial support for this (publication/project/research) by the Fulbright U.S.
-Student Program, which is sponsored by the U.S.
-Department of State and German-American Fulbright Commission.
+A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S.
+Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission.
 Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
@@
 
-M.R.\ and U.S.S.\ thanks the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
+M.R.\ and U.S.S.\ thank the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
 

### 2024-10-24 src/tex/ms.tex
fix typo

@@
 In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
-Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string ofa a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 

### 2024-10-24 src/tex/ms.tex
address humans as volunteers

--- 2024-10-24 src/tex/ms.tex
@@
 
-We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists (volunteers) in our study on average.
 However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
@@
 However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.
+Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human volunteers) chemists.
 
@@
 
-To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
+To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human volunteers) chemists of different specializations.
 In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
@@
     We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework.
-To provide a baseline, we built a web application that we used to survey experts in chemistry.
+To provide a baseline, we built a web application that we used to survey experts (human volunteers) in chemistry.
     The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the worst, best, and average performance of the experts (human volunteers) in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
@@
 For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
-We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
+We then computed the percentage of questions the models or experts (human volunteers) answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
 In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly).
@@
  For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
- While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
- For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
+ For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 On the other hand, there are still striking limitations.
-For very relevant topics the answers models provide are wrong.
+For very relevant topics, the answers that models provide are wrong.
 On top of that, many models are not able to reliably estimate their own limitations.

### 2024-10-24 src/tex/appendix.tex
also address micha's comments

@@
 The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
-However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
 
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
 
@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.

### 2024-10-24 src/tex/ms.tex
also address micha's comments

--- 2024-10-24 src/tex/ms.tex
@@
     \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
-    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
+    The framework's foundation is the benchmark corpus that consists of thousands of questions and answers that we manually or semi-automatically compiled from various sources (see \Cref{sec:curation}).
     Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
@@
 
-\paragraph{\enquote{Tiny} subset}
+\paragraph{\chembenchmini}
 It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
-To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
+To address this, we also provide a subset (\chembenchmini, encompassing \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
@@
     \includegraphics{figures/human_subset_performance.pdf}
-    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly.
+    \caption{\textbf{Performance of models and humans on \chembenchmini.} The figure shows the percentage of questions that the models answered correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
@@
     \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+    \caption{\textbf{Performance of the models and humans on the different topics on \chembenchmini.} The radar plot shows the performance of the models and humans on the different topics of \chembenchmini.
 The performance is measured as the fraction of questions that were answered completely correctly by the models.
     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
     }

### 2024-10-24 src/tex/ms.tex
change participants to volunteers

--- 2024-10-24 src/tex/ms.tex
@@
 To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human volunteers) chemists of different specializations.
-In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
+In parts of the survey, the volunteers were also allowed to use tools such as web search to create a realistic setting.
 Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
@@
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
-Our human participants answered all the questions in this subset.
+Our human volunteers answered all the questions in this subset.
 
@@
 \paragraph{Study design}
-Human participants were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
-Questions were shown in random order and participants were not allowed to skip questions.
-For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
-Prior to answering questions, participants were asked to provide information about their education and experience in chemistry.
+Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
+Questions were shown in random order and volunteers were not allowed to skip questions.
+For a subset of the questions, the volunteers were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
+Prior to answering questions, volunteers were asked to provide information about their education and experience in chemistry.
 The study was conducted in English.
 
-\paragraph{Participants}
+\paragraph{Volunteers}
 Users were open to reporting about their experience in chemistry.
@@
 \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
-For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.
+For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.
 
@@
 The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe).
-Informed consent was obtained from all participants.
+Informed consent was obtained from all volunteers.
 

### 2024-10-24 src/tex/appendix.tex
finish michas comments

@@
 Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition.
+    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, and intuition.
     \item \emph{Cover a range of difficulty levels}.
@@
 \clearpage
-\subsection{Model performance}
+\subsection{Model performance} \label{sec:model_performance_app}
 We also evaluated the model performance on the entire \chembench corpus.
@@
 The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
     }
@@
     \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals).}
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} }
     \script{correlate_with_molecule_features.py}
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
+A score of one would mean that a volunteer answered all questions correctly.
+A score of zero would mean that no question was answered correctly.}
     \label{fig:human_score_distribution}
@@
 \subsection{Impact of sampling temperature}
-We also investigated the impact of sampling temperature on the performance of the models.
-\Cref{fig:temperature_impact} shows that, generally, the performance of models tends to degrease with increasing temperature.
+We also investigated the impact of sampling temperature (i.e.
+temperature 0 means always sampling the most likely token, higher temperatures introduce some randomness in the generation process) on the performance of the models.
+\Cref{fig:temperature_impact} shows that, generally, the performance of models tends to decrease with increasing temperature.
 

### 2024-10-24 src/scripts/plot_temperature_diffs.py

@@ -63,8 +63,8 @@ def compute_scores_subset(data, title):
     sns.lineplot(x="Temperature", y="Score", hue="Model", data=df_melted,
         palette=color_palette, markers="", ci=None, legend=False)
 
-    plt.xlabel("Temperature")
-    plt.ylabel("Score")
+    plt.xlabel("temperature")
+    plt.ylabel("score")
     # Place the legend outside the plot to the right
     plt.legend(loc="center left", bbox_to_anchor=(1, 0.5), fontsize="small")
 

### 2024-10-24 src/tex/ms.tex
fix suggested changes

@@
 
-To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human volunteers) chemists of different specializations.
+To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
 In parts of the survey, the volunteers were also allowed to use tools such as web search to create a realistic setting.
-Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
-We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
 
@@
     We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework.
-To provide a baseline, we built a web application that we used to survey experts (human volunteers) in chemistry.
+To provide a baseline, we built a web application that we used to survey experts in chemistry.
     The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
@@
 For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
-We then computed the percentage of questions the models or experts (human volunteers) answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
+We then computed the percentage of questions the models or experts answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
 In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly).

### 2024-10-24 src/tex/appendix.tex
suggestions from gil

--- 2024-10-24 src/tex/appendix.tex
@@
     \includegraphics{figures/overall_performance.pdf}
-    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
+    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered correctly by the models.
 Scores computed on the entire \chembench corpus.}
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+The performance is measured as the fraction of questions answered correctly by the models.
+    A score of 1 indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
     }
@@
     \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none of the questions were answered correctly.
         We see that the performance of the models varies significantly between the different data sources.
@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
         We see that the performance of the models varies significantly between the different data sources.
@@
 \Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
-We see that the performance of the models correlates with their size for the models of the LLama-3 and Llama-3.1 herd of models.
+We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.
 
@@
 In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
-\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).
+\LaTeX\xspace equations and chemical equations were rendered using MathJax \Cref{fig:screenshots}.
 
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
@@
 
-We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
+We also recorded the time humans took to answer the questions \cref{fig:human_timing}. This time is the time from the question being displayed to the human to the human submitting the answer.
 
@@
 \subsection{Leaderboard}
+\label{sec:leaderboard}
 Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}

### 2024-10-24 src/tex/ms.tex
suggestions from gil

@@
 
-We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists (volunteers) in our study on average.
+We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
 However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
-Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
+Still, it is essential to realize that LLMs' user base is wider than that of chemistry and materials science experts who can critically reflect on every output these models produce.
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
-Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human volunteers) chemists.
+Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.
 
@@
 This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
-On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.
+On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts \autocite{wu2018moleculenet}.
 
@@
 To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
-In parts of the survey, the volunteers were also allowed to use tools such as web search to create a realistic setting.
+In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
+Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
+We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
 
@@
 To provide a baseline, we built a web application that we used to survey experts in chemistry.
-    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
+    The results of the evaluations are then compiled in publicly accessible leaderboards \ref{sec:leaderboard}, which we propose as a foundation for evaluating future models.
     }
@@
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
-Our human volunteers answered all the questions in this subset.
+Our human participants answered all the questions in this subset.
 
@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the experts (human volunteers) in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
@@
 This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
-This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
+This indicates that there is still room for improvement in chemical \gls{llm}s by training them on more specialized data sources or by integrating them with specialized databases.
 
 In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
-This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.
+This is in line with observations in other domains but also indicates that chemical \gls{llm}s could, to some extent, be further improved by scaling them up.
 
@@
 For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
-We then computed the percentage of questions the models or experts answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
+We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
 In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly).
@@
     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+The performance is measured as the fraction of questions that were answered correctly by the models.
     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
     }
@@
  For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
- While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
- For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
+ For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 On the other hand, there are still striking limitations.
-For very relevant topics, the answers that models provide are wrong.
+For very relevant topics the answers models provide are wrong.
 On top of that, many models are not able to reliably estimate their own limitations.
@@
 \paragraph{Study design}
-Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
-Questions were shown in random order and volunteers were not allowed to skip questions.
-For a subset of the questions, the volunteers were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
-Prior to answering questions, volunteers were asked to provide information about their education and experience in chemistry.
+Human participants were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
+Questions were shown in random order and participants were not allowed to skip questions.
+For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
+Prior to answering questions, participants were asked to provide information about their education and experience in chemistry.
 The study was conducted in English.
 
-\paragraph{Volunteers}
+\paragraph{Participants}
 Users were open to reporting about their experience in chemistry.
@@
 \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
-For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.
+For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.
 
@@
 The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe).
-Informed consent was obtained from all volunteers.
+Informed consent was obtained from all participants.
 

### 2024-10-24 src/tex/ms.tex
feat: incorporate feedback from Macjonathan and Micha (#66)

@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
-Still, it is essential to realize that LLMs' user base is wider than that of chemistry and materials science experts who can critically reflect on every output these models produce.
+Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
 This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
-On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts \autocite{wu2018moleculenet}.
+On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.
 
@@
 To provide a baseline, we built a web application that we used to survey experts in chemistry.
-    The results of the evaluations are then compiled in publicly accessible leaderboards \ref{sec:leaderboard}, which we propose as a foundation for evaluating future models.
+    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
     }
@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
 
-\paragraph{\enquote{Tiny} subset}
+\paragraph{\chembenchmini}
 It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
-To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
+To address this, we also provide a subset (\chembenchmini, encompassing \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
@@
 This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
-This indicates that there is still room for improvement in chemical \gls{llm}s by training them on more specialized data sources or by integrating them with specialized databases.
+This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
 
 In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
-This is in line with observations in other domains but also indicates that chemical \gls{llm}s could, to some extent, be further improved by scaling them up.
+This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.
 
@@
     \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
-The performance is measured as the fraction of questions that were answered correctly by the models.
+    \caption{\textbf{Performance of the models and humans on the different topics on \chembenchmini.} The radar plot shows the performance of the models and humans on the different topics of \chembenchmini.
+The performance is measured as the fraction of questions that were answered completely correctly by the models.
     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
+    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
     }
@@
 
-In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
-Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 
  These findings also shine an interesting light on the value of textbook-inspired questions.
- A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
+ A subset of the questions in \chembench are based on textbooks targeted at undergraduate students.
  On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
@@
 \paragraph{Confidence estimates} One might wonder whether the models can estimate if they can answer a question correctly.
-If they could do so, incorrect answers would be less problematic as one could detect when an answer is incorrect.
+If they could do so, incorrect answers would be less problematic.
 
@@
  While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
- For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ For example, for the questions about the \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 These questions assess the ability to establish a \enquote{preference}, such as favoring a specific molecule.
-Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that lasts several years within a chemist's career.
+Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that takes several years within a chemist's career.
 Our data corpus is adapted from the published dataset by \textcite{Choung_2023}, which consists of more than 5000 question-answer pairs about chemical intuition.
@@
 Out of those,
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} has a bachelor's degree.
 For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.
@@
 The performance metrics reported for models in the main text are computed on the same questions the humans answered.
-Metrics for the entire corpus are reported in the appendix.
+Metrics for the entire corpus are reported in the appendix (\Cref{sec:model_performance_app}).
 
@@
 In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question.
-We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
+We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
 the H/P statements of chemicals).

### 2024-10-24 src/tex/ms.tex
Update src/tex/ms.tex

@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the experts (human volunteers) in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the worst, best, and average performance of the experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.

### 2024-10-24 src/tex/ms.tex
Update src/tex/ms.tex

@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
 
-\paragraph{\enquote{Tiny} subset}
+\paragraph{\chembenchmini}
 It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
-To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
+To address this, we also provide a subset (\chembenchmini, \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
@@
  While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
- For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ For example, for the questions about the \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 
-\paragraph{Volunteers}
+\paragraph{Human volunteers}
 Users were open to reporting about their experience in chemistry.
 Overall, \variable{output/num_users_with_education_info.txt} did so.
-Out of those,
-\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+Out of those, \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} has a bachelor's degree.
 For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.

### 2024-10-24 src/tex/appendix.tex
feat: plot chembench corpus and tiny subset

--- 2024-10-24 src/tex/appendix.tex
@@
 The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
-However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
 
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.
+
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \enquote{tiny} subset.
+The relative performance difference between both is actually quite similar across most models.
+This makes \enquote{tiny} subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
 
@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered completely correctly by the models.
@@
     \script{analyze_performance_per_source.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/corpus_human_comparison.pdf}
+    \caption{\textbf{
+        Performance of the models on the \chembench corpus and \enquote{tiny} subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \enquote{tiny} subset.
+        We see that the model ranking remains fairly consistent across both sets}  
+    }
+    \label{fig:performance_corpus_and_tiny}
+    \script{corpus_humanset_performance.py}
 \end{figure}
@@
 In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
-\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).
+\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}.
 

### 2024-10-25 src/tex/appendix.tex

--- 2024-10-24 src/tex/appendix.tex
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However smaller but well-designed models like can compete with larger models.
+For example \GemmaTwoNineBIt has an overall accuracy of 0.47 on \ChemBench corpus, which is only 18\% less accurate than the larger \LlamaThreeOneEightBInstruct model.
+Clear progression in model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
+Newer versions consistently outperform their predecessors indicating improvements in scientific understanding of these model families.
+Basic tasks are consistently solved across all models and even top model show a drop in performance on the advanced tasks except for \oone model.
+Scores of knowledge intensive are generally lower than both Calculation and Reasoning. Reasoning tends to be the strongest skill among three.
+
+Based on these findings some scope for improving these models could be to focus on enhancing knowledge-based training.
+Possibly through improved pre-training on chemistry-specific texts and or with chemistry knowledge base integration.
+Models show strongest performance in reasoning tasks, this suggests current training approaches are good at developing logical/reasoning capabilities.
+Larger models consistently perform better, However, there's evidence that well trained smaller models can compete with larger models.
+Performance drops significantly for advanced tasks, this suggests including more advanced chemistry problems with complex, multi-step solutions in training or finetuning.
+
+\clearpage
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-25 src/tex/appendix.tex

@@
 
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \enquote{tiny} subset.
+The relative performance difference between both is actually quite similar across most models.
+This makes \enquote{tiny} subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+
 \begin{figure}[htb]
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However smaller but well-designed models like can compete with larger models.
-For example \GemmaTwoNineBIt has an overall accuracy of 0.47 on \ChemBench corpus, which is only 18\% less accurate than the larger \LlamaThreeOneEightBInstruct model.
-Clear progression in model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
-Newer versions consistently outperform their predecessors indicating improvements in scientific understanding of these model families.
-Basic tasks are consistently solved across all models and even top model show a drop in performance on the advanced tasks except for \oone model.
-Scores of knowledge intensive are generally lower than both Calculation and Reasoning. Reasoning tends to be the strongest skill among three.
-
-Based on these findings some scope for improving these models could be to focus on enhancing knowledge-based training.
-Possibly through improved pre-training on chemistry-specific texts and or with chemistry knowledge base integration.
-Models show strongest performance in reasoning tasks, this suggests current training approaches are good at developing logical/reasoning capabilities.
-Larger models consistently perform better, However, there's evidence that well trained smaller models can compete with larger models.
-Performance drops significantly for advanced tasks, this suggests including more advanced chemistry problems with complex, multi-step solutions in training or finetuning.
-
-\clearpage
-
-\clearpage
 \subsection{Leaderboard}

### 2024-10-25 src/tex/appendix.tex

@@
 
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \enquote{tiny} subset.
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
 The relative performance difference between both is actually quite similar across most models.
-This makes \enquote{tiny} subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
 
@@
     \caption{\textbf{
-        Performance of the models on the \chembench corpus and \enquote{tiny} subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \enquote{tiny} subset.
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \chembenchmini subset.
         We see that the model ranking remains fairly consistent across both sets}  

### 2024-10-25 src/tex/appendix.tex
feat: incorporate feedback from Santiago

@@
 
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is actually quite similar across most models.
-This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
-
 \begin{figure}[htb]
@@
     \script{analyze_performance_per_source.py}
-\end{figure}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/corpus_human_comparison.pdf}
-    \caption{\textbf{
-        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \chembenchmini subset.
-        We see that the model ranking remains fairly consistent across both sets}  
-    }
-    \label{fig:performance_corpus_and_tiny}
-    \script{corpus_humanset_performance.py}
 \end{figure}
@@
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
-    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
+    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity).
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
@@
 A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.}
+A score of zero would mean that no question was answered correctly.
+    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
+    }
     \label{fig:human_score_distribution}

### 2024-10-25 src/tex/ms.tex
feat: incorporate feedback from Santiago

@@
 Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language.}\autocite{White_2023}
-This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions,\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
+This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions,\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform experiments in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}
 
@@
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
-     The line plot shows the average fraction of correctly answered questions for each confidence level.
-The bar plot shows the distribution of confidence estimates.
+     The line plot shows the average fraction of correctly answered questions for each confidence level. The bar plot shows the distribution of confidence estimates.
      A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.
+The dashed black line indicates this ideal behavior, which would be monotonically increasing correctness with higher
+     levels of confidence.
      We find that most models are not well-calibrated and provide misleading confidence estimates.
@@
 Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text.
-For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
+For example, all the \gls{smiles} representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
 Our prompting strategy consistently adheres to these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations, and physical units (by either adding or removing wrappers).
-This step can be easily customized in our codebase.
+This step can be easily customized in our codebase and we provide presets for the models we evaluated.
 
@@
 \section*{Acknowledgements}
-This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
-We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
+This work was supported by the Carl Zeiss Foundation, and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
+
+In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
+
+Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
+
 K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
-K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
-Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
+
+K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo runs via access to the \gls{api}.
+We also thank Stability.AI for the access to its HPC cluster.
+
 A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S.

### 2024-10-25 src/tex/ms.tex
clarify what we mean with systems

@@
 
-Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
+Since many widely utilized \gls{llm} systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
 This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems.

### 2024-10-25 src/tex/appendix.tex
remove unused text and reorder figures and tables according to Martino's suggestion
feat: add Nawaf's section
co-authored-by: n0w0f <pvt.nawaf@gmail.com>

@@
 This also implies that future baselines can be built on top of our infrastructure if saved in the same format.
+
+\subsubsection{Curation workflow}
+Questions were added via pull requests to the \chembench repository on GitHub.
+This allowed for a manual review of each question by expert reviewers (with backgrounds in chemistry, materials science, chemical engineering, and physics).
+The reviews were conducted directly on the GitHub platform, where our entire curation history is also publicly available.
+
+The general guidelines followed by the reviewer are the following:
+
+\begin{itemize}
+    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
+    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
+    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
+    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
+    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
+    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
+They should add value by expanding the breadth of the dataset.
+That is, questions already multiple (>10) times in the corpus in a similar form are rejected.
+\end{itemize}
+
+Reviewers also solved the questions to verify the answers.
+They also performed web searches to ensure questions were not easily found online.
+The reviewers often guided the revision process to ensure the question aligned with the guidelines.
+Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question.
+Reviewers also provide feedback on the skill and difficulty annotations.
+
+In addition to the manual review, we also performed automated checks to ensure the quality of the questions.
+The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
+
+\subsubsection{Composition}
 
@@
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
-    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity).
+    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules.
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
@@
 A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.
-    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
-    }
+A score of zero would mean that no question was answered correctly.}
     \label{fig:human_score_distribution}

### 2024-10-25 src/tex/appendix.tex
add missing comma
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-25 src/tex/appendix.tex
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of 0.47 on \ChemBench corpus, which is only 18\% less accurate than the larger \LlamaThreeOneEightBInstruct model.
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
+Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
+Scores of knowledge-intensive are generally lower than those of Calculation and Reasoning. Reasoning tends to be the strongest skill among the three.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+Models show the strongest performance in reasoning tasks, which suggests that current training approaches are good at developing logical/reasoning capabilities.
+Larger models consistently perform better.
+However, there's evidence that well-trained smaller models can compete with larger models.
+Performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
+
+\clearpage
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-25 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>
feat: include feedback from Santiago (#71)

--- 2024-10-25 src/tex/appendix.tex
@@
 \end{figure}
-
-%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
-%We see that the questions are generally complex to read.
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
-It is calculated based on the average number of syllables per word and words per sentence.
-The higher the score, the easier the text is to read.
-The distribution of the questions' scores is shown in the histogram.
-}
-%     \label{fig:flesch_kincaid_reading_ease}
-%     \script{wordcloud.py}
-% \end{figure}
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-%     \label{fig:question_count_barplot_mcq_vs_general}
-%     \script{plot_statistics.py}
-% \end{figure}
 
@@
 
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
-
 \begin{figure}[htb]
@@
 
-% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
-% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/model_overall_cal.pdf}
-%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-%     \script{requires_calculation_plot.py}
-%     \label{fig:requires_cal}
-% \end{figure}
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 \end{figure}
-
-% \begin{table}
-%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
-}
-%     \begin{tabularx}{\textwidth}{XXXX XXXX}
-%         \toprule
-%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\
-%         \midrule
-%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
-%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
-%         \midrule
-%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
-%         \midrule
-%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
-%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
-%         \bottomrule
-%     \end{tabularx}
-%     \label{tab:correlation_coefficients}
-% \end{table}
 
@@
 The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer.
-We used \LlamaThreeSeventyBInstruct accessed via the Groq API.
+We used \LlamaThreeSeventyBInstruct, accessed via the Groq API.
 \Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.

### 2024-10-25 src/tex/appendix.tex
chore: add additional observation

--- 2024-10-25 src/tex/appendix.tex
@@
 
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
+The relative performance difference between both is actually quite similar across most models.
+This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases. 
+An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+
 \begin{figure}[htb]
@@
     \script{analyze_performance_per_source.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/corpus_human_comparison.pdf}
+    \caption{\textbf{
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \chembenchmini subset.
+        We see that the model ranking remains fairly consistent across both sets}  
+    }
+    \label{fig:performance_corpus_and_tiny}
+    \script{corpus_humanset_performance.py}
 \end{figure}
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of 0.47 on \ChemBench corpus, which is only 18\% less accurate than the larger \LlamaThreeOneEightBInstruct model.
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
-Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
-Scores of knowledge-intensive are generally lower than those of Calculation and Reasoning. Reasoning tends to be the strongest skill among the three.
-
-Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-Models show the strongest performance in reasoning tasks, which suggests that current training approaches are good at developing logical/reasoning capabilities.
-Larger models consistently perform better.
-However, there's evidence that well-trained smaller models can compete with larger models.
-Performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
-
-\clearpage
-
-\clearpage
 \subsection{Leaderboard}

### 2024-10-25 src/scripts/compute_basic_statistics.py
fix: make source count consistent

@@ -23,13 +23,13 @@ def load_data():
         f.write(str(len(non_mcq_questions)) + "\endinput")
 
 
-    automatically_generated = df[df["is_semiautomatically_generated"]] 
+    automatically_generated = df[df["is_semiautomatically_generated"]]
 
-    with open(output / "automatically_generated.txt", "w") as f:    
+    with open(output / "automatically_generated.txt", "w") as f:
         f.write(str(len(automatically_generated)) + "\endinput")
 
-    manually_generated = df[~df["is_semiautomatically_generated"]]  
-    with open(output / "manually_generated.txt", "w") as f: 
+    manually_generated = df[~df["is_semiautomatically_generated"]]
+    with open(output / "manually_generated.txt", "w") as f:
         f.write(str(len(manually_generated)) + "\endinput")
 
     dai_data = glob(

### 2024-10-25 src/tex/appendix.tex
chore: update the format of the examples table

@@
 Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, and intuition.
+    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition.
     \item \emph{Cover a range of difficulty levels}.
@@
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-
+\begin{table}
+    \centering
+    \small
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Question Type & \% & Example Questions  \\
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%)} \\
             \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+            A. X-ray fluorescence analysis \\
+            B. Differential pulse polarography \\
+            C. Flame-atomic absorption spectroscopy \\
+            D. Gas chromatography with flame ionization detector \\
+            E. Hydride generation atomic absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
             \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \midrule
+            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \midrule
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+            A. \ce{F-} forms a stable bonds to the metal ions \\
+            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \midrule
+            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+            A. $E_1$ \\
+            B. $E_{cb}$ \\
+            C. $S_N1$ \\
+            D. $S_N2$ \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \midrule
+            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
+            A. \ce{C60} \\
+            B. \ce{CH4} \\
+            C. \ce{Fe(CO)5} \\
+            D. \ce{H2+} \\
+            E. \ce{NaCl} \\
             \midrule
-        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \midrule
+            Which of the following statements is true about the different types of ideal reactors? \\
+            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \midrule
+            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-        & & A. Pindolol has a greater therapeutic range than propranolol \\
-        & & B. Pindolol has a longer half-life than propranolol \\
-        & & C. Pindolol has intrinsic activity \\
-        & & D. Pindolol is more lipophilic than propranolol \\
-        & & E. Pindolol is more potent than propranolol \\
-        \midrule
-        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-        & & A. X-ray fluorescence analysis \\
-        & & B. Differential pulse polarography \\
-        & & C. Flame-atomic absorption spectroscopy \\
-        & & D. Gas chromatography with flame ionization detector \\
-        & & E. Hydride generation atomic absorption spectroscopy \\
-        \midrule
-        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \ce{HF} over \ce{HCl} for this purpose?
-\\
-        & & A. \ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
-        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
-        \midrule
-        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
-        \midrule
-        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
-\\
-        & & A. \ce{C60} \\
-        & & B. \ce{CH4} \\
-        & & C. \ce{Fe(CO)5} \\
-        & & D. \ce{H2+} \\
-        & & E. \ce{NaCl} \\
-        \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            A. Pindolol has a greater therapeutic range than propranolol \\
+            B. Pindolol has a longer half-life than propranolol \\
+            C. Pindolol has intrinsic activity \\
+            D. Pindolol is more lipophilic than propranolol \\
+            E. Pindolol is more potent than propranolol \\
             \bottomrule
+        \end{tabularx}
+    }
+    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
     \label{tab:chembench_corpus_topic}
-\end{xltabular}
+\end{table}
 
@@
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
-
+\begin{table}
+    \centering
+    \small
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Required Skills & Number of Questions & Example Question \\
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
             \midrule
-        Knowledge & 919 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
             \midrule
-        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By atomic force microscopy \\
-        & & B. By transform infrared spectroscopy \\
-        & & C. By measuring a photoluminescence spectrum \\
-        & & D. By absorption spectroscopy \\
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
             \midrule
-        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+            A. By atomic force microscopy \\
+            B. By transform infrared spectroscopy \\
+            C. By measuring a photoluminescence spectrum \\
+            D. By absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
+            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
             \midrule
-        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
@@
             \bottomrule
-
+        \end{tabularx}
+    }
+    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
+Note that the total count in this table is bigger than the \chembench corpus.
+This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
+Reasoning and Calculation.} \\
     \label{tab:chembench_corpus_cognitive}
-\end{xltabular}
+\end{table}
 
@@
 \clearpage
-\subsection{Model performance} \label{sec:model_performance_app}
+\subsection{Model performance}
 We also evaluated the model performance on the entire \chembench corpus.
@@
 The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
     }
@@
 
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
-
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is actually quite similar across most models.
-This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases. 
-An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.
 
@@
     \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
-    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} }
+    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals).}
     \script{correlate_with_molecule_features.py}
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
-A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.}
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
     \label{fig:human_score_distribution}
@@
 \subsection{Impact of sampling temperature}
-We also investigated the impact of sampling temperature (i.e.
-temperature 0 means always sampling the most likely token, higher temperatures introduce some randomness in the generation process) on the performance of the models.
-\Cref{fig:temperature_impact} shows that, generally, the performance of models tends to decrease with increasing temperature.
+We also investigated the impact of sampling temperature on the performance of the models.
+\Cref{fig:temperature_impact} shows that, generally, the performance of models tends to degrease with increasing temperature.
 

### 2024-10-25 src/tex/appendix.tex
fix suggestions from kevin

--- 2024-10-25 src/tex/appendix.tex
@@
 The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
-However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
 
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
     \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+The performance is measured as the fraction of questions answered correctly by the models.
+    A score of 1 indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
     }
@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
         We see that the performance of the models varies significantly between the different data sources.
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
-A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.}
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
     \label{fig:human_score_distribution}
@@
 
-We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
+We also recorded the time humans took to answer the questions (\Cref{fig:human_timing}). This time is the time from the question being displayed to the human to the human submitting the answer.
 

### 2024-10-25 src/tex/ms.tex
fix suggestions from kevin

--- 2024-10-24 src/tex/ms.tex
@@
 To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
-Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
+Still, it is essential to realize that the user base of \glspl{llm} is broader than that of chemistry and materials science experts who can critically reflect on every output these models produce.
 For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
@@
     \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
-    The framework's foundation is the benchmark corpus that consists of thousands of questions and answers that we manually or semi-automatically compiled from various sources (see \Cref{sec:curation}).
+    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
     Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
@@
 To provide a baseline, we built a web application that we used to survey experts in chemistry.
-    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
+    The results of the evaluations are then compiled in publicly accessible leaderboards (\Cref{sec:leaderboard}), which we propose as a foundation for evaluating future models.
     }
@@
 For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).
-In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
+In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations).
 We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}
 
-\paragraph{\chembenchmini}
+\paragraph{\enquote{Tiny} subset}
 It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
 For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
-To address this, we also provide a subset (\chembenchmini, \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
+To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus.
 While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
-Our human volunteers answered all the questions in this subset.
+Our human participants answered all the questions in this subset.
 
@@
     \includegraphics{figures/human_subset_performance.pdf}
-    \caption{\textbf{Performance of models and humans on \chembenchmini.} The figure shows the percentage of questions that the models answered correctly.
+    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly.
 We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
@@
 In this figure, we show the percentage of questions that the models answered correctly.
-Moreover, we show the worst, best, and average performance of the experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
+Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
 Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
 Many other models also outperform the average human performance.
-Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
+Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
 
@@
 This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
-This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
+This indicates that there is still room for improvement in chemical \glspl{llm} by training them on more specialized data sources or by integrating them with specialized databases.
 
 In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
-This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.
+This is in line with observations in other domains but also indicates that chemical \glspl{llm} could, to some extent, be further improved by scaling them up.
 
@@
     \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
-    \caption{\textbf{Performance of the models and humans on the different topics on \chembenchmini.} The radar plot shows the performance of the models and humans on the different topics of \chembenchmini.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus.
+The performance is measured as the fraction of questions that were answered correctly by the models.
     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
     }
@@
 
-In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
-Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 
  These findings also shine an interesting light on the value of textbook-inspired questions.
- A subset of the questions in \chembench are based on textbooks targeted at undergraduate students.
+ A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
  On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
@@
 \paragraph{Confidence estimates} One might wonder whether the models can estimate if they can answer a question correctly.
-If they could do so, incorrect answers would be less problematic.
+If they could do so, incorrect answers would be less problematic as one could detect when an answer is incorrect.
 
@@
  For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
- While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
- For example, for the questions about the \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
+ While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
+ For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.
 
@@
 These questions assess the ability to establish a \enquote{preference}, such as favoring a specific molecule.
-Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that takes several years within a chemist's career.
+Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that lasts several years within a chemist's career.
 Our data corpus is adapted from the published dataset by \textcite{Choung_2023}, which consists of more than 5000 question-answer pairs about chemical intuition.
@@
 \paragraph{Study design}
-Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
-Questions were shown in random order and volunteers were not allowed to skip questions.
-For a subset of the questions, the volunteers were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
-Prior to answering questions, volunteers were asked to provide information about their education and experience in chemistry.
+Human participants were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
+Questions were shown in random order and participants were not allowed to skip questions.
+For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
+Prior to answering questions, participants were asked to provide information about their education and experience in chemistry.
 The study was conducted in English.
 
-\paragraph{Human volunteers}
+\paragraph{Participants}
 Users were open to reporting about their experience in chemistry.
 Overall, \variable{output/num_users_with_education_info.txt} did so.
-Out of those, \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} has a bachelor's degree.
-For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.
+Out of those,
+\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
+For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.
 
@@
 The performance metrics reported for models in the main text are computed on the same questions the humans answered.
-Metrics for the entire corpus are reported in the appendix (\Cref{sec:model_performance_app}).
+Metrics for the entire corpus are reported in the appendix.
 
@@
 In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question.
-We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
+We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
 the H/P statements of chemicals).
@@
 \section*{Data and code availability}
-The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench}.
+The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench}.
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
@@
 Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
-A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S.
-Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission.
+A.A.\ gratefully acknowledges financial support for this (publication/project/research) by the Fulbright U.S.
+Student Program, which is sponsored by the U.S.
+Department of State and German-American Fulbright Commission.
 Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
@@
 
-M.R.\ and U.S.S.\ thank the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
+M.R.\ and U.S.S.\ thanks the \enquote{Deutsche Forschungsgemeinschaft} for funding under the regime of the priority programme SPP 2363 \enquote{Utilization and Development of Machine Learning for Molecular Applications – Molecular Machine Learning} (SCHU 1229/63-1; project number 497115849).
 

### 2024-10-25 src/tex/appendix.tex
Update src/tex/appendix.tex
feat: plot chembench corpus and tiny subset (#69)
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-25 src/tex/appendix.tex
@@
 
+\subsubsection{Curation workflow}
+Questions were added via pull requests to the \chembench repository on GitHub.
+This allowed for a manual review of each question by expert reviewers (with backgrounds in chemistry, materials science, chemical engineering, and physics).
+The reviews were conducted directly on the GitHub platform, where our entire curation history is also publicly available.
+
+The general guidelines followed by the reviewer are the following:
+
+\begin{itemize}
+    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
+    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
+    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
+    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
+    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
+    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
+They should add value by expanding the breadth of the dataset.
+That is, questions already multiple (>10) times in the corpus in a similar form are rejected.
+\end{itemize}
+
+Reviewers also solved the questions to verify the answers.
+They also performed web searches to ensure questions were not easily found online.
+The reviewers often guided the revision process to ensure the question aligned with the guidelines.
+Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question.
+Reviewers also provide feedback on the skill and difficulty annotations.
+
+In addition to the manual review, we also performed automated checks to ensure the quality of the questions.
+The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
+
+\subsubsection{Composition}
+
 \Cref{fig:cb_humanset} shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
 \Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is actually quite similar across most models.
+The relative performance difference between both is quite similar across most models.
 This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+
 An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.

### 2024-10-25 src/tex/appendix.tex
fix latex syntax
feat: skillset distribution plot

@@
 
-In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+The plot shows an even distribution of the required skills in the \chembench corpus.
+\Cref{tab:cb_skillset} shows an example question for each skill category.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
+    \caption{\textbf{Composition of the required skills considered in the \ChemBench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
+    \label{fig:cb_skillset}
+\end{figure}
 

### 2024-10-25 src/tex/appendix.tex
fix: remove non-model-scores from plot (#75)

--- 2024-10-25 src/tex/appendix.tex
@@
     \caption{\textbf{
-        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \chembenchmini subset.
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
         We see that the model ranking remains fairly consistent across both sets}  
-    }
     \label{fig:performance_corpus_and_tiny}

### 2024-10-25 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-25 src/tex/appendix.tex
@@
 In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
-The plot shows an even distribution of the required skills in the \chembench corpus.
+The plot shows a balanced distribution of the required skills in the \chembench corpus.
 \Cref{tab:cb_skillset} shows an example question for each skill category.
@@
     \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
-    \caption{\textbf{Composition of the required skills considered in the \ChemBench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
+    \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
     \label{fig:cb_skillset}

### 2024-10-25 src/tex/appendix.tex
add table for models struggle to answer

--- 2024-10-25 src/tex/appendix.tex
@@
 
-\subsubsection{Curation workflow}
-Questions were added via pull requests to the \chembench repository on GitHub.
-This allowed for a manual review of each question by expert reviewers (with backgrounds in chemistry, materials science, chemical engineering, and physics).
-The reviews were conducted directly on the GitHub platform, where our entire curation history is also publicly available.
-
-The general guidelines followed by the reviewer are the following:
-
-\begin{itemize}
-    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
-    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
-    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
-    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
-    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
-    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
-They should add value by expanding the breadth of the dataset.
-That is, questions already multiple (>10) times in the corpus in a similar form are rejected.
-\end{itemize}
-
-Reviewers also solved the questions to verify the answers.
-They also performed web searches to ensure questions were not easily found online.
-The reviewers often guided the revision process to ensure the question aligned with the guidelines.
-Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question.
-Reviewers also provide feedback on the skill and difficulty annotations.
-
-In addition to the manual review, we also performed automated checks to ensure the quality of the questions.
-The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
-
-\subsubsection{Composition}
-
 \Cref{fig:cb_humanset} shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
 
+%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
+%We see that the questions are generally complex to read.
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+It is calculated based on the average number of syllables per word and words per sentence.
+The higher the score, the easier the text is to read.
+The distribution of the questions' scores is shown in the histogram.
+}
+%     \label{fig:flesch_kincaid_reading_ease}
+%     \script{wordcloud.py}
+% \end{figure}
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+%     \label{fig:question_count_barplot_mcq_vs_general}
+%     \script{plot_statistics.py}
+% \end{figure}
+
 The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+\begin{table}
+    \centering
+    \small
     \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-
+    \label{tab:chembench_corpus_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Question Type & \% & Example Questions  \\
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%)} \\
             \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+            A. X-ray fluorescence analysis \\
+            B. Differential pulse polarography \\
+            C. Flame-atomic absorption spectroscopy \\
+            D. Gas chromatography with flame ionization detector \\
+            E. Hydride generation atomic absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
             \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \midrule
+            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \midrule
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+            A. \ce{F-} forms a stable bonds to the metal ions \\
+            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \midrule
+            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+            A. $E_1$ \\
+            B. $E_{cb}$ \\
+            C. $S_N1$ \\
+            D. $S_N2$ \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \midrule
+            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
+            A. \ce{C60} \\
+            B. \ce{CH4} \\
+            C. \ce{Fe(CO)5} \\
+            D. \ce{H2+} \\
+            E. \ce{NaCl} \\
             \midrule
-        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \midrule
+            Which of the following statements is true about the different types of ideal reactors? \\
+            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \midrule
+            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-        & & A. Pindolol has a greater therapeutic range than propranolol \\
-        & & B. Pindolol has a longer half-life than propranolol \\
-        & & C. Pindolol has intrinsic activity \\
-        & & D. Pindolol is more lipophilic than propranolol \\
-        & & E. Pindolol is more potent than propranolol \\
-        \midrule
-        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-        & & A. X-ray fluorescence analysis \\
-        & & B. Differential pulse polarography \\
-        & & C. Flame-atomic absorption spectroscopy \\
-        & & D. Gas chromatography with flame ionization detector \\
-        & & E. Hydride generation atomic absorption spectroscopy \\
-        \midrule
-        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
-Why would one choose \ce{HF} over \ce{HCl} for this purpose?
-\\
-        & & A. \ce{F-} forms a stable bonds to the metal ions \\
-        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
-        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
-        \midrule
-        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
-        \midrule
-        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
-\\
-        & & A. \ce{C60} \\
-        & & B. \ce{CH4} \\
-        & & C. \ce{Fe(CO)5} \\
-        & & D. \ce{H2+} \\
-        & & E. \ce{NaCl} \\
-        \midrule
-        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            A. Pindolol has a greater therapeutic range than propranolol \\
+            B. Pindolol has a longer half-life than propranolol \\
+            C. Pindolol has intrinsic activity \\
+            D. Pindolol is more lipophilic than propranolol \\
+            E. Pindolol is more potent than propranolol \\
             \bottomrule
-    \label{tab:chembench_corpus_topic}
-\end{xltabular}
+        \end{tabularx}
+    }
+\end{table}
 
@@
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
-
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
+Note that the total count in this table is bigger than the \chembench corpus.
+This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
+Reasoning and Calculation.} \\
+    \label{tab:chembench_corpus_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Required Skills & Number of Questions & Example Question \\
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
             \midrule
-        Knowledge & 919 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
             \midrule
-        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By atomic force microscopy \\
-        & & B. By transform infrared spectroscopy \\
-        & & C. By measuring a photoluminescence spectrum \\
-        & & D. By absorption spectroscopy \\
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
             \midrule
-        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+            A. By atomic force microscopy \\
+            B. By transform infrared spectroscopy \\
+            C. By measuring a photoluminescence spectrum \\
+            D. By absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
+            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
             \midrule
-        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
@@
             \bottomrule
+        \end{tabularx}
+    }
+\end{table}
 
-    \label{tab:chembench_corpus_cognitive}
-\end{xltabular}
+\normalsize
+
+In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
+
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
+    \label{tab:chembench_corpus_models_failed_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
+            \midrule
+            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
+            A. DESI-mass spectrometry \\
+            B. XPS \\
+            C. atomic force microscopy \\
+            D. scanning tunneling microscopy \\
+            E. secondary ion mass spectrometry \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
+            \midrule
+            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
+            A. \ce{SCl6} does not exist because of steric effects \\
+            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
+            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
+            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
+            \midrule
+            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
+            \midrule
+            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
+            A. All dimension of the material must be smaller than exciton Bohr radius \\
+            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
+            C. The movement od charge carriers is limited in all spatial dimensions \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
+            \midrule
+            What makes acrylates more reactive than typical hydrocarbons? \\
+            A. The carbonyl group \\
+            B. The double bond \\
+            C. The hydroxyl group \\
+            D. The vinyl group \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
+            \midrule
+            Why are polar molecules more soluble in water than non-polar molecules? \\
+            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
+            B. The entropy of the system increases when non-polar molecules dissolve in water \\
+            C. Water has a dipole and can easier isolate polar molecules from each other \\
+            D. Water makes stable dipole-dipole interactions with polar molecules \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
+            \midrule
+            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
+            A. Broader product diameter distribution than pellets or extrudates \\
+            B. Facilitates reactor upscaling \\
+            C. It is also called snow-ball \\
+            D. Not suitable to prepare egg-shell catalysts \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
+            \midrule
+            What determines the flammability of polyols? \\
+            A. Their chemical structure \\
+            B. Their low volatility \\
+            C. Their odor threshold \\
+            D. Their reaction with acids \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
+\normalsize
+
+To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
+The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
+    \label{tab:chembench_corpus_models_failed_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+            \midrule
+            Which of the following are true for phosphoric acid? \\
+            A. Is a weak acid, so it is not toxic \\
+            B. Should never be tasted \\
+            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+            \midrule
+            Which of the following elements has the largest atomic radius? \\
+            A. \ce{Fr} \\
+            B. \ce{He} \\
+            C. \ce{Li} \\
+            D. \ce{Rb} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
+            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
+\\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
 
@@
 The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
-However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
-
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is quite similar across most models.
-This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
-
-An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/corpus_human_comparison.pdf}
-    \caption{\textbf{
-        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
-        We see that the model ranking remains fairly consistent across both sets}  
-    \label{fig:performance_corpus_and_tiny}
-    \script{corpus_humanset_performance.py}
-\end{figure}
+% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
+% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
 
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/model_overall_cal.pdf}
+%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+%     \script{requires_calculation_plot.py}
+%     \label{fig:requires_cal}
+% \end{figure}
 
@@
 
+% \begin{table}
+%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+}
+%     \begin{tabularx}{\textwidth}{XXXX XXXX}
+%         \toprule
+%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\
+%         \midrule
+%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+%         \midrule
+%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+%         \midrule
+%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{tab:correlation_coefficients}
+% \end{table}
+
 \clearpage
@@
 The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer.
-We used \LlamaThreeSeventyBInstruct, accessed via the Groq API.
+We used \LlamaThreeSeventyBInstruct accessed via the Groq API.
 \Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
-A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.
-    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
-    }
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
     \label{fig:human_score_distribution}

### 2024-10-25 src/tex/acronymns.tex
add note about leaderboard
[showyourwork]
[showyourwork] overleaf sync
more cleanup

@@
 \newacronym{mcq}{MCQ}{multiple-choice question}
-\newacronym{rest}{REST}{Representational State Transfer}
+\newacronym{rest}{REST}{representational state transfer}
 \newacronym{orm}{ORM}{object relational mapping}
 \newacronym{dai}{DAI}{daily allowed intake}
-\newacronym{ghs}{GHS}{Globally Harmonized System of Classification and Labelling of Chemicals}
+\newacronym{ghs}{GHS}{globally harmonized system of classification and labelling of chemicals}
 \newacronym{who}{WHO}{World Health Organization}
@@
 \newacronym{helm}{HELM}{Holistic Evaluation of Language Models}
-\newacronym{smiles}{SMILES}{Simplified Molecular Input Line-Entry System}
+\newacronym{smiles}{SMILES}{simplified molecular input line-entry system}
 \newacronym{pca}{PCA}{principal component analysis}
 \newacronym{iupac}{IUPAC}{International Union of Pure and Applied Chemistry}
+\newacronym{json}{JSON}{JavaScript object notation}
+\newacronym{rag}{RAG}{retrieval augmented generation}
+\newacronym{ece}{ECE}{expected calibration error}

### 2024-10-25 src/tex/appendix.tex

--- 2024-10-25 src/tex/appendix.tex
@@
 The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
-However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 \LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).
+In the applications, human participants were presented with molecules as rendered drawings and SMILES strings.
+\LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).
 
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
-    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
-A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.
-    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
-    }
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly.
+    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
     \label{fig:human_score_distribution}

### 2024-10-25 src/tex/authors.tex
fix some references
feat: cleanup text for submission (#62)

@@
 % Question answerers
+\credit{Anagha Aneesh}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
 \credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
@@
 \author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
-\author[1,2 $\star$]{Martiño~Ríos-García~\orcidlink{0000-0003-1507-4048}}
+\author[1,3 $\star$]{Martiño~Ríos-García~\orcidlink{0000-0003-1507-4048}}
 \author[ ]{Benedict~Emoekabu}
@@
 
+\author[1]{Anagha~Aneesh~\orcidlink{0009-0001-0275-2586}}
 \author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
@@
 \author[9]{Amir~Mohammad~Elahi~\orcidlink{0009-0001-5907-101X}}
-\author[1]{Hani~M.~Elbeheiry~\orcidlink{0000-0002-5205-2852}}
+\author[10]{Hani~M.~Elbeheiry~\orcidlink{0000-0002-5205-2852}}
 \author[3]{María~Victoria~Gil~\orcidlink{0000-0003-4894-4660}}
@@
 \author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
-\author[10, 11]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
+\author[11, 12]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
 \author[1]{Jakob~Meyer}
-\author[12]{Santiago~Miret~\orcidlink{0000-0002-4787-2757}}
+\author[13]{Santiago~Miret~\orcidlink{0000-0002-4787-2757}}
 \author[1]{Jan~Matthias~Peschel~\orcidlink{0009-0002-4787-2757}}
-\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
-\author[1, 13]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
+\author[1, 14]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
+\author[1, 14]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}
 
-\author[1, 11]{Johanna~Schreiber~\orcidlink{0009-0000-0991-8967}}
-\author[1,2, 10, 13]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
-\author[1, 13]{Leanne~M.~Stafast~\orcidlink{0009-0008-5604-261X}}
-\author[14]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
+\author[1, 14]{Johanna~Schreiber~\orcidlink{0009-0000-0991-8967}}
+\author[1,2, 10, 14]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
+\author[1, 14]{Leanne~M.~Stafast~\orcidlink{0009-0008-5604-261X}}
+\author[15]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}
 
 % Senior authors
-\author[15,16]{Michael~Pieler~\orcidlink{0000-0001-9186-7045}}
+\author[16,17]{Michael~Pieler~\orcidlink{0000-0001-9186-7045}}
 \author[5, 6]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
-\author[1,2, 10, 13, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
+\author[1,2, 11, 14, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}
 
@@
 \affil[9]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
-\affil[10]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[11]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
-\affil[12]{Intel Labs}
-\affil[13]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
+\affil[10]{Institute for Inorganic and Analytical Chemistry (IAAC), Friedrich Schiller University Jena, Humboldtstrasse 8, 07743 Jena, Germany}
+
+\affil[11]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[12]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
+\affil[13]{Intel Labs}
+\affil[14]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}
 
 \affil[14]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
-\affil[15]{OpenBioML.org}
-\affil[16]{Stability.AI}
+\affil[16]{OpenBioML.org}
+\affil[17]{Stability.AI}
 

### 2024-10-25 src/tex/appendix.tex
Update src/tex/appendix.tex
chore: update the format of the examples table (#74)
Co-authored-by: sreekanth221998 <sreekanthkuncha@gmail.com>
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+\begin{table}
+    \centering
+    \small
     \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-
+    \label{tab:chembench_corpus_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Question Type & \% & Example Questions  \\
+            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
             \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+            A. X-ray fluorescence analysis \\
+            B. Differential pulse polarography \\
+            C. Flame-atomic absorption spectroscopy \\
+            D. Gas chromatography with flame ionization detector \\
+            E. Hydride generation atomic absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
             \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \midrule
+            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \midrule
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+            A. \ce{F-} forms a stable bonds to the metal ions \\
+            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \midrule
+            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+            A. $E_1$ \\
+            B. $E_{cb}$ \\
+            C. $S_N1$ \\
+            D. $S_N2$ \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \midrule
+            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
+            A. \ce{C60} \\
+            B. \ce{CH4} \\
+            C. \ce{Fe(CO)5} \\
+            D. \ce{H2+} \\
+            E. \ce{NaCl} \\
             \midrule
-        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \midrule
+            Which of the following statements is true about the different types of ideal reactors? \\
+            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \midrule
+            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.

### 2024-10-25 src/tex/ms.tex
fix: spillover line in overview dataset (#77)
feat: add comments about no data from benchs

--- 2024-10-25 src/tex/ms.tex
@@
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
+Related to this, in order to avoid data contamination we did not include data from other chemistry-focused benchmarks.
+However, if in the future we decide to include them, it should not be difficult to automatically do it.
 
@@
 Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.
+The chemical preference questions are the main corpus of the semi-automatically generated group with 1001 questions of the total 1752.}
     \variable{output/sources_table.tex}

### 2024-10-27 src/tex/appendix.tex
fix: change the location for data from other benchs

--- 2024-10-25 src/tex/appendix.tex
@@
 The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
+
+While adding questions from existing benchmarks seemed to be another good source of semi-automatically generated data, we prioritize diversity of the data and avoiding data contamination while addressing the guidelines above and in \Cref{sec:desired-properties}.
+Therefore, we decided to not include questions from other previous published chemistry-focused benchmarks into the \chembench corpus.
 
@@
 
-In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
-The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{tab:cb_skillset} shows an example question for each skill category.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
-    \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
-    \label{fig:cb_skillset}
-\end{figure}
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 

### 2024-10-27 src/tex/ms.tex
fix: change the location for data from other benchs

@@
 This requires that \Gls{llm} developers filter their training dataset for this canary string.\autocite{openai2024gpt4, srivastava2022beyond}
-Related to this, in order to avoid data contamination we did not include data from other chemistry-focused benchmarks.
-However, if in the future we decide to include them, it should not be difficult to automatically do it.
 
@@
 All questions have been rephrased, annotated, and reviewed before being added to the corpus.
-The chemical preference questions are the main corpus of the semi-automatically generated group with 1001 questions of the total 1752.}
+Note that the chemical preference questions are the main corpus of the semi-automatically generated group with 1001 questions of the total 1752.}
     \variable{output/sources_table.tex}

### 2024-10-27 src/tex/appendix.tex
try to make compile work

--- 2024-10-25 src/tex/appendix.tex
@@
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-    \label{tab:chembench_corpus_topic}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
-            \midrule
-            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-            A. X-ray fluorescence analysis \\
-            B. Differential pulse polarography \\
-            C. Flame-atomic absorption spectroscopy \\
-            D. Gas chromatography with flame ionization detector \\
-            E. Hydride generation atomic absorption spectroscopy \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+%\begin{table}
+%    \centering
+%    \small
+%    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
+%    \label{tab:chembench_corpus_topic}
+%    {\fontsize{8}{9}\selectfont
+%        \begin{tabularx}{\textwidth}{X}
+%            \toprule
+%            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
+%            \midrule
+%            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+%            A. X-ray fluorescence analysis \\
+%            B. Differential pulse polarography \\
+%            C. Flame-atomic absorption spectroscopy \\
+%            D. Gas chromatography with flame ionization detector \\
+%            E. Hydride generation atomic absorption spectroscopy \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+%            \midrule
+%            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
-            \midrule
-            Which of the following salts is an acidic salt? \\
-            A. \ce{NH4Cl} \\
-            B. \ce{Na2CO3} \\
-            C. \ce{NaH2PO4} \\
-            D. \ce{Zn(OH)Cl} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
-            \midrule
-            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
-            \midrule
-            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
-            A. \ce{F-} forms a stable bonds to the metal ions \\
-            B. \ce{HF} has a better water solubility than \ce{HCl} \\
-            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-            D. \ce{HF} is a weaker acid than \ce{HCl} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
-            \midrule
-            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
-            A. $E_1$ \\
-            B. $E_{cb}$ \\
-            C. $S_N1$ \\
-            D. $S_N2$ \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
-            \midrule
-            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+%            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+%            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
+%            \midrule
+%            Which of the following salts is an acidic salt? \\
+%            A. \ce{NH4Cl} \\
+%            B. \ce{Na2CO3} \\
+%            C. \ce{NaH2PO4} \\
+%            D. \ce{Zn(OH)Cl} \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+%            \midrule
+%            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+%            \midrule
+%            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+%            A. \ce{F-} forms a stable bonds to the metal ions \\
+%            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+%            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+%            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+%            \midrule
+%            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+%            A. $E_1$ \\
+%            B. $E_{cb}$ \\
+%            C. $S_N1$ \\
+%            D. $S_N2$ \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+%            \midrule
+%            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
 Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-            A. \ce{C60} \\
-            B. \ce{CH4} \\
-            C. \ce{Fe(CO)5} \\
-            D. \ce{H2+} \\
-            E. \ce{NaCl} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
-            \midrule
-            Which of the following statements is true about the different types of ideal reactors? \\
-            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
-            \midrule
-            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+%            A. \ce{C60} \\
+%            B. \ce{CH4} \\
+%            C. \ce{Fe(CO)5} \\
+%            D. \ce{H2+} \\
+%            E. \ce{NaCl} \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+%            \midrule
+%            Which of the following statements is true about the different types of ideal reactors? \\
+%            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+%            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+%            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+%            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+%            \midrule
+%            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-            A. Pindolol has a greater therapeutic range than propranolol \\
-            B. Pindolol has a longer half-life than propranolol \\
-            C. Pindolol has intrinsic activity \\
-            D. Pindolol is more lipophilic than propranolol \\
-            E. Pindolol is more potent than propranolol \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
+%            A. Pindolol has a greater therapeutic range than propranolol \\
+%            B. Pindolol has a longer half-life than propranolol \\
+%            C. Pindolol has intrinsic activity \\
+%            D. Pindolol is more lipophilic than propranolol \\
+%            E. Pindolol is more potent than propranolol \\
+%            \bottomrule
+%        \end{tabularx}
+%    }
+%\end{table}
 

### 2024-10-27 src/scripts/make_name_dir_map.py

@@ -1,8 +1,8 @@
-import fire 
-import os 
-from glob import glob 
+import fire
+import os
+from glob import glob
 import yaml
-import pickle 
+import pickle
 from pathlib import Path
 from utils import obtain_chembench_repo
 from paths import data
@@ -46,6 +46,13 @@ def build_name_dir_map(report_base_dir=None):
                 else:
                     model_name = yaml_data["model_name"] + "-T-one"
 
+            model_name = yaml_data["model_name"]
+            if model_name == 'Llama-2-70B Chat':
+                model_name = 'Llama-2-70B-Chat'
+            else:
+                raise ValueError(f"Model name {model_name} not recognized")
+
+
         name_dir_map[model_name] = os.path.abspath(report_subdir)
 
     with open(data / "name_to_dir_map.pkl", "wb") as f:
@@ -55,4 +62,3 @@ def build_name_dir_map(report_base_dir=None):
 
 if __name__ == "__main__":
     fire.Fire(build_name_dir_map)
-

### 2024-10-27 src/tex/appendix.tex
try to make compile work

@@
 This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.} \\
+Reasoning and Calculation.}
     \label{tab:chembench_corpus_cognitive}
@@
     \small
-    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
+    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.}
     \label{tab:chembench_corpus_models_failed_topic}
@@
     \small
-    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
+    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.}
     \label{tab:chembench_corpus_models_failed_cognitive}

### 2024-10-27 src/tex/appendix.tex
try to make compile work

@@
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.}
-    \label{tab:chembench_corpus_models_failed_topic}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
-            \midrule
-            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
-            A. DESI-mass spectrometry \\
-            B. XPS \\
-            C. atomic force microscopy \\
-            D. scanning tunneling microscopy \\
-            E. secondary ion mass spectrometry \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+%\begin{table}
+%    \centering
+%    \small
+%    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.}
+%    \label{tab:chembench_corpus_models_failed_topic}
+%    {\fontsize{8}{9}\selectfont
+%        \begin{tabularx}{\textwidth}{X}
+%            \toprule
+%            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
+%            \midrule
+%            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
+%            A. DESI-mass spectrometry \\
+%            B. XPS \\
+%            C. atomic force microscopy \\
+%            D. scanning tunneling microscopy \\
+%            E. secondary ion mass spectrometry \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
+%            \midrule
+%            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
-            \midrule
-            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
-            A. \ce{SCl6} does not exist because of steric effects \\
-            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
-            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
-            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
-            \midrule
-            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
-            \midrule
-            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
-            A. All dimension of the material must be smaller than exciton Bohr radius \\
-            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
-            C. The movement od charge carriers is limited in all spatial dimensions \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
-            \midrule
-            What makes acrylates more reactive than typical hydrocarbons? \\
-            A. The carbonyl group \\
-            B. The double bond \\
-            C. The hydroxyl group \\
-            D. The vinyl group \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
-            \midrule
-            Why are polar molecules more soluble in water than non-polar molecules? \\
-            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
-            B. The entropy of the system increases when non-polar molecules dissolve in water \\
-            C. Water has a dipole and can easier isolate polar molecules from each other \\
-            D. Water makes stable dipole-dipole interactions with polar molecules \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
-            \midrule
-            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
-            A. Broader product diameter distribution than pellets or extrudates \\
-            B. Facilitates reactor upscaling \\
-            C. It is also called snow-ball \\
-            D. Not suitable to prepare egg-shell catalysts \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
-            \midrule
-            What determines the flammability of polyols? \\
-            A. Their chemical structure \\
-            B. Their low volatility \\
-            C. Their odor threshold \\
-            D. Their reaction with acids \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
+%            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+%            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
+%            \midrule
+%            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
+%            A. \ce{SCl6} does not exist because of steric effects \\
+%            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
+%            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
+%            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
+%            \midrule
+%            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
+%            \midrule
+%            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
+%            A. All dimension of the material must be smaller than exciton Bohr radius \\
+%            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
+%            C. The movement od charge carriers is limited in all spatial dimensions \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
+%            \midrule
+%            What makes acrylates more reactive than typical hydrocarbons? \\
+%            A. The carbonyl group \\
+%            B. The double bond \\
+%            C. The hydroxyl group \\
+%            D. The vinyl group \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
+%            \midrule
+%            Why are polar molecules more soluble in water than non-polar molecules? \\
+%            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
+%            B. The entropy of the system increases when non-polar molecules dissolve in water \\
+%            C. Water has a dipole and can easier isolate polar molecules from each other \\
+%            D. Water makes stable dipole-dipole interactions with polar molecules \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
+%            \midrule
+%            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
+%            A. Broader product diameter distribution than pellets or extrudates \\
+%            B. Facilitates reactor upscaling \\
+%            C. It is also called snow-ball \\
+%            D. Not suitable to prepare egg-shell catalysts \\
+%            \midrule
+%            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
+%            \midrule
+%            What determines the flammability of polyols? \\
+%            A. Their chemical structure \\
+%            B. Their low volatility \\
+%            C. Their odor threshold \\
+%            D. Their reaction with acids \\
+%            \bottomrule
+%        \end{tabularx}
+%    }
+%\end{table}
 

### 2024-10-27 src/tex/appendix.tex
try to make compile work

@@
 \\
-            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
-            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
+            A. [START\_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END\_SMILES] \\
+            B. [START\_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END\_SMILES] \\
             \midrule
@@
             \midrule
-            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23}{V}.
 What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?

### 2024-10-27 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-27 src/tex/appendix.tex
@@
 
+While adding questions from existing benchmarks might seem to be another good source of semi-automatically generated data, we prioritized the diversity of the data and avoided data contamination while addressing the guidelines above and in \Cref{sec:desired-properties}.
+However, even though we decided not to include questions from other previously published chemistry-focused benchmarks into the \chembench corpus, the framework is flexible enough to be readily extended with questions from other benchmarks.
+
 \subsubsection{Composition}
@@
 
-%\begin{table}
-%    \centering
-%    \small
-%    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
-%    \label{tab:chembench_corpus_topic}
-%    {\fontsize{8}{9}\selectfont
-%        \begin{tabularx}{\textwidth}{X}
-%            \toprule
-%            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
-%            \midrule
-%            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-%            A. X-ray fluorescence analysis \\
-%            B. Differential pulse polarography \\
-%            C. Flame-atomic absorption spectroscopy \\
-%            D. Gas chromatography with flame ionization detector \\
-%            E. Hydride generation atomic absorption spectroscopy \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
-%            \midrule
-%            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
+
+        \toprule
+        Question Type & \% & Example Questions  \\
+        \midrule
+        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-%            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-%            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
-%            \midrule
-%            Which of the following salts is an acidic salt? \\
-%            A. \ce{NH4Cl} \\
-%            B. \ce{Na2CO3} \\
-%            C. \ce{NaH2PO4} \\
-%            D. \ce{Zn(OH)Cl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
-%            \midrule
-%            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
-%            \midrule
-%            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
-%            A. \ce{F-} forms a stable bonds to the metal ions \\
-%            B. \ce{HF} has a better water solubility than \ce{HCl} \\
-%            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-%            D. \ce{HF} is a weaker acid than \ce{HCl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
-%            \midrule
-%            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
-%            A. $E_1$ \\
-%            B. $E_{cb}$ \\
-%            C. $S_N1$ \\
-%            D. $S_N2$ \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
-%            \midrule
-%            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+        \midrule
+        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
+        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+        \midrule
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
 \\
-%            A. \ce{C60} \\
-%            B. \ce{CH4} \\
-%            C. \ce{Fe(CO)5} \\
-%            D. \ce{H2+} \\
-%            E. \ce{NaCl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
-%            \midrule
-%            Which of the following statements is true about the different types of ideal reactors? \\
-%            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-%            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-%            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-%            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
-%            \midrule
-%            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+        & & A. $E_1$ \\
+        & & B. $E_{cb}$ \\
+        & & C. $S_N1$ \\
+        & & D. $S_N2$ \\
+        \midrule
+        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-%            A. Pindolol has a greater therapeutic range than propranolol \\
-%            B. Pindolol has a longer half-life than propranolol \\
-%            C. Pindolol has intrinsic activity \\
-%            D. Pindolol is more lipophilic than propranolol \\
-%            E. Pindolol is more potent than propranolol \\
-%            \bottomrule
-%        \end{tabularx}
-%    }
-%\end{table}
+        & & A. Pindolol has a greater therapeutic range than propranolol \\
+        & & B. Pindolol has a longer half-life than propranolol \\
+        & & C. Pindolol has intrinsic activity \\
+        & & D. Pindolol is more lipophilic than propranolol \\
+        & & E. Pindolol is more potent than propranolol \\
+        \midrule
+        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+        & & A. X-ray fluorescence analysis \\
+        & & B. Differential pulse polarography \\
+        & & C. Flame-atomic absorption spectroscopy \\
+        & & D. Gas chromatography with flame ionization detector \\
+        & & E. Hydride generation atomic absorption spectroscopy \\
+        \midrule
+        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
+Why would one choose \ce{HF} over \ce{HCl} for this purpose?
+\\
+        & & A. \ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
+        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
+        \midrule
+        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
+        \midrule
+        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+\\
+        & & A. \ce{C60} \\
+        & & B. \ce{CH4} \\
+        & & C. \ce{Fe(CO)5} \\
+        & & D. \ce{H2+} \\
+        & & E. \ce{NaCl} \\
+        \midrule
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+        \bottomrule
+    \label{tab:chembench_corpus_topic}
+\end{xltabular}
 
@@
 
-In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
-The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{tab:cb_skillset} shows an example question for each skill category.
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
-    \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
-    \label{fig:cb_skillset}
-\end{figure}
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
-Note that the total count in this table is bigger than the \chembench corpus.
-This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.}
-    \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
         \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+        Required Skills & Number of Questions & Example Question \\
         \midrule
-            Which of the following salts is an acidic salt? \\
-            A. \ce{NH4Cl} \\
-            B. \ce{Na2CO3} \\
-            C. \ce{NaH2PO4} \\
-            D. \ce{Zn(OH)Cl} \\
+        Knowledge & 919 & Which of the following salts is an acidic salt? \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
         \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        & & A. By atomic force microscopy \\
+        & & B. By transform infrared spectroscopy \\
+        & & C. By measuring a photoluminescence spectrum \\
+        & & D. By absorption spectroscopy \\
         \midrule
-            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-            A. By atomic force microscopy \\
-            B. By transform infrared spectroscopy \\
-            C. By measuring a photoluminescence spectrum \\
-            D. By absorption spectroscopy \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
+        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
         \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-            \midrule
-            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
@@
         \bottomrule
-        \end{tabularx}
-    }
-\end{table}
 
-\normalsize
-
-In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
-
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.}
-    \label{tab:chembench_corpus_models_failed_topic}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
-            \midrule
-            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
-            A. DESI-mass spectrometry \\
-            B. XPS \\
-            C. atomic force microscopy \\
-            D. scanning tunneling microscopy \\
-            E. secondary ion mass spectrometry \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
-            \midrule
-            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
-            A. \ce{SCl6} does not exist because of steric effects \\
-            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
-            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
-            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
-            \midrule
-            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
-            \midrule
-            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
-            A. All dimension of the material must be smaller than exciton Bohr radius \\
-            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
-            C. The movement od charge carriers is limited in all spatial dimensions \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
-            \midrule
-            What makes acrylates more reactive than typical hydrocarbons? \\
-            A. The carbonyl group \\
-            B. The double bond \\
-            C. The hydroxyl group \\
-            D. The vinyl group \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
-            \midrule
-            Why are polar molecules more soluble in water than non-polar molecules? \\
-            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
-            B. The entropy of the system increases when non-polar molecules dissolve in water \\
-            C. Water has a dipole and can easier isolate polar molecules from each other \\
-            D. Water makes stable dipole-dipole interactions with polar molecules \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
-            \midrule
-            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
-            A. Broader product diameter distribution than pellets or extrudates \\
-            B. Facilitates reactor upscaling \\
-            C. It is also called snow-ball \\
-            D. Not suitable to prepare egg-shell catalysts \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
-            \midrule
-            What determines the flammability of polyols? \\
-            A. Their chemical structure \\
-            B. Their low volatility \\
-            C. Their odor threshold \\
-            D. Their reaction with acids \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
-
-\normalsize
-
-To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
-The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.}
-    \label{tab:chembench_corpus_models_failed_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
-            \midrule
-            Which of the following are true for phosphoric acid? \\
-            A. Is a weak acid, so it is not toxic \\
-            B. Should never be tasted \\
-            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
-            \midrule
-            Which of the following elements has the largest atomic radius? \\
-            A. \ce{Fr} \\
-            B. \ce{He} \\
-            C. \ce{Li} \\
-            D. \ce{Rb} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
-            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-            \midrule
-            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
-What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
-\\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
+    \label{tab:chembench_corpus_cognitive}
+\end{xltabular}
 

### 2024-10-28 src/tex/appendix.tex
fix: solve formatting problems in examples table

--- 2024-10-27 src/tex/appendix.tex
@@
 
-%\begin{table}
-%    \centering
-%    \small
-%    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
-%    \label{tab:chembench_corpus_topic}
-%    {\fontsize{8}{9}\selectfont
-%        \begin{tabularx}{\textwidth}{X}
-%            \toprule
-%            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
-%            \midrule
-%            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-%            A. X-ray fluorescence analysis \\
-%            B. Differential pulse polarography \\
-%            C. Flame-atomic absorption spectroscopy \\
-%            D. Gas chromatography with flame ionization detector \\
-%            E. Hydride generation atomic absorption spectroscopy \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
-%            \midrule
-%            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
+    \label{tab:chembench_corpus_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
+            \midrule
+            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+            A. X-ray fluorescence analysis \\
+            B. Differential pulse polarography \\
+            C. Flame-atomic absorption spectroscopy \\
+            D. Gas chromatography with flame ionization detector \\
+            E. Hydride generation atomic absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-%            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-%            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
-%            \midrule
-%            Which of the following salts is an acidic salt? \\
-%            A. \ce{NH4Cl} \\
-%            B. \ce{Na2CO3} \\
-%            C. \ce{NaH2PO4} \\
-%            D. \ce{Zn(OH)Cl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
-%            \midrule
-%            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
-%            \midrule
-%            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
-%            A. \ce{F-} forms a stable bonds to the metal ions \\
-%            B. \ce{HF} has a better water solubility than \ce{HCl} \\
-%            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-%            D. \ce{HF} is a weaker acid than \ce{HCl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
-%            \midrule
-%            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
-%            A. $E_1$ \\
-%            B. $E_{cb}$ \\
-%            C. $S_N1$ \\
-%            D. $S_N2$ \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
-%            \midrule
-%            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+            A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
+            \midrule
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \midrule
+            What is the oxidation number of the metal in the compound  \ce{[ZrF_{7}]3-} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \midrule
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose HF over \ce{HCl} for this purpose? \\
+            A. \ce{F-} forms a stable bonds to the metal ions \\
+            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \midrule
+            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+            A. $E_1$ \\
+            B. $E_{cb}$ \\
+            C. $S_N1$ \\
+            D. $S_N2$ \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \midrule
+            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
 Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-%            A. \ce{C60} \\
-%            B. \ce{CH4} \\
-%            C. \ce{Fe(CO)5} \\
-%            D. \ce{H2+} \\
-%            E. \ce{NaCl} \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
-%            \midrule
-%            Which of the following statements is true about the different types of ideal reactors? \\
-%            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-%            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-%            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-%            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
-%            \midrule
-%            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+            A. \ce{C60} \\
+            B. \ce{CH4} \\
+            C. \ce{Fe(CO)5} \\
+            D. \ce{H2+} \\
+            E. \ce{NaCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \midrule
+            Which of the following statements is true about the different types of ideal reactors? \\
+            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \midrule
+            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-%            A. Pindolol has a greater therapeutic range than propranolol \\
-%            B. Pindolol has a longer half-life than propranolol \\
-%            C. Pindolol has intrinsic activity \\
-%            D. Pindolol is more lipophilic than propranolol \\
-%            E. Pindolol is more potent than propranolol \\
-%            \bottomrule
-%        \end{tabularx}
-%    }
-%\end{table}
+            A. Pindolol has a greater therapeutic range than propranolol \\
+            B. Pindolol has a longer half-life than propranolol \\
+            C. Pindolol has intrinsic activity \\
+            D. Pindolol is more lipophilic than propranolol \\
+            E. Pindolol is more potent than propranolol \\
+            \bottomrule
+        \end{tabularx}
+    }
+ \end{table}
 
@@
 
-\normalsize
+% \normalsize
 
-To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
+% To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
 The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.}
-    \label{tab:chembench_corpus_models_failed_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
-            \midrule
-            Which of the following are true for phosphoric acid? \\
-            A. Is a weak acid, so it is not toxic \\
-            B. Should never be tasted \\
-            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
-            \midrule
-            Which of the following elements has the largest atomic radius? \\
-            A. \ce{Fr} \\
-            B. \ce{He} \\
-            C. \ce{Li} \\
-            D. \ce{Rb} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+% \begin{table}
+%     \centering
+%     \small
+%     \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.}
+%     \label{tab:chembench_corpus_models_failed_cognitive}
+%     {\fontsize{8}{9}\selectfont
+%         \begin{tabularx}{\textwidth}{X}
+%             \toprule
+%             \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+%             \midrule
+%             Which of the following are true for phosphoric acid? \\
+%             A. Is a weak acid, so it is not toxic \\
+%             B. Should never be tasted \\
+%             C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
+%             \midrule
+%             \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+%             \midrule
+%             Which of the following elements has the largest atomic radius? \\
+%             A. \ce{Fr} \\
+%             B. \ce{He} \\
+%             C. \ce{Li} \\
+%             D. \ce{Rb} \\
+%             \midrule
+%             \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+%             \midrule
+%             Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
-            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-            \midrule
-            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+%             A. [START\_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END\_SMILES] \\
+%             B. [START\_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END\_SMILES] \\
+%             \midrule
+%             \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+%             \midrule
+%             The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23}{V}.
 What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
 \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
+%             \bottomrule
+%         \end{tabularx}
+%     }
+% \end{table}
 

### 2024-10-28 src/tex/ms.tex
fix: fix orcid and add acknowledgments (#83)
Co-authored-by: vgvinter <vicky@INCAR-162.campus.csic.es>

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V2.pdf}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
 Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.
-Note that the chemical preference questions are the main corpus of the semi-automatically generated group with 1001 questions of the total 1752.}
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
     \variable{output/sources_table.tex}
@@
 
+M.R.G.\ and M.V.G.\ acknowledge financial support from the Spanish Agencia Estatal de Investigaci\'{o}n (AEI) through grants TED2021-131693B-I00 and CNS2022-135474, funded by MICIU/AEI/10.13039/501100011033 and by the European Union NextGenerationEU/PRTR.
+M.V.G.\ acknowledges support from the Spanish National Research Council (CSIC) through Programme for internationalization i-LINK 2023 (Project ILINK23047).
+
 A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S.

### 2024-10-28 src/tex/appendix.tex
fix: change the reasoning example

@@
             \midrule
-            What is the oxidation number of the metal in the compound  \ce{[ZrF_{7}]3-} \\
+            What is the oxidation number of the metal in the compound \ce{[ZrF_{7}]3-}? \\
             \midrule
@@
 
-In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+In addition, as shown in \Cref{fig:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 The plot shows a balanced distribution of the required skills in the \chembench corpus.
@@
             \midrule
-            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-            A. By atomic force microscopy \\
-            B. By transform infrared spectroscopy \\
-            C. By measuring a photoluminescence spectrum \\
-            D. By absorption spectroscopy \\
+            In an electrochemical system, the standard potential for the hydrogen evolution reaction \\ce{2H+ + 2e- -> H2} is \\pu{-0.07 V}, and for \\ce{Br-} oxidation is \\pu{1.02 V} vs.
+\\ce{Ag/AgBr}.
+During an experiment, it's observed that the onset potential for hydrogen evolution has shifted towards a higher potential, while the onset potential for \\ce{Br-} oxidation has shifted towards a lower potential, compared to their respective standard potentials.
+Which of the following explanations is most likely correct?
+            A. The electrode material has changed, causing catalytic effects that alter the onset potentials. \\
+            B. There is probably a change in temperature, which affects reaction kinetics. \\
+            C. This is due to changes in concentration of reactants and products, following the Nernst equation. \\
+            D. This observation is due to experimental error. \\
             \midrule
@@
 
-In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
 The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement

### 2024-10-28 src/tex/appendix.tex
fix: correct the failed examples table

@@
 
-%\begin{table}
-%    \centering
-%    \small
-%    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.}
-%    \label{tab:chembench_corpus_models_failed_topic}
-%    {\fontsize{8}{9}\selectfont
-%        \begin{tabularx}{\textwidth}{X}
-%            \toprule
-%            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
-%            \midrule
-%            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
-%            A. DESI-mass spectrometry \\
-%            B. XPS \\
-%            C. atomic force microscopy \\
-%            D. scanning tunneling microscopy \\
-%            E. secondary ion mass spectrometry \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
-%            \midrule
-%            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
+    \label{tab:chembench_corpus_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
+            A. Cyclohexene \\
+            B. Cyclooctene \\
+            C. Decalin \\
+            D. Ethyl ether \\
+            E. Isopropyl ether \\
+            F. Para-dioxane \\
+            G. Tetralin \\
+            \midrule
+            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
+            \midrule
+            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
+            A. The intersection with y-axis shift to higher numbers \\
+            B. The intersection with y-axis shift to lower numbers \\
+            C. The slope decreases \\
+            D. The slope increases \\
+            \midrule
+            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
+A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
+What is the number of moles of acid in the sample?
 \\
-%            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-%            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
-%            \midrule
-%            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
-%            A. \ce{SCl6} does not exist because of steric effects \\
-%            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
-%            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
-%            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
-%            \midrule
-%            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
-%            \midrule
-%            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
-%            A. All dimension of the material must be smaller than exciton Bohr radius \\
-%            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
-%            C. The movement od charge carriers is limited in all spatial dimensions \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
-%            \midrule
-%            What makes acrylates more reactive than typical hydrocarbons? \\
-%            A. The carbonyl group \\
-%            B. The double bond \\
-%            C. The hydroxyl group \\
-%            D. The vinyl group \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
-%            \midrule
-%            Why are polar molecules more soluble in water than non-polar molecules? \\
-%            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
-%            B. The entropy of the system increases when non-polar molecules dissolve in water \\
-%            C. Water has a dipole and can easier isolate polar molecules from each other \\
-%            D. Water makes stable dipole-dipole interactions with polar molecules \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
-%            \midrule
-%            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
-%            A. Broader product diameter distribution than pellets or extrudates \\
-%            B. Facilitates reactor upscaling \\
-%            C. It is also called snow-ball \\
-%            D. Not suitable to prepare egg-shell catalysts \\
-%            \midrule
-%            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
-%            \midrule
-%            What determines the flammability of polyols? \\
-%            A. Their chemical structure \\
-%            B. Their low volatility \\
-%            C. Their odor threshold \\
-%            D. Their reaction with acids \\
-%            \bottomrule
-%        \end{tabularx}
-%    }
-%\end{table}
-
-% \normalsize
-
-% To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
-The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
-% \begin{table}
-%     \centering
-%     \small
-%     \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.}
-%     \label{tab:chembench_corpus_models_failed_cognitive}
-%     {\fontsize{8}{9}\selectfont
-%         \begin{tabularx}{\textwidth}{X}
-%             \toprule
-%             \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
-%             \midrule
-%             Which of the following are true for phosphoric acid? \\
-%             A. Is a weak acid, so it is not toxic \\
-%             B. Should never be tasted \\
-%             C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
-%             \midrule
-%             \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
-%             \midrule
-%             Which of the following elements has the largest atomic radius? \\
-%             A. \ce{Fr} \\
-%             B. \ce{He} \\
-%             C. \ce{Li} \\
-%             D. \ce{Rb} \\
-%             \midrule
-%             \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-%             \midrule
-%             Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
+            \midrule
+            What is hydrofluoric acid used for? \\
+            A. For cleaning facades \\
+            B. For etching glass \\
+            C. For filling accumulators \\
+            D. For soldering \\
+            \midrule
+            How many chiral centers are present in the molecular structure of citric acid? \\
+            \midrule
+            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
+            A. 10 \\
+            B. 14 \\
+            C. 18 \\
+            D. 23 \\
+            E. 9 \\
+            \midrule
+            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
+What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
 \\
-%             A. [START\_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END\_SMILES] \\
-%             B. [START\_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END\_SMILES] \\
-%             \midrule
-%             \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-%             \midrule
-%             The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23}{V}.
-What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
-\\
-%             \bottomrule
-%         \end{tabularx}
-%     }
-% \end{table}
+            \midrule
+            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
+            A. 10.0\% \\ 
+            B. 6.6\% \\
+            C. 7.0\% \\
+            D. 94\% \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
 

### 2024-10-28 src/tex/ms.tex
fix: correct automatic count link

--- 2024-10-27 src/tex/ms.tex
@@
 All questions have been rephrased, annotated, and reviewed before being added to the corpus.
-Note that the chemical preference questions are the main corpus of the semi-automatically generated group with 1001 questions of the total 1752.}
+Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.}
     \variable{output/sources_table.tex}

### 2024-10-28 src/tex/appendix.tex
small fix of text

@@
 
+%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
+%We see that the questions are generally complex to read.
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+It is calculated based on the average number of syllables per word and words per sentence.
+The higher the score, the easier the text is to read.
+The distribution of the questions' scores is shown in the histogram.
+}
+%     \label{fig:flesch_kincaid_reading_ease}
+%     \script{wordcloud.py}
+% \end{figure}
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+%     \label{fig:question_count_barplot_mcq_vs_general}
+%     \script{plot_statistics.py}
+% \end{figure}
+
 The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
-    \label{tab:chembench_corpus_topic}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
+
         \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
+        Question Type & \% & Example Questions  \\
         \midrule
-            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
-            A. X-ray fluorescence analysis \\
-            B. Differential pulse polarography \\
-            C. Flame-atomic absorption spectroscopy \\
-            D. Gas chromatography with flame ionization detector \\
-            E. Hydride generation atomic absorption spectroscopy \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
         \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
+        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
+        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
         \midrule
-            Which of the following salts is an acidic salt? \\
-            A. \ce{NH4Cl} \\
-            B. \ce{Na2CO3} \\
-            C. \ce{NaH2PO4} \\
-            D. \ce{Zn(OH)Cl} \\
+        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
+\\
+        & & A. $E_1$ \\
+        & & B. $E_{cb}$ \\
+        & & C. $S_N1$ \\
+        & & D. $S_N2$ \\
         \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
-            \midrule
-            What is the oxidation number of the metal in the compound \ce{[ZrF_{7}]3-}? \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
-            \midrule
-            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose HF over \ce{HCl} for this purpose? \\
-            A. \ce{F-} forms a stable bonds to the metal ions \\
-            B. \ce{HF} has a better water solubility than \ce{HCl} \\
-            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
-            D. \ce{HF} is a weaker acid than \ce{HCl} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
-            \midrule
-            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
-            A. $E_1$ \\
-            B. $E_{cb}$ \\
-            C. $S_N1$ \\
-            D. $S_N2$ \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
-            \midrule
-            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
-Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
-\\
-            A. \ce{C60} \\
-            B. \ce{CH4} \\
-            C. \ce{Fe(CO)5} \\
-            D. \ce{H2+} \\
-            E. \ce{NaCl} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
-            \midrule
-            Which of the following statements is true about the different types of ideal reactors? \\
-            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
-            \midrule
-            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 \\
-            A. Pindolol has a greater therapeutic range than propranolol \\
-            B. Pindolol has a longer half-life than propranolol \\
-            C. Pindolol has intrinsic activity \\
-            D. Pindolol is more lipophilic than propranolol \\
-            E. Pindolol is more potent than propranolol \\
+        & & A. Pindolol has a greater therapeutic range than propranolol \\
+        & & B. Pindolol has a longer half-life than propranolol \\
+        & & C. Pindolol has intrinsic activity \\
+        & & D. Pindolol is more lipophilic than propranolol \\
+        & & E. Pindolol is more potent than propranolol \\
+        \midrule
+        Analytical Chemistry & 5.8 & Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+        & & A. X-ray fluorescence analysis \\
+        & & B. Differential pulse polarography \\
+        & & C. Flame-atomic absorption spectroscopy \\
+        & & D. Gas chromatography with flame ionization detector \\
+        & & E. Hydride generation atomic absorption spectroscopy \\
+        \midrule
+        Materials Science & 3.1 & For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact.
+Why would one choose \ce{HF} over \ce{HCl} for this purpose?
+\\
+        & & A. \ce{F-} forms a stable bonds to the metal ions \\
+        & & B. \ce{HF} has a better water solubility than \ce{HCl} \\
+        & & C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+        & & D. \ce{HF} is a weaker acid than \ce{HCl} \\
+        \midrule
+        General Chemistry & 5.3 & Which of the following salts is an acidic salt? \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
+        \midrule
+        Physical Chemistry & 6.3 & The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
+\\
+        & & A. \ce{C60} \\
+        & & B. \ce{CH4} \\
+        & & C. \ce{Fe(CO)5} \\
+        & & D. \ce{H2+} \\
+        & & E. \ce{NaCl} \\
+        \midrule
+        Inorganic Chemistry & 3.3 & What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
         \bottomrule
-        \end{tabularx}
-    }
- \end{table}
+    \label{tab:chembench_corpus_topic}
+\end{xltabular}
 
@@
 
-In addition, as shown in \Cref{fig:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
-The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{tab:cb_skillset} shows an example question for each skill category.
+In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
-    \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
-    \label{fig:cb_skillset}
-\end{figure}
+\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
-Note that the total count in this table is bigger than the \chembench corpus.
-This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.}
-    \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
         \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+        Required Skills & Number of Questions & Example Question \\
         \midrule
-            Which of the following salts is an acidic salt? \\
-            A. \ce{NH4Cl} \\
-            B. \ce{Na2CO3} \\
-            C. \ce{NaH2PO4} \\
-            D. \ce{Zn(OH)Cl} \\
+        Knowledge & 919 & Which of the following salts is an acidic salt? \\
+        & & A. \ce{NH4Cl} \\
+        & & B. \ce{Na2CO3} \\
+        & & C. \ce{NaH2PO4} \\
+        & & D. \ce{Zn(OH)Cl} \\
         \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+        & & A. By atomic force microscopy \\
+        & & B. By transform infrared spectroscopy \\
+        & & C. By measuring a photoluminescence spectrum \\
+        & & D. By absorption spectroscopy \\
         \midrule
-            In an electrochemical system, the standard potential for the hydrogen evolution reaction \\ce{2H+ + 2e- -> H2} is \\pu{-0.07 V}, and for \\ce{Br-} oxidation is \\pu{1.02 V} vs.
-\\ce{Ag/AgBr}.
-During an experiment, it's observed that the onset potential for hydrogen evolution has shifted towards a higher potential, while the onset potential for \\ce{Br-} oxidation has shifted towards a lower potential, compared to their respective standard potentials.
-Which of the following explanations is most likely correct?
-            A. The electrode material has changed, causing catalytic effects that alter the onset potentials. \\
-            B. There is probably a change in temperature, which affects reaction kinetics. \\
-            C. This is due to changes in concentration of reactants and products, following the Nernst equation. \\
-            D. This observation is due to experimental error. \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
+        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
         \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-            \midrule
-            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
@@
         \bottomrule
-        \end{tabularx}
-    }
-\end{table}
 
-\normalsize
-
-In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
-
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
     \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
-            A. Cyclohexene \\
-            B. Cyclooctene \\
-            C. Decalin \\
-            D. Ethyl ether \\
-            E. Isopropyl ether \\
-            F. Para-dioxane \\
-            G. Tetralin \\
-            \midrule
-            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
-            \midrule
-            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
-            A. The intersection with y-axis shift to higher numbers \\
-            B. The intersection with y-axis shift to lower numbers \\
-            C. The slope decreases \\
-            D. The slope increases \\
-            \midrule
-            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
-A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
-What is the number of moles of acid in the sample?
-\\
-            \midrule
-            What is hydrofluoric acid used for? \\
-            A. For cleaning facades \\
-            B. For etching glass \\
-            C. For filling accumulators \\
-            D. For soldering \\
-            \midrule
-            How many chiral centers are present in the molecular structure of citric acid? \\
-            \midrule
-            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
-            A. 10 \\
-            B. 14 \\
-            C. 18 \\
-            D. 23 \\
-            E. 9 \\
-            \midrule
-            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
-What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
-\\
-            \midrule
-            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
-            A. 10.0\% \\ 
-            B. 6.6\% \\
-            C. 7.0\% \\
-            D. 94\% \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
+\end{xltabular}
 
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is quite similar across most models.
-This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
-
-An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
-
 \begin{figure}[htb]
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/corpus_human_comparison.pdf}
-    \caption{\textbf{
-        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
-        We see that the model ranking remains fairly consistent across both sets}
-    \label{fig:performance_corpus_and_tiny}
-    \script{corpus_humanset_performance.py}
-\end{figure}
+% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
+% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
 
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/model_overall_cal.pdf}
+%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+%     \script{requires_calculation_plot.py}
+%     \label{fig:requires_cal}
+% \end{figure}
 
@@
 \end{figure}
+
+% \begin{table}
+%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules.
+}
+%     \begin{tabularx}{\textwidth}{XXXX XXXX}
+%         \toprule
+%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\
+%         \midrule
+%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
+%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
+%         \midrule
+%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
+%         \midrule
+%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
+%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{tab:correlation_coefficients}
+% \end{table}
 
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
+Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
+Scores of knowledge-intensive are generally lower than those of Calculation and Reasoning. Reasoning tends to be the strongest skill among the three.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+Models show the strongest performance in reasoning tasks, which suggests that current training approaches are good at developing logical/reasoning capabilities.
+Larger models consistently perform better.
+However, there's evidence that well-trained smaller models can compete with larger models.
+Performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-28 src/tex/appendix.tex
Update appendix.tex

@@
 For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo).
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
 Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.

### 2024-10-28 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 
-\oone model consistently leads across both \ChemBench corpus and \chembenchmini.
+\oone consistently leads across both \ChemBench corpus and \chembenchmini.
 \ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
@@
 Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
-Scores of knowledge-intensive are generally lower than those of Calculation and Reasoning. Reasoning tends to be the strongest skill among the three.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
 

### 2024-10-28 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-Models show the strongest performance in reasoning tasks, which suggests that current training approaches are good at developing logical/reasoning capabilities.
-Larger models consistently perform better.
-However, there's evidence that well-trained smaller models can compete with larger models.
-Performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
 

### 2024-10-28 src/tex/appendix.tex
Update appendix.tex

@@
 For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
 One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
@@
 Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, this type of knowledge should be integrated with sound machine learning approaches.
+For example, \GalacticaOneTwentyB showed the lowest perforamnce in our study, even though it was specifically trained on scientific text.
+This suggest that the knowledge embedded in papers was not leveraged correctly.
+
 Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.

### 2024-10-28 src/tex/appendix.tex
Update appendix.tex
feat: add comments about no data from benchs (#80)
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-28 src/tex/appendix.tex
@@
 
+\subsubsection{Curation workflow}
+Questions were added via pull requests to the \chembench repository on GitHub.
+This allowed for a manual review of each question by expert reviewers (with backgrounds in chemistry, materials science, chemical engineering, and physics).
+The reviews were conducted directly on the GitHub platform, where our entire curation history is also publicly available.
+
+The general guidelines followed by the reviewer are the following:
+
+\begin{itemize}
+    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
+    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
+    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
+    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
+    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
+    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
+They should add value by expanding the breadth of the dataset.
+That is, questions already multiple (>10) times in the corpus in a similar form are rejected.
+\end{itemize}
+
+Reviewers also solved the questions to verify the answers.
+They also performed web searches to ensure questions were not easily found online.
+The reviewers often guided the revision process to ensure the question aligned with the guidelines.
+Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question.
+Reviewers also provide feedback on the skill and difficulty annotations.
+
+In addition to the manual review, we also performed automated checks to ensure the quality of the questions.
+The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
+
+While adding questions from existing benchmarks might seem to be another good source of semi-automatically generated data, we prioritized the diversity of the data and avoided data contamination while addressing the guidelines above and in \Cref{sec:desired-properties}.
+However, even though we decided not to include questions from other previously published chemistry-focused benchmarks into the \chembench corpus, the framework is flexible enough to be readily extended with questions from other benchmarks.
+
+\subsubsection{Composition}
+
 \Cref{fig:cb_humanset} shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
 
-%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
-%We see that the questions are generally complex to read.
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
-%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
-It is calculated based on the average number of syllables per word and words per sentence.
-The higher the score, the easier the text is to read.
-The distribution of the questions' scores is shown in the histogram.
-}
-%     \label{fig:flesch_kincaid_reading_ease}
-%     \script{wordcloud.py}
-% \end{figure}
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
-%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
-%     \label{fig:question_count_barplot_mcq_vs_general}
-%     \script{plot_statistics.py}
-% \end{figure}
-
 The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
+\begin{table}
+    \centering
+    \small
     \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
-
+    \label{tab:chembench_corpus_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Question Type & \% & Example Questions  \\
+            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
             \midrule
-        Chemical Preference & 35.0 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            Which of the following analytical methods is most appropriate for performing a survey analysis of a solid sample containing various metals? \\
+            A. X-ray fluorescence analysis \\
+            B. Differential pulse polarography \\
+            C. Flame-atomic absorption spectroscopy \\
+            D. Gas chromatography with flame ionization detector \\
+            E. Hydride generation atomic absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-        & & B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-        Technical Chemistry & 1.5 & Which of the following statements is true about the different types of ideal reactors? \\
-        & & A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
-        & & B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
-        & & C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
-        & & D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
             \midrule
-        Organic Chemistry & 15.5 & What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]?
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \midrule
+            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \midrule
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+            A. \ce{F-} forms a stable bonds to the metal ions \\
+            B. \ce{HF} has a better water solubility than \ce{HCl} \\
+            C. \ce{HF} has a higher boiling point than \ce{HCl} \\
+            D. \ce{HF} is a weaker acid than \ce{HCl} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \midrule
+            What is the reaction mechanism that describes the following reaction (represented using reaction SMILES) [START\_RXNSMILES]CCCl.CO[Na]>>[Na]Cl.CCOC\-[END\_RXNSMILES]? \\
+            A. $E_1$ \\
+            B. $E_{cb}$ \\
+            C. $S_N1$ \\
+            D. $S_N2$ \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \midrule
+            The Born-Oppenheimer (BO) approximation is widely used in computational chemistry, but its accuracy can vary depending on the system.
+Among the following options, for which system is the Born-Oppenheimer approximation likely to be least applicable?
 \\
-        & & A. $E_1$ \\
-        & & B. $E_{cb}$ \\
-        & & C. $S_N1$ \\
-        & & D. $S_N2$ \\
+            A. \ce{C60} \\
+            B. \ce{CH4} \\
+            C. \ce{Fe(CO)5} \\
+            D. \ce{H2+} \\
+            E. \ce{NaCl} \\
             \midrule
-        Toxicity/Safety & 24.2 & Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \midrule
+            Which of the following statements is true about the different types of ideal reactors? \\
+            A. In a batch reactor, the composition is uniformly mixed and remains the same throughout the reactor and at the exit \\
+            B. In a batch reactor, the fluid passes through the reactor with no mixing of earlier and later entering fluid \\
+            C. In a mixed flow reactor, the composition changes with time but is uniform everywhere within the reactor \\
+            D. In a plug flow reactor, the fluid moves in single flow through the reactor with no mixing and no overtaking \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \midrule
+            Pindolol and propranolol are (relatively nonselective) antagonists at $\beta_1$- and $\beta_2$-adrenoceptors.
 However, pindolol is a partial agonist, whereas propranolol is a pure antagonist.
@@
 
-In addition, as shown in \Cref{tab:chembench_corpus_cognitive} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+The plot shows a balanced distribution of the required skills in the \chembench corpus.
+\Cref{tab:cb_skillset} shows an example question for each skill category.
 
-\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
-        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
+    \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
+    \label{fig:cb_skillset}
+\end{figure}
 
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
+Note that the total count in this table is bigger than the \chembench corpus.
+This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
+Reasoning and Calculation.} \\
+    \label{tab:chembench_corpus_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
-        Required Skills & Number of Questions & Example Question \\
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
             \midrule
-        Knowledge & 919 & Which of the following salts is an acidic salt? \\
-        & & A. \ce{NH4Cl} \\
-        & & B. \ce{Na2CO3} \\
-        & & C. \ce{NaH2PO4} \\
-        & & D. \ce{Zn(OH)Cl} \\
+            Which of the following salts is an acidic salt? \\
+            A. \ce{NH4Cl} \\
+            B. \ce{Na2CO3} \\
+            C. \ce{NaH2PO4} \\
+            D. \ce{Zn(OH)Cl} \\
             \midrule
-        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-        & & A. By atomic force microscopy \\
-        & & B. By transform infrared spectroscopy \\
-        & & C. By measuring a photoluminescence spectrum \\
-        & & D. By absorption spectroscopy \\
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
             \midrule
-        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+            A. By atomic force microscopy \\
+            B. By transform infrared spectroscopy \\
+            C. By measuring a photoluminescence spectrum \\
+            D. By absorption spectroscopy \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
 Which of the following two candidates would you prefer for further development?
 \\
-        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
-        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
+            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
+            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
             \midrule
-        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution.
 The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}.
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
-
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
 
@@
 
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
+The relative performance difference between both is quite similar across most models.
+This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+
+An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+
 \begin{figure}[htb]
@@
 
-% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
-% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/corpus_human_comparison.pdf}
+    \caption{\textbf{
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
+        We see that the model ranking remains fairly consistent across both sets}  
+    \label{fig:performance_corpus_and_tiny}
+    \script{corpus_humanset_performance.py}
+\end{figure}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/model_overall_cal.pdf}
-%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-%     \script{requires_calculation_plot.py}
-%     \label{fig:requires_cal}
-% \end{figure}
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
-Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
-Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
-
-Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
-However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
-
-\clearpage
 \subsection{Leaderboard}

### 2024-10-28 src/tex/ms.tex
feat: add comments about no data from benchs (#80)
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-28 src/tex/ms.tex
@@
 Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.
+Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.}
     \variable{output/sources_table.tex}

### 2024-10-29 src/tex/appendix.tex
keep chembenchmini version

--- 2024-10-28 src/tex/appendix.tex
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions answered correctly by the models.
+    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
+    }
+    \label{fig:all_questions_models_completely_correct_radar_overall}
+    \script{analyze_model_reports.py}
+\end{figure}
+
+To further investigate the performance of the models, we also compared the performance on different data sources.
+Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
+In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic.pdf}
+    \caption{\textbf{Fraction of correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none of the questions were answered correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
+
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
+The relative performance difference between both is quite similar across most models.
+This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+
+An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic_tiny.pdf}
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic_tiny}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/corpus_human_comparison.pdf}
+    \caption{\textbf{
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
+        We see that the model ranking remains fairly consistent across both sets}  
+    \label{fig:performance_corpus_and_tiny}
+    \script{corpus_humanset_performance.py}
+\end{figure}
 
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
-    }
-    \label{fig:all_questions_models_completely_correct_radar_overall}
-    \script{analyze_model_reports.py}
-\end{figure}
-
-To further investigate the performance of the models, we also compared the performance on different data sources.
-Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
-In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
-Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic}
-    \script{analyze_performance_per_source.py}
-\end{figure}
-
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered completely correctly by the models.
-A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
-        We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic_tiny}
-    \script{analyze_performance_per_source.py}
-\end{figure}
-
-% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
-% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
-
-% \begin{figure}
-%     \centering
-%     \includegraphics{figures/model_overall_cal.pdf}
-%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
-%     \script{requires_calculation_plot.py}
-%     \label{fig:requires_cal}
-% \end{figure}
-
 \clearpage
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
+    \caption{\textbf{Distribution of human scores.}The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
 A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.}
+A score of zero would mean that no question was answered correctly.
+    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.}
     \label{fig:human_score_distribution}
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
-Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
-Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
-
-Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
-However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
-
-\clearpage
 \subsection{Leaderboard}
+\label{sec:leaderboard}
 Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}

### 2024-10-29 src/tex/ms.tex
keep chembenchmini version

--- 2024-10-28 src/tex/ms.tex
@@
 This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
-This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.
+This indicates that there is still room for improvement of chemical \glspl{llm} by training them on more specialized data sources or by integrating them with specialized databases.
 
 In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}).
-This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.
+This is in line with observations in other domains but also indicates that chemical \glspl{llm} could, to some extent, be further improved by scaling them up.
 
@@
     \caption{\textbf{Performance of the models and humans on the different topics on \chembenchmini.} The radar plot shows the performance of the models and humans on the different topics of \chembenchmini.
-The performance is measured as the fraction of questions that were answered completely correctly by the models.
+The performance is measured as the fraction of questions that were answered correctly by the models.
     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
-    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
+    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
     }
@@
 Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.
-Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.}
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
     \variable{output/sources_table.tex}

### 2024-10-29 src/tex/appendix.tex
change tiny to chembenchmini

@@
     \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+    \caption{\textbf{Fraction of completely correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
 The performance is measured as the fraction of questions answered correctly by the models.

### 2024-10-29 src/tex/appendix.tex
chore: change the example for reasoning

--- 2024-10-28 src/tex/appendix.tex
@@
 
-While adding questions from existing benchmarks might seem to be another good source of semi-automatically generated data, we prioritized the diversity of the data and avoided data contamination while addressing the guidelines above and in \Cref{sec:desired-properties}.
-However, even though we decided not to include questions from other previously published chemistry-focused benchmarks into the \chembench corpus, the framework is flexible enough to be readily extended with questions from other benchmarks.
-
 \subsubsection{Composition}
@@
     \small
-    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
+    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
     \label{tab:chembench_corpus_topic}
@@
 \\
-            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
@@
             \midrule
-            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
+            What is the oxidation number of the metal in the compound \ce{[ZrF_{7}]3-}? \\
             \midrule
@@
             \midrule
-            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose HF over \ce{HCl} for this purpose? \\
             A. \ce{F-} forms a stable bonds to the metal ions \\
@@
             \midrule
-            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
-            A. By atomic force microscopy \\
-            B. By transform infrared spectroscopy \\
-            C. By measuring a photoluminescence spectrum \\
-            D. By absorption spectroscopy \\
+            You make the following experimental observations: The sample consisted of small colorless/white crystals, wine-red transparent crystals, turquoise transparent crystals, and violet flakes.
+Most components dissolved in cold water, with turquoise crystals dissolving upon heating.
+Dimethylglyoxime produced a raspberry-red color with dissolved turquoise crystals.
+Violet flakes turned blue on contact and showed red/violet colors in flame tests.
+An oxidation melt turned green.
+No red coloration occurred with \ce{KSCN}, but a blue organic phase formed with acetone.
+Potassium ferrocyanide showed no changes.
+\ce{KOH} produced a brown/black precipitate.
+Alizarin-S testing resulted in red coloration after acidification.
+Heating with cobalt nitrate produced a bluish color without luster.
+Lead dioxide and sulfuric acid oxidation formed a violet solution.
+No changes occurred in chromate oxidation or titanium detection attempts.
+\ce{BaCl2} formed precipitates in both original solution and soda extract.
+Lunge's reagent showed red coloration around zinc granules.
+Silver nitrate formed white precipitates that dissolved in \ce{NH3} and reappeared with \ce{HNO3} acidification.
+The separation procedure began with the aqueous solution being treated with \ce{(NH4)2CO3} solution until a persistent turbidity appeared, followed by addition of dilute \ce{HCl} until the turbidity disappeared.
+The pH was then adjusted to 5-6 with \ce{NH3} (2 M) and heated on a water bath.
+Three spatula tips of urotropin were added, resulting in a white, rather flocculent precipitate.
+This precipitate was centrifuged off and washed with warm water.
+In the Urotropin group, the residue was dissolved in dilute \ce{HCl} and diluted with water.
+An alkaline precipitation was performed, resulting in a brown coloration even after repeated washing of the urotropin precipitation.
+This precipitate dissolved in 2 M \ce{HCl}, but showed no change when treated with either \ce{KSCN} or potassium ferrocyanide.
+In the \ce{(NH4)2S} group, addition of dimethylglyoxim produced a very flocculent red precipitate that rose to the top of the test tube.
+When \ce{H2S} was introduced to the centrifugate, a black precipitation occurred.
+The precipitates did not completely dissolve in \ce{HCl}.
+Treatment with \ce{KSCN} resulted in a blue organic phase when layered with acetone.
+An alkaline precipitation produced a brown/black residue.
+No precipitation was observed when \ce{H2S} was introduced to the supernatant liquid after the alkaline precipitation.
+What ions are in the sample?
+\\
+            A. \ce{Al^{3+}} \\
+            B. \ce{Co^{2+}} \\
+            C. \ce{Cr^{3+}} \\
+            D. \ce{Fe^{3+}} \\
+            E. \ce{Mn^{2+}} \\
+            F. \ce{Ni^{2+}} \\
+            G. \ce{Zn^{2+}} \\
             \midrule
@@
 \end{table}
-
 \normalsize
 
-In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
 The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
@@
     \small
-    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
-    \label{tab:chembench_corpus_models_failed_topic}
+    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
+    \label{tab:chembench_corpus_cognitive}
     {\fontsize{8}{9}\selectfont
@@
             \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
+            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
+            A. Cyclohexene \\
+            B. Cyclooctene \\
+            C. Decalin \\
+            D. Ethyl ether \\
+            E. Isopropyl ether \\
+            F. Para-dioxane \\
+            G. Tetralin \\
             \midrule
-            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
-            A. DESI-mass spectrometry \\
-            B. XPS \\
-            C. atomic force microscopy \\
-            D. scanning tunneling microscopy \\
-            E. secondary ion mass spectrometry \\
+            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
             \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
+            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
+            A. The intersection with y-axis shift to higher numbers \\
+            B. The intersection with y-axis shift to lower numbers \\
+            C. The slope decreases \\
+            D. The slope increases \\
             \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
+            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
+A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
+What is the number of moles of acid in the sample?
 \\
-            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
+            What is hydrofluoric acid used for? \\
+            A. For cleaning facades \\
+            B. For etching glass \\
+            C. For filling accumulators \\
+            D. For soldering \\
             \midrule
-            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
-            A. \ce{SCl6} does not exist because of steric effects \\
-            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
-            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
-            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
+            How many chiral centers are present in the molecular structure of citric acid? \\
             \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
+            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
+            A. 10 \\
+            B. 14 \\
+            C. 18 \\
+            D. 23 \\
+            E. 9 \\
             \midrule
-            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
+What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
+\\
             \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
-            \midrule
-            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
-            A. All dimension of the material must be smaller than exciton Bohr radius \\
-            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
-            C. The movement od charge carriers is limited in all spatial dimensions \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
-            \midrule
-            What makes acrylates more reactive than typical hydrocarbons? \\
-            A. The carbonyl group \\
-            B. The double bond \\
-            C. The hydroxyl group \\
-            D. The vinyl group \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
-            \midrule
-            Why are polar molecules more soluble in water than non-polar molecules? \\
-            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
-            B. The entropy of the system increases when non-polar molecules dissolve in water \\
-            C. Water has a dipole and can easier isolate polar molecules from each other \\
-            D. Water makes stable dipole-dipole interactions with polar molecules \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
-            \midrule
-            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
-            A. Broader product diameter distribution than pellets or extrudates \\
-            B. Facilitates reactor upscaling \\
-            C. It is also called snow-ball \\
-            D. Not suitable to prepare egg-shell catalysts \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
-            \midrule
-            What determines the flammability of polyols? \\
-            A. Their chemical structure \\
-            B. Their low volatility \\
-            C. Their odor threshold \\
-            D. Their reaction with acids \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
-
-\normalsize
-
-To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
-The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
-    \label{tab:chembench_corpus_models_failed_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
-            \midrule
-            Which of the following are true for phosphoric acid? \\
-            A. Is a weak acid, so it is not toxic \\
-            B. Should never be tasted \\
-            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
-            \midrule
-            Which of the following elements has the largest atomic radius? \\
-            A. \ce{Fr} \\
-            B. \ce{He} \\
-            C. \ce{Li} \\
-            D. \ce{Rb} \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
-            \midrule
-            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
-Which of the following two candidates would you prefer for further development?
-\\
-            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
-            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
-            \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
-            \midrule
-            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
-What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
-\\
+            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
+            A. 10.0\% \\ 
+            B. 6.6\% \\
+            C. 7.0\% \\
+            D. 94\% \\
             \bottomrule

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 \end{figure}
+
+%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
+%We see that the questions are generally complex to read.
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
+%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read.
+It is calculated based on the average number of syllables per word and words per sentence.
+The higher the score, the easier the text is to read.
+The distribution of the questions' scores is shown in the histogram.
+}
+%     \label{fig:flesch_kincaid_reading_ease}
+%     \script{wordcloud.py}
+% \end{figure}
+
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
+%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
+%     \label{fig:question_count_barplot_mcq_vs_general}
+%     \script{plot_statistics.py}
+% \end{figure}
 
@@
     \includegraphics{figures/overall_performance.pdf}
-    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered correctly by the models.
+    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models.
 Scores computed on the entire \chembench corpus.}
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
-
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
-    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
-The performance is measured as the fraction of questions answered correctly by the models.
-    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
-    }
-    \label{fig:all_questions_models_completely_correct_radar_overall}
-    \script{analyze_model_reports.py}
-\end{figure}
-
-To further investigate the performance of the models, we also compared the performance on different data sources.
-Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
-In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
-Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic.pdf}
-    \caption{\textbf{Fraction of correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered correctly by the models.
-A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none of the questions were answered correctly.
-        We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic}
-    \script{analyze_performance_per_source.py}
-\end{figure}
-
-\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
-
-\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
-The relative performance difference between both is quite similar across most models.
-This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
-
-An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/performance_per_topic_tiny.pdf}
-    \caption{\textbf{Fraction of completely correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
-The performance is measured as the fraction of questions answered correctly by the models.
-A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
-        We see that the performance of the models varies significantly between the different data sources.
-For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
-However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
-    }
-    \label{fig:performance_per_topic_tiny}
-    \script{analyze_performance_per_source.py}
-\end{figure}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics{figures/corpus_human_comparison.pdf}
-    \caption{\textbf{
-        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
-        We see that the model ranking remains fairly consistent across both sets}  
-    \label{fig:performance_corpus_and_tiny}
-    \script{corpus_humanset_performance.py}
-\end{figure}
 
@@
 
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
+    }
+    \label{fig:all_questions_models_completely_correct_radar_overall}
+    \script{analyze_model_reports.py}
+\end{figure}
+
+To further investigate the performance of the models, we also compared the performance on different data sources.
+Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
+In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic.pdf}
+    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic_tiny.pdf}
+    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered completely correctly by the models.
+A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic_tiny}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
+% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).
+
+% \begin{figure}
+%     \centering
+%     \includegraphics{figures/model_overall_cal.pdf}
+%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
+%     \script{requires_calculation_plot.py}
+%     \label{fig:requires_cal}
+% \end{figure}
+
 \clearpage
@@
     \script{plot_human_score_distribution.py}
-    \caption{\textbf{Distribution of human scores.}The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
+    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
     Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.
 A score of one would mean that a volunteer answered all questions correctly.
-A score of zero would mean that no question was answered correctly.
-    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.}
+A score of zero would mean that no question was answered correctly.}
     \label{fig:human_score_distribution}
@@
 
-We also recorded the time humans took to answer the questions (\Cref{fig:human_timing}). This time is the time from the question being displayed to the human to the human submitting the answer.
+We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
 
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
+Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, this type of knowledge should be integrated with sound machine-learning approaches.
+For example, \GalacticaOneTwentyB showed the lowest performance in our study, even though it was specifically trained on scientific text.
+This suggests that the knowledge embedded in papers and textbooks has not been fully leveraged.
+
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
+
+\clearpage
 \subsection{Leaderboard}
-\label{sec:leaderboard}
 Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}

### 2024-10-29 src/tex/appendix.tex
Update appendix.tex

@@
 However, this type of knowledge should be integrated with sound machine-learning approaches.
-For example, \GalacticaOneTwentyB showed the lowest performance in our study, even though it was specifically trained on scientific text.
-This suggests that the knowledge embedded in papers and textbooks has not been fully leveraged.
+For example, even \PaperQATwo could not outperform the \gls{llm} using the tools, even though the model now would have access to literature evidence.
+This suggests we might need an overhaul of how we approach agent building, and how we train models for science applications, since clearly the valuable information stored in textbooks and papers, that can be used to answer questions is not fully leveraged by any of the state-of-the-art models and systems.
 

### 2024-10-29 src/tex/appendix.tex
fix changes from tiny to chembenchmini

--- 2024-10-29 src/tex/appendix.tex
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
+    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus.
+The performance is measured as the fraction of questions answered correctly by the models.
+    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
+    }
+    \label{fig:all_questions_models_completely_correct_radar_overall}
+    \script{analyze_model_reports.py}
+\end{figure}
+
+To further investigate the performance of the models, we also compared the performance on different data sources.
+Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
+In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
+Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic.pdf}
+    \caption{\textbf{Fraction of correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none of the questions were answered correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.
+
+\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset.
+The relative performance difference between both is quite similar across most models.
+This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.
+
+An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/performance_per_topic_tiny.pdf}
+    \caption{\textbf{Fraction of correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources.
+The performance is measured as the fraction of questions answered correctly by the models.
+A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
+        We see that the performance of the models varies significantly between the different data sources.
+For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans.
+However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
+    }
+    \label{fig:performance_per_topic_tiny}
+    \script{analyze_performance_per_source.py}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/corpus_human_comparison.pdf}
+    \caption{\textbf{
+        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
+        We see that the model ranking remains fairly consistent across both sets}  
+    \label{fig:performance_corpus_and_tiny}
+    \script{corpus_humanset_performance.py}
+\end{figure}
 
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
-
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
-Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
-Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
-
-Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-However, this type of knowledge should be integrated with sound machine-learning approaches.
-For example, \GalacticaOneTwentyB showed the lowest performance in our study, even though it was specifically trained on scientific text.
-This suggests that the knowledge embedded in papers and textbooks has not been fully leveraged.
-
-Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
-However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
-
-\clearpage
 \subsection{Leaderboard}
+\label{sec:leaderboard}
 Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>
Update appendix.tex

--- 2024-10-29 src/tex/appendix.tex
@@
 One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
-Newer versions consistently outperform their predecessors, indicating improvements in scientific understanding of these model families.
+Newer versions consistently outperform their predecessors.
 Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
@@
 However, this type of knowledge should be integrated with sound machine-learning approaches.
-For example, even \PaperQATwo could not outperform the \gls{llm} using the tools, even though the model now would have access to literature evidence.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
 This suggests we might need an overhaul of how we approach agent building, and how we train models for science applications, since clearly the valuable information stored in textbooks and papers, that can be used to answer questions is not fully leveraged by any of the state-of-the-art models and systems.

### 2024-10-29 src/tex/ms.tex
recompile

@@
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, such as for agents (more discussion in \Cref{sec:react-environment}), it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
 Although current systems might be far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.
@@
 Questions based on lecture notes or URL have been curated based on content presented in those resources.
-All questions have been rephrased, annotated, and reviewed before being added to the corpus.}
+All questions have been rephrased, annotated, and reviewed before being added to the corpus.
+}
+    %Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.
     \variable{output/sources_table.tex}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
-However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions should be included in training or finetuning.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
 

### 2024-10-29 src/tex/appendix.tex
add variables

@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of 0.48 on \ChemBench corpus, which is only 10\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 

### 2024-10-29 src/tex/appendix.tex

--- 2024-10-29 src/tex/appendix.tex
@@
     \small
-    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
+    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
     \label{tab:chembench_corpus_topic}
@@
 \\
-            A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
-            B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
@@
             \midrule
-            What is the oxidation number of the metal in the compound \ce{[ZrF_{7}]3-}? \\
+            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\
             \midrule
@@
             \midrule
-            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose HF over \ce{HCl} for this purpose? \\
+            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\
             A. \ce{F-} forms a stable bonds to the metal ions \\
@@
 
-In addition, as shown in \Cref{fig:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 The plot shows a balanced distribution of the required skills in the \chembench corpus.
@@
 This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.}
+Reasoning and Calculation.} \\
     \label{tab:chembench_corpus_cognitive}
@@
             \midrule
-            You make the following experimental observations: The sample consisted of small colorless/white crystals, wine-red transparent crystals, turquoise transparent crystals, and violet flakes.
-Most components dissolved in cold water, with turquoise crystals dissolving upon heating.
-Dimethylglyoxime produced a raspberry-red color with dissolved turquoise crystals.
-Violet flakes turned blue on contact and showed red/violet colors in flame tests.
-An oxidation melt turned green.
-No red coloration occurred with \ce{KSCN}, but a blue organic phase formed with acetone.
-Potassium ferrocyanide showed no changes.
-\ce{KOH} produced a brown/black precipitate.
-Alizarin-S testing resulted in red coloration after acidification.
-Heating with cobalt nitrate produced a bluish color without luster.
-Lead dioxide and sulfuric acid oxidation formed a violet solution.
-No changes occurred in chromate oxidation or titanium detection attempts.
-\ce{BaCl2} formed precipitates in both original solution and soda extract.
-Lunge's reagent showed red coloration around zinc granules.
-Silver nitrate formed white precipitates that dissolved in \ce{NH3} and reappeared with \ce{HNO3} acidification.
-The separation procedure began with the aqueous solution being treated with \ce{(NH4)2CO3} solution until a persistent turbidity appeared, followed by addition of dilute \ce{HCl} until the turbidity disappeared.
-The pH was then adjusted to 5-6 with \ce{NH3} (2 M) and heated on a water bath.
-Three spatula tips of urotropin were added, resulting in a white, rather flocculent precipitate.
-This precipitate was centrifuged off and washed with warm water.
-In the Urotropin group, the residue was dissolved in dilute \ce{HCl} and diluted with water.
-An alkaline precipitation was performed, resulting in a brown coloration even after repeated washing of the urotropin precipitation.
-This precipitate dissolved in 2 M \ce{HCl}, but showed no change when treated with either \ce{KSCN} or potassium ferrocyanide.
-In the \ce{(NH4)2S} group, addition of dimethylglyoxim produced a very flocculent red precipitate that rose to the top of the test tube.
-When \ce{H2S} was introduced to the centrifugate, a black precipitation occurred.
-The precipitates did not completely dissolve in \ce{HCl}.
-Treatment with \ce{KSCN} resulted in a blue organic phase when layered with acetone.
-An alkaline precipitation produced a brown/black residue.
-No precipitation was observed when \ce{H2S} was introduced to the supernatant liquid after the alkaline precipitation.
-What ions are in the sample?
-\\
-            A. \ce{Al^{3+}} \\
-            B. \ce{Co^{2+}} \\
-            C. \ce{Cr^{3+}} \\
-            D. \ce{Fe^{3+}} \\
-            E. \ce{Mn^{2+}} \\
-            F. \ce{Ni^{2+}} \\
-            G. \ce{Zn^{2+}} \\
+            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
+            A. By atomic force microscopy \\
+            B. By transform infrared spectroscopy \\
+            C. By measuring a photoluminescence spectrum \\
+            D. By absorption spectroscopy \\
             \midrule
@@
 \end{table}
+
 \normalsize
 
-In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
 The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
@@
     \small
-    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
-    \label{tab:chembench_corpus_cognitive}
+    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
+    \label{tab:chembench_corpus_models_failed_topic}
     {\fontsize{8}{9}\selectfont
@@
             \toprule
-            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
-            A. Cyclohexene \\
-            B. Cyclooctene \\
-            C. Decalin \\
-            D. Ethyl ether \\
-            E. Isopropyl ether \\
-            F. Para-dioxane \\
-            G. Tetralin \\
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
             \midrule
-            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
+            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
+            A. DESI-mass spectrometry \\
+            B. XPS \\
+            C. atomic force microscopy \\
+            D. scanning tunneling microscopy \\
+            E. secondary ion mass spectrometry \\
             \midrule
-            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
-            A. The intersection with y-axis shift to higher numbers \\
-            B. The intersection with y-axis shift to lower numbers \\
-            C. The slope decreases \\
-            D. The slope increases \\
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
             \midrule
-            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
-A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
-What is the number of moles of acid in the sample?
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
 \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-            What is hydrofluoric acid used for? \\
-            A. For cleaning facades \\
-            B. For etching glass \\
-            C. For filling accumulators \\
-            D. For soldering \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
             \midrule
-            How many chiral centers are present in the molecular structure of citric acid? \\
+            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
+            A. \ce{SCl6} does not exist because of steric effects \\
+            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
+            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
+            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
             \midrule
-            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
-            A. 10 \\
-            B. 14 \\
-            C. 18 \\
-            D. 23 \\
-            E. 9 \\
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
             \midrule
-            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
-What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
+            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
+            \midrule
+            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
+            A. All dimension of the material must be smaller than exciton Bohr radius \\
+            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
+            C. The movement od charge carriers is limited in all spatial dimensions \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
+            \midrule
+            What makes acrylates more reactive than typical hydrocarbons? \\
+            A. The carbonyl group \\
+            B. The double bond \\
+            C. The hydroxyl group \\
+            D. The vinyl group \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
+            \midrule
+            Why are polar molecules more soluble in water than non-polar molecules? \\
+            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
+            B. The entropy of the system increases when non-polar molecules dissolve in water \\
+            C. Water has a dipole and can easier isolate polar molecules from each other \\
+            D. Water makes stable dipole-dipole interactions with polar molecules \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
+            \midrule
+            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
+            A. Broader product diameter distribution than pellets or extrudates \\
+            B. Facilitates reactor upscaling \\
+            C. It is also called snow-ball \\
+            D. Not suitable to prepare egg-shell catalysts \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
+            \midrule
+            What determines the flammability of polyols? \\
+            A. Their chemical structure \\
+            B. Their low volatility \\
+            C. Their odor threshold \\
+            D. Their reaction with acids \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
+\normalsize
+
+To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
+The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
+    \label{tab:chembench_corpus_models_failed_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+            \midrule
+            Which of the following are true for phosphoric acid? \\
+            A. Is a weak acid, so it is not toxic \\
+            B. Should never be tasted \\
+            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+            \midrule
+            Which of the following elements has the largest atomic radius? \\
+            A. \ce{Fr} \\
+            B. \ce{He} \\
+            C. \ce{Li} \\
+            D. \ce{Rb} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
 \\
+            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
+            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
             \midrule
-            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
-            A. 10.0\% \\ 
-            B. 6.6\% \\
-            C. 7.0\% \\
-            D. 94\% \\
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
+\\
             \bottomrule
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
+Newer versions consistently outperform their predecessors.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, this type of knowledge should be integrated with sound machine-learning approaches.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
+This suggests we might need an overhaul of how we approach agent building, and how we train models for science applications, since clearly the valuable information stored in textbooks and papers, that can be used to answer questions is not fully leveraged by any of the state-of-the-art models and systems.
+
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-29 src/tex/appendix.tex
include RLHF comment
fix typo
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-However, this type of knowledge should be integrated with sound machine-learning approaches.
+However, innovation is also needed to incorporate this information into the systems.
 For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
-This suggests we might need an overhaul of how we approach agent building, and how we train models for science applications, since clearly the valuable information stored in textbooks and papers, that can be used to answer questions is not fully leveraged by any of the state-of-the-art models and systems.
+This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant systems and should be coupled with more domain-specific databases.
+However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks.
 
@@
 However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
+
+The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
 

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Martiño Rios-Garcia <147555961+MrtinoRG@users.noreply.github.com>

@@
 
+As commented above, \PaperQATwo was outperformed by the \gls{llm} powering it.
+Similarly, \ClaudeThreeFiveSonnet only slightly improved the results when equipped with some specific tools, while \GPTFourO performance decreased in this same tool environment (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
+Therefore, a robust chemical agent environment optimized to solve general chemical problems, instead of focusing on bibliographic search or lab work, can be useful in the future.
+
 \clearpage

### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
add specific question comments

The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).
Models also struggle with identifying the relevant chemical trends across various questions (e.g., ordering the entropy of compounds).

\gls{llm} also struggle with open-ended questions. Albeit in our study, we only investigated numerical open-ended questions, thus a part of these would require multi-step calculations, which can be challenging for \gls{llm}.

### 2024-10-29 src/tex/appendix.tex
chore: add example url in quideline (#85)

--- 2024-10-29 src/tex/appendix.tex
@@
     \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
+    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain (example \url{https://github.com/lamalab-org/chem-bench/pull/391#discussion_r1679276714})
     \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
-    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
     \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
@@
 \clearpage
-\subsection{Summary of trends and recommendations}
-
-\oone consistently leads across both \ChemBench corpus and \chembenchmini.
-\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
-These models show robust performance across all skills and difficulty levels.
-
-Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
-However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
-
-One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
-Newer versions consistently outperform their predecessors.
-Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
-
-Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
-However, innovation is also needed to incorporate this information into the systems.
-For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
-This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant systems and should be coupled with more domain-specific databases.
-However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks.
-
-Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
-However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
-
-The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
-
-\clearpage
 \subsection{Leaderboard}

### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
small fix
add reference

@@
 The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).
-Models also struggle with identifying the relevant chemical trends across various questions (e.g., ordering the entropy of compounds).
+Models also struggle with identifying the relevant chemical trends (e.g., ordering the entropy of compounds).
 
-\gls{llm} also struggle with open-ended questions.
-Albeit in our study, we only investigated numerical open-ended questions, thus a part of these would require multi-step calculations, which can be challenging for \gls{llm}.
+\glspl{llm} also struggle with open-ended questions.
+Albeit in our study, we only investigated numerical open-ended questions, thus a part of these would require multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}

### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
update
fix wording

--- 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
@@
 The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).
-Models also struggle with identifying the relevant chemical trends across various questions (e.g., ordering the entropy of compounds).
 
-\gls{llm} also struggle with open-ended questions.
-Albeit in our study, we only investigated numerical open-ended questions, thus a part of these would require multi-step calculations, which can be challenging for \gls{llm}.
+\glspl{llm} also struggle with questions that do are not \glspl{mcq}.
+Albeit in our study, we only investigated numerical outputs that can be parsed and scored.
+Thus, a part of these would involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}

### 2024-10-29 src/tex/appendix.tex
feat: add appendix section about React runs

@@
     \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
-    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain (example \url{https://github.com/lamalab-org/chem-bench/pull/391#discussion_r1679276714})
     \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
+    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
     \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
+Models with \enquote{T-one} in the name were run for Temperature=1, which allow to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to better answer the questions.
+However, we limit those systems to a maximum of ten calls to the LLM.
+This constraint led the systems to often not find the correct answer within the specified number of calls.
+In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
     \resizebox{\textwidth}{!}{
@@
 \clearpage
+\subsection{Tool augmented models} \label{sec:react-environment}
+In addition to directly prompting LLMs, we also investigated the performance of tool-augmented systems.
+For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or CoT\autocite{wei2023cot}).
+For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv API, the Wikipedia access, and a web search (using Brave browser API).
+We based the election of these tools based in the most used tools by the humans (see \Cref{fig:tool_use}).
+Additionally, we added two specific tools to convert IUPAC names to SMILES, and SMILES to IUPAC names.
+These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configurations.
+We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten LLM calls.
+
+While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools. 
+If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.
+On the other hand, the scores decreased question asking about the number of isomers or information contained in pictograms.
+For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
+By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
+This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.
+
+\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}

### 2024-10-29 src/tex/ms.tex
feat: add a hint in the conclusions

@@
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, such as for the ReAct systems (more discussion in \cref{sec:react-environment}), it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
 Although current systems might be far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.

### 2024-10-29 src/tex/appendix.tex
feat: add reference to the performance table

@@
 
-While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools. 
+While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
 If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
     \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
-Models with \enquote{T-one} in the name were run for Temperature=1, which allow to study the temperature effect in the benchmark.
-Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to better answer the questions.
-However, we limit those systems to a maximum of ten calls to the LLM.
-This constraint led the systems to often not find the correct answer within the specified number of calls.
+Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
+However, we limit those systems to ten calls to the \gls{llm}.
+This constraint often led the systems to not find the correct answer within the specified number of calls.
 In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
-Models with \enquote{T-one} in the name were run for Temperature=1, which allow to study the temperature effect in the benchmark.
-Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to better answer the questions.
-However, we limit those systems to a maximum of ten calls to the LLM.
-This constraint led the systems to often not find the correct answer within the specified number of calls.
-In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 \clearpage
-\subsection{Tool augmented models} \label{sec:react-environment}
-In addition to directly prompting LLMs, we also investigated the performance of tool-augmented systems.
-For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or CoT\autocite{wei2023cot}).
-For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv API, the Wikipedia access, and a web search (using Brave browser API).
-We based the election of these tools based in the most used tools by the humans (see \Cref{fig:tool_use}).
-Additionally, we added two specific tools to convert IUPAC names to SMILES, and SMILES to IUPAC names.
-These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configurations.
-We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten LLM calls.
-
-While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools. 
-If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.
-On the other hand, the scores decreased question asking about the number of isomers or information contained in pictograms.
-For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
-By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
-This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.
-
-\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone consistently leads across both \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
+Newer versions consistently outperform their predecessors.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, innovation is also needed to incorporate this information into the systems.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
+This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant systems and should be coupled with more domain-specific databases.
+However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks.
+
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
+
+The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
+
+As commented above, \PaperQATwo was outperformed by the \gls{llm} powering it.
+Similarly, \ClaudeThreeFiveSonnet only slightly improved the results when equipped with some specific tools, while \GPTFourO performance decreased in this same tool environment (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
+Therefore, a robust chemical agent framework optimized to solve general chemical problems instead of focusing on bibliographic search or lab work can be useful in the future.
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-29 src/tex/appendix.tex
fix: remove repetitive paragraph

@@
 This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant systems and should be coupled with more domain-specific databases.
-However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks.
+However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
 
@@
 The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
-
-As commented above, \PaperQATwo was outperformed by the \gls{llm} powering it.
-Similarly, \ClaudeThreeFiveSonnet only slightly improved the results when equipped with some specific tools, while \GPTFourO performance decreased in this same tool environment (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
-Therefore, a robust chemical agent framework optimized to solve general chemical problems instead of focusing on bibliographic search or lab work can be useful in the future.
 

### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
Update src/tex/sections/patterns_in_wrong_questions.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 
-\glspl{llm} also struggle with questions that do are not \glspl{mcq}.
-Albeit in our study, we only investigated numerical outputs that can be parsed and scored.
-Thus, a part of these would involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}
+\glspl{llm} also struggles with questions that require numerical outputs. Some of these involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 \subsection{Tool augmented models} \label{sec:react-environment}
-In addition to directly prompting LLMs, we also investigated the performance of tool-augmented systems.
+In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
 For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or CoT\autocite{wei2023cot}).

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or CoT\autocite{wei2023cot}).
+For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
 For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv API, the Wikipedia access, and a web search (using Brave browser API).

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
-For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv API, the Wikipedia access, and a web search (using Brave browser API).
+For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
 We based the election of these tools based in the most used tools by the humans (see \Cref{fig:tool_use}).

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
-We based the election of these tools based in the most used tools by the humans (see \Cref{fig:tool_use}).
-Additionally, we added two specific tools to convert IUPAC names to SMILES, and SMILES to IUPAC names.
+We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
+Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
 These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configurations.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
-These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configurations.
-We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten LLM calls.
+These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
+We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
 

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
-Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
-Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
-However, we limit those systems to ten calls to the \gls{llm}.
-This constraint often led the systems to not find the correct answer within the specified number of calls.
-In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 \clearpage
-\subsection{Tool augmented models} \label{sec:react-environment}
-In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
-For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
-We based the election of these tools based in the most used tools by the humans (see \Cref{fig:tool_use}).
-Additionally, we added two specific tools to convert IUPAC names to SMILES, and SMILES to IUPAC names.
-These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configurations.
-We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten LLM calls.
-
-While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
-If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.
-On the other hand, the scores decreased question asking about the number of isomers or information contained in pictograms.
-For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
-By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
-This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.
-
-\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone consistently leads across both the \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
+Newer versions consistently outperform their predecessors.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, innovation is also needed to incorporate this information into the systems.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
+This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant systems and should be coupled with more domain-specific databases.
+However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
+
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
+
+The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 Newer versions consistently outperform their predecessors.
-Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions. Performance on reasoning-intensive questions in \chembench tends to be the highest.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions.
 
@@
 
-The post-training (e.g., \gls{rlhf}) we currently do does not equip models with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
+The post-training (e.g., \gls{rlhf}) of the current models does not equip them with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
 

### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
Update src/tex/sections/patterns_in_wrong_questions.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
-The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).
+The models seem unable to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule, or identifying a symmetry group).
 
-\glspl{llm} also struggles with questions that require numerical outputs. Some of these involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}
+\glspl{llm} also struggles with questions in \chembench that require numerical outputs.
+Some of these involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 
-While for \ClaudeThreeFiveSonnet the overall performance increased (less than 1\%), they show a sloped decrease for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
+While for \ClaudeThreeFiveSonnet the overall performance increased (by less than 1\%), we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}).
 If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 While for \ClaudeThreeFiveSonnet the overall performance increased (by less than 1\%), we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}).
-If we study the results by each type of question, we observe an improvement for questions regarding electron count or point group questions.
-On the other hand, the scores decreased question asking about the number of isomers or information contained in pictograms.
+If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
+However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
 For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
-For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
+For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions. 
 By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

@@
 For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions. 
-By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
+By studying the reasoning path for these cases, we found that the error results from the models responding in a format that the LangChain framework with the default ReAct loop cannot handle. 
 This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex

--- 2024-10-29 src/tex/appendix.tex
@@
 \clearpage
-\subsection{Tool augmented models} \label{sec:react-environment}
-In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
-For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
-We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
-Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
-These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
-We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
-
-While for \ClaudeThreeFiveSonnet the overall performance increased (by less than 1\%), we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}).
-If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
-However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
-For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
-By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
-This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.
-
-\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}
@@
 \clearpage
+\subsection{Summary of trends and recommendations}
+
+\oone consistently leads across both the \ChemBench corpus and \chembenchmini.
+\ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
+These models show robust performance across all skills and difficulty levels.
+
+Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct).
+However, smaller but well-designed models can compete with larger models.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+
+One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT).
+Newer versions consistently outperform their predecessors.
+Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions.
+
+Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration.
+However, innovation is also needed to incorporate this information into the systems.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
+This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant papers and should be coupled with more domain-specific databases.
+However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks (see \Cref{sec:react-environment} for further discussion about the ReAct environment).
+
+Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams.
+However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.
+
+The post-training (e.g., \gls{rlhf}) of the current models does not equip them with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
+
+\clearpage
 \subsection{Leaderboard}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 By studying the reasoning path for these cases, we found that the error results from the models responding in a format that the LangChain framework with the default ReAct loop cannot handle. 
-This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.
+This indicates that agent frameworks need optimization to be more robust. It involves not only equipping the \glspl{llm} with tools but also necessitates engineering efforts to create robust systems.
 

### 2024-10-29 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

--- 2024-10-29 src/tex/ms.tex
@@
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, such as for agents (more discussion in \cref{sec:react-environment}), it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
 Although current systems might be far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex
Co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>

@@
 
-While for \ClaudeThreeFiveSonnet the overall performance increased (by less than 1\%), we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}).
+While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
 If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.

### 2024-10-29 src/tex/appendix.tex
feat: add Langchain citation

--- 2024-10-29 src/tex/appendix.tex
@@
 These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
-We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
+We implemented the systems using Langchain\autocite{Chase_LangChain_2022} with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
 

### 2024-10-29 src/tex/appendix.tex
feat: add React text (#86)
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>
Co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>

--- 2024-10-29 src/tex/appendix.tex
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
+Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
+However, we limit those systems to ten calls to the \gls{llm}.
+This constraint often led the systems to not find the correct answer within the specified number of calls.
+In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
     \resizebox{\textwidth}{!}{
@@
 \clearpage
+\subsection{Tool augmented models} \label{sec:react-environment}
+In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
+For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
+For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
+We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
+Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
+These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
+We implemented the systems using Langchain\autocite{Chase_LangChain_2022} with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
+
+While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
+If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
+However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
+For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions. 
+By studying the reasoning path for these cases, we found that the error results from the models responding in a format that the LangChain framework with the default ReAct loop cannot handle. 
+This indicates that agent frameworks need optimization to be more robust. It involves not only equipping the \glspl{llm} with tools but also necessitates engineering efforts to create robust systems.
+
+\clearpage
 \subsection{Confidence estimates} \label{sec:confidence_estimates}

### 2024-10-29 src/tex/appendix.tex
recompile

@@
 \subsection{Related work}
-Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
+Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024}, \textcite{rein2023gpqagraduatelevelgoogleproofqa} fail to comply with most of the requirements stipulated above.
 While these benchmarks could provide valuable insights in the short term, they cannot follow the rapid additions to the \gls{llm} space.

### 2024-10-29 src/tex/ms.tex
recompile

--- 2024-10-29 src/tex/ms.tex
@@
 
-We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
+We curated more than 2,700 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
 However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
@@
 All questions have been rephrased, annotated, and reviewed before being added to the corpus.
-Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.}
+}
+    %Note that the chemical preference questions are the main corpus of the semi-automatically generated group with \variable{output/question_count_per_dir/json_file_counts_preference.txt} questions of the total \variable{output/automatically_generated.txt}.
     \variable{output/sources_table.tex}

### 2024-10-29 src/tex/appendix.tex
recompile

@@
 
-While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
+While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \gls{llm} without tools (\Cref{tab:performance_table}).
 If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.

### 2024-10-29 src/tex/appendix.tex
recompile

@@
 
-\oone consistently leads across both the \ChemBench corpus and \chembenchmini.
+\oone consistently leads across both the \chembench corpus and \chembenchmini.
 \ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 

### 2024-10-29 src/tex/ms.tex

--- 2024-10-29 src/tex/ms.tex
@@
 
-While our findings indicate many areas for further improvement of \gls{llm}-based systems, such as for agents (more discussion in \cref{sec:react-environment}), it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
+While our findings indicate many areas for further improvement of \gls{llm}-based systems, such as for agents (more discussion in \Cref{sec:react-environment}), it is also important to realize that clearly defined metrics have been the key to the progress of many fields of \gls{ml}, such as computer vision.
 Although current systems might be far from reasoning like a chemist, our \chembench framework will be a stepping stone for developing systems that might come closer to this goal.

### 2024-10-30 src/tex/appendix.tex
fix some crefs

@@
 The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{tab:cb_skillset} shows an example question for each skill category.
+\Cref{fig:cb_skillset} shows an example question for each skill category.
 
@@
 In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
+The following \Cref{tab:chembench_corpus_cognitive} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
 
@@
 An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
 

### 2024-10-30 src/tex/appendix.tex
chore: reapply reference changes by Adrian

@@
 The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{fig:cb_skillset} shows an example question for each skill category.
+\Cref{tab:chembench_corpus_cognitive} shows an example question for each skill category.
 
@@
 In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-The following \Cref{tab:chembench_corpus_cognitive} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
+\Cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
 

### 2024-10-30 src/tex/appendix.tex
fix: update circle plots

--- 2024-10-29 src/tex/appendix.tex
@@
 \subsection{Related work}
-Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024}, \textcite{rein2023gpqagraduatelevelgoogleproofqa} fail to comply with most of the requirements stipulated above.
+Existing benchmarks such as those from \textcite{guo2023large}, \textcite{sun2023scieval}, \textcite{Schulze_Balhorn_2024}, \textcite{Cai_2024} fail to comply with most of the requirements stipulated above.
 While these benchmarks could provide valuable insights in the short term, they cannot follow the rapid additions to the \gls{llm} space.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V2.pdf}
     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
             \toprule
-            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
-            A. Cyclohexene \\
-            B. Cyclooctene \\
-            C. Decalin \\
-            D. Ethyl ether \\
-            E. Isopropyl ether \\
-            F. Para-dioxane \\
-            G. Tetralin \\
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
             \midrule
-            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
+            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
+            A. DESI-mass spectrometry \\
+            B. XPS \\
+            C. atomic force microscopy \\
+            D. scanning tunneling microscopy \\
+            E. secondary ion mass spectrometry \\
             \midrule
-            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
-            A. The intersection with y-axis shift to higher numbers \\
-            B. The intersection with y-axis shift to lower numbers \\
-            C. The slope decreases \\
-            D. The slope increases \\
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
             \midrule
-            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
-A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
-What is the number of moles of acid in the sample?
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
 \\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
             \midrule
-            What is hydrofluoric acid used for? \\
-            A. For cleaning facades \\
-            B. For etching glass \\
-            C. For filling accumulators \\
-            D. For soldering \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
             \midrule
-            How many chiral centers are present in the molecular structure of citric acid? \\
+            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
+            A. \ce{SCl6} does not exist because of steric effects \\
+            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
+            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
+            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
             \midrule
-            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
-            A. 10 \\
-            B. 14 \\
-            C. 18 \\
-            D. 23 \\
-            E. 9 \\
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
             \midrule
-            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
-What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
+            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
+            \midrule
+            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
+            A. All dimension of the material must be smaller than exciton Bohr radius \\
+            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
+            C. The movement od charge carriers is limited in all spatial dimensions \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
+            \midrule
+            What makes acrylates more reactive than typical hydrocarbons? \\
+            A. The carbonyl group \\
+            B. The double bond \\
+            C. The hydroxyl group \\
+            D. The vinyl group \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
+            \midrule
+            Why are polar molecules more soluble in water than non-polar molecules? \\
+            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
+            B. The entropy of the system increases when non-polar molecules dissolve in water \\
+            C. Water has a dipole and can easier isolate polar molecules from each other \\
+            D. Water makes stable dipole-dipole interactions with polar molecules \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
+            \midrule
+            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
+            A. Broader product diameter distribution than pellets or extrudates \\
+            B. Facilitates reactor upscaling \\
+            C. It is also called snow-ball \\
+            D. Not suitable to prepare egg-shell catalysts \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
+            \midrule
+            What determines the flammability of polyols? \\
+            A. Their chemical structure \\
+            B. Their low volatility \\
+            C. Their odor threshold \\
+            D. Their reaction with acids \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
+\normalsize
+
+To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
+The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
+    \label{tab:chembench_corpus_models_failed_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+            \midrule
+            Which of the following are true for phosphoric acid? \\
+            A. Is a weak acid, so it is not toxic \\
+            B. Should never be tasted \\
+            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+            \midrule
+            Which of the following elements has the largest atomic radius? \\
+            A. \ce{Fr} \\
+            B. \ce{He} \\
+            C. \ce{Li} \\
+            D. \ce{Rb} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
 \\
+            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
+            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
             \midrule
-            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
-            A. 10.0\% \\
-            B. 6.6\% \\
-            C. 7.0\% \\
-            D. 94\% \\
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
+\\
             \bottomrule
@@
 
-While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \gls{llm} without tools (\Cref{tab:performance_table}).
+While for \ClaudeThreeFiveSonnet the overall performance did not change, we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
 If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
@@
 
-\oone consistently leads across both the \chembench corpus and \chembenchmini.
+\oone consistently leads across both the \ChemBench corpus and \chembenchmini.
 \ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers.
@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 

### 2024-10-30 src/tex/ms.tex
fix: update circle plots

--- 2024-10-29 src/tex/ms.tex
@@
 \section*{Data and code availability}
-The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench}.
+The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench} and archived on Zenodo under \href{10.5281/zenodo.14010212}{https://zenodo.org/records/14010212}.
 The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.

### 2024-10-30 src/tex/sections/human_subset_selection.tex
Update src/tex/sections/human_subset_selection.tex
feat: cleanup text for submission (#62)

--- 2024-10-18 src/tex/sections/human_subset_selection.tex
@@
-\paragraph{Question selection}
+\paragraph{Question selection} \label{sec:subset-selection}
 
-Several design choices were made when selecting a human subset.
-Firstly, from the full dataset, we keep all the questions that are labeled as advanced.
-In this way, we can obtain a deeper insight about the capabilities of \glspl{llm} on advanced tasks
-when compared to actual chemists.
+ Several design choices were made when selecting \chembenchmini.
+Firstly, from the full dataset, we kept all the questions labeled as advanced.
+In this way, we can obtain a deeper insight into the capabilities of \glspl{llm} on advanced tasks when compared to actual chemists.
 Secondly, we sample a maximum of three questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry).
-Thirdly, we do not include any intuition questions
-in this subset, because the correct answers are subjective, and would hinder an objective evaluation of the chemists against \glspl{llm}.
-In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for the human study.
+Thirdly, we do not include any intuition questions in this subset because the intended use of \chembenchmini is to provide a fast and fair evaluation of \glspl{llm} independent of any human baseline.
+In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for \chembenchmini.
 Then, this set is divided into two subsets based on the aforementioned combinations.

### 2024-10-30 src/tex/appendix.tex
chore: remove examples

--- 2024-10-30 src/tex/appendix.tex
@@
 
-In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-\Cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
-
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
-    \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
-            A. Cyclohexene \\
-            B. Cyclooctene \\
-            C. Decalin \\
-            D. Ethyl ether \\
-            E. Isopropyl ether \\
-            F. Para-dioxane \\
-            G. Tetralin \\
-            \midrule
-            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
-            \midrule
-            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
-            A. The intersection with y-axis shift to higher numbers \\
-            B. The intersection with y-axis shift to lower numbers \\
-            C. The slope decreases \\
-            D. The slope increases \\
-            \midrule
-            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
-A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
-What is the number of moles of acid in the sample?
-\\
-            \midrule
-            What is hydrofluoric acid used for? \\
-            A. For cleaning facades \\
-            B. For etching glass \\
-            C. For filling accumulators \\
-            D. For soldering \\
-            \midrule
-            How many chiral centers are present in the molecular structure of citric acid? \\
-            \midrule
-            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
-            A. 10 \\
-            B. 14 \\
-            C. 18 \\
-            D. 23 \\
-            E. 9 \\
-            \midrule
-            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
-What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
-\\
-            \midrule
-            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
-            A. 10.0\% \\
-            B. 6.6\% \\
-            C. 7.0\% \\
-            D. 94\% \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
-
-\normalsize
-
 \clearpage
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
-
-\begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
-Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
-Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
-However, we limit those systems to ten calls to the \gls{llm}.
-This constraint often led the systems to not find the correct answer within the specified number of calls.
-In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table.tex}
-    }
-    \label{tab:performance_table}
-\end{table}
-
-\begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
-    \resizebox{\textwidth}{!}{
-    \variable{output/performance_table_human_subset.tex}
-    }
-    \label{tab:performance_table_human_subset}
-\end{table}
 
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
+Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
+However, we limit those systems to ten calls to the \gls{llm}.
+This constraint often led the systems to not find the correct answer within the specified number of calls.
+In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
     \resizebox{\textwidth}{!}{
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{

### 2024-10-30 src/tex/appendix.tex
chore: cleanup
chore: update skill composition circular
fix some typos

--- 2024-10-30 src/tex/appendix.tex
@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt}\% on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 
@@
 However, innovation is also needed to incorporate this information into the systems.
-For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is used for directing the tools, even though the agent has access to literature evidence.
 This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant papers and should be coupled with more domain-specific databases.

### 2024-10-30 src/tex/appendix.tex
fix: apply suggestions from code review

@@
 However, smaller but well-designed models can compete with larger models.
-For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt}\% on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
+For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \chembench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.
 

### 2024-10-30 src/tex/appendix.tex
Update appendix.tex

--- 2024-10-30 src/tex/appendix.tex
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V2-flat.pdf}
     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
 
-In addition, as shown in \Cref{fig:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
+In addition, as shown in \Cref{tab:cb_skillset} the \chembench corpus can be divided considering the different skills needed to solve the questions.
 The plot shows a balanced distribution of the required skills in the \chembench corpus.
-\Cref{tab:chembench_corpus_cognitive} shows an example question for each skill category.
+\Cref{tab:cb_skillset} shows an example question for each skill category.
 
@@
 \end{table}
+
+\normalsize
+
+In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
+
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
+    \label{tab:chembench_corpus_models_failed_topic}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
+            \midrule
+            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
+            A. DESI-mass spectrometry \\
+            B. XPS \\
+            C. atomic force microscopy \\
+            D. scanning tunneling microscopy \\
+            E. secondary ion mass spectrometry \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
+            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
+            \midrule
+            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
+            A. \ce{SCl6} does not exist because of steric effects \\
+            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
+            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
+            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
+            \midrule
+            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
+            \midrule
+            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
+            A. All dimension of the material must be smaller than exciton Bohr radius \\
+            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
+            C. The movement od charge carriers is limited in all spatial dimensions \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
+            \midrule
+            What makes acrylates more reactive than typical hydrocarbons? \\
+            A. The carbonyl group \\
+            B. The double bond \\
+            C. The hydroxyl group \\
+            D. The vinyl group \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
+            \midrule
+            Why are polar molecules more soluble in water than non-polar molecules? \\
+            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
+            B. The entropy of the system increases when non-polar molecules dissolve in water \\
+            C. Water has a dipole and can easier isolate polar molecules from each other \\
+            D. Water makes stable dipole-dipole interactions with polar molecules \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
+            \midrule
+            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
+            A. Broader product diameter distribution than pellets or extrudates \\
+            B. Facilitates reactor upscaling \\
+            C. It is also called snow-ball \\
+            D. Not suitable to prepare egg-shell catalysts \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
+            \midrule
+            What determines the flammability of polyols? \\
+            A. Their chemical structure \\
+            B. Their low volatility \\
+            C. Their odor threshold \\
+            D. Their reaction with acids \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
+\normalsize
+
+To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them.
+The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
+    \label{tab:chembench_corpus_models_failed_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+            \midrule
+            Which of the following are true for phosphoric acid? \\
+            A. Is a weak acid, so it is not toxic \\
+            B. Should never be tasted \\
+            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+            \midrule
+            Which of the following elements has the largest atomic radius? \\
+            A. \ce{Fr} \\
+            B. \ce{He} \\
+            C. \ce{Li} \\
+            D. \ce{Rb} \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \midrule
+            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality).
+Which of the following two candidates would you prefer for further development?
+\\
+            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
+            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
+            \midrule
+            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \midrule
+            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}.
+What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V?
+\\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
 \normalsize
@@
 However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
+Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
+However, we limit those systems to ten calls to the \gls{llm}.
+This constraint often led the systems to not find the correct answer within the specified number of calls.
+In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table.tex}
+    }
+    \label{tab:performance_table}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset}
+\end{table}
 
@@
 An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores.
-A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
+A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.
 
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.
-Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
-Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
-However, we limit those systems to ten calls to the \gls{llm}.
-This constraint often led the systems to not find the correct answer within the specified number of calls.
-In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{
@@
 \begin{table}
-    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
     \resizebox{\textwidth}{!}{

### 2024-10-30 src/tex/appendix.tex
fix: update circular plots (#92)
Co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>

--- 2024-10-30 src/tex/appendix.tex
@@
 
-\begin{table}
+\begin{xltabular}{\textwidth}{X}
     \centering
-    \small
     \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
@@
 This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.}
+Reasoning and Calculation.} \\
     \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
             \toprule
@@
             \bottomrule
-        \end{tabularx}
-    }
-\end{table}
-\normalsize
-
-In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
-\Cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
-
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
-    \label{tab:chembench_corpus_cognitive}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
-            \toprule
-            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
-            A. Cyclohexene \\
-            B. Cyclooctene \\
-            C. Decalin \\
-            D. Ethyl ether \\
-            E. Isopropyl ether \\
-            F. Para-dioxane \\
-            G. Tetralin \\
-            \midrule
-            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
-            \midrule
-            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
-            A. The intersection with y-axis shift to higher numbers \\
-            B. The intersection with y-axis shift to lower numbers \\
-            C. The slope decreases \\
-            D. The slope increases \\
-            \midrule
-            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
-A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
-What is the number of moles of acid in the sample?
-\\
-            \midrule
-            What is hydrofluoric acid used for? \\
-            A. For cleaning facades \\
-            B. For etching glass \\
-            C. For filling accumulators \\
-            D. For soldering \\
-            \midrule
-            How many chiral centers are present in the molecular structure of citric acid? \\
-            \midrule
-            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
-            A. 10 \\
-            B. 14 \\
-            C. 18 \\
-            D. 23 \\
-            E. 9 \\
-            \midrule
-            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
-What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
-\\
-            \midrule
-            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
-            A. 10.0\% \\
-            B. 6.6\% \\
-            C. 7.0\% \\
-            D. 94\% \\
-            \bottomrule
-        \end{tabularx}
-    }
-\end{table}
-
-\normalsize
+\end{xltabular}
 
@@
     \centering
-    \includegraphics{figures/corpus_human_comparison.pdf}
+    \includegraphics[width=\textwidth]{figures/corpus_human_comparison.pdf}
     \caption{\textbf{
@@
 However, innovation is also needed to incorporate this information into the systems.
-For example, even \PaperQATwo could not outperform the \gls{llm} it is used for directing the tools, even though the agent has access to literature evidence.
+For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
 This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant papers and should be coupled with more domain-specific databases.

### 2024-10-30 src/tex/appendix.tex
fix typos

@@
     \item \emph{Cover a diverse set of topics}.
-Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
-    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, and intuition.
+Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities}, the topics covered by a chemistry benchmark must be very diverse.
+    \item \emph{Cover diverse skills}. To holistically judge performance, it is important to cover questions that rely on reasoning, calculation, knowledge, and intuition.
     \item \emph{Cover a range of difficulty levels}.
@@
 A benchmark should contain questions that are impossible to solve with current models.
-If current models can solve all questions, the benchmark provides no useful signal.
+The benchmark provides no useful signal if current models can solve all questions.
 \end{itemize}
@@
 The reviewers often guided the revision process to ensure the question aligned with the guidelines.
-Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question.
+Questions that don't meet the criteria are either rejected or suggested for revision, and most often, they are modified to a new question.
 Reviewers also provide feedback on the skill and difficulty annotations.
@@
 
-The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong to.
+The corpus of the questions in \chembench, as shown in \Cref{tab:chembench_corpus_topic}, can be divided according to which chemical topic they belong.
 
@@
 \begin{xltabular}{\textwidth}{X}
-    \centering
-    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
+    \caption{\textbf{Examples for each required skill considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
 Note that the total count in this table is bigger than the \chembench corpus.
-This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.} \\
-    \label{tab:chembench_corpus_cognitive}
+This is because the same question can be annotated with two different skills, e.g., Reasoning and Calculation.}    \label{tab:chembench_corpus_cognitive} \\
             \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
+            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
+            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
+            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
+            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt} Questions} \\
             \midrule
@@
 We also evaluated the model performance on the entire \chembench corpus.
-\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.
+\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions the models answered correctly.
 
@@
     \includegraphics{figures/overall_performance.pdf}
-    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered correctly by the models.
+    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions the models answered correctly.
 Scores computed on the entire \chembench corpus.}
@@
 \Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
-%The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.
 
@@
     \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
-    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity).
+    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals regardless of molecular complexity).
 For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
@@
 
-\Cref{tab:refusal_counts_and_parsing} lists how many refusals were detected for the reponses of different models on \chembench.
+\Cref{tab:refusal_counts_and_parsing} lists how many refusals were detected for the responses of different models on \chembench.
 Overall, we find that refusals do not majorly affect the performance measured by \chembench.
@@
 
-In our parsing workflow, we use pipelines based on regular expressions to extract the answers.
+Our parsing workflow uses pipelines based on regular expressions to extract the answers.
 In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions.
@@
 The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
-    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
+    The regex-based parsing is elaborate and can handle special cases such as scientific notation or Roman numerals.
     If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback.
@@
 They could also report what tools they used for answering questions.
-As \Cref{fig:tool_use} shows, the most commonly tool was some form of web search (which, according to the freetext responses, often was a multi-step process).
+As \Cref{fig:tool_use} shows, the most common tool was some form of web search (which, according to the free text responses, often was a multi-step process).
 
@@
 In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
-For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
+For this, we investigated the two models that showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}).
 For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
@@
 Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
-These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
+These conversion tools allow us to understand better how the agents perform for specific questions in this agent environment configuration.
 We implemented the systems using Langchain\autocite{Chase_LangChain_2022} with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.
@@
 If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
-However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
+However, the scores decreased for questions about the number of isomers or \gls{ghs} pictograms.
 For the specific questions about converting \gls{iupac} names to \gls{smiles}, the results decreased notably despite the models having access to specific tools prepared for those questions.

### 2024-10-30 src/tex/ms.tex
Update ms.tex
fix: update circular plots (#92)
Co-authored-by: Adrian Mirza <adrianmirza81@gmail.com>
fix typos

@@
 
-However, we possess only a limited systematic understanding of the chemical capabilities of LLMs, which would be required to improve models and mitigate potential harms.
+However, we possess only a limited systematic understanding of the chemical capabilities of LLMs, which would be required to improve models and mitigate potential harm.
 Here, we introduce \enquote{\chembench,} an automated framework for evaluating the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of chemists.
@@
 We curated more than 2,700 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
-However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.
+However, the models struggle with some basic tasks and provide overconfident predictions.
 
@@
     \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
-    The framework's foundation is the benchmark corpus that consists of thousands of questions and answers that we manually or semi-automatically compiled from various sources (see \Cref{sec:curation}).
+    The framework's foundation is the benchmark corpus comprising thousands of questions and answers that we manually or semi-automatically compiled from various sources (see \Cref{sec:curation}).
     Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
@@
 
-To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel manually crafted questions, over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
+To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel, manually crafted questions over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
 For quality assurance, all questions have been reviewed by at least two scientists in addition to the original curator and automated checks.
@@
 We also classify the questions based on what skills are required to answer them.
-Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
-Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty by the annotator.
+Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition, or a combination of these.
+Moreover, the annotator also classifies the questions by difficulty to allow for a more nuanced evaluation of the models' capabilities.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V2.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V3-flat.pdf}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
 Many other models also outperform the average human performance.
-Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.
+Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open-source models can be competitive with the best proprietary models also in chemical settings.
 
-Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts.
+Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}); that is, they did not memorize the relevant facts.
 Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo.
 This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
-This indicates that there is still room for improvement of chemical \glspl{llm} by training them on more specialized data sources or by integrating them with specialized databases.
+This indicates that there is still room for improving chemical \glspl{llm} by training them on more specialized data sources or integrating them with specialized databases.
 
@@
 
-In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
-Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
+In the subfield of analytical chemistry, the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
+Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
 
@@
  For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
- While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
+ While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for the general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
 
- We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
+ We also gain insight into the models' struggles with chemical reasoning tasks by examining their performance as a function of molecular descriptors.
  If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
  However, we find that the models' performance does not correlate with complexity indicators (see \Cref{sec:molecular_features}).
- This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on proximity of the molecules to the training data.\autocite{mccoy2023embersautoregressionunderstandinglarge}
+ This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on the proximity of the molecules to the training data.\autocite{mccoy2023embersautoregressionunderstandinglarge}
 
@@
 Despite chemists demonstrating a reasonable level of interrater agreement, our models largely fail to align with expert chemists' preferences.
-Their performance is often indistinguishable from random guessing, even though these same models excel in various other tasks in \chembench (\Cref{tab:performance_table_human_subset}).
+Their performance is often indistinguishable from random guessing, even though these same models excel in other tasks in \chembench (\Cref{tab:performance_table_human_subset}).
 This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.
@@
 
-To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly (see \Cref{sec:confidence_estimates} for details on methodology and comparison to logit-based approaches).
+To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly (see \Cref{sec:confidence_estimates} for details on the methodology and comparison to logit-based approaches).
 
@@
 Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
-For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
+For instance, while models perform well on many textbook questions, they struggle with questions requiring more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
 Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
@@
 Detailed sources are available in the source data on GitHub.
-Questions without source have been curated completely from scratch.
-Questions based on lecture notes or URL have been curated based on content presented in those resources.
+Questions without a source have been curated completely from scratch.
+Questions based on lecture notes or URLs have been curated based on content presented in those resources.
 All questions have been rephrased, annotated, and reviewed before being added to the corpus.
@@
 Our prompting strategy consistently adheres to these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations, and physical units (by either adding or removing wrappers).
-This step can be easily customized in our codebase and we provide presets for the models we evaluated.
+This step can be easily customized in our codebase, and we provide presets for the models we evaluated.
 
@@
 \paragraph{Models}
-For all models we performed inference using greedy decoding (i.e., temperature 0).
-We used the \gls{api} endpoints provided by the model developers and the ones provided by Groq.
+For all models, we performed inference using greedy decoding (i.e., temperature 0).
+We used the \gls{api} endpoints provided by the model developers and those provided by Groq.
 \PaperQATwo was used (in August 2024) via an \gls{api} provided by FutureHouse.
@@
 We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases than other prompting strategies, which might be more suitable when implemented in systems.
-In addition, this approach captures semantic uncertainty which is not the same as the probability of a token given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).
-On top of that, many propietary models do not provide access to the logits, making this approach more general.
-In \Cref{sec:confidence_estimates} we provide more details and comparisons with a logit-based approach.
+In addition, this approach captures semantic uncertainty, which is not the same as the probability of a token being given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).
+On top of that, many proprietary models do not provide access to the logits, making this approach more general.
+In \Cref{sec:confidence_estimates}, we provide more details and comparisons with a logit-based approach.
 
@@
 \paragraph{Study design}
-Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations.
-Questions were shown in random order and volunteers were not allowed to skip questions.
+Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}), which rendered chemicals and equations.
+Questions were shown in random order, and volunteers were not allowed to skip questions.
 For a subset of the questions, the volunteers were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions.
@@
 \subsection{Data annotation}\label{sec:meth-topic}
-In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question.
-We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g.
-the H/P statements of chemicals).
+In the curation of our dataset, we manually assigned difficulty levels and required skills to each question.
+We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g., the H/P statements of chemicals).
 Reasoning is required if answering a question requires multiple reasoning steps.
-Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.
+Basic questions only require those skills up to the high school level. Advanced questions would require an expert multiple minutes up to hours to answer.
 

### 2024-10-30 src/scripts/make_required_skills_table.py

@@ -9,7 +9,7 @@ from chembench.analysis import (
 )
 
 from utils import obtain_chembench_repo
-from paths import tex, data
+from paths import tex, data, output
 
 def load_and_count(datafolder):
     data_paths = construct_name_to_path_dict(datafolder)
@@ -27,7 +27,7 @@ def load_and_count(datafolder):
         keywords = data["keywords"]
         for keyword in keywords:
             if "requires" in keyword:
-                if keyword == "requires-reasoning": 
+                if keyword == "requires-reasoning":
                     reasoning += 1
                 if keyword == "requires-knowledge":
                     knowledge += 1
@@ -35,8 +35,8 @@ def load_and_count(datafolder):
                     calculation += 1
                 if keyword == "requires-intuition":
                     intuition += 1
-                
-    
+
+
     results = {
         "requires-reasoning": reasoning,
         "requires-knowledge": knowledge,
@@ -53,6 +53,18 @@ if __name__ == "__main__":
     )
 
     data = load_and_count(datafolder)
+
+    with open(output / "reasoning_count.txt", 'w') as f:
+        f.write(str(data["requires-reasoning"]) + "\endinput")
+
+    with open(output / "knowledge_count.txt", 'w') as f:
+        f.write(str(data["requires-knowledge"]) + "\endinput")
+
+    with open(output / "calculation_count.txt", 'w') as f:
+        f.write(str(data["requires-calculation"]) + "\endinput")
+
+    with open(output / "intuition_count.txt", 'w') as f:
+        f.write(str(data["requires-intuition"]) + "\endinput")
 
     # Save the data variable to a pickle file
     with open(os.path.join(data, 'requires_data.pkl'), 'wb') as f:

### 2024-10-30 src/tex/appendix.tex
fix: add flattened pdfs

--- 2024-10-30 src/tex/appendix.tex
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V2-flat.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_HUMANSET_FIG_V2.pdf}
     \caption{\textbf{Composition of the human subset.} The circular plot shows the distribution of topics and required skills in the human subset of the \chembench corpus.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V2.pdf}
     \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
@@
 
-\begin{xltabular}{\textwidth}{X}
+\begin{table}
     \centering
+    \small
     \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.
@@
 This is because the same question can be annotated with two different \enquote{requires} skills, i.e.
-Reasoning and Calculation.} \\
+Reasoning and Calculation.}
     \label{tab:chembench_corpus_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
             \toprule
@@
             \bottomrule
-\end{xltabular}
+        \end{tabularx}
+    }
+\end{table}
+\normalsize
+
+In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers.
+\Cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement
+
+\begin{table}
+    \centering
+    \small
+    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
+    \label{tab:chembench_corpus_cognitive}
+    {\fontsize{8}{9}\selectfont
+        \begin{tabularx}{\textwidth}{X}
+            \toprule
+            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
+            A. Cyclohexene \\
+            B. Cyclooctene \\
+            C. Decalin \\
+            D. Ethyl ether \\
+            E. Isopropyl ether \\
+            F. Para-dioxane \\
+            G. Tetralin \\
+            \midrule
+            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
+            \midrule
+            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
+            A. The intersection with y-axis shift to higher numbers \\
+            B. The intersection with y-axis shift to lower numbers \\
+            C. The slope decreases \\
+            D. The slope increases \\
+            \midrule
+            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid).
+A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize.
+What is the number of moles of acid in the sample?
+\\
+            \midrule
+            What is hydrofluoric acid used for? \\
+            A. For cleaning facades \\
+            B. For etching glass \\
+            C. For filling accumulators \\
+            D. For soldering \\
+            \midrule
+            How many chiral centers are present in the molecular structure of citric acid? \\
+            \midrule
+            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
+            A. 10 \\
+            B. 14 \\
+            C. 18 \\
+            D. 23 \\
+            E. 9 \\
+            \midrule
+            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter.
+What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion?
+\\
+            \midrule
+            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
+            A. 10.0\% \\
+            B. 6.6\% \\
+            C. 7.0\% \\
+            D. 94\% \\
+            \bottomrule
+        \end{tabularx}
+    }
+\end{table}
+
+\normalsize
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/corpus_human_comparison.pdf}
+    \includegraphics{figures/corpus_human_comparison.pdf}
     \caption{\textbf{

### 2024-10-30 src/tex/ms.tex
fix changes

@@
  To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
- This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
+ This is facilitated by the tooling that \chembench provides, allowing contributors to automatically add new models in an open science fashion.
 
@@
 Our parsing workflow is multistep and primarily based on regular expressions.
-In the case of instruction-tuned models, we first identify the \texttt{[ANSWER]}\texttt{[\textbackslash ANSWER]} environment we prompt the model to report the answer in.
+In the case of instruction-tuned models, we first identify the \texttt{[ANSWER]}\texttt{[\textbackslash ANSWER]} environment we prompt the model to report the answer.
 In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.

### 2024-10-31 src/tex/appendix.tex
fix changes
topic counts

--- 2024-10-30 src/tex/appendix.tex
@@
             \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
+            \multicolumn{1}{c}{\textbf{Analytical Chemistry} \variable{output/total_analytical.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
+            \multicolumn{1}{c}{\textbf{Chemical Preference} \variable{output/total_chemical_preference.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} \variable{output/total_general.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} \variable{output/total_inorganic.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
+            \multicolumn{1}{c}{\textbf{Materials Science} \variable{output/total_materials_science.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} \variable{output/total_organic.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} \variable{output/total_physical.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} \variable{output/total_technical.txt}} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} \variable{output/total_toxicity_safety.txt}} \\
             \midrule
@@
             \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt} Questions} \\
+            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt}\hspace{1.5em} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt} Questions} \\
+            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt}\hspace{1.5em} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt} Questions} \\
+            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt}\hspace{1.5em} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt} Questions} \\
+            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt}\hspace{1.5em} Questions} \\
             \midrule

### 2024-10-31 src/scripts/make_topic_performance_tables.py

@@ -190,31 +190,49 @@ def compile_df(results: dict, human_subset=False) -> pd.DataFrame:
 
     if not human_subset:
         with open(output / "total_analytical.txt", "w") as f:
-            f.write(f"{total_analytical}" + "\endinput")
+            percentage = count_analytical / total_analytical
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_analytical} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_chemical_preference.txt", "w") as f:
-            f.write(f"{total_chemical_preference}" + "\endinput")
+            percentage = count_chemical_preference / total_chemical_preference
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_chemical_preference} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_general.txt", "w") as f:
-            f.write(f"{total_general}" + "\endinput")
+            percentage = count_general / total_general
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_general} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_inorganic.txt", "w") as f:
-            f.write(f"{total_inorganic}" + "\endinput")
+            percentage = count_inorganic / total_inorganic
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_inorganic} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_materials_science.txt", "w") as f:
-            f.write(f"{total_materials_science}" + "\endinput")
+            percentage = count_materials_science / total_materials_science
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_materials_science} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_organic.txt", "w") as f:
-            f.write(f"{total_organic}" + "\endinput")
+            percentage = count_organic / total_organic
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
+            f.write(f"{total_organic} Questions {percentage_str}" + "\endinput")
 
         with open(output / "total_physical.txt", "w") as f:
-            f.write(f"{total_physical}" + "\endinput")
+            percentage = count_physical / total_physical
+            percentage_str = f"({percentage:.2%})".replace("%", "\%")
[…]

### 2024-10-31 src/data/chem-bench-main/reports/claude3/run_benchmark.py
finalize
Former-commit-id: 474f22beb6dea804eda460c8045f354a385aa7b6

@@ -1,0 +1,96 @@
+import pickle
+import time
+from glob import glob
+from pathlib import Path
+from typing import List
+
+import anthropic
+from dotenv import load_dotenv
+
+from chembench.baseline import Generation, Generations
+from chembench.evaluate import ChemBenchmark
+from chembench.prompter import PrompterBuilder
+from chembench.utils import (
+    enable_caching,
+    enable_logging,
+    get_folder_name_string_from_path,
+)
+
+load_dotenv("../../.env", override=True)
+
+client = anthropic.Anthropic()
+
+
+def generate(
+    prompt,
+):
+    try:
+        message = client.messages.create(
+            model="claude-3-opus-20240229",
+            max_tokens=1000,
+            temperature=0,
+            system="Today is March 4, 2024.",
+            messages=[
+                {
+                    "role": "user",
+                    "content": [{"type": "text", "text": prompt}],
+                }
+            ],
+        )
+        time.sleep(5)
+        return message.content[0].text
+    except Exception as e:
+        raise ValueError(e)
+
+
+class Claude3:
+    def __init__(self): ...
+    def generate(self, prompt: List[str]) -> Generations:
+        generations = []
+        for prompt_ in prompt:
+            generation = generate(prompt_)
+            generation = [Generation(text=generation)]
+            generations.append(generation)
+
+        return Generations(generations=generations)
+
+
+if __name__ == "__main__":
+    enable_logging()
+    enable_caching()
+    already_done = list(glob("reports/**/*.json", recursive=True))
+    to_do = glob("../../data/**/*.json", recursive=True)
+    already_done_names = []
+    for name in already_done:
+        path = Path(name).stem.split("-")
+        folder = path[0]
+        stem = path[1]
+        already_done_names.append(f"{folder}-{stem}")
+    number_initial = len(to_do)
+
+    print(f"Found {len(to_do)} to do, {len(already_done_names)} already done")
+    to_do = [x for x in to_do if get_folder_name_string_from_path(x) not in already_done_names]
+    print(f"Now running on {len(to_do)}")
+    assert len(to_do) > 0
+
+    assert len(to_do) <= number_initial
+
+    benchmark = ChemBenchmark.from_json_paths(
+        to_do,
+        skip_errors=True,
+    )
+    model = Claude3()
+    prompter = prompter = PrompterBuilder.from_model_object(
+        model=model,
[…]

### 2024-10-31 src/data/chem-bench-main/reports/confidence_estimates/run_confidence_estimates.py

@@ -1,0 +1,175 @@
+import os
+from glob import glob
+from typing import List
+import anthropic
+import fire
+import pandas as pd
+from dotenv import load_dotenv
+from langchain_community.chat_models import ChatAnthropic
+from langchain_openai import ChatOpenAI
+from loguru import logger
+from chembench.baseline import Generation, Generations
+from chembench.prompter import ConfidenceInstructionPrompter
+from chembench.task import Task
+from chembench.extractor import LLMExtractor
+from chembench.utils import (
+    LangChainChatModelWrapper,
+    enable_caching,
+    enable_logging,
+    passthrough,
+    remove_ce,
+    remove_math,
+    remove_pu,
+    remove_rxnsmiles,
+    remove_smiles,
+)
+import groq
+
+load_dotenv("../../.env", override=True)
+all_tasks = glob("../../data/**/*.json", recursive=True)
+report_directory_root = "confidence_estimates"
+if not os.path.exists(report_directory_root):
+    os.makedirs(report_directory_root)
+
+load_dotenv("../../.env", override=True)
+client = anthropic.Anthropic()
+groq_client = groq.Groq()
+
+
+def generate(prompt):
+    try:
+        message = client.messages.create(
+            model="claude-3-5-sonnet-20240620",
+            max_tokens=1000,
+            temperature=0,
+            system="Today is October 15, 2024.",
+            messages=[
+                {
+                    "role": "user",
+                    "content": [{"type": "text", "text": prompt}],
+                }
+            ],
+        )
+        return message.content[0].text
+    except Exception:
+        return "ERROR OCCURRED. PLEASE TRY AGAIN."
+
+
+class Claude3:
+    def __init__(self): ...
+
+    def generate(self, prompt: List[str]) -> Generations:
+        generations = []
+        for prompt_ in prompt:
+            generation = generate(prompt_)
+            generation = [Generation(text=generation)]
+            generations.append(generation)
+        return Generations(generations=generations)
+
+
+class GroqLlama:
+    def __init__(self):
+        self.client = groq_client
+
+    def generate(self, prompt: List[str]) -> Generations:
+        generations = []
+        llm_extractor = LLMExtractor(
+            model_type="groq",
+            seed=42,
+            permute_options=False,
+            temperature=0,
+        )
+        for prompt_ in prompt:
+            try:
+                completion = self.client.chat.completions.create(
+                    model="llama-3.1-8b-instant",
+                    messages=[
[…]

### 2024-10-31 src/data/chem-bench-main/reports/gpt-3.5-turbo/run_benchmark.py

@@ -1,0 +1,46 @@
+import pickle
+from glob import glob
+from pathlib import Path
+
+from dotenv import load_dotenv
+
+from chembench.evaluate import ChemBenchmark
+from chembench.prompter import PrompterBuilder
+from chembench.utils import (
+    enable_caching,
+    enable_logging,
+    get_folder_name_string_from_path,
+)
+
+load_dotenv("../../.env", override=True)
+
+
+if __name__ == "__main__":
+    enable_logging()
+    enable_caching()
+    already_done = list(glob("reports/**/*.json", recursive=True))
+    to_do = glob("../../data/**/*.json", recursive=True)
+    already_done_names = []
+    for name in already_done:
+        path = Path(name).stem.split("-")
+        folder = path[0]
+        stem = path[1]
+        already_done_names.append(f"{folder}-{stem}")
+    number_initial = len(to_do)
+
+    print(f"Found {len(to_do)} to do, {len(already_done_names)} already done")
+    to_do = [x for x in to_do if get_folder_name_string_from_path(x) not in already_done_names]
+    print(f"Now running on {len(to_do)}")
+    assert len(to_do) > 0
+
+    assert len(to_do) <= number_initial
+
+    benchmark = ChemBenchmark.from_json_paths(
+        to_do,
+        skip_errors=True,
+    )
+    prompter = PrompterBuilder.from_preset("openai/gpt-3.5-turbo", model_kwargs={"temperature": 0})
+
+    result = benchmark.bench(prompter)
+    with open("results.pkl", "wb") as f:
+        pickle.dump(result, f)

### 2024-10-31 src/data/chem-bench-main/reports/gpt-4/run_benchmark.py

@@ -1,0 +1,44 @@
+from chembench.evaluate import ChemBenchmark
+from chembench.prompter import PrompterBuilder
+from glob import glob
+from chembench.utils import (
+    enable_logging,
+    enable_caching,
+    get_folder_name_string_from_path,
+)
+import pickle
+from dotenv import load_dotenv
+from pathlib import Path
+
+load_dotenv("../../.env", override=True)
+
+
+if __name__ == "__main__":
+    enable_logging()
+    enable_caching()
+    already_done = list(glob("reports/**/*.json", recursive=True))
+    to_do = glob("../../data/**/*.json", recursive=True)
+    already_done_names = []
+    for name in already_done:
+        path = Path(name).stem.split("-")
+        folder = path[0]
+        stem = path[1]
+        already_done_names.append(f"{folder}-{stem}")
+    number_initial = len(to_do)
+
+    print(f"Found {len(to_do)} to do, {len(already_done_names)} already done")
+    to_do = [x for x in to_do if get_folder_name_string_from_path(x) not in already_done_names]
+    print(f"Now running on {len(to_do)}")
+    assert len(to_do) > 0
+
+    assert len(to_do) <= number_initial
+
+    benchmark = ChemBenchmark.from_json_paths(
+        to_do,
+        skip_errors=True,
+    )
+    prompter = PrompterBuilder.from_preset("openai/gpt-4", model_kwargs={"temperature": 0})
+
+    result = benchmark.bench(prompter)
+    with open("results.pkl", "wb") as f:
+        pickle.dump(result, f)

### 2024-10-31 src/data/chem-bench-main/reports/humans/test_scorepipe.py

@@ -1,0 +1,102 @@
+import ast
+import json
+
+import pandas as pd
+from fire import Fire
+from fuzzywuzzy import process
+from generate_reports import ReportGenerator
+from loguru import logger
+
+from chembench.task import Task
+
+
+def get_lookup() -> pd.DataFrame:
+    """Reads the question lookup CSV file and removes duplicates.
+
+    Returns:
+        A DataFrame containing the question lookup data.
+    """
+    questionsLookup = pd.read_csv("questionLookup.csv")
+    questionsLookup = questionsLookup.drop_duplicates(subset="examples")
+    return questionsLookup
+
+
+def get_task_from_qid(qid: str) -> Task:
+    """Gets a Task object for a given question ID.
+
+    Args:
+        qid: The question ID.
+
+    Returns:
+        A Task object for the question.
+    """
+    questiondata = pd.read_csv("questions.csv")
+    questionsLookup = get_lookup()
+    questionsLookup["examples"] = questionsLookup["examples"].apply(lambda x: ast.literal_eval(x) if isinstance(x, str) else x)
+
+    question_details = questiondata.loc[questiondata["id"] == qid]
+
+    qexample = json.loads(question_details["example"].item())
+    try:
+        qrow = questionsLookup[questionsLookup["examples"] == qexample]
+        qpath = qrow["path"].item()
+    except ValueError:
+        logger.warning(f"Failed to get qname for {qexample}. Attempting to handle this...")
+        closest, score = process.extractOne(qexample["input"], [q["input"] for q in questionsLookup["examples"]])
+        if score < 95:
+            logger.warning(f"Failed to find a close match for {qexample}. Skipping...")
+
+        logger.info(f"Closest match for {qexample} is {closest} with a score of {score}")
+        qrow = questionsLookup.iloc[[q["input"] for q in questionsLookup["examples"]].index(closest)]
+        qpath = qrow["path"]
+    task = Task.from_json(qpath)
+    return task
+
+
+def get_correct_answer_from_qid(qid: str) -> str:
+    """Gets the correct answer for a given question ID.
+
+    Args:
+        qid: The question ID.
+
+    Returns:
+        The correct answer for the question.
+    """
+    example = get_task_from_qid(qid)._examples[0]
+    is_classification = "target_scores" in example
+    if not is_classification:
+        correct_answer = example["target"]
+    else:
[…]

### 2024-10-31 src/data/chem-bench-main/scripts/cluster_questions.py

@@ -1,0 +1,405 @@
+import json
+from pathlib import Path
+
+import fire
+import pandas as pd
+import plotly.express as px
+from chembench.types import PathType
+from fastcore.xtras import loads
+from loguru import logger
+
+SHORT_KEYWORDS = {
+    "analytical-chemistry": "AC",
+    "biochemistry": "BC",
+    "inorganic-chemistry": "IC",
+    "toxicity/safety": "T/S",
+    "organic-chemistry": "OC",
+    "physical-chemistry": "PC",
+    "photochemistry": "PhC",
+    "solid-state-chemistry": "SSC",
+    "technical-chemistry": "TC",
+    "polymer-chemistry": "PolyC",
+    "computational-chemistry": "CC",
+    "quantum-chemistry": "QC",
+    "electrochemistry": "EC",
+    "chemical-preference": "ChemPref",
+    "general-chemistry": "GC",
+    "lab-observations": "LabObs",
+    "materials-science": "MS",
+    "pharmacology": "Pharm",
+}
+
+FULL_NAMES_LEGEND = {
+    "AC": "Analytical Chemistry",
+    "BC": "Biochemistry",
+    "IC": "Inorganic Chemistry",
+    "T/S": "Toxicity/Safety",
+    "OC": "Organic Chemistry",
+    "PC": "Physical Chemistry",
+    "PhC": "Photochemistry",
+    "SSC": "Solid State Chemistry",
+    "TC": "Technical Chemistry",
+    "PolyC": "Polymer Chemistry",
+    "CC": "Computational Chemistry",
+    "QC": "Quantum Chemistry",
+    "EC": "Electrochemistry",
+    "ChemPref": "Chemical Preference",
+    "GC": "General Chemistry",
+    "LabObs": "Lab Observations",
+    "MS": "Materials Science",
+    "Pharm": "Pharmacology",
+    "RK": "Knowledge",
+    "RR": "Reasoning",
+    "RI": "Intuition",
+    "RC": "Calculation",
+}
+
+REQUIRES_TAGS_SHORT = {
+    "requires-knowledge": "RK",
+    "requires-reasoning": "RR",
+    "requires-intuition": "RI",
+    "requires-calculation": "RC",
+}
+
+
+def find_all_json_files(path_to_data: PathType):
+    return Path(path_to_data).rglob("**/*.json")
+
+
+def keyword_filter(keyword_list: list[str]):
+    """
+    `exclude` defines words that are not main categories or can
+    interfere with the classification
+
[…]

### 2024-10-31 src/data/chem-bench-main/scripts/split_files.py

@@ -1,0 +1,42 @@
+import json
+import os
+from glob import glob
+
+import fire
+from tqdm import tqdm
+
+
+def rewrite_file(file, remove_original: bool = False) -> None:
+    with open(file, "r") as f:
+        data = json.load(f)
+
+    examples = data["examples"]
+
+    # if there is more than one example, we need to split them up
+    # one example per file, keeping everything else the same
+    if len(examples) > 1:
+        for i, example in enumerate(examples):
+            new_data = data.copy()
+            new_data["examples"] = [example]
+            new_file = file.replace(".json", f"_{i}.json")
+            with open(new_file, "w") as f:
+                json.dump(new_data, f, indent=2)
+        if remove_original:
+            os.remove(file)
+
+
+def rewrite_files(files: str, remove_original: bool = False) -> None:
+    files = glob(files)
+    for file in tqdm(files):
+        rewrite_file(file, remove_original=remove_original)
+
+
+def rewrite_folder(folder: str, remove_original: bool = False) -> None:
+    # recursively rewrite all files in a folder
+    files = glob(os.path.join(folder, "**/*.json"), recursive=True)
+    for file in tqdm(files):
+        rewrite_file(file, remove_original=remove_original)
+
+
+if __name__ == "__main__":
+    fire.Fire(rewrite_folder)

### 2024-10-31 src/data/chem-bench-main/src/chembench/__init__.py

@@ -1,0 +1,4 @@
+from loguru import logger
+
+logger.disable("chembench")
+__version__ = "0.2.0-dev"

### 2024-10-31 src/data/chem-bench-main/src/chembench/prompter.py

@@ -1,0 +1,632 @@
+import os
+from abc import ABC, abstractmethod
+from collections import defaultdict
+from typing import Any, Dict, List, Literal, Optional, Tuple
+
+import numpy as np
+from fastcore.basics import basic_repr
+from loguru import logger
+
+from chembench.constant import (
+    FLOATQ_REGEX_TEMPLATE_1,
+    MCQ_REGEX_TEMPLATE_1,
+    MCQ_REGEX_TEMPLATE_2,
+    NUM_REGEX,
+)
+from chembench.extractor import LLMExtractor
+from chembench.metrics import _AGGREGATION_MAP, _METRIC_MAP, classification_scores
+from chembench.report import Report
+from chembench.settings import MODEL_POST_PROCESS_ENV
+from chembench.task import Task
+from chembench.utils import (
+    LangChainChatModelWrapper,
+    create_multiple_choice_confidence_prompt,
+    create_multiple_choice_prompt,
+    create_multiple_choice_prompt_completion,
+    create_multiple_choice_regex,
+    detect_refusal,
+    general_prompt,
+    general_prompt_completion,
+    general_prompt_confidence,
+    passthrough,
+    post_process_prompts,
+    prepare_general_answer,
+    prepare_mcq_answer,
+    remove_ce,
+    remove_math,
+    remove_pu,
+    remove_rxnsmiles,
+    remove_smiles,
+    run_regex,
+)
+from chembench.watermark import get_watermark
+
+
+class BasePrompter(ABC):
+    """Prompters take a task and build prompts for the LLM.
+
+    For this reason, Prompters have specific methods to build prompts
+    that might be different for different LLMs.
+
+    The Prompters extract information from the tasks and use
+    it to also implement the `report` method that will prompt the LLM
+    and compute the metrics of interest.
+
+    For this, we will parse the output in a comparatively robust way,
+    for example, taking into account various ways of writing scientific
+    numbers.
+
+    Classes that inherit from this class should implement the following
+    methods:
+
+    - `_prompts_with_choices`
+    - `_prompts_general`
+
+    And also set the following attributes:
+
+    - `_mcq_regex_template`
+    - `_floatq_regex_template`
+
+
+    Args:
+        model: LLM to prompt.
+        post_process_ce: Function to post process chemical equations.
+        post_process_math: Function to post process math equations.
+        post_process_pu: Function to post process physical units.
+        post_process_smiles: Function to post process SMILES.
[…]

### 2024-10-31 src/data/chem-bench-main/src/chembench/utils.py

@@ -1,0 +1,765 @@
+import random
+import re
+import sys
+from pathlib import Path
+from typing import Dict, List, Optional, Tuple, Union
+
+import pint
+from fastcore.xtras import loads
+from loguru import logger
+
+from chembench.constant import (
+    COMPLETION_TEMPLATE_1,
+    COMPLETION_TEMPLATE_2,
+    CONFIDENCE_SCORE_ESTIMATE_1,
+    CONFIDENCE_SCORE_ESTIMATE_2,
+    COT_PROMPT,
+    FLOATQ_REGEX_TEMPLATE_1,
+    INSTRUCT_TEMPLATE_1,
+    INSTRUCT_TEMPLATE_2,
+    LATEX_ENV_REGEX,
+    MCQ_REGEX_TEMPLATE_1,
+    NUM_REGEX,
+    NUM_WORDS_REGEX,
+)
+from chembench.extractor import LLMExtractor
+from chembench.types import PathType
+
+
+def permute_options(options, seed):
+    random.seed(seed)
+    options = list(options)  # Convert dict_keys or other iterable to list
+    permutation_indices = random.sample(range(len(options)), len(options))
+    permuted_options = [options[i] for i in permutation_indices]
+    return permuted_options
+
+
+def available_models():
+    """
+    Returns a list of available models.
+    """
+    from chembench.settings import MODEL_POST_PROCESS_ENV
+
+    return [*MODEL_POST_PROCESS_ENV]
+
+
+def print_available_models():
+    print("Currently available models:")
+    for model_index, model in enumerate(available_models()):
+        print(f"    {model_index + 1} {model}")
+
+
+def get_folder_name_string_from_path(path: Union[str, Path]):
+    p = Path(path)
+    parts = p.parts
+    return f"{parts[-2]}-{p.stem}"
+
+
+def load_json(filename: PathType):
+    with open(filename, "r") as handle:
+        return loads(handle.read())
+
+
+def unescape(string: str) -> str:
+    # To be improved
+    return string.replace("//", "/")
+
+
+def convert_to_number(matched_value: str) -> float:
+    """Helper Function to convert string to number using pint
+
+    Args:
+        matched_value: String which is to be converted to number
+
+    Returns:
+        float: converted number
+    """
+
+    ureg = pint.UnitRegistry()
+
+    matched_value = matched_value.replace("x", "*").replace("\\times", "*").replace("\\u00d7", "*")  # Replace "x" with "*" pint cant handle x
+    try:
[…]

### 2024-10-31 src/data/chem-bench-main/tests/test_prompter.py

@@ -1,0 +1,113 @@
+import random
+from unittest.mock import patch
+
+import pytest
+
+from chembench.extractor import LLMExtractor
+from chembench.prompter import PrompterBuilder
+from chembench.task import Task
+
+SEED = random.randint(1, 100)
+
+
+class MockResponse:
+    def __init__(self, text):
+        self.generations = [[MockGeneration(text)]]
+
+
+class MockGeneration:
+    def __init__(self, text):
+        self.text = text
+
+
+@pytest.mark.requires_api_key
+def test_llm_extraction_number():
+    completion = "The correct answer is [ANSWER]42[/ANSWER]."
+
+    with patch("langchain_community.chat_models.ChatAnthropic.generate") as mock_generate:
+        mock_generate.return_value = MockResponse(text="[ANSWER]42[/ANSWER]")
+        extractor = LLMExtractor()
+        result = extractor.extract(completion)
+        assert result == 42.0
+
+
+@pytest.mark.requires_api_key
+def test_llm_extraction_mcq_multiple():
+    completion = "The correct answers are [ANSWER]A, B, C[/ANSWER]."
+
+    with patch("langchain_community.chat_models.ChatAnthropic.generate") as mock_generate:
+        mock_generate.return_value = MockResponse(text="[ANSWER]A, B, C[/ANSWER]")
+        extractor = LLMExtractor()
+        result = extractor.extract(completion, mcq=True)
+        assert result == ["A", "B", "C"]
+
+
+@pytest.mark.requires_api_key
+def test_llm_extraction_no_answer():
+    completion = "There is no correct answer."
+
+    with patch("langchain_community.chat_models.ChatAnthropic.generate") as mock_generate:
+        mock_generate.return_value = MockResponse(text="[ANSWER][/ANSWER]")
+        extractor = LLMExtractor()
+        result = extractor.extract(completion)
+        assert result is None
+
+
+@pytest.mark.requires_api_key
+def test_llm_extraction_mcq_example():
+    completion = "The correct answer is [ANSWER]A,B[/ANSWER]."
+    example = {
+        "input": "Which of the following compounds is linear?",
+        "target_scores": {
+            "\\ce{XeF2}": 1,
+            "\\ce{I3-}": 1,
+            "\\ce{SO2}": 0,
+            "\\ce{CO2}": 1,
+            "\\ce{H2S}": 0,
+        },
+    }
+
[…]

### 2024-10-31 src/data/chem-bench-main/val/run_validation.py

@@ -1,0 +1,98 @@
+import json
+import os
+from glob import glob
+from pprint import pprint
+
+import fire
+import yaml
+from loguru import logger
+from model import Metadata, Task
+
+
+def validate_task_file(task_file):
+    with open(task_file, "r") as f:
+        loaded_dict = json.load(f)
+        task = Task(**loaded_dict)
+    return task
+
+
+def _validate_number_of_questions(data_dir, report_dir):
+    number_of_questions = len(glob(f"{data_dir}/**/*.json", recursive=True))
+    number_of_reports = len(glob(f"{report_dir}/*.json", recursive=True))
+
+    if number_of_questions != number_of_reports:
+        logger.warning(f"Number of questions ({number_of_questions}) does not match number of reports ({number_of_reports}) in {report_dir}")
+
+
+def validate_number_of_questions_in_report_dirs(data_dir, report_root):
+    """
+    Validates the number of questions in report directories.
+
+    Args:
+        data_dir (str): The directory containing the data.
+        report_root (str): The root directory of the reports.
+
+    Raises:
+        ValueError: If the report root does not exist or is not a directory.
+
+    Returns:
+        None
+    """
+    if not os.path.exists(report_root):
+        raise ValueError(f"Report root {report_root} does not exist")
+
+    if not os.path.isdir(report_root):
+        raise ValueError(f"Report root {report_root} is not a directory")
+
+    report_dirs = glob(f"{report_root}/**/reports/**/*.json", recursive=True)
+    report_dirs = {os.path.dirname(report_dir) for report_dir in report_dirs}
+    for report_dir in report_dirs:
+        if os.path.isdir(report_dir) and "human" not in report_dir:
+            _validate_number_of_questions(data_dir, report_dir)
+
+
+def validate_task_dir(task_dir):
+    """
+    Validates a directory containing task files.
+
+    Args:
+        task_dir (str): The path to the directory containing task files.
+
+    Raises:
+        Exception: If any task file fails validation.
+    """
+    task_files = glob(f"{task_dir}/**/*.json", recursive=True)
+    has_failure = False
+    for task_file in task_files:
+        try:
+            validate_task_file(task_file)
+        except Exception as e:
+            print(f"Validation failed for {task_file}")
+            pprint(e.errors())
+            has_failure = True
+
+    if has_failure:
+        raise Exception("Validation failed")
+
+
+def validate_metadata_file(path: str) -> None:
[…]

### 2024-10-31 src/tex/appendix.tex
fix: skillset composition fig
fix: skillset composition fig (#95)

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_FINAL_SKILL_flat.pdf}
     \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
@@
 Note that the total count in this table is bigger than the \chembench corpus.
-This is because the same question can be annotated with two different skills, e.g., Reasoning and Calculation.}    \label{tab:chembench_corpus_cognitive} \\
+This is because the same question can be annotated with two different skills, e.g., Reasoning and Calculation.}
+    \label{tab:chembench_corpus_cognitive} \\
             \toprule

### 2024-10-31 src/tex/appendix.tex
finalize
Former-commit-id: 9c194006e3e56a88e4dab9f3326e35b18ae5f44e

--- 2024-10-31 src/tex/appendix.tex
@@
 
-\begin{table}
-    \centering
-    \small
-    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
-    \label{tab:chembench_corpus_topic}
-    {\fontsize{8}{9}\selectfont
-        \begin{tabularx}{\textwidth}{X}
+\begin{xltabular}{\textwidth}{X}
+    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \label{tab:chembench_corpus_topic} \\
             \toprule
@@
             \bottomrule
-        \end{tabularx}
-    }
- \end{table}
+ \end{xltabular}
 

### 2024-10-31 src/tex/ms.tex
finalize
Former-commit-id: 11a53db44d0f06d0787405d905e82b891564698b
Former-commit-id: c1e40cd182cc9386e21f59b66357fc4c2a19bec8

--- 2024-10-30 src/tex/ms.tex
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V3-flat.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_OVERVIEW_FIG_V3-flat.png}
     \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them.
@@
  To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
- This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
+ This is facilitated by the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
 

### 2024-10-31 src/tex/appendix.tex
finalize
Former-commit-id: 11a53db44d0f06d0787405d905e82b891564698b
Former-commit-id: c1e40cd182cc9386e21f59b66357fc4c2a19bec8
fix: skillset composition pdf -> png

--- 2024-10-31 src/tex/appendix.tex
@@
             \toprule
-            \multicolumn{1}{c}{\textbf{Analytical Chemistry} \variable{output/total_analytical.txt}} \\
+            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Chemical Preference} \variable{output/total_chemical_preference.txt}} \\
+            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{General Chemistry} \variable{output/total_general.txt}} \\
+            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} \variable{output/total_inorganic.txt}} \\
+            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Materials Science} \variable{output/total_materials_science.txt}} \\
+            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Organic Chemistry} \variable{output/total_organic.txt}} \\
+            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Physical Chemistry} \variable{output/total_physical.txt}} \\
+            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Technical Chemistry} \variable{output/total_technical.txt}} \\
+            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%)} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Toxicity/Safety} \variable{output/total_toxicity_safety.txt}} \\
+            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%)} \\
             \midrule
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_SKILL_FIG_V1.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_FINAL_SKILL.png}
     \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
@@
             \toprule
-            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt}\hspace{1.5em} Questions} \\
+            \multicolumn{1}{c}{\textbf{Knowledge} \variable{output/knowledge_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt}\hspace{1.5em} Questions} \\
+            \multicolumn{1}{c}{\textbf{Reasoning} \variable{output/reasoning_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt}\hspace{1.5em} Questions} \\
+            \multicolumn{1}{c}{\textbf{Intuition} \variable{output/intuition_count.txt} Questions} \\
             \midrule
@@
             \midrule
-            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt}\hspace{1.5em} Questions} \\
+            \multicolumn{1}{c}{\textbf{Calculation} \variable{output/calculation_count.txt} Questions} \\
             \midrule

### 2024-10-31 src/tex/appendix.tex
add per-topic performance table
Former-commit-id: 3ea98c52c9a66c89e3190485c1a18c39c5c8a91f

--- 2024-10-31 src/tex/appendix.tex
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/CB_FINAL_SKILL_flat.pdf}
+    \includegraphics[width=\textwidth]{figures/CB_FINAL_SKILL_flat.png}
     \caption{\textbf{Composition of the required skills considered in the \chembench corpus.} The circular plot shows the distribution of required skills in the \chembench corpus.}
@@
     \label{tab:performance_table_human_subset}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different topics.
+Models with \enquote{T-one} in the name were run for a temperature of 1, which allows us to study the temperature effect in the benchmark.
+Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or a calculator to answer the questions better.
+However, we limit those systems to ten calls to the \gls{llm}.
+This constraint often led the systems to not find the correct answer within the specified number of calls.
+In this case, we consider the answer as incorrect (see \Cref{sec:react-environment}).}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_topic_table.tex}
+    }
+    \label{tab:performance_table_topic}
+\end{table}
+
+\begin{table}
+    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered correctly by the models for different topics.}
+    \resizebox{\textwidth}{!}{
+    \variable{output/performance_topic_table_human_subset.tex}
+    }
+    \label{tab:performance_table_human_subset_topic}
 \end{table}

### 2024-10-31 src/tex/appendix.tex
update formatting
Former-commit-id: 501f51ac185563ad2cbff39867d808e4817163eb
last cleanup
Former-commit-id: 1cdaf1565bd60c21aa867e8086408b4de3b56b2f

@@
 \normalsize
+
+\clearpage
 
@@
     \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
-    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models.
+    \caption{\textbf{Reliability diagram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models.
 Only logprobs of the tokens corresponding to the answers were considered.
-Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
+Linear probability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).
 The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities.

### 2024-10-31 src/tex/ms.tex
last cleanup
Former-commit-id: 1cdaf1565bd60c21aa867e8086408b4de3b56b2f

@@
      \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
-     \caption{\textbf{Calibration and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model.
+     \caption{\textbf{Reliability and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model.
 We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
@@
 
-\input{sections/human_subset_selection.tex}
+\paragraph{Question selection} \label{sec:subset-selection}
+
+ Several design choices were made when selecting \chembenchmini.
+Firstly, from the full dataset, we kept all the questions labeled as advanced.
+In this way, we can obtain a deeper insight into the capabilities of \glspl{llm} on advanced tasks when compared to actual chemists.
+Secondly, we sample a maximum of three questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry).
+Thirdly, we do not include any intuition questions in this subset because the intended use of \chembenchmini is to provide a fast and fair evaluation of \glspl{llm} independent of any human baseline.
+In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for \chembenchmini.
+Then, this set is divided into two subsets based on the aforementioned combinations.
+One of the question subsets allows tool use, and the other does not.
 

### 2024-10-31 src/tex/appendix.tex
prepare for latex
Former-commit-id: c6a611cc108b38d941430242ce74a65e47dc7aa3

@@
 \begin{itemize}
-    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
-    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
-    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
-    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain (example \url{https://github.com/lamalab-org/chem-bench/pull/391#discussion_r1679276714})
-    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
-    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
+    \item \textit{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
+    \item \textit{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
+    \item \textit{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
+    \item \textit{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain (example \url{https://github.com/lamalab-org/chem-bench/pull/391#discussion_r1679276714})
+    \item \textit{Out of Scope:} Questions outside the realm of chemistry are rejected.
+    \item \textit{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines.
 They should add value by expanding the breadth of the dataset.
@@
 In addition to the manual review, we also performed automated checks to ensure the quality of the questions.
-The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.
+The schemas, \LaTeX\xspace templating, and other formatting aspects are checked automatically using GitHub Actions.
 

### 2024-10-31 src/tex/ms.tex
prepare for latex
Former-commit-id: c6a611cc108b38d941430242ce74a65e47dc7aa3

@@
 These findings reveal LLMs' impressive chemical capabilities while emphasizing the need for further research to improve their safety and usefulness.
-They also suggest adapting chemistry education and developing better evaluation methods for LLMs in chemistry.
+They also suggest adapting chemistry education and show the value of benchmarking frameworks for evaluating LLMs in specific domains.
 \end{abstract}
@@
  On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
- For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
+ For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5__Sonnet_.txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
  While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for the general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.
@@
  For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
- For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
+ For example, for the questions about the safety profile of compounds, \GPTFour reported a confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
  While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
